### 2016-06-11 gruntz.tex
Fix mistakes in the Gruntz algorithm example
1. how to choose subexpression omega (this part was: wrong)
2. the rewrite stage (was: non-exsistent), I hope
now it's clear how to rewrite elements of the mrv set,
and this is not a magic step for our readers (or worse,
for "authors" of the SymPy) anymore.
3. mention the rewrite stage in comment about the
algorithm termination.
4. some minor fixes, i.e. "The goal is to calculate" was
dropped, the Gruntz algorithm IS about calculation limits
at the infinity.

SymPy calculates limits using the Gruntz algorithm, as described in%
~\cite{Gruntz1996limits}. The basic idea is as follows: any limit can be
converted to a limit $\lim\limits_{x\to\infty} f(x)$ by substitutions like
$x\to{1\over x}$. Then the subexpression $\omega$ (that converges
to zero as $x\to\infty$ faster than all other subexpressions) is identified in
$f(x)$, and $f(x)$ is expanded into a series with respect to $\omega$. Any
positive powers of $\omega$ converge to zero (while negative powers indicate
an infinite limit) and any constant term independent of
$\omega$ determines the limit. When a constant term still dependends on
$x$ the Gruntz algorithm is applied again until a final numerical value
is obtained as the limit.

To determine the most rapidly varying subexpression, the comparability classes
must first be defined, by calculating $L$:
\begin{equation}
L\equiv \lim_{x\to\infty} {\log |f(x)| \over \log |g(x)|}
\end{equation}
The relations $<$, $>$, and $\sim$ are defined as follows: $f>g$ when
$L=\pm\infty$ (it is said that $f$ is more rapidly varying than $g$, i.e., $f$
goes to $\infty$ or $0$ faster than $g$), $f<g$ when $L=0$ ($f$ is less
rapidly varying than $g$) and $f\sim g$ when $L\neq 0,\pm\infty$ (both $f$ and
$g$ are bounded from above and below by suitable integral powers of the
other). Note that if $f > g$, then $f > g^n$ for any $n$. Here
are some examples of comparability classes:
\[2 < x < e^x < e^{x^2} < e^{e^x}\]
\[2\sim 3\sim -5\]
\[x\sim x^2\sim x^3\sim {1\over x}\sim x^m\sim -x\]
\[e^x\sim e^{-x}\sim e^{2x}\sim e^{x+e^{-x}}\]
\[f(x)\sim{1\over f(x)}\]

The Gruntz algorithm is now illustrated with the following example:
\begin{equation}
    \label{gruntz_example_fn}
f(x) = e^{x+2e^{-x}} - e^x + {1\over x} \,.
\end{equation}
First, the set of most rapidly varying subexpressions is determined ---
the so-called \textit{mrv set}. For~\eqref{gruntz_example_fn}, the mrv set
$\{e^x, e^{-x}, e^{x+2e^{-x}}\}$ is obtained. These are all subexpressions of%
~\eqref{gruntz_example_fn} and they all belong to the same comparability
class. This calculation can be done using SymPy as follows:

% dict_keys output order varies
% no-doctest
\begin{verbatim}
>>> from sympy.series.gruntz import mrv
>>> mrv(exp(x+2*exp(-x))-exp(x) + 1/x, x)[0].keys()
dict_keys([exp(x + 2*exp(-x)), exp(x), exp(-x)])
\end{verbatim}

Next, an arbitrary item $\omega$ is taken from mrv set that converges to zero for
$x\to\infty$ and doesn't have subexpressions in the given mrv set.  In our case,
only item $\omega=e^{-x}$ can be accepted.  If such a term is not
present in the mrv set (i.e., all terms converge to infinity instead of zero),
the relation $f(x)\sim {1\over f(x)}$ can be used.

The next step is to rewrite the mrv set in terms of $\omega=g(x)$.  Every element
$f(x)$ of the mrv set is rewritten as $A \omega^c$, where
\begin{equation}
\label{gruntz_rewrite}
c = \lim\limits_{x\to\infty} \frac{\log{f(x)}}{\log{g(x)}},
\qquad
A = e^{\log f - c \log g}
\end{equation}
Note that this step include calculation of more simple limits, for instance
\begin{equation}
	\lim\limits_{x\to\infty} \frac{\log{e^{x + 2 e^{-x}}}}{\log e^{-x}}=
	\lim\limits_{x\to\infty} \frac{x + 2 e^{-x}}{-x} = -1
\end{equation}
In our example we obtain the rewritten mrv set: $\{{1\over\omega},
\omega, {1\over\omega}e^{2\omega}\}$. This can be done in the SymPy with
\begin{verbatim}
>>> from sympy.series.gruntz import mrv, rewrite
>>> m = mrv(exp(x+2*exp(-x))-exp(x) + 1/x, x)
>>> w = Symbol('w')
>>> rewrite(m[1], m[0], x, w)[0]
     2*w
1   e      1
- + ---- - -
x    w     w
\end{verbatim}
Then the rewritten subexpressions are
substituted back into $f(x)$ in~(\ref{gruntz_example_fn})
and the result is expanded with respect to~$\omega$:
\begin{equation}
    \label{gruntz_example_fn2}
f(x) = {1\over x}-{1\over\omega}+{1\over\omega}e^{2\omega}
     = 2+{1\over x} + 2\omega + O(\omega^2)
\end{equation}

Since $\omega$ is from the mrv set, then in the limit as $x\to\infty$,
$\omega\to0$, and so $2\omega + O(\omega^2) \to 0$ in~\eqref{gruntz_example_fn2}:
\begin{equation}
f(x) = {1\over x}-{1\over\omega}+{1\over\omega}e^{2\omega}
    = 2+{1\over x} + 2\omega + O(\omega^2)
    \to 2 + {1\over x}
\end{equation}

In our example the result ($2+{1\over x}$) still depends on $x$, so the above procedure is
repeated until just a value independent of $x$ is obtained. This is the final
limit. In the above case the limit is $2$, as can be
verified by SymPy:

\begin{verbatim}
>>> limit(exp(x+2*exp(-x))-exp(x) + 1/x, x, oo)
2
\end{verbatim}

In general, when $f(x)$ is expanded in terms of $\omega$, the following is obtained:
\begin{equation}
f(x) = \underbrace{O\left({1\over \omega^3}\right)}_\infty
    + \underbrace{C_{-2}(x)\over \omega^2}_\infty
    + \underbrace{C_{-1}(x)\over \omega}_\infty
    + {C_{0}(x)}
    + \underbrace{C_{1}(x)\omega}_0
    + \underbrace{O(\omega^2)}_0
\end{equation}
The positive powers of $\omega$ are zero. If there are any negative powers of
$\omega$, then the result of the limit is infinity, otherwise the limit is
equal to $\lim\limits_{x\to\infty} C_0(x)$. The expression $C_0(x)$ is always
simpler than original $f(x)$, same is true for limits, arising in the
rewrite stage~(\ref{gruntz_rewrite}), so the algorithm converges. A proof of this and further
details on the algorithm are given in Gruntz's PhD thesis~\cite{Gruntz1996limits}.

### 2016-06-11 supplement.tex
Remove custom modifications from wlpeerj.cls

%% Submissions for peer-review must enable line-numbering 
%% using the lineno option in the \documentclass command.
%%
%% Preprints and camera-ready submissions do not need 
%% line numbers, and should have this option removed.
%%
%% Please note that the line numbering option requires
%% version 1.1 or newer of the wlpeerj.cls file.

\documentclass[fleqn,10pt,lineno,numbers]{wlpeerj} % for journal submissions
% \documentclass[fleqn,10pt]{wlpeerj} % for preprint submissions

%\usepackage[numbers]{natbib}

\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[scaled=0.8]{DejaVuSansMono}

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage[all]{xy}
\usepackage{amsmath}
\usepackage{caption}
\graphicspath{ {images/} }

% Makes quote characters in monospace font not be curly
\usepackage{upquote}

\usepackage{amsmath}
\usepackage{url}
\usepackage{hyperref}

% this is required for all the \url{} commands in the bib file
%\usepackage{hyperref}

% for nice units
\usepackage{siunitx}

% for images: png, pdf, etc
\usepackage{graphicx}

% for nice table formatting, i.e., /toprule, /midrule, etc
\usepackage{booktabs}

% to allow for \verb++ declarations in captions.
\usepackage{cprotect}

% to allow usage of \mathbb symbols
\usepackage{amssymb}

\usepackage{longtable}

\usepackage{listings}

\begin{document}

\flushbottom
\thispagestyle{empty}%
\vskip-36pt%
{\raggedright\sffamily\bfseries\fontsize{20}{25}\selectfont SymPy: Symbolic Computing in Python\par}%
\vskip10pt
{\raggedright\sffamily\fontsize{12}{16}\selectfont  Supplementary material\par}
\vskip25pt%

As in the paper, all examples in the supplement assume that the following
has been run:
\begin{verbatim}
>>> from sympy import *
>>> x, y, z = symbols('x y z')
\end{verbatim}

\section{Limits: The Gruntz Algorithm}

\input{gruntz.tex}

% Series module (Formal Power Series, Fourier Series)
\section{Series}

\input{series.tex}

% Logic module
\section{Logic}

\input{logic}

\section{Diophantine Equations}

\input{diophantine}

% Sets
\section{Sets}
\input{sets}

\section{Category Theory}
\input{categories}

\section{SymPy Gamma}\label{sympy-gamma}

\input{gamma}

\section{SymPy Live}\label{sympy-live}

\input{live}

\section{Comparison with Mathematica}

\input{comparison_with_mma}

\section{Other Projects that use SymPy}

\input{other_projects_that_use_sympy}

\section{Tensors}

\input{tensors}

\section{Numerical simplification}

\input{nsimplify}

\section{Examples}

\input{examples}

\section{References}

\bibliography{paper}

\end{document}

### 2016-06-11 gruntz.tex
Fix issues, mentioned in the pr thread (for 89ed7f7)

@@
 Next, an arbitrary item $\omega$ is taken from mrv set that converges to zero for
-$x\to\infty$ and doesn't have subexpressions in the given mrv set.  In our case,
-only item $\omega=e^{-x}$ can be accepted.  If such a term is not
+$x\to\infty$ and doesn't have subexpressions in the given mrv set.  If such a term is not
 present in the mrv set (i.e., all terms converge to infinity instead of zero),
-the relation $f(x)\sim {1\over f(x)}$ can be used.
+the relation $f(x)\sim {1\over f(x)}$ can be used.  In the
+considered case, only item $\omega=e^{-x}$ can be accepted.
 
@@
 \end{equation}
-Note that this step include calculation of more simple limits, for instance
+Note that this step includes calculation of more simple limits, for instance
 \begin{equation}
@@
 \end{equation}
-In our example we obtain the rewritten mrv set: $\{{1\over\omega},
-\omega, {1\over\omega}e^{2\omega}\}$. This can be done in the SymPy with
+In this example we obtain the rewritten mrv set: $\{{1\over\omega},
+\omega, {1\over\omega}e^{2\omega}\}$. This can be done in SymPy with
 \begin{verbatim}
@@
 >>> rewrite(m[1], m[0], x, w)[0]
-     2*w
-1   e      1
-- + ---- - -
-x    w     w
+1/x + exp(2*w)/w - 1/w
 \end{verbatim}
@@
 
-In our example the result ($2+{1\over x}$) still depends on $x$, so the above procedure is
+In this example the result ($2+{1\over x}$) still depends on $x$, so the above procedure is
 repeated until just a value independent of $x$ is obtained. This is the final

### 2016-06-15 architecture.tex
Note the SymPy and mpmath versions used for the paper

Software architecture is of central importance in any large software
project because it establishes predictable patterns of usage and
development~\cite{Shaw1996}. This section describes the essential
structural components of SymPy, provides justifications for the design
decisions that have been made, and gives example
user-facing code as appropriate.

\subsection{Basic Usage}

% symbols, various ways to declare them

The following statement
imports all SymPy functions into the global Python namespace.
From here on, all examples in this paper assume that this statement has been
executed.\footnote{All examples in this paper use SymPy version 1.0.}

\begin{verbatim}
>>> from sympy import *
\end{verbatim}

Symbolic variables, called symbols, must be defined and assigned to
Python variables before they can be used. This is typically done through the
\texttt{symbols} function, which may create multiple symbols in a single
function call. For instance,
\begin{verbatim}
>>> x, y, z = symbols('x y z')
\end{verbatim}
creates three symbols representing variables named $x$, $y$, and $z$. In this
particular instance, these symbols are all assigned to Python variables of the
same name. However, the user is free to assign them to different
Python variables, while representing the same symbol, such as
\texttt{a, b, c = symbols(\textquotesingle{}x y z\textquotesingle{})}.
In order to minimize potential confusion, though, all examples in this paper will
assume that
the symbols \verb|x|, \verb|y|, and \verb|z| have been assigned to Python variables
identical to their symbolic names.

Expressions are created from symbols using Python's mathematical syntax.  For
instance, the following Python code creates the expression $(x^2 - 2x + 3)/y$.

\begin{verbatim}
>>> (x**2 - 2*x + 3)/y
(x**2 - 2*x + 3)/y
\end{verbatim}

Importantly, SymPy expressions are immutable. This simplifies the design of
SymPy by allowing expression interning. It also enables expressions to be
hashed and stored in Python dictionaries, thereby permitting features
such as caching.

%% I volunteer to write this section. --Aaron
%%
%% Representing symbolic expressions using Python objects

\subsection{The Core}

A computer algebra system (CAS) represents mathematical expressions as data
structures.  For example, the mathematical expression $x + y$ is represented as
a tree with three nodes, $+$, $x$, and $y$, where $x$ and $y$ are ordered
children of $+$.  As users manipulate
mathematical expressions with traditional mathematical syntax, the CAS
manipulates the underlying data structures.  Automated optimizations and
computations such as integration, simplification, etc.\ are all functions that
consume and produce expression trees.

In SymPy every symbolic expression is an instance of a Python \texttt{Basic}
class, a superclass of all SymPy types providing common methods to all SymPy
tree-elements, such as traversals.  The children of a node in the
tree are held in the \texttt{args} attribute.  A terminal or leaf node in the
expression tree has empty \texttt{args}.

For example, consider the expression $xy + 2$:
\begin{verbatim}
>>> expr = x*y + 2
\end{verbatim}
By order of operations, the parent of the expression tree for \texttt{expr} is
an addition, so it is of type \texttt{Add}. The child nodes of \texttt{expr} are
\texttt{2} and \texttt{x*y}.
\begin{verbatim}
>>> type(expr)
<class 'sympy.core.add.Add'>
>>> expr.args
(2, x*y)
\end{verbatim}

Descending further down into the expression tree yields the full expression. For
example, the next child node (given by \texttt{expr.args[0]}) is
\texttt{2}. Its class is \texttt{Integer}, and it has an empty \texttt{args}
tuple, indicating that it is a leaf node.
\begin{verbatim}
>>> expr.args[0]
2
>>> type(expr.args[0])
<class 'sympy.core.numbers.Integer'>
>>> expr.args[0].args
()
\end{verbatim}

A useful way to view an expression tree is using the \texttt{srepr} function, which
returns a string representation of an expression as valid Python code
with all the nested class constructor calls to create the given expression.
\begin{verbatim}
>>> srepr(expr)
"Add(Mul(Symbol('x'), Symbol('y')), Integer(2))"
\end{verbatim}

Every SymPy expression satisfies a key identity invariant:
\begin{verbatim}
expr.func(*expr.args) == expr
\end{verbatim}
This means that expressions are
rebuildable from their \texttt{args}.\footnote{\texttt{expr.func} is used
instead of \texttt{type(expr)} to allow the function of an expression to be
distinct from its actual Python class. In most cases the two are the same.}
Note that in SymPy the \texttt{==} operator represents exact
structural equality, not mathematical equality. This allows testing if any two
expressions are equal to one another as expression trees. For example, even
though ${(x + 1)}^2$ and $x^2 + 2x + 1$ are equal mathematically, SymPy gives
\begin{verbatim}
>>> (x + 1)**2 == x**2 + 2*x + 1
False
\end{verbatim}
because they are different as expression trees (the former is a \verb|Pow|
object and the latter is an \verb|Add| object).

Python allows classes to override mathematical operators. The Python
interpreter translates the above \texttt{x*y + 2} to, roughly,
\verb|(x.__mul__(y)).__add__(2)|. Both \texttt{x} and \texttt{y}, returned
from the \texttt{symbols} function, are \texttt{Symbol} instances. The
\texttt{2} in the expression is processed by Python as a literal, and is
stored as Python's built in \texttt{int} type. When \texttt{2} is passed to the
\verb|__add__| method of \texttt{Symbol}, it is converted to the SymPy type
\verb|Integer(2)| before being stored in the resulting expression tree. In
this way, SymPy expressions can be built in the natural way using Python
operators and numeric literals.

%%
%% Assumptions
\subsection{Assumptions}
\label{sec:assumptions}

SymPy performs logical inference through its assumptions system. The
assumptions system allows users to specify that symbols have certain common
mathematical properties, such as being positive, imaginary, or integral. SymPy
is careful to never perform simplifications on an expression unless the
assumptions allow them. For instance, the identity $\sqrt{t^2} = t$ holds if
$t$ is nonnegative ($t\ge 0$). However, for general complex $t$, no
such identity holds.

By default, SymPy performs all calculations assuming that symbols are
complex valued. This assumption makes it easier to treat mathematical problems
in full generality.
\begin{verbatim}
>>> t = Symbol('t')
>>> sqrt(t**2)
sqrt(t**2)
\end{verbatim}

By assuming the most general case, that symbols are complex by default, SymPy
avoids performing mathematically invalid operations. However, in many cases
users will wish to simplify expressions containing terms like $\sqrt{t^2}$.

Assumptions are set on \texttt{Symbol} objects when they are created. For
instance \texttt{Symbol(\textquotesingle{}t\textquotesingle{}, positive=True)} will create a symbol named
\texttt{t} that is assumed to be positive.
\begin{verbatim}
>>> t = Symbol('t', positive=True)
>>> sqrt(t**2)
t
\end{verbatim}

Some of the common assumptions that SymPy allows are \texttt{positive},
\texttt{negative}, \texttt{real}, \texttt{nonpositive}, \texttt{integer},
\texttt{prime} and \texttt{commutative}.\footnote{If $A$ and
$B$ are Symbols created with \texttt{commutative=False} then SymPy will keep
$A\cdot B$ and $B\cdot A$ distinct.} Assumptions on any object can be checked with the
\verb|is_|\texttt{\textit{assumption}} attributes, like \verb|t.is_positive|.

Assumptions are only needed to restrict a domain so that certain
simplifications can be performed. They are not required to make the domain match
the input of a function. For instance, one can create the object
$\sum_{n=0}^m f(n)$ as \verb|Sum(f(n), (n, 0, m))| without setting
\texttt{integer=True} when creating the Symbol object \texttt{n}.

The assumptions system additionally has deductive capabilities. The
assumptions use a three-valued logic using the Python built in objects
\texttt{True}, \texttt{False}, and \texttt{None}.  Note that \texttt{False}
is returned if the SymPy object doesn't or can't have the assumption.
For example, both \texttt{I.is\_real} and \texttt{I.is\_prime} return
\texttt{False} for the imaginary unit \texttt{I}.

\texttt{None} represents the
``unknown'' case. This could mean that given assumptions do not unambiguously
specify the truth of an attribute. For instance,
\texttt{Symbol(\textquotesingle{}x\textquotesingle{}, real=True).is\_positive}
will give \texttt{None} because a real symbol might be positive or negative.
The \texttt{None} could also mean that not enough is known or implemented to
compute the given fact. For instance, \texttt{(pi + E).is\_irrational} gives
\texttt{None}, because determining whether $\pi + e$ is rational or irrational
is an open problem in mathematics~\cite{lang1966introduction}.

Basic implications between the facts are used to deduce assumptions. For
instance, the assumptions system knows that being an integer implies being
rational, so \texttt{Symbol(\textquotesingle{}x\textquotesingle{}, integer=True).is\_rational} returns
\texttt{True}. Furthermore, expressions compute the assumptions on themselves
based on the assumptions of their arguments. For instance, if \texttt{x} and
\texttt{y} are both created with \texttt{positive=True}, then
\texttt{(x + y).is\_positive} will be \texttt{True} whereas
\texttt{(x - y).is\_positive} will be \texttt{None}.

%% TODO: describe how assumptions are implemented

%%
%% Extensibility
\subsection{Extensibility}

While the core of SymPy is relatively small, it has been extended to a wide variety
of domains by a broad range of contributors.  This is due in part because the
same language, Python, is used both for the internal implementation and the
external usage by users.  All of the extensibility capabilities available to
users are also utilized by SymPy itself. This eases the transition pathway from
SymPy user to SymPy developer.

The typical way to create a custom SymPy object is to subclass an existing
SymPy class, usually  \texttt{Basic}, \texttt{Expr}, or
\texttt{Function}. All SymPy classes used for expression trees\footnote{Some
  internal classes, such as those used in the polynomial module, do not follow
  this rule for efficiency reasons.} should be subclasses of the base class
\texttt{Basic}, which defines some basic methods for symbolic expression
trees. \texttt{Expr} is the subclass for mathematical expressions that can be
added and multiplied together. Instances of \texttt{Expr} typically represent
complex numbers, but may also include other ``rings'' like matrix expressions.
Not all SymPy classes are subclasses of \texttt{Expr}. For instance, logic expressions such
as \verb|And(x, y)| are subclasses of \texttt{Basic} but not of \texttt{Expr}.

The \texttt{Function} class is a subclass of \texttt{Expr} which makes it
easier to define mathematical functions called with arguments. This includes
named functions like $\sin(x)$ and $\log(x)$ as well as undefined functions
like $f(x)$. Subclasses of \texttt{Function} should define a
class method \texttt{eval}, which returns a canonical form of the function
application (usually an instance of some other class, i.e. a \texttt{Number})
or \texttt{None}, if for given arguments that function should not be
automatically evaluated.

Many SymPy functions perform various evaluations down the expression tree.
Classes define their behavior in such functions by defining a relevant
\verb|_eval_|\texttt{\textit{*}} method. For instance, an object can indicate
to the \texttt{diff} function how to take the derivative of itself by defining
the \verb|_eval_derivative(self, x)| method, which may in turn call
\texttt{diff} on its \texttt{args}. (Subclasses of \texttt{Function} should
implement \texttt{fdiff} method instead, it returns the derivative of the function
without considering the chain rule.) The most common
\verb|_eval_|\texttt{\textit{*}} methods relate to the assumptions:
\verb|_eval_is_|\texttt{\textit{assumption}} is used to deduce
\textit{assumption} on the object.

As an example of the notions presented in this section,
Listing~\ref{fig:gamma-example} presents a minimal version of the gamma
function $\Gamma(x)$ from SymPy, which evaluates itself on positive integer
arguments, has the positive and real assumptions defined, can be rewritten in
terms of factorial with \verb|gamma(x).rewrite(factorial)|, and can be
differentiated. \texttt{self.func} is used throughout
instead of referencing \texttt{gamma} explicitly so that potential subclasses
of \texttt{gamma} can reuse the methods.

\lstset{
  basicstyle=\ttfamily,
}

\begin{lstlisting}[caption={A minimal implementation of \texttt{sympy.gamma}.},label=fig:gamma-example]
from sympy import Integer, Function, floor, factorial, polygamma

class gamma(Function)
    @classmethod
    def eval(cls, arg):
        if isinstance(arg, Integer) and arg.is_positive:
            return factorial(arg - 1)

    def _eval_is_positive(self):
        x = self.args[0]
        if x.is_positive:
            return True
        elif x.is_noninteger:
            return floor(x).is_even

    def _eval_is_real(self):
        x = self.args[0]
        # noninteger means real and not integer
        if x.is_positive or x.is_noninteger:
            return True

    def _eval_rewrite_as_factorial(self, z):
        return factorial(z - 1)

    def fdiff(self, argindex=1):
        from sympy.core.function import ArgumentIndexError
        if argindex == 1:
            return self.func(self.args[0])*polygamma(0, self.args[0])
        else:
            raise ArgumentIndexError(self, argindex)
\end{lstlisting}
The gamma function implemented in SymPy has many more capabilities than the
above listing, such as evaluation at rational points and series expansion.

### 2016-06-15 numerics.tex

Floating point numbers in SymPy are implemented by the \texttt{Float} class,
which represents an arbitrary-precision binary floating-point number by
storing its value and precision (in bits). This representation is distinct
from the Python built-in \texttt{float} type, which is a wrapper around
machine \texttt{double} types and uses a fixed precision (53-bit).

Because Python \texttt{float} literals are limited in precision, strings
should be used to input precise decimal values:
\begin{verbatim}
>>> Float(1.1)
1.10000000000000
>>> Float(1.1, 30)   # precision equivalent to 30 digits
1.10000000000000008881784197001
>>> Float("1.1", 30)
1.10000000000000000000000000000
\end{verbatim}
The \texttt{evalf} method converts a constant symbolic expression to a
\texttt{Float} with the specified precision, here 25 digits:
\begin{verbatim}
>>> (pi + 1).evalf(25)
4.141592653589793238462643
\end{verbatim}
\texttt{Float} numbers do not track their accuracy,
and should be used with caution within symbolic expressions
since familiar dangers of floating-point arithmetic apply~\cite{goldberg1991every}.
A notorious case is that of catastrophic cancellation:
\begin{verbatim}
>>> cos(exp(-100)).evalf(25) - 1
0
\end{verbatim}
Applying the \texttt{evalf} method to the whole expression solves
this problem. Internally, \texttt{evalf} estimates the number of accurate
bits of the floating-point
approximation for each sub-expression, and adaptively increases the
working precision until the estimated accuracy of the
final result matches the sought number of decimal digits:
\begin{verbatim}
>>> (cos(exp(-100)) - 1).evalf(25)
-6.919482633683687653243407e-88
\end{verbatim}
The \texttt{evalf} method works with complex numbers and supports
more complicated expressions, such as
special functions, infinite series, and integrals.
The internal error tracking does not provide rigorous error bounds
(in the sense of interval arithmetic) and cannot be used to accurately track
uncertainty in measurement data;
the sole purpose is to mitigate loss of accuracy that typically occurs
when converting symbolic expressions to numerical values.

\subsection{The mpmath library}

The implementation of arbitrary-precision floating-point arithmetic is
supplied by the mpmath library. Originally, it was developed as a SymPy
module but has subsequently been moved to a standalone pure-Python package.
The basic datatypes in mpmath are \texttt{mpf} and \texttt{mpc}, which
respectively act as multiprecision substitutes for Python's \texttt{float} and
\texttt{complex}. The floating-point precision is controlled by a global
context:\footnote{All examples in this section use mpmath version 0.19.}

% doctest printer doesn't display "mpf"
% no-doctest
\begin{verbatim}
>>> import mpmath
>>> mpmath.mp.dps = 30    # 30 digits of precision
>>> mpmath.mpf("0.1") + mpmath.exp(-50)
mpf('0.100000000000000000000192874984794')
>>> print(_)   # pretty-printed
0.100000000000000000000192874985
\end{verbatim}

For pure numerical computing, it is convenient to use mpmath directly
with \texttt{from mpmath import *}.  Nevertheless, it is best to avoid such an
import statement when using SymPy simultaneously, since the names of numerical
functions such as \texttt{exp} will collide the symbolic counterparts
in SymPy.

Like SymPy, mpmath is a pure Python library.
Internally, mpmath represents a floating-point number
${(-1)}^s x \cdot 2^y$ by a tuple $(s, x, y, b)$ where
$x$ and $y$ are arbitrary-size Python integers
and the redundant integer $b$ stores the bit length of $x$ for quick access.
If GMPY~\cite{GMPY} is installed, mpmath automatically uses
the \texttt{gmpy.mpz} type for~$x$, and GMPY methods
for rounding-related operations, improving performance.

The mpmath library supports
special functions, root-finding, linear algebra, polynomial approximation,
and numerical computation of limits, derivatives, integrals, infinite
series, and ODE solutions. All features work in arbitrary precision
and use algorithms that allow computing hundreds of digits rapidly
(except in degenerate cases).

The double exponential (tanh-sinh) quadrature is used for numerical
integration by default. For smooth integrands, this algorithm usually
converges extremely rapidly, even when the integration interval is infinite
or singularities are present at the endpoints~\cite{takahasi1974double,bailey2005comparison}.
However, for good performance, singularities
in the middle of the interval must be specified
by the user.
To evaluate slowly converging limits and infinite series, mpmath
automatically tries Richardson extrapolation and the
Shanks transformation
(Euler-Maclaurin summation can also be used)~\cite{BenderOrszag1999}.
A function to evaluate oscillatory integrals by means of convergence
acceleration is also available.

A wide array of higher mathematical functions are implemented
with full support for complex values of all parameters and arguments,
including complete and incomplete gamma functions,
Bessel functions, orthogonal polynomials, elliptic functions and integrals,
zeta and polylogarithm functions,
the generalized hypergeometric function, and the Meijer G-function.
The Meijer G-function instance
$G_{1, 3}^{3, 0}\left(0 ; \tfrac{1}{2}, -1, - \tfrac{3}{2} | x \right)$
is a good test case~\cite{Toth2007}; past versions of both Maple and
Mathematica produced incorrect numerical values for large $x > 0$.
Here, mpmath automatically removes an internal singularity
and compensates for cancellations (amounting to 656 bits
of precision when $x = 10000$), giving correct values:
% doctest printer doesn't display "mpf"
% no-doctest
\begin{verbatim}
>>> mpmath.mp.dps = 15
>>> mpmath.meijerg([[],[0]],[[-0.5,-1,-1.5],[]],10000)
mpf('2.4392576907199564e-94')
\end{verbatim}

Equivalently, with SymPy's interface this function can be evaluated as:
\begin{verbatim}
>>> meijerg([[],[0]],[[-S(1)/2,-1,-S(3)/2],[]],10000).evalf()
2.43925769071996e-94
\end{verbatim}

Symbolic integration and summation often produces hypergeometric
and Meijer G-function closed forms (see Subsection~\ref{sec:calculus});
numerical evaluation of such special functions is a useful complement
to direct numerical integration and summation.

### 2016-08-04 rebuttal.tex
Draft for rebuttal letter (style proposition)

\documentclass[answers,12pt]{exam}
\usepackage{xcolor}

\unframedsolutions
\shadedsolutions
\renewcommand{\solutiontitle}{}

\begin{document}

Dear Editor, \bigskip

Please read below our answer about the
questions made by you and reviewers. I hope that you agree
with all our comments. \bigskip

Best regards,\bigskip

The authors

\pagebreak

\section{Editor}

\begin{questions}
\question Spell foo as bar
\begin{solution}
 We opted to follow the ABC style of ...
\end{solution}
\question  Include reference to Really, Important and Stuff (2010)
\begin{solution}
Section 2 has been updated and the discussion ...
\end{solution}
\end{questions}

\section{Reviewer 1}

\begin{questions}
\question one
\begin{solution}
 We opted to follow the ABC style of ...
\end{solution}
\question  two
\begin{solution}
Section 2 has been updated and the discussion ...
\end{solution}
\end{questions}

\section{Reviewer 2}

\begin{questions}
\question one
\begin{solution}
 We opted to follow the ABC style of ...
\end{solution}
\question  two
\begin{solution}
Section 2 has been updated and the discussion ...
\end{solution}
\end{questions}

\section{Reviewer 3}

\begin{questions}
\question one
\begin{solution}
 We opted to follow the ABC style of ...
\end{solution}
\question  two
\begin{solution}
Section 2 has been updated and the discussion ...
\end{solution}
\end{questions}

\end{document}

### 2016-08-04 architecture.tex
Clarify commutative assumption
This is point 1 of the reviewer 1.

Now it should be clear, that symbols assumed
to be commutative, unless _both_ specified to
be non-commutative.

Not sure how much else we should add about the
rest of this point.  I.e. how to declare commutators.

@@
 \texttt{negative}, \texttt{real}, \texttt{nonpositive}, \texttt{integer},
-\texttt{prime} and \texttt{commutative}.\footnote{If $A$ and
-$B$ are Symbols created with \texttt{commutative=False} then SymPy will keep
-$A\cdot B$ and $B\cdot A$ distinct.} Assumptions on any object can be checked with the
-\verb|is_|\texttt{\textit{assumption}} attributes, like \verb|t.is_positive|.
+\texttt{prime} and \texttt{commutative}. Assumptions on any object can
+be checked with the \verb|is_|\texttt{\textit{assumption}} attributes,
+like \verb|t.is_positive|.
 
@@
 assumptions use a three-valued logic using the Python built in objects
-\texttt{True}, \texttt{False}, and \texttt{None}.  Note that \texttt{False}
+\texttt{True}, \texttt{False}, and \texttt{None}.\footnote{Special
+case is the \texttt{commutative} assumption: it can't be \texttt{None}.
+Unless a Symbol is created with \texttt{commutative=False}, it's
+assumed to commute with any other object with respect
+to the multiplication operation.} Note that \texttt{False}
 is returned if the SymPy object doesn't or can't have the assumption.

### 2016-08-04 architecture.tex
Warn about using wildcard imports
This part was removed from mpmath section and added
to the introduction, as footnote.

This address point 2 of the reviewer 2.

@@
 The following statement
-imports all SymPy functions into the global Python namespace.
+imports all SymPy functions into the global Python namespace.\footnote{
+Nevertheless, it is best to avoid such wildcard import statements
+in production code, as they make it unclear which names are present
+in the namespace and injected names could clash with already existing
+imports from another package.  For example, both SymPy and standard
+Python \texttt{math} library define the \texttt{exp} function.}
 From here on, all examples in this paper assume that this statement has been

### 2016-08-04 numerics.tex
Warn about using wildcard imports
This part was removed from mpmath section and added
to the introduction, as footnote.

This address point 2 of the reviewer 2.

@@
 
-For pure numerical computing, it is convenient to use mpmath directly
-with \texttt{from mpmath import *}.  Nevertheless, it is best to avoid such an
-import statement when using SymPy simultaneously, since the names of numerical
-functions such as \texttt{exp} will collide the symbolic counterparts
-in SymPy.
-
 Like SymPy, mpmath is a pure Python library.

### 2016-08-04 architecture.tex
Add a note about dotprint
I hope, this addresses 3rd point of the
reviewer 2 (Lines 151-153).

@@
 A useful way to view an expression tree is using the \texttt{srepr} function, which
-returns a string representation of an expression as valid Python code
+returns a string representation of an expression as valid Python code\footnote{
+The \texttt{dotprint} function from the \texttt{sympy.printing.dot} package
+prints output to dot format, which can be rendered with Graphviz to
+visualize expression trees graphically.}
 with all the nested class constructor calls to create the given expression.

### 2016-08-04 architecture.tex
Minor rewrite of Line 216, suggested by reviewer 2
This address point 5 of review 2.
This footnote is not about "expression trees"
Perhaps, this address point 6 of reviewer 2 (the phrase 'Expression
tree' is cited but this is not the first time you've used it.)

@@
 While the core of SymPy is relatively small, it has been extended to a wide variety
-of domains by a broad range of contributors.  This is due in part because the
-same language, Python, is used both for the internal implementation and the
-external usage by users.  All of the extensibility capabilities available to
+of domains by a broad range of contributors.
+This is due, in part, to the fact that the same language, Python,
+is used both for the internal implementation and the external usage by users.
+All of the extensibility capabilities available to
 users are also utilized by SymPy itself. This eases the transition pathway from
@@
 SymPy class, usually  \texttt{Basic}, \texttt{Expr}, or
-\texttt{Function}. All SymPy classes used for expression trees\footnote{Some
+\texttt{Function}. All SymPy classes\footnote{Some
   internal classes, such as those used in the polynomial module, do not follow
-  this rule for efficiency reasons.} should be subclasses of the base class
+  this rule for efficiency reasons.} used for expression trees should be subclasses of the base class
 \texttt{Basic}, which defines some basic methods for symbolic expression

### 2016-08-05 simplification.tex

% polynomial expressions

% functions

% expand( ), factor( ), collect( ), together( ), apart( )
%% maybe a table best suits this part.

% simplification: simplify, sqrt denest, fu, trigsimp

The generic way to simplify an expression is by calling the \texttt{simplify}
function.
It must be emphasized that simplification is not a rigorously defined
mathematical operation~\cite{Carette2004understanding}.
The \texttt{simplify} function applies several simplification routines along
with heuristics to make the output expression ``simple''.\footnote{The
\texttt{measure} parameter of the \texttt{simplify}
function lets the user specify the Python function used to determine how complex an
expression is.  The default measure function returns the total number of
operations in the expression.}

It is often preferable to apply more directed simplification functions. These
apply very specific rules to the input expression and are typically able to make
guarantees about the output. For instance, the \texttt{factor} function,
given a polynomial with rational coefficients in several variables,
is guaranteed to
produce a factorization into irreducible factors. Table~\ref{simplify-table}
lists common simplification functions.

\begin{longtable}[htbc]{lp{0.83\linewidth}}
\caption{Some SymPy Simplification Functions\label{simplify-table}}\\
\toprule
\verb|expand| & expand the expression \\
\verb|factor| & factor a polynomial into irreducibles \\
\verb|collect| & collect polynomial coefficients \\
\verb|cancel| & rewrite a rational function as $p/q$ with common factors
canceled \\
\verb|apart| & compute the partial fraction decomposition of a rational function
\\
\verb|trigsimp| & simplify trigonometric expressions~\cite{fu2006automated} \\
\verb|hyperexpand| & expand hypergeometric functions~\cite{Roach1996hyper,roach1997meijerg} \\
\bottomrule
\end{longtable}

### 2016-08-09 rebuttal.tex
Add editor comments to the rebuttal

@@
 \usepackage{xcolor}
+\usepackage{hyperref}
 
@@
 
-Dear Editor, \bigskip
+Dear Editors, \bigskip
 
@@
 
-\section{Editor}
+\section{Comments from Editor}
 
 \begin{questions}
-\question Spell foo as bar
+\question I found the \url{http://live.sympy.org/} site very convenient for trying things
+out in SymPy, but it is only mentioned in the supplement, which might be
+overlooked by readers. I suggest moving that section into the main paper, as
+it provides a great way to play along with the examples while reading the
+paper.
 \begin{solution}
- We opted to follow the ABC style of ...
+  % TODO
 \end{solution}
-\question  Include reference to Really, Important and Stuff (2010)
+\question The Basic Usage section omits what I think is an important point:
+how does one distinguish between evaluating exp(1) in Python and exp(1) in
+SymPy? In other words, how are symbolic constants specified?
 \begin{solution}
-Section 2 has been updated and the discussion ...
+  % TODO. I'm unclear what the editor is asking here.
+\end{solution}
+\question The first
+thing I tried after seeing Table 2 (simplication functions) did not work:
+
+\begin{verbatim}
+>>> trigsimp (exp( Matrix(2, 2, [0, -y, y, 0])) )
+\end{verbatim}
+fails to recognize cos and
+sin. Am I expecting too much?
+\begin{solution}
+  This indeed does not work. The implementation in \verb|trigsimp| primarily
+  thus far has focused on transforming trigonometric functions into other
+  trigonometric functions. The ability to simplify complex exponentials is
+  something that we would like to work.
+  \href{https://github.com/sympy/sympy/issues/11459}{Issue 11459} in our
+  public issue tracker tracks this feature.
+\end{solution}
+\question Section 3.6: are eigenvalues and singular
+values included? If not, why not?
+\begin{solution}
+  % TODO
+\end{solution}
+\question Like one of the referees, I expected to
+see MPFR mentioned in Section 4.1, at least to mention the pros and cons and
+say why it isn’t used.
+\begin{solution}
+  % TODO
+\end{solution}
+\question Section 4.1: please state whether the syntax for
+functions in mpmath is identical, or not, to that in Sympy.
+\begin{solution}
+  % TODO
+\end{solution}
+\question Unless I have
+missed something, a weakness of the paper is that speed is mentioned only
+twice, on lines 496 and 666. My assumption is that SymPy is slow compared with
+its commercial competitors. Please comment on the speed issue, and not just in
+the conclusions.
+\begin{solution}
+  % TODO
+\end{solution}
+\question Editors are needed for [20].
+\begin{solution}
+  % TODO: I do not know how to make bibtex (with apalike) show the editor
+  % here.
 \end{solution}

### 2016-08-09 rebuttal.tex
subsubsection -> subsection

@@
 
-\section{Reviewer 1}
+\section{Reviewer 1 (James Davenport)}
+\subsection{Basic reporting}
+\begin{questions}
+\question Generally good. A part that confused me is the assertion (footnote 3) that "If
+A and B are Symbols created with commutative=False then SymPy will keep A·B
+and B·A distinct." Does that mean that BOTH of them must be created this way,
+and that A and x (if x is created normally) will commute? Is there any way to
+declare commutators? How does one guarantee that other pieces of code, e.g.
+Gaussian elimination, respect non-commutativity?
+\begin{solution}
+  % TODO.
+\end{solution}
 
+\question References reasonable, though [8], an excellent reference, makes almost
+precisely the opposite point to that for which it is cited --- ``simplification
+is not well-defined''. I think the authors are trying to follow [8]'s
+definition of simplification.
+\begin{solution}
+  % TODO. I agree that [8] is a bad reference, because we are not actually
+  % using their definition. The aim was just to have a reference for
+  % "simplification is not well-defined", but that paper aims to define it.
+  % What is a better reference here? Perhaps Moses [19] from the Carette
+  % paper? Perhaps we can remove the reference (the referenced statement is
+  % "it must be emphasized that simplification is not a rigorously defined
+  % mathematical operation").
+\end{solution}
+\end{questions}
+
+\subsection{Experimental design}
 \begin{questions}
-\question one
+\question The referee is asked to comment whether the research is withinthe scope of the
+journal, defined as "PeerJ is an Open Access, peer-reviewed, scholarly
+journal. It considers articles in the Biological Sciences, Medical Sciences,
+and Health Sciences. PeerJ does not publish in the Physical Sciences, the
+Mathematical Sciences, the Social Sciences, or the Humanities (except where
+articles in those areas have clear applicability to the core areas of
+Biological, Medical or Health sciences)."
+
+Hence I fear a software description on the maths/cmputing boundary with no
+cited applications is out of scope.
+
 \begin{solution}
- We opted to follow the ABC style of ...
+  The editor has confirmed that the paper is in scope for PeerJ Computer Science.
 \end{solution}
-\question  two
+\end{questions}
+
+\subsection{Validity of the findings}
+\begin{questions}
+\question There are no findings as such, since this is a software description, not an
+experimental paper.
+\end{questions}
+
+\subsection{Comments for the author}
+\begin{questions}
+\question Nice paper - pity it seems totally out of scope. Why not J. Symbolic
+Computation or some such?
 \begin{solution}
-Section 2 has been updated and the discussion ...
+  See above.
 \end{solution}

### 2016-08-09 rebuttal.tex
Add comments from reviewer 2 to the rebuttal

@@
 
-\section{Reviewer 2}
+\section{Reviewer 2 (Michael Croucher)}
+
+\subsection{Basic reporting}
+No Comments.
+
+\subsection{Experimental design}
+No Comments
+
+\subsection{Validity of the findings}
+\begin{questions}
+\question There is a minor problem with the supplementary notebook
+
+The final cell is
+
+\begin{verbatim}
+>>> circuit_plot(fourier, nqubits=3);
+plt.savefig('./images/circuitplot-qft.pdf', format='pdf')
+\end{verbatim}
+
+This fails if the user does not have an images folder.
+\begin{solution}
+We have added
+
+\begin{verbatim}
+%mkdir -p './images'
+\end{verbatim}
+
+to the cell before the call to \verb|savefig|.
+\end{solution}
+
+\question I suggest removing the >>> before each line. It looks strange in a notebook.
+\begin{solution}
+We have done this.
+\end{solution}
+\end{questions}
+\subsection{Comments for the author}
+Sympy is a superb package and I am happy to see that there is now a paper that describes its current state. Here are a few comments that you may wish to consider before publication.
 
 \begin{questions}
-\question one
+\question Line 105
+It is generally frowned upon to import all symbols from a Python module in this manner.
+I understand why you are doing it in this paper, it makes subsequent sympy commands less verbose.
+It might, however, encourage bad practice. It may also lead to a poor user experience for newbies.
+
+For example, say I had the following code
+
+\begin{verbatim}
+𝔬mνmpyimportsin,arraytest=array(123)sin(test)
+\end{verbatim}
+[sic]
+
+I decide that I want to use sympy for something and, following your example, do
+
+\begin{verbatim}
+𝔬mνmpyimportsin,array𝔬msympyimport⋅test=array(123)sin(test)x,y,z=symbols(′xyz′)
+\end{verbatim}
+[sic]
+
+The code will break because sympy has it's own sin that gets imported that doesn't work on numpy arrays. As a newbie, I might not know this.
+I tested using Sympy 1.0 and Python 3.
+
 \begin{solution}
- We opted to follow the ABC style of ...
+
 \end{solution}
-\question  two
+\question Later in the text, a similar assertion is made by you (line 485) in reference to a different package that may break sympy.
+
+\question Lines 151-153
+Minor comment: srepr is a useful command.
+As someone who also uses Mathematica, I wonder if there is a sympy version of the TreeForm command which produces a visualisation of the expression tree, or maybe an output format that I could pass to a graph library for visualisation?
+
 \begin{solution}
-Section 2 has been updated and the discussion ...
+
+\end{solution}
+\question Section 2.3 - Assumptions
+Is there a way of the user listing all available assumptions?
+
+\begin{solution}
+
+\end{solution}
+\question Section 2.4
+Line 216 - This is due in part because the same language, Python, is used both for the internal implementation and the external usage by users.
+
+This reads badly. Perhaps the following might be better?
+
+This is due, in part, to the fact that the same language, Python, is used both for the internal implementation and the external usage by users.
+
+\begin{solution}
+
+\end{solution}
+\question Line 221:
+the phrase 'Expression tree' is cited but this is not the first time you've used it. Perhaps cite earlier? Line 129 perhaps?
+
+\begin{solution}
+
+\end{solution}
+\question Page 8: Footnote 5
+The line reads
+The measure parameter of the simplify function lets specify the Python function used to determine how
+
+should this be..
+
+The measure parameter of the simplify function lets the user specify the Python function used to determine how
+
+\begin{solution}
+
+\end{solution}
+\question Section 4.1
+Should mpmath be cited here?
+
+\begin{solution}
+
 \end{solution}

### 2016-08-09 rebuttal.tex
Modify the import * footnote, and note it in the rebuttal

@@
 \begin{solution}
+We have used \verb|import *| on purpose, as we felt that explicitly importing
+names would unnecessarily distract from the content of the paper, but the
+reviewer is absolutely right that this is bad practice for actual code. We
+have added a footnote:
 
+\begin{quote}\texttt{import *} has been used here to aid the
+  readability of the paper, but is best to avoid such wildcard import
+  statements in production code, as they make it unclear which names are
+  present in the namespace. Furthermore, imported names could clash with
+  already existing imports from another package. For example, SymPy, the
+  standard Python \texttt{math} library, and NumPy all define the \texttt{exp}
+  function, but only the SymPy one will work with SymPy symbolic expressions.
+\end{quote}
 \end{solution}
@@
 The line reads
+
 The measure parameter of the simplify function lets specify the Python function used to determine how

### 2016-08-09 architecture.tex
Modify the import * footnote, and note it in the rebuttal

@@
 
-The following statement
-imports all SymPy functions into the global Python namespace.\footnote{
-Nevertheless, it is best to avoid such wildcard import statements
-in production code, as they make it unclear which names are present
-in the namespace and injected names could clash with already existing
-imports from another package.  For example, both SymPy and standard
-Python \texttt{math} library define the \texttt{exp} function.}
+The following statement imports all SymPy functions into the global Python
+namespace.\footnote{\texttt{import *} has been used here to aid the
+  readability of the paper, but is best to avoid such wildcard import
+  statements in production code, as they make it unclear which names are
+  present in the namespace. Furthermore, imported names could clash with
+  already existing imports from another package. For example, SymPy, the
+  standard Python \texttt{math} library, and NumPy all define the \texttt{exp}
+  function, but only the SymPy one will work with SymPy symbolic expressions.}
 From here on, all examples in this paper assume that this statement has been

### 2016-08-09 rebuttal.tex
More on import * in the rebuttal

@@
 
+Later in the text, a similar assertion is made by you (line 485) in reference to a different package that may break sympy.
+
 \begin{quote}\texttt{import *} has been used here to aid the
@@
 \end{quote}
-\end{solution}
-\question Later in the text, a similar assertion is made by you (line 485) in reference to a different package that may break sympy.
+
+We have also removed the reference to \texttt{from mpmath import *} from line 485.
 

### 2016-08-10 rebuttal.tex
Fix rebuttal LaTeX
Add note about dotprint to rebuttal letter
Add note about assumptions to the rebuttal
We still need to potentially add something to the paper for this.

@@
 \usepackage{hyperref}
+\usepackage{upquote}
+\usepackage{textcomp}
 
@@
 We have also removed the reference to \texttt{from mpmath import *} from line 485.
+\end{solution}
 
@@
 \begin{solution}
+We have added a footnote about the \texttt{dotprint} function (footnote 3),
+which outputs expressions in the dot format that can be rendered with Graphviz.
+\end{solution}
 
-\end{solution}
 \question Section 2.3 - Assumptions
@@
 \begin{solution}
-
+There is no simple way to access this in code (the simplest way is
+\texttt{sys.\allowbreak{}modules[\allowbreak{}\textquotesingle{}sympy.core.assumptions\textquotesingle{}].\allowbreak{}\_assume\_defined}, which works
+around some technical issues with Python's import system, and uses a
+private method). The recommended list is in the SymPy documentation, at
+\url{http://docs.sympy.org/latest/modules/core.html#module-sympy.core.assumptions}.
 \end{solution}

### 2016-08-10 rebuttal.tex
fix code examples from reviewer 1

@@
 see MPFR mentioned in Section 4.1, at least to mention the pros and cons and
-say why it isn’t used.
+say why it isn't used.
 \begin{solution}
@@
 \begin{solution}
-  % TODO.
+  We have clarified the footnote. Both expressions must be set as
+  \texttt{commutative=False}. Internally, in \texttt{Mul}, the ``commutative
+  part'' of an expression is pulled out to the front and canonically ordered,
+  and the ``noncommutative part'' is not reordered.
+
+Commutators are implemented in the \texttt{sympy.physics.quantum} module and
+are discussed in section 5.2.
+
+Different algorithms do generally require special consideration for
+noncommutative expressions to be correct. If an algorithm implicitly assumes
+that expressions commute, it may return incorrect results when given a
+noncommutative expression.
 \end{solution}
@@
 
-\question I suggest removing the >>> before each line. It looks strange in a notebook.
+\question I suggest removing the \verb|>>>| before each line. It looks strange in a notebook.
 \begin{solution}
@@
 It might, however, encourage bad practice. It may also lead to a poor user experience for newbies.
+\label{rev1point1}
 
@@
 \begin{verbatim}
-𝔬mνmpyimportsin,arraytest=array(123)sin(test)
+from numpy import sin, array
+test=array(123)
+sin(test)
 \end{verbatim}
-[sic]
 
@@
 \begin{verbatim}
-𝔬mνmpyimportsin,array𝔬msympyimport⋅test=array(123)sin(test)x,y,z=symbols(′xyz′)
+from numpy import sin, array
+from sympy import *
+x,y,z=symbols(′xyz′)
+test=array(123)
+sin(test)
 \end{verbatim}
-[sic]
 
@@
 I tested using Sympy 1.0 and Python 3.
+
+Later in the text, a similar assertion is made by you (line 485) in reference to a different package that may break sympy.
 
@@
 have added a footnote:
-
-Later in the text, a similar assertion is made by you (line 485) in reference to a different package that may break sympy.
 
@@
 
-\question Lines 151-153
+\question Lines 151--153
 Minor comment: srepr is a useful command.
@@
 
-\question Section 2.3 - Assumptions
+\question Section 2.3: Assumptions
 Is there a way of the user listing all available assumptions?
@@
 \question Section 2.4
-Line 216 - This is due in part because the same language, Python, is used both for the internal implementation and the external usage by users.
+Line 216: This is due in part because the same language, Python, is used both for the internal implementation and the external usage by users.
 
@@
 \begin{solution}
-
+We have changed the sentence as suggested.
 \end{solution}
 \question Line 221:
-the phrase 'Expression tree' is cited but this is not the first time you've used it. Perhaps cite earlier? Line 129 perhaps?
+the phrase `Expression tree' is cited but this is not the first time you've used it. Perhaps cite earlier? Line 129 perhaps?
 
 \begin{solution}
-
+The footnote here was in reference to the SymPy classes. We have moved the
+footnote later in the sentence, to ``\texttt{Basic}'', as suggested
+by reviewer 3, point~\ref{rev3point16}.
 \end{solution}
@@
 \begin{solution}
-
+We have changed the sentence as suggested.
 \end{solution}
@@
 \begin{solution}
-
+We have added a citation for mpmath.
 \end{solution}
@@
 
-\section{Reviewer 3}
+\section{Reviewer 3 (Vedran Sego)}
 
 \begin{questions}
-\question one
+
+\question Explaining what Python is (lines 70--72) should go before talking about SymPy as a ``CAS written in Python''.
+Further, the paper assumes a moderate familiarity with Python (for example, Python's console, OOP, and exceptions), and this should be specified.
+There should be a short note on the used Python console (\texttt{>>>} is the prompt, with the results of computation following immediately in the lines after it).
+The citation [25] from line 65 should be moved next to ``Python'' in line 70.
 \begin{solution}
- We opted to follow the ABC style of ...
+  % TODO: It seems better to start the paper with "SymPy is" rather than
+  % "Python is". The reviewer perhaps wants us to not mention Python until the
+  % second paragraph, but Python is important enough to SymPy's design that it
+  % warrants mentioning in the first sentence, in my opinion. --Aaron
 \end{solution}
-\question  two
+
+\question Line 73 has outdated information.
+Sage was renamed to SageMath and it no longer aims only at pure mathematics but also at algebra, numerical analysis, etc.
+The reference [40] should be replaced by a more up-to-date one.
 \begin{solution}
-Section 2 has been updated and the discussion ...
+
 \end{solution}
+
+\question The plural ``CASs'' is usually written as ``CASes'' or ``CAS's'', with the latter being somewhat problematic due to it looking like it implies possession.
+\begin{solution}
+  % TODO: I recall we looked this up and the spelling consensus agreed with
+  % the way we have it. We should double check (perhaps we can email the
+  % editor to ask if the journal has a convention).
+\end{solution}
+
+\question Line 88 mentions ``printers'', but it doesn't state what they are, which is confusing for those readers that are yet to learn the concept in section 3.4.
+\begin{solution}
+We have changed ``printers'' to ``display formatters'', which is the
+terminology used by Jupyter.
+\end{solution}
+
+\question Also in line 88, Jupyter's citation [30] is actually about IPython and should be replaced by a more up-to-date version.
+\begin{solution}
+We have changed the citation to
+
+Kluyver, T., Ragan-Kelley, B., P{\'e}rez, F.,
+Granger, B., Bussonnier, M., Frederic, J., Kelley,
+K., Hamrick, J., Grout,
+J., Corlay, S., et al. (2016). Jupyter notebooks---a publishing format for
+reproducible computational workflows. In \textit{Positioning and Power in Academic
+Publishing: Players, Agents and Agendas: Proceedings of the 20th International
+Conference on Electronic Publishing}, page 87. IOS Press.
+
+as recommended by the Jupyter developers at \url{https://github.com/jupyter/jupyter/issues/190}.
+\end{solution}
+
+\question The word ``software'' in line 91 is ambiguous; ``library'' or ``package'' would make a better choice.
+\begin{solution}
+We have changed ``software'' to ``library''.
+\end{solution}
+
+\question Lines 91--96: ``we discuss/look at/etc'' is the preferred form, instead of ``section discusses/looks at/etc''.
+\begin{solution}
+  % TODO: We had decided not to use first person. We should perhaps email the editor
+  % to ask about this.
+\end{solution}
+
+\question The paragraph in lines 103--105 should be moved to the introduction,
+and the footnote from line 104 should be added to that paragraph as a
+full-blown sentence, expanded by all the relevant technical information
+(Python version, OS, \ldots). Given that the end of life for Python 2 is
+2020., a comment on whether all the presented examples work in both Python 2
+and 3 should be included as well. Further, emphasise that wildcard imports,
+``\texttt{import *}'', should almost never be used in programs (see PEP 8, the
+item ``wildcard imports''). The same goes for the \texttt{import} mentioned in lines 484--487.
+\begin{solution}
+We have moved the footnote inline, and have noted that the examples should
+work with Python 2.7, 3.2, 3.3, 3.4, or 3.5, with any operating system that
+supports Python.
+
+We have added a footnote discussing the concerns with \texttt{import *} (see
+reviewer 3, point~\ref{rev1point1}).
+\end{solution}
+
+\question Line 119 should be removed, as it is basically a copy of the previous line.
+\begin{solution}
+Line 119 is the output of line 118, so now showing it would
+be incorrect.
+
+The aim here was simply to show that the input expression remains unevaluated.
+We have added a sentence before the example to note this.
+This particular example was chosen because it shows the basic syntax for
+addition, subtraction, multiplication, division, and exponentiation.
+\end{solution}
+
+\question In line 121 the word ``stored'' should be replaced by ``used as keys''.
+\begin{solution}
+We have made the suggested change.
+\end{solution}
+
+\question What do the authors mean by ``thereby permitting features such as caching'' in line 122? Caching can be done for mutable types as well, just not through hashing.
+\begin{solution}
+
+\end{solution}
+
+\question There is no need to repeat ``(CAS)'' in line 124, as it was already given in line 64.
+\begin{solution}
+
+\end{solution}
+
+\question In the same line, the word ``represents'' should be replaced by ``stores''.
+\begin{solution}
+  % I don't personally agree with this, but it's a minor point. --Aaron
+\end{solution}
+
+\question In line 184, ``symbols are'' should be replaced by ``\texttt{t} is'' (the general rule is already given in line 179).
+\begin{solution}
+
+\end{solution}
+
+\question The code in line 210 should be made into its own line (like a displaymath formula), for typesetting reasons and better readability.
+\begin{solution}
+
+\end{solution}
+
+\question The footnote 4 should be moved from line 221 to line 222, right after ``Basic''.
+\label{rev3point16}
+\begin{solution}
+We have made the suggested change.
+\end{solution}
+
+\question The part ``which defines some basic methods for symbolic expression trees'' should be removed from line 222, as it was already given in line 130.
+\begin{solution}
+
+\end{solution}
+
+\question In line 225, the sentence ``Not all SymPy classes are subclasses of \texttt{Expr}.'' sounds confusing as a reader new to SymPy wouldn't expect, for example, symbols to inherit ``Expr''.
+It would be better to expand this, for example ``Most of the SymPy classes (including \texttt{Symbol}) are subclasses of \texttt{Expr}, but there are exceptions to this rule''.
+\begin{solution}
+
+\end{solution}
+
+\question The title ``Features'' in line 276 is ambiguous, as a ``feature''
+has no precise meaning in Python (or even software libraries in general). It
+should be replaced by ``Packages and modules'' or a similar more precise
+wording. The same goes for ``feature'' in most other places in the paper (for
+example, the caption of Table 1.\ and line 495). It would be very useful to also include actual names of the packages/modules in Table 1., as well as in any section covering those packages/modules.
+\begin{solution}
+
+\end{solution}
+
+\question The sets support listing is unnaturally split in two by the ``This includes\ldots'' sentence which would fit better in parentheses.
+\begin{solution}
+
+\end{solution}
+
+\question Line 355, add a sentence explaining that in SymPy \texttt{str == repr}, because in Python \texttt{repr} is used to get an unambiguous valid Python code representation, while the return value of \texttt{str} is meant to be human-readable.
+\begin{solution}
+
+\end{solution}
+
+\question Lines 359 and 379: what is 2D text representation? It seems that ``2D'' shouldn't be here.
+\begin{solution}
+
+\end{solution}
+
+\question Line 427: every dictionary is a ``dictionary of keys''. This should be a dictionary with coordinate tuples as keys associated with the appropriate values.
+\begin{solution}
+
+\end{solution}
+
+\question Section 4 would benefit from an introduction, and lines 440-- should become a new subsection 4.1. (named ``Float'' or ``Real numbers support'' or similar).
+\begin{solution}
+
+\end{solution}
+
+\question I suggest a better example for lines 459--460: a computation of $(e^{100} +1)− e^{100}$:
+
+\begin{verbatim}
+   >>> (exp(100)+1).evalf() - exp(100).evalf()
+   0
+   >>> ((exp(100)+1) - exp(100)).evalf()
+   1.00000000000000
+   >>> (exp(100)+1) - exp(100)
+   1
+\end{verbatim}
+
+or two different ways to compute the 100th Fibonacci number:
+
+\begin{verbatim}
+   >>> phi = (1+sqrt(5))/2
+   >>> psi = (1-sqrt(5))/2
+   >>> ((phi**100-psi**100)/sqrt(5)).evalf()-fibonacci(100)
+   65536.0000000000
+   >>> ((phi**100-psi**100)/sqrt(5) - fibonacci(100)).evalf()
+   0.e-104
+\end{verbatim}
+
+Obviously, like your own example, these are problematic because a part of the computation is relying on Python's builtin floats.
+Please include a comment on whether symbolic computation (i.e., applying \texttt{evalf()} on the whole expression) always avoids these errors or not, possibly with an example when it doesn't resolve this problem.
+\begin{solution}
+
+\end{solution}
+
+\question The footnote from line 477 should be moved as a sentence in its own right to the introduction, with other technical specifications.
+\begin{solution}
+
+\end{solution}
+
+\question In line 495, the word ``solving'' seems more appropriate than ``solutions''.
+\begin{solution}
+
+\end{solution}
+
+\question In line 504 ``is'' should be used instead of ``are'' (because ``array'' is singular).
+\begin{solution}
+
+\end{solution}
+
+\question In line 518, ``produces'' should become plural.
+\begin{solution}
+
+\end{solution}
+
+\question The title of section 5, ``Domain specific submodules'', seems inappropriate because the section only covers Physics package (not ``submodules'').
+It should either be expanded with a short introduction listing other domain specific packages, or it should be renamed to ``Physics package''.
+\begin{solution}
+
+\end{solution}
+
+\question The word ``symbolics'' in line 528 should be removed (as almost everything in the paper deals with symbolic computation).
+\begin{solution}
+
+\end{solution}
+
+\question In line 530, sympy.physics.vector is a module, not a package.
+\begin{solution}
+
+\end{solution}
+
+\question It is unclear what ``both of these objects'' refer to in line 532. My guess is vectors and dyadic objects, but this should be reworded to make it more clear.
+\begin{solution}
+
+\end{solution}
+
+\question In lines 543 and 545, ``rad'' should be removed. Radians are assumed when no other measure (like degrees) is given.
+\begin{solution}
+
+\end{solution}
+
+\question In lines 567--568, ``performing symbolic quantum mechanics'' makes no sense. This should probably be ``computations'', ``solving problems related to'', etc.
+\begin{solution}
+
+\end{solution}
+
+\question The sentence ``SymPy expressions are immutable trees of Python objects.'' doesn't belong in the conclusion. This can be moved to the appropriate place when discussing SymPy's architecture.
+\begin{solution}
+
+\end{solution}
+
+\question All ``submodules'' should be replaced by ``modules'' (examples: lines 657, 662).
+\begin{solution}
+
+\end{solution}
+
+\question The sentences in lines 657--661 should swap places, because ``areas of mathematics'' are discrete mathematics, concrete mathematics, etc., while simplifying expressions, performing common calculus operations, pretty print- ing expressions, etc.\ belongs to common operations (``other areas'' is also fine, albeit slightly wrong).
+\begin{solution}
+
+\end{solution}
+
+\question In line 662 ``classical mechanics and quantum mechanics'' are listed as the only example of the support for specific domains, as in section 5, which leaves the impression that physics the only one.
+Either more domains should be listed, or it should be reworded to recognize the fact that there are no others.
+\begin{solution}
+
+\end{solution}
+
+\question Lines 670--678 contain explanations what some of the authors' institutions are.
+It is customary to use acknowledgements to thank people and institutions, while institutions' details should be provided in the authors' footnotes in the documents' head.
+\begin{solution}
+
+\end{solution}
+
+\question The citation [5] in line 691 is missing identification data, probably a URL\@.
+\begin{solution}
+
+\end{solution}
+
+\question The citation [21] in line 728 should have ``2D'' instead of ``2d''. More specific comments and correction suggestions for the supplement fol- low.
+\begin{solution}
+
+\end{solution}
+
+\question The supplement should have an introduction, explaining what is being covered in it in general and in each section.
+\begin{solution}
+
+\end{solution}
+
+\question Since the Guntz algorithm is covered in depth, it would be good to
+include how an interested reader can see SymPy's steps of computation:
+
+\begin{verbatim}
+     >>> import os
+     >>> os.environ['SYMPY_DEBUG'] = 'True'
+     >>> from sympy import *
+     >>> x = symbols('x')
+     >>> limit(sin(x)/x, x, 0)
+     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
+     DEBUG: returned None
+     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
+     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
+     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
+     DEBUG: returned None
+     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
+     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
+     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
+     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
+     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
+     DEBUG: returned None
+     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
+     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
+     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
+     DEBUG: returned None
+     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
+     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
+     DEBUG: parsing of expression [(1, 1, None, None)] with symbol _w
+     DEBUG: returned None
+     limitinf(x*sin(1/x), x) = 1
+     +-mrv_leadterm(_p*sin(1/_p), _p) = (1, 0)
+     | +-mrv(_p*sin(1/_p), _p) = ({_p: _Dummy_14}, {}, _Dummy_14*sin(1/_Dummy_14))
+     | | +-mrv(_p, _p) = ({_p: _Dummy_14}, {}, _Dummy_14)
+     | | +-mrv(sin(1/_p), _p) = ({_p: _Dummy_15}, {}, sin(1/_Dummy_15))
+     | |   +-mrv(1/_p, _p) = ({_p: _Dummy_15}, {}, 1/_Dummy_15)
+     | |     +-mrv(_p, _p) = ({_p: _Dummy_15}, {}, _Dummy_15)
+     | +-rewrite(_Dummy_14*sin(1/_Dummy_14), {exp(_p): _Dummy_14}, {}, _p, _w)
+             = (sin(_w)/_w, -_p)
+     | | +-sign(_p, _p) = 1
+     | | +-limitinf(1, _p) = 1
+     | +-calculate_series(sin(_w)/_w, _w) = 1
+     |   +-limitinf(_w, _w) = oo
+     |   | +-mrv_leadterm(_w, _w) = (1, -1)
+     |   | | +-mrv(_w, _w) = ({_w: _Dummy_18}, {}, _Dummy_18)
+     |   | | +-rewrite(_Dummy_18, {exp(_w): _Dummy_18}, {}, _w, _w) = (1/_w, -_w)
+     |   | | | +-sign(_w,_w)=1
+     |   | | | +-limitinf(1, _w) = 1
+     |   | | +-calculate_series(1/_w, _w) = 1/_w
+     |   | +-sign(-1, _w) = -1
+     |   | +-sign(1, _w) = 1
+     |   +-limitinf(_w, _w) = oo
+     |   +-limitinf(_w, _w) = oo
+     |   +-limitinf(_w, _w) = oo
+     +-sign(0, _p) = 0
+     +-limitinf(1, _p) = 1
+\end{verbatim}
+
+Emphasize that the environment variable \texttt{SYMPY\_DEBUG} must be set before importing SymPy for the first time.
+\begin{solution}
+
+\end{solution}
+
+\question In line 183, syntax \texttt{symbols(\textquotesingle{}a:d\textquotesingle{})} is used without being previously defined or explained.
+\begin{solution}
+
+\end{solution}
+
+\question Figure 2 should not be a screenshot, but rather proper code with \LaTeX-rendered results and explanations, as was done in the rest of the paper. A
+ shortened link to SymPy Gamma for the example's expression can be included
+ for the user to try this for themselves.
+\begin{solution}
+
+\end{solution}
+
 \end{questions}

### 2016-08-10 rebuttal.tex
Unindent Gruntz debugging example

@@
 >>> from sympy import *
+sympy/external/importtools.py:145: UserWarning: gmpy2 module is not installed
+  warnings.warn("%s module is not installed" % module, UserWarning)
 >>> x = symbols('x')
@@
 +-mrv_leadterm(_p*sin(1/_p), _p) = (1, 0)
-     | +-mrv(_p*sin(1/_p), _p) = ({_p: _Dummy_14}, {}, _Dummy_14*sin(1/_Dummy_14))
-     | | +-mrv(_p, _p) = ({_p: _Dummy_14}, {}, _Dummy_14)
-     | | +-mrv(sin(1/_p), _p) = ({_p: _Dummy_15}, {}, sin(1/_Dummy_15))
-     | |   +-mrv(1/_p, _p) = ({_p: _Dummy_15}, {}, 1/_Dummy_15)
-     | |     +-mrv(_p, _p) = ({_p: _Dummy_15}, {}, _Dummy_15)
-     | +-rewrite(_Dummy_14*sin(1/_Dummy_14), {exp(_p): _Dummy_14}, {}, _p, _w)
-             = (sin(_w)/_w, -_p)
+| +-mrv(_p*sin(1/_p), _p) = ({_p: _Dummy_19}, {}, _Dummy_19*sin(1/_Dummy_19))
+| | +-mrv(_p, _p) = ({_p: _Dummy_19}, {}, _Dummy_19)
+| | +-mrv(sin(1/_p), _p) = ({_p: _Dummy_20}, {}, sin(1/_Dummy_20))
+| |   +-mrv(1/_p, _p) = ({_p: _Dummy_20}, {}, 1/_Dummy_20)
+| |     +-mrv(_p, _p) = ({_p: _Dummy_20}, {}, _Dummy_20)
+| +-rewrite(_Dummy_19*sin(1/_Dummy_19), {exp(_p): _Dummy_19}, {}, _p, _w) = (sin(_w)/_w, -_p)
 | | +-sign(_p, _p) = 1
@@
 |   | +-mrv_leadterm(_w, _w) = (1, -1)
-     |   | | +-mrv(_w, _w) = ({_w: _Dummy_18}, {}, _Dummy_18)
-     |   | | +-rewrite(_Dummy_18, {exp(_w): _Dummy_18}, {}, _w, _w) = (1/_w, -_w)
-     |   | | | +-sign(_w,_w)=1
+|   | | +-mrv(_w, _w) = ({_w: _Dummy_23}, {}, _Dummy_23)
+|   | | +-rewrite(_Dummy_23, {exp(_w): _Dummy_23}, {}, _w, _w) = (1/_w, -_w)
+|   | | | +-sign(_w, _w) = 1
 |   | | | +-limitinf(1, _w) = 1
@@
 |   +-limitinf(_w, _w) = oo
-     |   +-limitinf(_w, _w) = oo
 +-sign(0, _p) = 0
 +-limitinf(1, _p) = 1
+1
 \end{verbatim}

### 2016-08-15 rebuttal.tex
Respond to the rest of reviewer 2 comments
Fix some latex formatting

--- 2016-08-10 rebuttal.tex
@@
 \begin{solution}
-
+The footnote here was in reference to the SymPy classes. We have moved the
+footnote earlier in the sentence, to ``SymPy classes''.
 \end{solution}

### 2016-08-15 architecture.tex
Remove commutative footnote (it's an unimportant detail)

@@
 assumptions use a three-valued logic using the Python built in objects
-\texttt{True}, \texttt{False}, and \texttt{None}.\footnote{Special
-case is the \texttt{commutative} assumption: it can't be \texttt{None}.
-Unless a Symbol is created with \texttt{commutative=False}, it's
-assumed to commute with any other object with respect
-to the multiplication operation.} Note that \texttt{False}
+\texttt{True}, \texttt{False}, and \texttt{None}. Note that \texttt{False}
 is returned if the SymPy object doesn't or can't have the assumption.

### 2016-08-15 architecture.tex
Re-add commutative footnote, with clarifications for the reviewer

@@
 \texttt{negative}, \texttt{real}, \texttt{nonpositive}, \texttt{integer},
-\texttt{prime} and \texttt{commutative}. Assumptions on any object can
-be checked with the \verb|is_|\texttt{\textit{assumption}} attributes,
-like \verb|t.is_positive|.
+\texttt{prime} and \texttt{commutative}\footnote{SymPy assumes that two
+  expressions $A$ and $B$ commute with each other multiplicatively, that is,
+  $A\cdot B=B\cdot A$, unless they both have \texttt{commutative=False}. Many
+  algorithms in SymPy require special consideration to work correctly with
+  noncommutative products.}. Assumptions on any object can be checked with the
+\verb|is_|\texttt{\textit{assumption}} attributes, like \verb|t.is_positive|.
 

### 2016-08-15 rebuttal.tex
Add reviewer comments (specific comments only) to the rebuttal

@@
 
-\section{Reviewer 3}
+\section{Reviewer 3 (Vedran Sego)}
 
 \begin{questions}
-\question one
+
+\question Explaining what Python is (lines 70–72) should go before talking about SymPy as a “CAS written in Python”.
+Further, the paper assumes a moderate familiarity with Python (for example, Python’s console, OOP, and exceptions), and this should be specified.
+There should be a short note on the used Python console (>>> is the prompt, with the results of computation following immediately in the lines after it).
+The citation [25] from line 65 should be moved next to “Python” in line 70.
 \begin{solution}
- We opted to follow the ABC style of ...
+
 \end{solution}
-\question  two
+
+\question Line 73 has outdated information.
+Sage was renamed to SageMath and it no longer aims only at pure mathematics but also at algebra, numerical analysis, etc.
+The reference [40] should be replaced by a more up-to-date one.
 \begin{solution}
-Section 2 has been updated and the discussion ...
+
 \end{solution}
+
+\question The plural “CASs” is usually written as “CASes” or “CAS’s”, with the latter being somewhat problematic due to it looking like it implies posses- sion.
+\begin{solution}
+
+\end{solution}
+
+\question Line 88 mentions “printers”, but it doesn’t state what they are, which is confusing for those readers that are yet to learn the concept in section 3.4.
+\begin{solution}
+
+\end{solution}
+
+\question Also in line 88, Jupyter’s citation [30] is actually about IPython and should be replaced by a more up-to-date version.
+\begin{solution}
+
+\end{solution}
+
+\question The word “software” in line 91 is ambiguous; “library” or “package” would make a better choice.
+\begin{solution}
+
+\end{solution}
+
+\question Lines 91–96: “we discuss/look at/etc” is the preferred form, instead of “section discusses/looks at/etc”.
+\begin{solution}
+
+\end{solution}
+
+\question The paragraph in lines 103–105 should be moved to the introduction, and the footnote from line 104 should be added to that paragraph as a full-blown sentence, expanded by all the relevant technical information (Python version, OS,...).
+Given that the end of life for Python 2 is 2020., a comment on whether all the presented examples work in both Python 2 and 3 should be included as well.
+Further, emphasise that wildcard imports, “import *”, should almost never be used in programs (see PEP 8, the item “wildcard imports”).
+The same goes for the import mentioned in lines 484–487.
+\begin{solution}
+
+\end{solution}
+
+\question Line 119 should be removed, as it is basically a copy of the previous line.
+\begin{solution}
+
+\end{solution}
+
+\question In line 121 the word “stored” should be replaced by “used as keys”.
+\begin{solution}
+
+\end{solution}
+
+\question Whatdotheauthorsmeanby“therebypermittingfeaturessuchascaching” in line 122? Caching can be done for mutable types as well, just not through hashing. 2
+\begin{solution}
+
+\end{solution}
+
+\question There is no need to repeat “(CAS)” in line 124, as it was already given in line 64.
+\begin{solution}
+
+\end{solution}
+
+\question In the same line, the word “represents” should be replaced by “stores”.
+\begin{solution}
+
+\end{solution}
+
+\question In line 184, “symbols are” should be replaced by “t is” (the general rule is already given in line 179).
+\begin{solution}
+
+\end{solution}
+
+\question The code in line 210 should be made into its own line (like a displaymath formula), for typesetting reasons and better readability.
+\begin{solution}
+
+\end{solution}
+
+\question The footnote 4 should be moved from line 221 to line 222, right after “Basic”.
+\begin{solution}
+
+\end{solution}
+
+\question The part “which defines some basic methods for symbolic expression trees” should be removed from line 222, as it was already given in line 130.
+\begin{solution}
+
+\end{solution}
+
+\question In line 225, the sentence “Not all SymPy classes are subclasses of Expr.” sounds confusing as a reader new to SymPy wouldn’t expect, for example, symbols to inherit “Expr”.
+It would be better to expand this, for example “Most of the SymPy classes (including Symbol) are subclasses of Expr, but there are exceptions to this rule”.
+\begin{solution}
+
+\end{solution}
+
+\question The title “Features” in line 276 is ambiguous, as a “feature” has no precise meaning in Python (or even software libraries in general).
+It should be replaced by “Packages and modules” or a similar more precise wording.
+The same goes for ‘feature” in most other places in the paper (for exam- ple, the caption of Table 1.
+and line 495).
+It would be very useful to also include actual names of the packages/modules in Table 1., as well as in any section covering those packages/modules.
+\begin{solution}
+
+\end{solution}
+
+\question The sets support listing is unnaturally split in two by the “This includes...” sentence which would fit better in parentheses.
+\begin{solution}
+
+\end{solution}
+
+\question Line 355, add a sentence explaining that in SymPy str == repr, because in Python repr is used to get an unambiguous valid Python code repre- sentation, while the return value of str is meant to be human-readable.
+\begin{solution}
+
+\end{solution}
+
+\question Lines 359 and 379: what is 2D text representation? It seems that “2D” shouldn’t be here.
+\begin{solution}
+
+\end{solution}
+
+\question Line 427: every dictionary is a “dictionary of keys”. This should be a dictionary with coordinate tuples as keys associated with the appropriate values.
+\begin{solution}
+
+\end{solution}
+
+\question Section 4 would benefit from an introduction, and lines 440– should be- come a new subsection 4.1. (named “Float” or “Real numbers support” or similar).
+\begin{solution}
+
+\end{solution}
+
+\question I suggest a better example for lines 459–460: a computation of (e100 +1)− e100 :
+\begin{solution}
+
+\end{solution}
+   >>> (exp(100)+1).evalf() - exp(100).evalf()
+   0
+   >>> ((exp(100)+1) - exp(100)).evalf()
+   1.00000000000000
+   >>> (exp(100)+1) - exp(100)
+   1
+or two different ways to compute the 100th Fibonacci number:
+   >>> phi = (1+sqrt(5))/2
+   >>> psi = (1-sqrt(5))/2
+   >>> ((phi**100-psi**100)/sqrt(5)).evalf()-fibonacci(100)
+   65536.0000000000
+   >>> ((phi**100-psi**100)/sqrt(5) - fibonacci(100)).evalf()
+   0.e-104
+Obviously, like your own example, these are problematic because a part of the computation is relying on Python’s builtin floats.
+Please include a comment on whether symbolic computation (i.e., applying evalf() on the whole expression) always avoids these errors or not, possibly with an example when it doesn’t resolve this problem.
+
+\question The footnote from line 477 should be moved as a sentence in its own right to the introduction, with other technical specifications.
+\begin{solution}
+
+\end{solution}
+
+\question In line 495, the word “solving” seems more appropriate than “solutions”.
+\begin{solution}
+
+\end{solution}
+
+\question In line 504 “is” should be used instead of “are” (because “array” is singular).
+\begin{solution}
+
+\end{solution}
+
+\question In line 518, “produces” should become plural.
+\begin{solution}
+
+\end{solution}
+
+\question The title of section 5, “Domain specific submodules”, seems inappropriate because the section only covers Physics package (not “submodules”).
+It should either be expanded with a short introduction listing other domain specific packages, or it should be renamed to “Physics package”.
+\begin{solution}
+
+\end{solution}
+
+\question The word “symbolics” in line 528 should be removed (as almost everything in the paper deals with symbolic computation).
+\begin{solution}
+
+\end{solution}
+
+\question In line 530, sympy.physics.vector is a module, not a package.
+\begin{solution}
+
+\end{solution}
+
+\question It is unclear what “both of these objects” refer to in line 532. My guess is vectors and dyadic objects, but this should be reworded to make it more clear.
+\begin{solution}
+
+\end{solution}
+
+\question In lines 543 and 545, “rad” should be removed. Radians are assumed when no other measure (like degrees) is given.
+\begin{solution}
+
+\end{solution}
+
+\question In lines 567–568, “performing symbolic quantum mechanics” makes no sense. This should probably be “computations”, “solving problems related to”, etc.
+\begin{solution}
+
+\end{solution}
+
+\question The sentence “SymPy expressions are immutable trees of Python objects.” doesn’t belong in the conclusion. This can be moved to the appropriate place when discussing SymPy’s architecture.
+\begin{solution}
+
+\end{solution}
+
+\question All “submodules” should be replaced by “modules” (examples: lines 657, 662).
+\begin{solution}
+
+\end{solution}
+
+\question The sentences in lines 657–661 should swap places, because “areas of mathematics” are discrete mathematics, concrete mathematics, etc., while simplifying expressions, performing common calculus operations, pretty print- ing expressions, etc.
+belongs to common operations (“other areas” is also fine, albeit slightly wrong).
+\begin{solution}
+
+\end{solution}
+
+\question In line 662 “classical mechanics and quantum mechanics” are listed as the only example of the support for specific domains, as in section 5, which leaves the impression that physics the only one.
+Either more domains should be listed, or it should be reworded to recognize the fact that there are no others.
+\begin{solution}
+
+\end{solution}
+
+\question Lines 670–678 contain explanations what some of the authors’ institu- tions are.
+It is customary to use acknowledgements to thank people and institutions, while institutions’ details should be provided in the authors’ footnotes in the documents’ head.
+\begin{solution}
+
+\end{solution}
+
+\question The citation [5] in line 691 is missing identification data, probably a URL.
+\begin{solution}
+
+\end{solution}
+
+\question The citation [21] in line 728 should have “2D” instead of “2d”. More specific comments and correction suggestions for the supplement fol- low.
+\begin{solution}
+
+\end{solution}
+
+\question The supplement should have an introduction, explaining what is being covered in it in general and in each section.
+\begin{solution}
+
+\end{solution}
+
+\question Since the Guntz algorithm is covered in depth, it would be good to
+include how an interested reader can see SymPy’s steps of computation:
+
+\begin{verbatim}
+     >>> import os
+     >>> os.environ['SYMPY_DEBUG'] = 'True'
+     >>> from sympy import *
+     >>> x = symbols(’x’)
+     >>> limit(sin(x)/x, x, 0)
+     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
+     DEBUG: returned None
+     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
+     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
+     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
+     DEBUG: returned None
+     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
+     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
+     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
+     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
+     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
+     DEBUG: returned None
+     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
+     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
+     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
+     DEBUG: returned None
+     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
+     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
+     DEBUG: parsing of expression [(1, 1, None, None)] with symbol _w
+     DEBUG: returned None
+     limitinf(x*sin(1/x), x) = 1
+     +-mrv_leadterm(_p*sin(1/_p), _p) = (1, 0)
+     | +-mrv(_p*sin(1/_p), _p) = ({_p: _Dummy_14}, {}, _Dummy_14*sin(1/_Dummy_14))
+     | | +-mrv(_p, _p) = ({_p: _Dummy_14}, {}, _Dummy_14)
+     | | +-mrv(sin(1/_p), _p) = ({_p: _Dummy_15}, {}, sin(1/_Dummy_15))
+     | |   +-mrv(1/_p, _p) = ({_p: _Dummy_15}, {}, 1/_Dummy_15)
+     | |     +-mrv(_p, _p) = ({_p: _Dummy_15}, {}, _Dummy_15)
+     | +-rewrite(_Dummy_14*sin(1/_Dummy_14), {exp(_p): _Dummy_14}, {}, _p, _w)
+             = (sin(_w)/_w, -_p)
+     | | +-sign(_p, _p) = 1
+     | | +-limitinf(1, _p) = 1
+     | +-calculate_series(sin(_w)/_w, _w) = 1
+     |   +-limitinf(_w, _w) = oo
+     |   | +-mrv_leadterm(_w, _w) = (1, -1)
+     |   | | +-mrv(_w, _w) = ({_w: _Dummy_18}, {}, _Dummy_18)
+     |   | | +-rewrite(_Dummy_18, {exp(_w): _Dummy_18}, {}, _w, _w) = (1/_w, -_w)
+| |||+-sign(_w,_w)=1
+     |   | | | +-limitinf(1, _w) = 1
+     |   | | +-calculate_series(1/_w, _w) = 1/_w
+     |   | +-sign(-1, _w) = -1
+     |   | +-sign(1, _w) = 1
+     |   +-limitinf(_w, _w) = oo
+     |   +-limitinf(_w, _w) = oo
+     |   +-limitinf(_w, _w) = oo
+     +-sign(0, _p) = 0
+     +-limitinf(1, _p) = 1
+\end{verbatim}
+
+Emphasize that the environment variable \texttt{SYMPY\_DEBUG} must be set before importing SymPy for the first time.
+\begin{solution}
+
+\end{solution}
+
+\question In line 183, syntax symbols(a:d) is used without being previously defined or explained.
+\begin{solution}
+
+\end{solution}
+
+\question Figure 2 should not be a screenshot, but rather proper code with LATEX-
+w rendered results and explanations, as was done in the rest of the paper. A
+ shortened link to SymPy Gamma for the example’s expression can be included
+ for the user to try this for themselves.
+\begin{solution}
+
+\end{solution}
+
 \end{questions}

### 2016-08-15 rebuttal.tex
Fix several LaTeX rendering issues in the rebuttal

--- 2016-08-15 rebuttal.tex
@@
 \question Line 221:
-the phrase 'Expression tree' is cited but this is not the first time you've used it. Perhaps cite earlier? Line 129 perhaps?
+the phrase `Expression tree' is cited but this is not the first time you've used it. Perhaps cite earlier? Line 129 perhaps?
 
@@
 
-\section{Reviewer 3}
+\section{Reviewer 3 (Vedran Sego)}
 
 \begin{questions}
-\question one
+
+\question Explaining what Python is (lines 70--72) should go before talking about SymPy as a ``CAS written in Python''.
+Further, the paper assumes a moderate familiarity with Python (for example, Python's console, OOP, and exceptions), and this should be specified.
+There should be a short note on the used Python console (\texttt{>>>} is the prompt, with the results of computation following immediately in the lines after it).
+The citation [25] from line 65 should be moved next to ``Python'' in line 70.
 \begin{solution}
- We opted to follow the ABC style of ...
+
 \end{solution}
-\question  two
+
+\question Line 73 has outdated information.
+Sage was renamed to SageMath and it no longer aims only at pure mathematics but also at algebra, numerical analysis, etc.
+The reference [40] should be replaced by a more up-to-date one.
 \begin{solution}
-Section 2 has been updated and the discussion ...
+
 \end{solution}
+
+\question The plural ``CASs'' is usually written as ``CASes'' or ``CAS's'', with the latter being somewhat problematic due to it looking like it implies possession.
+\begin{solution}
+
+\end{solution}
+
+\question Line 88 mentions ``printers'', but it doesn't state what they are, which is confusing for those readers that are yet to learn the concept in section 3.4.
+\begin{solution}
+
+\end{solution}
+
+\question Also in line 88, Jupyter's citation [30] is actually about IPython and should be replaced by a more up-to-date version.
+\begin{solution}
+
+\end{solution}
+
+\question The word ``software'' in line 91 is ambiguous; ``library'' or ``package'' would make a better choice.
+\begin{solution}
+
+\end{solution}
+
+\question Lines 91--96: ``we discuss/look at/etc'' is the preferred form, instead of ``section discusses/looks at/etc''.
+\begin{solution}
+
+\end{solution}
+
+\question The paragraph in lines 103--105 should be moved to the introduction,
+and the footnote from line 104 should be added to that paragraph as a
+full-blown sentence, expanded by all the relevant technical information
+(Python version, OS, \ldots). Given that the end of life for Python 2 is
+2020., a comment on whether all the presented examples work in both Python 2
+and 3 should be included as well. Further, emphasise that wildcard imports,
+``\texttt{import *}'', should almost never be used in programs (see PEP 8, the
+item ``wildcard imports''). The same goes for the \texttt{import} mentioned in lines 484--487.
+\begin{solution}
+
+\end{solution}
+
+\question Line 119 should be removed, as it is basically a copy of the previous line.
+\begin{solution}
+
+\end{solution}
+
+\question In line 121 the word ``stored'' should be replaced by ``used as keys''.
+\begin{solution}
+
+\end{solution}
+
+\question What do the authors mean by ``thereby permitting features such as caching'' in line 122? Caching can be done for mutable types as well, just not through hashing.
+\begin{solution}
+
+\end{solution}
+
+\question There is no need to repeat ``(CAS)'' in line 124, as it was already given in line 64.
+\begin{solution}
+
+\end{solution}
+
+\question In the same line, the word ``represents'' should be replaced by ``stores''.
+\begin{solution}
+
+\end{solution}
+
+\question In line 184, ``symbols are'' should be replaced by ``\texttt{t} is'' (the general rule is already given in line 179).
+\begin{solution}
+
+\end{solution}
+
+\question The code in line 210 should be made into its own line (like a displaymath formula), for typesetting reasons and better readability.
+\begin{solution}
+
+\end{solution}
+
+\question The footnote 4 should be moved from line 221 to line 222, right after ``Basic''.
+\begin{solution}
+
+\end{solution}
+
+\question The part ``which defines some basic methods for symbolic expression trees'' should be removed from line 222, as it was already given in line 130.
+\begin{solution}
+
+\end{solution}
+
+\question In line 225, the sentence ``Not all SymPy classes are subclasses of \texttt{Expr}.'' sounds confusing as a reader new to SymPy wouldn't expect, for example, symbols to inherit ``Expr''.
+It would be better to expand this, for example ``Most of the SymPy classes (including \texttt{Symbol}) are subclasses of \texttt{Expr}, but there are exceptions to this rule''.
+\begin{solution}
+
+\end{solution}
+
+\question The title ``Features'' in line 276 is ambiguous, as a ``feature''
+has no precise meaning in Python (or even software libraries in general). It
+should be replaced by ``Packages and modules'' or a similar more precise
+wording. The same goes for ``feature'' in most other places in the paper (for
+example, the caption of Table 1.\ and line 495). It would be very useful to also include actual names of the packages/modules in Table 1., as well as in any section covering those packages/modules.
+\begin{solution}
+
+\end{solution}
+
+\question The sets support listing is unnaturally split in two by the ``This includes\ldots'' sentence which would fit better in parentheses.
+\begin{solution}
+
+\end{solution}
+
+\question Line 355, add a sentence explaining that in SymPy \texttt{str == repr}, because in Python \texttt{repr} is used to get an unambiguous valid Python code representation, while the return value of \texttt{str} is meant to be human-readable.
+\begin{solution}
+
+\end{solution}
+
+\question Lines 359 and 379: what is 2D text representation? It seems that ``2D'' shouldn't be here.
+\begin{solution}
+
+\end{solution}
+
+\question Line 427: every dictionary is a ``dictionary of keys''. This should be a dictionary with coordinate tuples as keys associated with the appropriate values.
+\begin{solution}
+
+\end{solution}
+
+\question Section 4 would benefit from an introduction, and lines 440-- should become a new subsection 4.1. (named ``Float'' or ``Real numbers support'' or similar).
+\begin{solution}
+
+\end{solution}
+
+\question I suggest a better example for lines 459--460: a computation of $(e^{100} +1)− e^{100}$:
+
+\begin{verbatim}
+   >>> (exp(100)+1).evalf() - exp(100).evalf()
+   0
+   >>> ((exp(100)+1) - exp(100)).evalf()
+   1.00000000000000
+   >>> (exp(100)+1) - exp(100)
+   1
+\end{verbatim}
+
+or two different ways to compute the 100th Fibonacci number:
+
+\begin{verbatim}
+   >>> phi = (1+sqrt(5))/2
+   >>> psi = (1-sqrt(5))/2
+   >>> ((phi**100-psi**100)/sqrt(5)).evalf()-fibonacci(100)
+   65536.0000000000
+   >>> ((phi**100-psi**100)/sqrt(5) - fibonacci(100)).evalf()
+   0.e-104
+\end{verbatim}
+
+Obviously, like your own example, these are problematic because a part of the computation is relying on Python's builtin floats.
+Please include a comment on whether symbolic computation (i.e., applying \texttt{evalf()} on the whole expression) always avoids these errors or not, possibly with an example when it doesn't resolve this problem.
+\begin{solution}
+
+\end{solution}
+
+\question The footnote from line 477 should be moved as a sentence in its own right to the introduction, with other technical specifications.
+\begin{solution}
+
+\end{solution}
+
+\question In line 495, the word ``solving'' seems more appropriate than ``solutions''.
+\begin{solution}
+
+\end{solution}
+
+\question In line 504 ``is'' should be used instead of ``are'' (because ``array'' is singular).
+\begin{solution}
+
+\end{solution}
+
+\question In line 518, ``produces'' should become plural.
+\begin{solution}
+
+\end{solution}
+
+\question The title of section 5, ``Domain specific submodules'', seems inappropriate because the section only covers Physics package (not ``submodules'').
+It should either be expanded with a short introduction listing other domain specific packages, or it should be renamed to ``Physics package''.
+\begin{solution}
+
+\end{solution}
+
+\question The word ``symbolics'' in line 528 should be removed (as almost everything in the paper deals with symbolic computation).
+\begin{solution}
+
+\end{solution}
+
+\question In line 530, sympy.physics.vector is a module, not a package.
+\begin{solution}
+
+\end{solution}
+
+\question It is unclear what ``both of these objects'' refer to in line 532. My guess is vectors and dyadic objects, but this should be reworded to make it more clear.
+\begin{solution}
+
+\end{solution}
+
+\question In lines 543 and 545, ``rad'' should be removed. Radians are assumed when no other measure (like degrees) is given.
+\begin{solution}
+
+\end{solution}
+
+\question In lines 567--568, ``performing symbolic quantum mechanics'' makes no sense. This should probably be ``computations'', ``solving problems related to'', etc.
+\begin{solution}
+
+\end{solution}
+
+\question The sentence ``SymPy expressions are immutable trees of Python objects.'' doesn't belong in the conclusion. This can be moved to the appropriate place when discussing SymPy's architecture.
+\begin{solution}
+
+\end{solution}
+
+\question All ``submodules'' should be replaced by ``modules'' (examples: lines 657, 662).
+\begin{solution}
+
+\end{solution}
+
+\question The sentences in lines 657--661 should swap places, because ``areas of mathematics'' are discrete mathematics, concrete mathematics, etc., while simplifying expressions, performing common calculus operations, pretty print- ing expressions, etc.\ belongs to common operations (``other areas'' is also fine, albeit slightly wrong).
+\begin{solution}
+
+\end{solution}
+
+\question In line 662 ``classical mechanics and quantum mechanics'' are listed as the only example of the support for specific domains, as in section 5, which leaves the impression that physics the only one.
+Either more domains should be listed, or it should be reworded to recognize the fact that there are no others.
+\begin{solution}
+
+\end{solution}
+
+\question Lines 670--678 contain explanations what some of the authors' institutions are.
+It is customary to use acknowledgements to thank people and institutions, while institutions' details should be provided in the authors' footnotes in the documents' head.
+\begin{solution}
+
+\end{solution}
+
+\question The citation [5] in line 691 is missing identification data, probably a URL\@.
+\begin{solution}
+
+\end{solution}
+
+\question The citation [21] in line 728 should have ``2D'' instead of ``2d''. More specific comments and correction suggestions for the supplement fol- low.
+\begin{solution}
+
+\end{solution}
+
+\question The supplement should have an introduction, explaining what is being covered in it in general and in each section.
+\begin{solution}
+
+\end{solution}
+
+\question Since the Guntz algorithm is covered in depth, it would be good to
+include how an interested reader can see SymPy's steps of computation:
+
+\begin{verbatim}
+     >>> import os
+     >>> os.environ['SYMPY_DEBUG'] = 'True'
+     >>> from sympy import *
+     >>> x = symbols('x')
+     >>> limit(sin(x)/x, x, 0)
+     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
+     DEBUG: returned None
+     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
+     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
+     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
+     DEBUG: returned None
+     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
+     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
+     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
+     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
+     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
+     DEBUG: returned None
+     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
+     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
+     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
+     DEBUG: returned None
+     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
+     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
+     DEBUG: parsing of expression [(1, 1, None, None)] with symbol _w
+     DEBUG: returned None
+     limitinf(x*sin(1/x), x) = 1
+     +-mrv_leadterm(_p*sin(1/_p), _p) = (1, 0)
+     | +-mrv(_p*sin(1/_p), _p) = ({_p: _Dummy_14}, {}, _Dummy_14*sin(1/_Dummy_14))
+     | | +-mrv(_p, _p) = ({_p: _Dummy_14}, {}, _Dummy_14)
+     | | +-mrv(sin(1/_p), _p) = ({_p: _Dummy_15}, {}, sin(1/_Dummy_15))
+     | |   +-mrv(1/_p, _p) = ({_p: _Dummy_15}, {}, 1/_Dummy_15)
+     | |     +-mrv(_p, _p) = ({_p: _Dummy_15}, {}, _Dummy_15)
+     | +-rewrite(_Dummy_14*sin(1/_Dummy_14), {exp(_p): _Dummy_14}, {}, _p, _w)
+             = (sin(_w)/_w, -_p)
+     | | +-sign(_p, _p) = 1
+     | | +-limitinf(1, _p) = 1
+     | +-calculate_series(sin(_w)/_w, _w) = 1
+     |   +-limitinf(_w, _w) = oo
+     |   | +-mrv_leadterm(_w, _w) = (1, -1)
+     |   | | +-mrv(_w, _w) = ({_w: _Dummy_18}, {}, _Dummy_18)
+     |   | | +-rewrite(_Dummy_18, {exp(_w): _Dummy_18}, {}, _w, _w) = (1/_w, -_w)
+     |   | | | +-sign(_w,_w)=1
+     |   | | | +-limitinf(1, _w) = 1
+     |   | | +-calculate_series(1/_w, _w) = 1/_w
+     |   | +-sign(-1, _w) = -1
+     |   | +-sign(1, _w) = 1
+     |   +-limitinf(_w, _w) = oo
+     |   +-limitinf(_w, _w) = oo
+     |   +-limitinf(_w, _w) = oo
+     +-sign(0, _p) = 0
+     +-limitinf(1, _p) = 1
+\end{verbatim}
+
+Emphasize that the environment variable \texttt{SYMPY\_DEBUG} must be set before importing SymPy for the first time.
+\begin{solution}
+
+\end{solution}
+
+\question In line 183, syntax \texttt{symbols(\textquotesingle{}a:d\textquotesingle{})} is used without being previously defined or explained.
+\begin{solution}
+
+\end{solution}
+
+\question Figure 2 should not be a screenshot, but rather proper code with \LaTeX-rendered results and explanations, as was done in the rest of the paper. A
+ shortened link to SymPy Gamma for the example's expression can be included
+ for the user to try this for themselves.
+\begin{solution}
+
+\end{solution}
+
 \end{questions}

### 2016-08-16 rebuttal.tex
Add some todos (comments) to the rebuttal
Address reviewer 3, comment 9

@@
 \begin{solution}
-
+  % TODO: I recall we looked this up and the spelling consensus agreed with
+  % the way we have it. We should double check (perhaps we can email the
+  % editor to ask if the journal has a convention).
 \end{solution}
@@
 \begin{solution}
-
+  % See https://github.com/jupyter/jupyter/issues/190
 \end{solution}
@@
 \begin{solution}
-
+  % TODO: We had decided not to use first person. We should perhaps email the editor
+  % to ask about this.
 \end{solution}
@@
 \begin{solution}
+Line 119 is the output of line 118, so now showing it would
+be incorrect.
 
+The aim here was simply to show that the input expression remains unevaluated.
+We have added a sentence before the example to note this.
+This particular example was chosen because it shows the basic syntax for
+addition, subtraction, multiplication, division, and exponentiation.
 \end{solution}

### 2016-08-16 architecture.tex
Address reviewer 3, comment 9
Move the footnote as suggested by the reviewer
This is a better place for it.

@@
 instance, the following Python code creates the expression $(x^2 - 2x + 3)/y$.
+Note that the expression remains unevaluated: it is represented symbolically.
 
@@
 SymPy class, usually  \texttt{Basic}, \texttt{Expr}, or
-\texttt{Function}. All SymPy classes\footnote{Some
+\texttt{Function}. All SymPy classes used for expression trees should be subclasses of the base class
+\texttt{Basic}\footnote{Some
   internal classes, such as those used in the polynomial module, do not follow
-  this rule for efficiency reasons.} used for expression trees should be subclasses of the base class
-\texttt{Basic}, which defines some basic methods for symbolic expression
+  this rule for efficiency reasons.}, which defines some basic methods for symbolic expression
 trees. \texttt{Expr} is the subclass for mathematical expressions that can be

### 2016-08-16 rebuttal.tex
Move the footnote as suggested by the reviewer
This is a better place for it.
Add a comment
Address editor question about Matrix section

@@
 \begin{solution}
-  % TODO
+Yes, included.  Second example in that section address computation of
+eigenvalues.  Let us know if we should expand this.
 \end{solution}
@@
 The footnote here was in reference to the SymPy classes. We have moved the
-footnote earlier in the sentence, to ``SymPy classes''.
+footnote later in the sentence, to ``\texttt{Basic}'', as suggested
+by reviewer 3, point~\ref{rev3point16}.
 \end{solution}
@@
 \begin{solution}
-
+  % TODO: It seems better to start the paper with "SymPy is" rather than
+  % "Python is". The reviewer perhaps wants us to not mention Python until the
+  % second paragraph, but Python is important enough to SymPy's design that it
+  % warrants mentioning in the first sentence, in my opinion. --Aaron
 \end{solution}
@@
 \begin{solution}
-
+We have changed ``printers'' to ``display formatters'', which is the
+terminology used by Jupyter.
 \end{solution}
@@
 \begin{solution}
-  % See https://github.com/jupyter/jupyter/issues/190
+We have changed the citation to
+
+Kluyver, T., Ragan-Kelley, B., P{\'e}rez, F.,
+Granger, B., Bussonnier, M., Frederic, J., Kelley,
+K., Hamrick, J., Grout,
+J., Corlay, S., et al. (2016). Jupyter notebooks---a publishing format for
+reproducible computational workflows. In \textit{Positioning and Power in Academic
+Publishing: Players, Agents and Agendas: Proceedings of the 20th International
+Conference on Electronic Publishing}, page 87. IOS Press.
+
+as recommended by the Jupyter developers at \url{https://github.com/jupyter/jupyter/issues/190}.
 \end{solution}
@@
 \begin{solution}
-
+We have changed ``software'' to ``library''.
 \end{solution}
@@
 \begin{solution}
+We have moved the footnote inline, and have noted that the examples should
+work with Python 2.7, 3.2, 3.3, 3.4, or 3.5, with any operating system that
+supports Python.
 
+We have added a footnote discussing the concerns with \texttt{import *} (see
+reviewer 3, point~\ref{rev1point1}).
 \end{solution}
@@
 \begin{solution}
-
+We have made the suggested change.
 \end{solution}
@@
 \begin{solution}
-
+  % I don't personally agree with this, but it's a minor point. --Aaron
 \end{solution}
@@
 \question The footnote 4 should be moved from line 221 to line 222, right after ``Basic''.
+\label{rev3point16}
 \begin{solution}
-
+We have made the suggested change.
 \end{solution}

### 2016-08-16 rebuttal.tex
Answer the editor's question about mpmath syntax (vs sympy)

@@
 \begin{solution}
-  % TODO
+The mpmath library is not a SymPy submodule, but a separate library,
+that should be clear for a reader from the text.  From that, we believe,
+it should obvious that in general syntax (i.e. names for special
+functions, order of their arguments, etc) will be different, just
+as for any other external library.
 \end{solution}

### 2016-08-16 architecture.tex
Move wildcard imports to the introduction
Also mention supported Python versions

@@
 % symbols, various ways to declare them
-
-The following statement imports all SymPy functions into the global Python
-namespace.\footnote{\texttt{import *} has been used here to aid the
-  readability of the paper, but is best to avoid such wildcard import
-  statements in production code, as they make it unclear which names are
-  present in the namespace. Furthermore, imported names could clash with
-  already existing imports from another package. For example, SymPy, the
-  standard Python \texttt{math} library, and NumPy all define the \texttt{exp}
-  function, but only the SymPy one will work with SymPy symbolic expressions.}
-From here on, all examples in this paper assume that this statement has been
-executed.\footnote{All examples in this paper use SymPy version 1.0.}
-
-\begin{verbatim}
->>> from sympy import *
-\end{verbatim}
 
@@
 SymPy by allowing expression interning. It also enables expressions to be
-hashed and stored in Python dictionaries, thereby permitting features
+hashed and used as keys in Python dictionaries, thereby permitting features
 such as caching.

### 2016-08-16 introduction.tex
Move wildcard imports to the introduction
Also mention supported Python versions

SymPy is a full featured computer algebra system (CAS) written in the Python
programming language~\cite{lutz2013learning}.
It is free and open source software, licensed under the 3-clause BSD
license~\cite{rosen2005open}.
The SymPy project was started by Ond\v{r}ej \v{C}ert\'{\i}k in 2005, and it has
since grown to over 500 contributors. Currently, SymPy is
developed on GitHub using a bazaar community
model~\cite{raymond1999cathedral}. The accessibility of the codebase and the
open community model allow SymPy to rapidly respond to the needs of
users and developers.

Python is a dynamically typed programming language that has a focus on
ease of use and readability. Due in part to this focus, it has become a popular
language for scientific
computing and data science, with a broad ecosystem of
libraries~\cite{oliphant2007python}. SymPy is itself used by many libraries
and tools to support research within a variety of domains, such as
Sage~\cite{SAGE} (pure mathematics),
yt~\cite{2011ApJS..192....9T} (astronomy and astrophysics),
PyDy~\cite{gede2013constrained} (multibody dynamics), and
SfePy~\cite{cimrman2014sfepy} (finite elements).

Unlike many CASs, SymPy does not invent its own programming language. Python
itself is used both for the internal implementation and end user
interaction. By using the operator overloading functionality of Python, SymPy follows the embedded domain specific language paradigm proposed by Hudak~\cite{dsl-little-languages}.  The exclusive usage of a single programming language makes it easier
for people already familiar with that language to use or develop SymPy.
Simultaneously, it enables developers to focus on mathematics, rather than
language design.  SymPy officially supports Python 2.6, 2.7 and 3.2--3.5.

SymPy is designed with a strong focus on usability as a library.
Extensibility is important in its application program interface
(API) design. Thus, SymPy makes no attempt to extend
the Python language itself. The goal is for users of SymPy to be able to
include SymPy alongside other Python libraries in their workflow, whether that
be in an interactive environment or as a programmatic part in a larger system.

As a library, SymPy does not have a built-in graphical user interface (GUI).
However, SymPy exposes a rich interactive display system, and supports
registering display formatters with Jupyter~\cite{kluyver2016jupyter} frontends,
including the Notebook and Qt Console, which will render SymPy expressions
using MathJax~\cite{cervone2012mathjax} or \LaTeX{}.

The remainder of this paper discusses key components of the SymPy library.
Section~\ref{sec:architecture} discusses the architecture of SymPy.
Section~\ref{sec:features} enumerates the features of SymPy and takes a closer
look at some of the important ones. The section~\ref{sec:numerics}
looks at the numerical features of SymPy and its dependency library, mpmath.
Section~\ref{sec:domain_specific} looks at the domain specific physics
submodules for performing symbolic and numerical calculations in classical mechanics
and quantum mechanics. Conclusions and future directions for SymPy are given
in section~\ref{sec:conclusion}.  All examples in this paper use SymPy version 1.0.

The following statement imports all SymPy functions into the global Python
namespace.\footnote{\texttt{import *} has been used here to aid the
  readability of the paper, but is best to avoid such wildcard import
  statements in production code, as they make it unclear which names are
  present in the namespace. Furthermore, imported names could clash with
  already existing imports from another package. For example, SymPy, the
  standard Python \texttt{math} library, and NumPy all define the \texttt{exp}
  function, but only the SymPy one will work with SymPy symbolic expressions.}
From here on, all examples in this paper assume that this statement has been
executed:

\begin{verbatim}
>>> from sympy import *
\end{verbatim}

### 2016-08-16 rebuttal.tex
Fixed ref [20]
Fixed reviewer's text in 3.4
fix typo in solution 9 (reviewer 3)
Address questions 10 and 11 of reviewer 3

@@
 \begin{solution}
-  % TODO: I do not know how to make bibtex (with apalike) show the editor
-  % here.
+We have done this.
 \end{solution}
@@
 have added a footnote:
-
 \begin{quote}\texttt{import *} has been used here to aid the
@@
 \begin{solution}
-Line 119 is the output of line 118, so now showing it would
+Line 119 is the output of line 118, so not showing it would
 be incorrect.
@@
 \begin{solution}
-We have made the suggested change.
+See next answer.
 \end{solution}
@@
 \begin{solution}
-
+This paragraph was replaced by:
+\begin{quote}
+Importantly, SymPy expressions are immutable. This simplifies the design of
+SymPy by allowing expression interning. It also enables expressions to be
+hashed, that is used to implement caching in SymPy.
+\end{quote}
 \end{solution}

### 2016-08-16 architecture.tex
Address questions 10 and 11 of reviewer 3
q 12 of reviewer 3

@@
 SymPy by allowing expression interning. It also enables expressions to be
-hashed and used as keys in Python dictionaries, thereby permitting features
-such as caching.
+hashed, that is used to implement caching in SymPy.
 
@@
 
-A computer algebra system (CAS) represents mathematical expressions as data
+A computer algebra system represents mathematical expressions as data
 structures.  For example, the mathematical expression $x + y$ is represented as

### 2016-08-16 architecture.tex
Address questions 13 and 14 of reviewer 3

@@
 
-A computer algebra system represents mathematical expressions as data
+A computer algebra system stores mathematical expressions as data
 structures.  For example, the mathematical expression $x + y$ is represented as
@@
 
-By assuming the most general case, that symbols are complex by default, SymPy
+By assuming the most general case, that \texttt{t} is complex by default, SymPy
 avoids performing mathematically invalid operations. However, in many cases

### 2016-08-16 architecture.tex
Address question 15 of reviewer 3

@@
 instance, the assumptions system knows that being an integer implies being
-rational, so \texttt{Symbol(\textquotesingle{}x\textquotesingle{}, integer=True).is\_rational} returns
-\texttt{True}. Furthermore, expressions compute the assumptions on themselves
+rational.
+\begin{verbatim}
+>>> i = Symbol('i', integer=True)
+>>> i.is_rational
+True
+\end{verbatim}
+Furthermore, expressions compute the assumptions on themselves
 based on the assumptions of their arguments. For instance, if \texttt{x} and

### 2016-08-16 rebuttal.tex
q 12 of reviewer 3
Address questions 13 and 14 of reviewer 3
Address questions 16 and 17 of reviewer 3
footnote was moved above, to the section 2.2.

@@
 \begin{solution}
-
+That was removed.
 \end{solution}
@@
 \begin{solution}
-  % I don't personally agree with this, but it's a minor point. --Aaron
+Done.
 \end{solution}
@@
 \begin{solution}
-
+Done.
 \end{solution}
@@
 \begin{solution}
-
+Now we use verbatim environment here.
 \end{solution}
@@
 \begin{solution}
-
+This part was removed and footnote moved to the mentioned
+text in the section 2.2, "The Core".
 \end{solution}
@@
 \begin{solution}
-
+This paragraph now looks as:
+\begin{quote}
+The typical way to create a custom SymPy object is to subclass an existing
+SymPy class, usually  \texttt{Basic}, \texttt{Expr}, or
+\texttt{Function}. As it was stated before, almost all SymPy classes used for
+expression trees should be subclasses of the base class \texttt{Basic}.
+\texttt{Expr} is the subclass for mathematical expressions (including \texttt{Symbol}) that can be
+added and multiplied together. Instances of \texttt{Expr} typically represent
+complex numbers, but may also include other ``rings'' like matrix expressions.
+Not all SymPy classes are subclasses of \texttt{Expr}. For instance, logic expressions such
+as \verb|And(x, y)| are subclasses of \texttt{Basic} but not of \texttt{Expr}.
+\end{quote}
 \end{solution}

### 2016-08-16 architecture.tex
Address questions 16 and 17 of reviewer 3
footnote was moved above, to the section 2.2.

@@
 In SymPy every symbolic expression is an instance of a Python \texttt{Basic}
-class, a superclass of all SymPy types providing common methods to all SymPy
-tree-elements, such as traversals.  The children of a node in the
+class\footnote{Some
+  internal classes, such as those used in the polynomial module, do not follow
+  this rule for efficiency reasons.}, a superclass of all SymPy types providing
+common methods to all SymPy tree-elements, such as traversals.  The children of a node in the
 tree are held in the \texttt{args} attribute.  A terminal or leaf node in the
@@
 SymPy class, usually  \texttt{Basic}, \texttt{Expr}, or
-\texttt{Function}. All SymPy classes used for expression trees should be subclasses of the base class
-\texttt{Basic}\footnote{Some
-  internal classes, such as those used in the polynomial module, do not follow
-  this rule for efficiency reasons.}, which defines some basic methods for symbolic expression
-trees. \texttt{Expr} is the subclass for mathematical expressions that can be
+\texttt{Function}. As it was stated before, almost all SymPy classes used for
+expression trees should be subclasses of the base class \texttt{Basic}.
+\texttt{Expr} is the subclass for mathematical expressions (including \texttt{Symbol}) that can be
 added and multiplied together. Instances of \texttt{Expr} typically represent

### 2016-08-16 conclusion_and_future_work.tex
submodules -> modules, per question 37

SymPy is a robust computer algebra system that provides a wide spectrum of
features both in traditional computer algebra and in a plethora of scientific
disciplines. This allows SymPy to be used in a first-class way with other
Python projects, including the scientific Python stack. Unlike many other CASs, SymPy
is designed to be used in an extensible way: both as an end-user
application and as a library.

SymPy expressions are immutable trees of Python objects. SymPy uses Python both
as the internal language and the user language. This permits users to access to
the same
methods that the library implements in order to extend it for their needs.
Additionally, SymPy has a powerful assumptions
system for declaring and deducing mathematical properties of expressions.

SymPy has modules for many areas of mathematics. This includes functions for
simplifying expressions, performing common calculus operations, pretty printing
expressions, solving equations, and representing symbolic matrices. Other included
areas
are discrete math, concrete math, plotting, geometry, statistics,
polynomials, sets, series, vectors, combinatorics, group theory, code
generation, tensors, Lie algebras, cryptography, and special functions.
Additionally, SymPy contains modules targeting certain specific domains,
such as classical mechanics and quantum mechanics.  This breadth of domains has
been engendered by a strong and vibrant user community.
Anecdotally, these users likely chose SymPy because of its ease of access.

% Future work:

Some of the planned future work for SymPy includes work on improving code
generation, improvements to the speed of SymPy (one area of work in this
direction is \href{https://github.com/symengine/symengine}{SymEngine}, a
C++ symbolic manipulation library that is planned to be usable as a
alternative core for SymPy), improving the assumptions system, and improving the
solvers module.

% TODO: Maybe one sentence for each item

% Feel free to add stuff here.

### 2016-08-16 domain_specific.tex
Address point 32 of the reviewer 3

SymPy includes several packages that allow users to solve domain specific
problems. For example, a comprehensive physics package is included that is
useful for solving problems in mechanics, optics, and quantum
mechanics along with support for manipulating physical quantities with units.

\subsection{Classical Mechanics}
One of the core domains that SymPy suports is the physics of classical
mechanics. This is in turn separated into two distinct components:
vector algebra symbolics and mechanics.

\subsubsection{Vector Algebra}
% TODO This section requires some citations.

The \verb|sympy.physics.vector| submodule provides reference frame-, time-, and
space-aware vector and dyadic objects that allow for three-di\-men\-sional
operations such as addition, subtraction, scalar multiplication, inner and
outer products, and cross products. Both of these objects can be written in
very compact notation that make it easy to express the vectors and dyadics in
terms of multiple reference frames with arbitrarily defined relative
orientations. The vectors are used to specify the positions, velocities, and
accelerations of points; orientations, angular velocities, and angular
accelerations of reference frames; and forces and torques. The dyadics are
essentially reference frame-aware $3 \times 3$ tensors~\cite{tai1997generalized}.
The vector and dyadic
objects can be used for any one-, two-, or three-dimensional vector algebra, and
they provide a strong framework for building physics and engineering tools.

The following Python code demonstrates how a vector is created using
the orthogonal unit vectors of three reference frames that are oriented with
respect to each other, and the result of expressing the vector in the $A$
frame. The $B$ frame is oriented with respect to the $A$ frame using Z-X-Z
Euler Angles of magnitude $\pi$, $\frac{\pi}{2}$, and
$\frac{\pi}{3}$, respectively, whereas the $C$ frame is oriented
with respect to the $B$ frame through a simple rotation about the $B$ frame's
$X$ unit vector through $\frac{\pi}{2}$.

\begin{verbatim}
>>> from sympy.physics.vector import ReferenceFrame
>>> A = ReferenceFrame('A')
>>> B = ReferenceFrame('B')
>>> C = ReferenceFrame('C')
>>> B.orient(A, 'body', (pi, pi/3, pi/4), 'zxz')
>>> C.orient(B, 'axis', (pi/2, B.x))
>>> v = 1*A.x + 2*B.z + 3*C.y
>>> v
A.x + 2*B.z + 3*C.y
>>> v.express(A)
A.x + 5*sqrt(3)/2*A.y + 5/2*A.z
\end{verbatim}

\subsubsection{Mechanics}

The \verb|sympy.physics.mechanics| package utilizes the \texttt{sympy.\allowbreak{}physics.\allowbreak{}vector} package
to populate time-aware particle and rigid-body objects to fully describe the
kinematics and kinetics of a rigid multi-body system. These objects store all
of the information needed to derive the ordinary differential or differential
algebraic equations that govern the motion of the system, i.e., the equations
of motion. These equations of motion abide by Newton's laws of motion and can
handle arbitrary kinematic constraints or complex loads. The package
offers two automated methods for formulating the equations of motion based on
Lagrangian Dynamics~\cite{Lagrange1811} and Kane's Method~\cite{Kane1985}.
Lastly, there are automated linearization routines for constrained dynamical
systems~\cite{Peterson2014}.

\subsection{Quantum Mechanics}

The \verb|sympy.physics.quantum| package has extensive capabilities for
performing symbolic quantum mechanics, using Python objects to represent the
different mathematical objects relevant in quantum theory~\cite{Sakurai2010}:
states (bras and kets), operators (unitary, Hermitian, etc.), and basis sets, as
well as operations on these objects such as representations, tensor products,
inner products, outer products, commutators, and anticommutators. The base
objects are designed in the most general way possible to enable any particular
quantum system to be implemented by subclassing the base operators and defining
the relevant class methods to provide system-specific logic.

Symbolic quantum operators and states may be defined, and one can perform
a full range of operations with them.
\begin{verbatim}
>>> from sympy.physics.quantum import Commutator, Dagger, Operator
>>> from sympy.physics.quantum import Ket, qapply
>>> A = Operator('A')
>>> B = Operator('B')
>>> C = Operator('C')
>>> D = Operator('D')
>>> a = Ket('a')
>>> comm = Commutator(A, B)
>>> comm
[A,B]
>>> qapply(Dagger(comm*a)).doit()
-<a|*(Dagger(A)*Dagger(B) - Dagger(B)*Dagger(A))
\end{verbatim}
Commutators can be expanded using common commutator identities:
\begin{verbatim}
>>> Commutator(C+B, A*D).expand(commutator=True)
-[A,B]*D - [A,C]*D + A*[B,D] + A*[C,D]
\end{verbatim}

On top of this set of base objects, a number of specific quantum systems have
been implemented in a fully symbolic framework. These include:

\begin{itemize}

\item Many of the exactly solvable quantum systems, including simple harmonic
oscillator states and raising/lowering operators, infinite square well states,
and 3D position and momentum operators and states.

\item Second quantized formalism of non-relativistic many-body quantum
mechanics~\cite{FetterWalecka2003}.

\item Quantum angular momentum~\cite{Zare1991}. Spin operators and their
eigenstates can be represented in any basis and for any quantum numbers.
A rotation operator representing the Wigner-D matrix, which may be defined
symbolically or numerically, is also implemented to rotate spin eigenstates.
Functionality for coupling and uncoupling of arbitrary spin eigenstates is
provided, including symbolic representations of Clebsch-Gordon coefficients and
Wigner symbols.

\item Quantum information and computing~\cite{Nielsen2011}. Multidimensional
qubit states, and a full set of one- and two-qubit gates are provided and can
be represented symbolically or as matrices/vectors. With these building blocks,
it is possible to implement a number of basic quantum algorithms including the
quantum Fourier transform, quantum error correction, quantum teleportation,
Grover's algorithm, dense coding, etc. In addition, any quantum circuit may be
plotted using the \verb|circuit_plot| function (Figure~\ref{fig-circuitplot-qft}).

\end{itemize}

Here are a few short examples of the quantum information and computing capabilities
in \verb|sympy.physics.quantum|. Start with a simple four-qubit state and flip the second
qubit from the right using a Pauli-X gate:

\begin{verbatim}
>>> from sympy.physics.quantum.qubit import Qubit
>>> from sympy.physics.quantum.gate import XGate
>>> q = Qubit('0101')
>>> q
|0101>
>>> X = XGate(1)
>>> qapply(X*q)
|0111>
\end{verbatim}
Qubit states can also be used in adjoint operations, tensor products, inner/outer
products:
\begin{verbatim}
>>> Dagger(q)
<0101|
>>> ip = Dagger(q)*q
>>> ip
<0101|0101>
>>> ip.doit()
1
\end{verbatim}
Quantum gates (unitary operators) can be applied to transform these states and
then classical measurements can be performed on the results:
\begin{verbatim}
>>> from sympy.physics.quantum.qubit import measure_all
>>> from sympy.physics.quantum.gate import H, X, Y, Z
>>> c = H(0)*H(1)*Qubit('00')
>>> c
H(0)*H(1)*|00>
>>> q = qapply(c)
>>> measure_all(q)
[(|00>, 1/4), (|01>, 1/4), (|10>, 1/4), (|11>, 1/4)]
\end{verbatim}
\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=0.65]{images/circuitplot-qft}
\caption{The circuit diagram for a three-qubit quantum Fourier transform
generated by SymPy.}
\label{fig-circuitplot-qft}
\end{center}
\end{figure}
Lastly, the following example demonstrates creating a three-qubit quantum Fourier
transform, decomposing it into one- and two-qubit gates, and then generating a
circuit plot for the sequence of gates (see Figure~\ref{fig-circuitplot-qft}).
\begin{verbatim}
>>> from sympy.physics.quantum.qft import QFT
>>> from sympy.physics.quantum.circuitplot import circuit_plot
>>> fourier = QFT(0,3).decompose()
>>> fourier
SWAP(0,2)*H(0)*C((0),S(1))*H(1)*C((0),T(2))*C((1),S(2))*H(2)
>>> c = circuit_plot(fourier, nqubits=3)
\end{verbatim}

### 2016-08-16 rebuttal.tex
Address question 20 of reviewer 3
Make symbols usage consistent with the rest
Fixes question 45 of the reviewer 3
Cleanup acknowledgements section
Per question 40 of the reviewer 3
submodules -> modules, per question 37
Removed sentence, see question 36
Address point 32 of the reviewer 3
Address question 31 of the reviewer 3
Address questions 27-29, reviewer 3

@@
 \begin{solution}
-
+This suggestion was implemented.
 \end{solution}
@@
 \begin{solution}
-
+This was replaced with ``solving ODE''.
 \end{solution}
@@
 \begin{solution}
-
+This was fixed.
 \end{solution}
@@
 \begin{solution}
-
+Word ``package'' replaced with ``submodule''.
 \end{solution}
@@
 \begin{solution}
-
+These units were removed.
 \end{solution}
@@
 \begin{solution}
-
+This sentence was removed.
 \end{solution}
@@
 \begin{solution}
-
+This was done.
 \end{solution}
@@
 \begin{solution}
-
+These details were removed.
 \end{solution}
@@
 \begin{solution}
-
+This is fixed now.
 \end{solution}
 
-\question The citation [21] in line 728 should have ``2D'' instead of ``2d''. More specific comments and correction suggestions for the supplement fol- low.
+\question The citation [21] in line 728 should have ``2D'' instead of ``2d''.
 \begin{solution}
+This is fixed.
+\end{solution}
 
-\end{solution}
+More specific comments and correction suggestions for the supplement follow.
 
@@
 \begin{solution}
-
+Now we use \texttt{symbols('a b c d')}, like in other places.
 \end{solution}

### 2016-08-16 numerics.tex
Address questions 27-29, reviewer 3

@@
 The implementation of arbitrary-precision floating-point arithmetic is
-supplied by the mpmath library. Originally, it was developed as a SymPy
+supplied by the mpmath library~\cite{mpmath}. Originally, it was developed as a SymPy
 module but has subsequently been moved to a standalone pure-Python package.
@@
 and numerical computation of limits, derivatives, integrals, infinite
-series, and ODE solutions. All features work in arbitrary precision
+series, and solving ODE. All features work in arbitrary precision
 and use algorithms that allow computing hundreds of digits rapidly
@@
 
-A wide array of higher mathematical functions are implemented
+A wide array of higher mathematical functions is implemented
 with full support for complex values of all parameters and arguments,
@@
 
-Symbolic integration and summation often produces hypergeometric
+Symbolic integration and summation often produce hypergeometric
 and Meijer G-function closed forms (see Subsection~\ref{sec:calculus});

### 2016-08-16 rebuttal.tex
Address question 26 of the reviewer 3
Address question 25 of the reviewer 3

@@
 \begin{solution}
+First your example, unfortunately, doesn't work due to automatic
+simplification (see last input line --- \texttt{exp(100)} cancel).
+Second example is good, but our looks to be more simple.
 
+Applying \texttt{evalf()} has nothing related to \emph{symbolic} computation,
+that method is for numerical evaluation, that should be clear from
+the sentence below ``The evalf method converts a constant symbolic expression
+to a Float\dots''.  Numerical evaluation of the whole expression should
+always solve cancellation problem, as it was stated in the sentence
+``Applying the evalf method to the whole expression solves this problem.''
 \end{solution}
@@
 \begin{solution}
-
+Mpmath version now mentioned in the introduction.
 \end{solution}

### 2016-08-16 rebuttal.tex
Change "printers" to "display formatters" in the introduction
The concept of printer hasn't been introduced yet at this point. This
addresses a comment by reviewer 3.

--- 2016-08-16 rebuttal.tex
@@
 \begin{solution}
-Yes, included.  Second example in that section address computation of
-eigenvalues.  Let us know if we should expand this.
+  % TODO
 \end{solution}
@@
 \begin{solution}
-The mpmath library is not a SymPy submodule, but a separate library,
-that should be clear for a reader from the text.  From that, we believe,
-it should obvious that in general syntax (i.e. names for special
-functions, order of their arguments, etc) will be different, just
-as for any other external library.
+  % TODO
 \end{solution}
@@
 \begin{solution}
-We have done this.
+  % TODO: I do not know how to make bibtex (with apalike) show the editor
+  % here.
 \end{solution}
@@
 It might, however, encourage bad practice. It may also lead to a poor user experience for newbies.
-\label{rev1point1}
 
@@
 \begin{verbatim}
-from numpy import sin, array
-test=array(123)
-sin(test)
+𝔬mνmpyimportsin,arraytest=array(123)sin(test)
 \end{verbatim}
+[sic]
 
@@
 \begin{verbatim}
-from numpy import sin, array
-from sympy import *
-x,y,z=symbols(′xyz′)
-test=array(123)
-sin(test)
+𝔬mνmpyimportsin,array𝔬msympyimport⋅test=array(123)sin(test)x,y,z=symbols(′xyz′)
 \end{verbatim}
+[sic]
 
@@
 I tested using Sympy 1.0 and Python 3.
-
-Later in the text, a similar assertion is made by you (line 485) in reference to a different package that may break sympy.
 
@@
 have added a footnote:
+
+Later in the text, a similar assertion is made by you (line 485) in reference to a different package that may break sympy.
+
 \begin{quote}\texttt{import *} has been used here to aid the
@@
 footnote later in the sentence, to ``\texttt{Basic}'', as suggested
-by reviewer 3, point~\ref{rev3point16}.
+by reviewer 3 (point \ref{rev3point16}).
 \end{solution}
@@
 \begin{solution}
-We have changed the citation to
-
-Kluyver, T., Ragan-Kelley, B., P{\'e}rez, F.,
-Granger, B., Bussonnier, M., Frederic, J., Kelley,
-K., Hamrick, J., Grout,
-J., Corlay, S., et al. (2016). Jupyter notebooks---a publishing format for
-reproducible computational workflows. In \textit{Positioning and Power in Academic
-Publishing: Players, Agents and Agendas: Proceedings of the 20th International
-Conference on Electronic Publishing}, page 87. IOS Press.
-
-as recommended by the Jupyter developers at \url{https://github.com/jupyter/jupyter/issues/190}.
+  % See https://github.com/jupyter/jupyter/issues/190
 \end{solution}
@@
 \begin{solution}
-We have changed ``software'' to ``library''.
+
 \end{solution}
@@
 \begin{solution}
-We have moved the footnote inline, and have noted that the examples should
-work with Python 2.7, 3.2, 3.3, 3.4, or 3.5, with any operating system that
-supports Python.
 
-We have added a footnote discussing the concerns with \texttt{import *} (see
-reviewer 3, point~\ref{rev1point1}).
 \end{solution}
@@
 \begin{solution}
-Line 119 is the output of line 118, so not showing it would
+Line 119 is the output of line 118, so now showing it would
 be incorrect.
@@
 \begin{solution}
-See next answer.
+
 \end{solution}
@@
 \begin{solution}
-This paragraph was replaced by:
-\begin{quote}
-Importantly, SymPy expressions are immutable. This simplifies the design of
-SymPy by allowing expression interning. It also enables expressions to be
-hashed, that is used to implement caching in SymPy.
-\end{quote}
+
 \end{solution}
@@
 \begin{solution}
-That was removed.
+
 \end{solution}
@@
 \begin{solution}
-Done.
+  % I don't personally agree with this, but it's a minor point. --Aaron
 \end{solution}
@@
 \begin{solution}
-Done.
+
 \end{solution}
@@
 \begin{solution}
-Now we use verbatim environment here.
+
 \end{solution}
@@
 \begin{solution}
-This part was removed and footnote moved to the mentioned
-text in the section 2.2, "The Core".
+
 \end{solution}
@@
 \begin{solution}
-This paragraph now looks as:
-\begin{quote}
-The typical way to create a custom SymPy object is to subclass an existing
-SymPy class, usually  \texttt{Basic}, \texttt{Expr}, or
-\texttt{Function}. As it was stated before, almost all SymPy classes used for
-expression trees should be subclasses of the base class \texttt{Basic}.
-\texttt{Expr} is the subclass for mathematical expressions (including \texttt{Symbol}) that can be
-added and multiplied together. Instances of \texttt{Expr} typically represent
-complex numbers, but may also include other ``rings'' like matrix expressions.
-Not all SymPy classes are subclasses of \texttt{Expr}. For instance, logic expressions such
-as \verb|And(x, y)| are subclasses of \texttt{Basic} but not of \texttt{Expr}.
-\end{quote}
+
 \end{solution}
@@
      >>> from sympy import *
-sympy/external/importtools.py:145: UserWarning: gmpy2 module is not installed
-  warnings.warn("%s module is not installed" % module, UserWarning)
      >>> x = symbols('x')
@@
      +-mrv_leadterm(_p*sin(1/_p), _p) = (1, 0)
-| +-mrv(_p*sin(1/_p), _p) = ({_p: _Dummy_19}, {}, _Dummy_19*sin(1/_Dummy_19))
-| | +-mrv(_p, _p) = ({_p: _Dummy_19}, {}, _Dummy_19)
-| | +-mrv(sin(1/_p), _p) = ({_p: _Dummy_20}, {}, sin(1/_Dummy_20))
-| |   +-mrv(1/_p, _p) = ({_p: _Dummy_20}, {}, 1/_Dummy_20)
-| |     +-mrv(_p, _p) = ({_p: _Dummy_20}, {}, _Dummy_20)
-| +-rewrite(_Dummy_19*sin(1/_Dummy_19), {exp(_p): _Dummy_19}, {}, _p, _w) = (sin(_w)/_w, -_p)
+     | +-mrv(_p*sin(1/_p), _p) = ({_p: _Dummy_14}, {}, _Dummy_14*sin(1/_Dummy_14))
+     | | +-mrv(_p, _p) = ({_p: _Dummy_14}, {}, _Dummy_14)
+     | | +-mrv(sin(1/_p), _p) = ({_p: _Dummy_15}, {}, sin(1/_Dummy_15))
+     | |   +-mrv(1/_p, _p) = ({_p: _Dummy_15}, {}, 1/_Dummy_15)
+     | |     +-mrv(_p, _p) = ({_p: _Dummy_15}, {}, _Dummy_15)
+     | +-rewrite(_Dummy_14*sin(1/_Dummy_14), {exp(_p): _Dummy_14}, {}, _p, _w)
+             = (sin(_w)/_w, -_p)
      | | +-sign(_p, _p) = 1
@@
      |   | +-mrv_leadterm(_w, _w) = (1, -1)
-|   | | +-mrv(_w, _w) = ({_w: _Dummy_23}, {}, _Dummy_23)
-|   | | +-rewrite(_Dummy_23, {exp(_w): _Dummy_23}, {}, _w, _w) = (1/_w, -_w)
-|   | | | +-sign(_w, _w) = 1
+     |   | | +-mrv(_w, _w) = ({_w: _Dummy_18}, {}, _Dummy_18)
+     |   | | +-rewrite(_Dummy_18, {exp(_w): _Dummy_18}, {}, _w, _w) = (1/_w, -_w)
+     |   | | | +-sign(_w,_w)=1
      |   | | | +-limitinf(1, _w) = 1
@@
      |   +-limitinf(_w, _w) = oo
+     |   +-limitinf(_w, _w) = oo
      +-sign(0, _p) = 0
      +-limitinf(1, _p) = 1
-1
 \end{verbatim}

### 2016-08-16 introduction.tex
Address question 26 of the reviewer 3
Change "printers" to "display formatters" in the introduction
The concept of printer hasn't been introduced yet at this point. This
addresses a comment by reviewer 3.

@@
 Simultaneously, it enables developers to focus on mathematics, rather than
-language design.  SymPy officially supports Python 2.6, 2.7 and 3.2--3.5.
+language design.
 
@@
 However, SymPy exposes a rich interactive display system, and supports
-registering display formatters with Jupyter~\cite{kluyver2016jupyter} frontends,
+registering display formatters with Jupyter~\cite{perez2007ipython} frontends,
 including the Notebook and Qt Console, which will render SymPy expressions
@@
 
-The remainder of this paper discusses key components of the SymPy library.
+The remainder of this paper discusses key components of the SymPy software.
 Section~\ref{sec:architecture} discusses the architecture of SymPy.
@@
 and quantum mechanics. Conclusions and future directions for SymPy are given
-in section~\ref{sec:conclusion}.  All examples in this paper use SymPy version 1.0.
-
-The following statement imports all SymPy functions into the global Python
-namespace.\footnote{\texttt{import *} has been used here to aid the
-  readability of the paper, but is best to avoid such wildcard import
-  statements in production code, as they make it unclear which names are
-  present in the namespace. Furthermore, imported names could clash with
-  already existing imports from another package. For example, SymPy, the
-  standard Python \texttt{math} library, and NumPy all define the \texttt{exp}
-  function, but only the SymPy one will work with SymPy symbolic expressions.}
-From here on, all examples in this paper assume that this statement has been
-executed:
-
-\begin{verbatim}
->>> from sympy import *
-\end{verbatim}
+in section~\ref{sec:conclusion}.

### 2016-08-16 matrices.tex
Address question 23 of the reviewer 3

Besides being an important feature in its own right, computations on
matrices with symbolic entries are important for many algorithms
within SymPy.  The following code shows some basic usage of the
\texttt{Matrix} class.
\begin{verbatim}
>>> A = Matrix(2, 2, [x, x + y, y, x])
>>> A
Matrix([
[x, x + y],
[y,     x]])
\end{verbatim}

SymPy matrices support common symbolic linear algebra manipulations, including
matrix addition, multiplication, exponentiation, computing determinants,
solving linear systems, and computing inverses using LU decomposition, LDL
decomposition, Gauss-Jordan elimination, Cholesky decomposition, Moore-Penrose
pseudoinverse, and adjugate matrix.

All operations are performed symbolically. For instance, eigenvalues are computed
by generating the characteristic polynomial using the Berkowitz algorithm and
then solving it using polynomial routines.

\begin{verbatim}
>>> A.eigenvals()
{x - sqrt(y*(x + y)): 1, x + sqrt(y*(x + y)): 1}
\end{verbatim}

Internally these matrices store the elements as lists of lists, making
it a dense representation.\footnote{Similar to the polynomials module, dense
  here means that all entries are stored in memory, contrasted with a sparse
  representation where only nonzero entries are stored.} For storing sparse
matrices, the \verb|SparseMatrix| class can be used.

SymPy also supports matrices with symbolic dimension values. \verb|MatrixSymbol|
represents a matrix with dimensions $m\times n$, where $m$ and $n$ can be
symbolic. Matrix addition and multiplication, scalar operations, matrix inverse,
and transpose are stored symbolically as matrix expressions.

Block matrices are also implemented in SymPy. \verb|BlockMatrix| elements can
be any matrix expression, including explicit matrices, matrix symbols, and
other block matrices. All functionalities of matrix expressions are also
present in \verb|BlockMatrix|.

When symbolic matrices are combined with the assumptions module for logical
inference, they provide powerful reasoning over invertibility,
semi-definiteness, orthogonality, etc., which are valuable in the construction
of numerical linear algebra systems.

More examples for \verb|Matrix| and \verb|BlockMatrix| may be found in the
supplement.

### 2016-08-16 introduction.tex
Update the Jupyter citation
As recommended by the Jupyter developers at
https://github.com/jupyter/jupyter/issues/190.

@@
 However, SymPy exposes a rich interactive display system, and supports
-registering display formatters with Jupyter~\cite{perez2007ipython} frontends,
+registering display formatters with Jupyter~\cite{kluyver2016jupyter} frontends,
 including the Notebook and Qt Console, which will render SymPy expressions

### 2016-08-16 rebuttal.tex
Change "software" to "library" in the introduction
As per a comment by reviewer 3.
Mention SymPy Live in the introduction (per editor suggestion)

--- 2016-08-16 rebuttal.tex
@@
 \begin{solution}
-  % TODO
+This sentence was added to the introduction:
+\begin{quote}
+All examples could be tested on the \href{http://live.sympy.org}{SymPy
+Live} instance, that is an online Python shell, which uses the Google
+App Engine to execute SymPy code.
+\end{quote}
 \end{solution}
@@
 \begin{solution}
-
+This sentence was omitted, just as for \texttt{polys} module.
 \end{solution}

### 2016-08-16 introduction.tex
Change "software" to "library" in the introduction
As per a comment by reviewer 3.
Mention SymPy Live in the introduction (per editor suggestion)

--- 2016-08-16 introduction.tex
@@
 and quantum mechanics. Conclusions and future directions for SymPy are given
-in section~\ref{sec:conclusion}.  All examples in this paper use SymPy version 1.0.
+in section~\ref{sec:conclusion}.  All examples in this paper use SymPy
+version 1.0, examples in section~\ref{sec:mpmath} use mpmath
+version 0.19.
 
@@
 \end{verbatim}
+
+All examples could be tested on the \href{http://live.sympy.org}{SymPy
+Live} instance, that is an online Python shell, which uses the Google
+App Engine to execute SymPy code.

### 2016-08-16 rebuttal.tex
Fix some text in the rebuttal
It was in the response but it was supposed to be from the reviewer.
Fix some formatting

--- 2016-08-16 rebuttal.tex
@@
 
+Later in the text, a similar assertion is made by you (line 485) in reference to a different package that may break sympy.
+
 \begin{solution}
@@
 footnote later in the sentence, to ``\texttt{Basic}'', as suggested
-by reviewer 3 (point \ref{rev3point16}).
+by reviewer 3, point~\ref{rev3point16}.
 \end{solution}

### 2016-08-16 architecture.tex
Move note about version 1.0 to the text, and note what Python versions it works with
Also add a citation for NumPy from the other footnote.

--- 2016-08-16 architecture.tex
@@
 % symbols, various ways to declare them
+
+The following statement imports all SymPy functions into the global Python
+namespace.\footnote{\texttt{import *} has been used here to aid the
+  readability of the paper, but is best to avoid such wildcard import
+  statements in production code, as they make it unclear which names are
+  present in the namespace. Furthermore, imported names could clash with
+  already existing imports from another package. For example, SymPy, the
+  standard Python \texttt{math} library, and the popular numeric library
+  NumPy~\cite{van2011numpy} all define the \texttt{exp} function, but only the
+  SymPy one will work with SymPy symbolic expressions.} From here on, all
+examples in this paper assume that this statement has been executed. All
+examples in this paper use SymPy version 1.0, and should work with Python
+versions 2.7, 3.2, 3.3, 3.4, or 3.5 in any operating system that supports
+Python.
+
+\begin{verbatim}
+>>> from sympy import *
+\end{verbatim}
 
@@
 SymPy by allowing expression interning. It also enables expressions to be
-hashed, that is used to implement caching in SymPy.
+hashed and stored in Python dictionaries, thereby permitting features
+such as caching.
 
@@
 
-A computer algebra system stores mathematical expressions as data
+A computer algebra system (CAS) represents mathematical expressions as data
 structures.  For example, the mathematical expression $x + y$ is represented as
@@
 
-By assuming the most general case, that \texttt{t} is complex by default, SymPy
+By assuming the most general case, that symbols are complex by default, SymPy
 avoids performing mathematically invalid operations. However, in many cases

### 2016-08-16 rebuttal.tex
Move note about version 1.0 to the text, and note what Python versions it works with
Also add a citation for NumPy from the other footnote.
Replace "stored" with "used as keys" as suggested by reviewer 3
Address 2nd question of the editor (I hope)

--- 2016-08-16 rebuttal.tex
@@
 \begin{solution}
-  % TODO. I'm unclear what the editor is asking here.
+It's unclear what you are asking for.  Both Python standard library
+and SymPy have exp callable, so it depends which symbols are
+imported in your global namespace: exp function from the math
+library or exp class from the SymPy.  In later case a special
+symbolic constant will be produced.
+
+We have added a sentence to the section 2.2, that shows symbolic
+constants as an example of leaf nodes:
+\begin{quote}
+Symbols or symbolic constants, like $e$ or $\pi$ are another examples of
+leaf nodes.
+\begin{verbatim}
+>>> exp(1)
+E
+>>> exp(1).args
+()
+>>> x.args
+()
+\end{verbatim}
+\end{quote}
 \end{solution}

### 2016-08-16 introduction.tex
Address question 1 of the reviewer 3

@@
-SymPy is a full featured computer algebra system (CAS) written in the Python
-programming language~\cite{lutz2013learning}.
+SymPy is a full featured computer algebra system (CAS) written in the
+Python~\cite{lutz2013learning}
+programming language.
 It is free and open source software, licensed under the 3-clause BSD
@@
 From here on, all examples in this paper assume that this statement has been
-executed:
+executed\footnote{Three greater-than signs denote standard
+Python prompt for interactive session.}:
 

### 2016-08-16 acknowledgements.tex

% People should add their funding agency acknowledgements here.
The author of this paper Francesco Bonazzi thanks the Deutsche
Forschungsgemeinschaft (DFG) for its financial support via the International
Research Training Group 1524 "Self-Assembled Soft Matter Nano-Structures at
Interfaces."

### 2016-08-16 diophantine.tex

Diophantine equations play a central role in number theory.  A
Diophantine equation has the form, $f(x_1, x_2, \dotsc, x_n) = 0$ where $n \geq
2$ and $x_1, x_2, \dotsc, x_n$ are integer variables. If there are $n$
integers $a_1, a_2, \dotsc, a_n$ such that $x_1 = a_1, x_2 = a_2, \dotsc, x_n =
a_n$ satisfies the above equation, the equation is said to be solvable.

Currently, the following five types of Diophantine equations can be solved using
SymPy's Diophantine module ($a_1, \ldots, a_{n+1}$, $a$, $b$, $c$, $d$, $e$,
$f$, and $k$ are explicitly given rational constants):

\begin{itemize}
    \item Linear Diophantine equations: $a_1x_1 + a_2x_2 + \dotsb{} + a_{n}x_{n} = b$
    \item General binary quadratic equation: $ax^2 + bxy + cy^2 + dx + ey + f = 0$
    \item Homogeneous ternary quadratic equation: $ax^2 + by^2 + cz^2 + dxy + eyz + fzx = 0$
    \item Extended Pythagorean equation: $a_{1}x_{1}^2 + a_{2}x_{2}^2 + \dotsb{} + a_{n}x_{n}^2 = a_{n+1}x_{n+1}^2$
    \item General sum of squares: $x_{1}^2 + x_{2}^2 + \dotsb{} + x_{n}^2 = k$
\end{itemize}

The \verb|diophantine| function factors the equation it is given (if
possible), solves each factor separately, and combines the results to give
a final solution set.
The following examples illustrate some of the
basic functionalities of the Diophantine module.

\begin{verbatim}
>>> from sympy.solvers.diophantine import *
>>> diophantine(2*x + 3*y - 5)
set([(3*t_0 - 5, -2*t_0 + 5)])

>>> diophantine(2*x + 4*y - 3)
set()

>>> diophantine(x**2 - 4*x*y + 8*y**2 - 3*x + 7*y - 5)
set([(2, 1), (5, 1)])

>>> diophantine(x**2 - 4*x*y + 4*y**2 - 3*x + 7*y - 5)
set([(-2*t**2 - 7*t + 10, -t**2 - 3*t + 5)])

>>> diophantine(3*x**2 + 4*y**2 - 5*z**2 + 4*x*y - 7*y*z + 7*z*x)
set([(-16*p**2 + 28*p*q + 20*q**2,
3*p**2 + 38*p*q - 25*q**2,
4*p**2 - 24*p*q + 68*q**2)])

>>> x1, x2, x3, x4, x5, x6 = symbols('x1, x2, x3, x4, x5, x6')
>>> diophantine(9*x1**2 + 16*x2**2 + x3**2 + 49*x4**2 + 4*x5**2 - 25*x6**2)
set([(70*t1**2 + 70*t2**2 + 70*t3**2 + 70*t4**2 - 70*t5**2, 105*t1*t5,
420*t2*t5, 60*t3*t5, 210*t4*t5,
42*t1**2 + 42*t2**2 + 42*t3**2 + 42*t4**2 + 42*t5**2)])

>>> a, b, c, d = symbols('a b c d')
>>> diophantine(a**2 + b**2 + c**2 + d**2 - 23)
set([(2, 3, 3, 1)])
\end{verbatim}

### 2016-08-16 features.tex

% Features to discuss in-depth:

Although SymPy's extensive feature set cannot be covered in-depth in this paper,
calculus and other bedrock areas are discussed in their own subsections.
Additionally,
Table~\ref{features-table} gives a compact listing of all major capabilities
present in the SymPy codebase. This grants a sampling from the breadth of
topics and application domains that SymPy services. Unless stated otherwise,
all features noted in Table~\ref{features-table} are symbolic in nature.
Numeric features are discussed in Section~\ref{sec:numerics}.

\begin{longtable}[htbc]{p{0.20\linewidth}p{0.73\linewidth}}
\caption{SymPy Features and Descriptions\label{features-table}}\\
\toprule
\textbf{Feature} & \textbf{Description} \\
\midrule
Calculus & Algorithms for computing derivatives, integrals, and limits.\\

Category Theory & Representation of objects, morphisms, and diagrams. Tools
for drawing diagrams with Xy-pic.\\

Code Generation & Generation of compilable and executable code in a
variety of different programming languages from expressions directly. Target
languages include C, Fortran, Julia, JavaScript,
Mathematica, MATLAB and Octave, Python, and Theano.\\

Combinatorics \& Group Theory & Permutations, combinations,
partitions, subsets, various permutation groups (such as polyhedral, Rubik,
symmetric, and others), Gray codes~\cite{Nijenhuis1978combinatorial},
and Prufer sequences~\cite{biggs1976graph}.\\

Concrete Math & Summation, products, tools for determining whether summation
and product expressions are convergent, absolutely convergent, hypergeometric,
and for determining other properties; computation of Gosper's normal form~\cite{petkovvsek1996bak} for two univariate polynomials.\\

Cryptography & Block and stream ciphers, including shift, Affine,
substitution, Vigen\`{e}re's, Hill's, bifid, RSA, Kid RSA,
linear-feedback shift registers, and Elgamal encryption.\\

Differential Geometry & Representations of manifolds, metrics, tensor
products, and coordinate systems in Riemannian and pseudo-Riemannian
geometries~\cite{FunctionalDifferentialGeometry}.\\
% TODO: Someone verify that this is a good summary of the diffgeom module

Geometry & Representations of 2D geometrical entities, such as lines and
circles. Enables queries on these entities, such as asking the area of an
ellipse, checking for collinearity of a set of
points, or finding the intersection between objects.\\

Lie Algebras & Representations of Lie algebras and root systems.\\

Logic & Boolean expressions, equivalence testing, satisfiability, and normal
forms.\\

Matrices & Tools for creating matrices of symbols and expressions.
Both sparse and dense representations, as well as symbolic linear
algebraic operations (e.g., inversion and factorization), are
supported.\\

Matrix Expressions & Matrices with symbolic dimensions (unspecified entries).
Block matrices.\\

Number Theory & Prime number generation, primality testing, integer
factorization, continued fractions, Egyptian fractions, modular arithmetic,
quadratic residues, partitions, binomial and multinomial coefficients,
prime number tools, hexidecimal digits of $\pi$, and integer factorization. \\

Plotting & Hooks for visualizing expressions via matplotlib~\cite{Hunter:2007}
or as text drawings when lacking a graphical back-end. 2D function plotting,
3D function
plotting, and 2D implicit function plotting are supported.\\

Polynomials & Polynomial algebras over various coefficient domains.
Functionality ranges from simple operations (e.g., polynomial division) to
advanced computations (e.g., Gr\"obner bases~\cite{adams1994introduction} and multivariate
factorization over algebraic number domains).\\

Printing & Functions for printing SymPy expressions in the terminal with ASCII
or Unicode characters and converting SymPy expressions to \LaTeX{} and
MathML.\\

Quantum Mechanics & Quantum states, bra--ket notation, operators, basis sets,
representations, tensor products, inner products, outer products, commutators,
anticommutators, and specific quantum system implementations.\\

Series & Series expansion, sequences, and limits of sequences.
This includes Taylor, Laurent, and Puiseux series as well as special series, such
as Fourier and formal power series.\\

Sets & Representations of empty, finite, and infinite sets (including
special sets such as for all natural, integer, and complex numbers). Operations
on sets such as union, intersection, Cartesian product, and building sets from
other sets are supported.\\

Simplification & Functions for manipulating and simplifying expressions.
Includes algorithms for simplifying hypergeometric functions, trigonometric
expressions, rational functions, combinatorial functions, square root
denesting, and common subexpression elimination.\\

Solvers & Functions for symbolically solving equations, systems
of equations, both linear and non-linear, inequalities, ordinary differential
equations, partial differential equations, Diophantine equations, and
recurrence relations.\\

Special Functions & Implementations of a number of well known special functions,
including Dirac delta, Gamma, Beta, Gauss error functions, Fresnel integrals,
Exponential integrals, Logarithmic integrals, Trigonometric integrals, Bessel,
Hankel, Airy, B-spline, Riemann Zeta, Dirichlet eta, polylogarithm, Lerch
transcendent, hypergeometric, elliptic integrals, Mathieu, Jacobi polynomials,
Gegenbauer polynomial, Chebyshev polynomial, Legendre polynomial, Hermite
polynomial, Laguerre polynomial, and
spherical harmonic functions.\\

Statistics & Support for a random variable type as well as the ability to
declare this variable from prebuilt distribution functions such as
Normal, Exponential, Coin, Die, and other custom distributions~\cite{StatsMRocklin}.\\

Tensors & Symbolic manipulation of indexed objects.\\

Vectors & Basic operations on vectors and differential calculus with respect
to 3D Cartesian coordinate systems.\\
\bottomrule

\end{longtable}

\subsection{Simplification}

\input{simplification}

\subsection{Calculus}
\label{sec:calculus}
\input{calculus}

\subsection{Polynomials}
\input{polys}

\subsection{Printers}

\input{printers.tex}

% Solvers (regular equations, maybe also mention other types of solvers like ODEs/recurrence/Diophantine)
\subsection{Solvers}
\input{solvers.tex}

% Matrices (worth emphasizing that they are symbolic)
\subsection{Matrices}

\input{matrices}

### 2016-08-17 rebuttal.tex
Address question 1 of the reviewer 3
Address question 46 of reviewer 3

@@
 \begin{solution}
-  % TODO: It seems better to start the paper with "SymPy is" rather than
-  % "Python is". The reviewer perhaps wants us to not mention Python until the
-  % second paragraph, but Python is important enough to SymPy's design that it
-  % warrants mentioning in the first sentence, in my opinion. --Aaron
+Citation was moved and footnote for Python interactive prompt was
+added where it was used first time.
+We believe that the paper assumes minimal familiarity with Python and
+that should be covered either with citation [25] or with the Python tutorial.
 \end{solution}
@@
 \begin{solution}
+We believe this is not a good idea.  SymPy Gamma currently
+don't export output as Jupyter notebook or other external
+format, but HTML.
 
+URL to the project site was added.
 \end{solution}

### 2016-08-17 gamma.tex
Address question 46 of reviewer 3

\href{http://gamma.sympy.org}{SymPy Gamma} is a simple web application that runs on Google App Engine.
It executes and displays the results of SymPy expressions as well as
additional related computations, in a fashion similar to that of
Wolfram\textbar{}Alpha. For instance, entering an integer will display
its prime factors, digits in the base-10 expansion, and a factorization
diagram. Entering a function will display its docstring; in general,
entering an arbitrary expression will display its derivative, integral,
series expansion, plot, and roots.

SymPy Gamma also has several features beyond just computing the
results using SymPy.

\begin{itemize}
\item
  SymPy Gamma displays integration and differentiation steps in detail, which
  can be viewed in Figure~\ref{fig:integralsteps}:\par
  {
    \centering
    \includegraphics[width=0.7\textwidth]{integral_steps.png}
    \captionof{figure}{Integral steps of $\tan (x)$}\label{fig:integralsteps}
    \par
  }
\item
  SymPy Gamma displays the factor tree diagrams for different numbers.
\item
  SymPy Gamma saves user search queries, and offers many such similar features
  for free, which Wolfram\textbar{}Alpha only offers to its paid users.
\end{itemize}
Every input query from the user on SymPy Gamma is first parsed by its
own parser capable of handling several different forms of function names
which SymPy as a library does not support. For instance, SymPy Gamma
supports queries like \texttt{sin\ x}, whereas SymPy will only recognise
\verb|sin(x)|.

This parser converts the input query to the equivalent SymPy readable code,
which is then processed by SymPy, and the result is finally printed
with the built-in \LaTeX{} output and rendered by the SymPy Gamma web application.

### 2016-08-17 gruntz.tex
Address question 44 from the reviewer 3

@@
 limit. In the above case the limit is $2$, as can be
-verified by SymPy:
+verified by SymPy\footnote{To see intermediate steps, discussed above, interested
+reader can switch on debugging output by setting
+\texttt{SYMPY\_DEBUG} environment variable, before importing anything from
+the SymPy namespace.}:
 

### 2016-08-17 gruntz.tex
SYMPY_DEBUG needs to be set specifically to True

--- 2016-06-11 gruntz.tex
@@
 limit. In the above case the limit is $2$, as can be
-verified by SymPy:
+verified by SymPy\footnote{To see intermediate steps, discussed above, interested
+readers can switch on debugging output by setting
+the environment variable \texttt{SYMPY\_DEBUG=True}, before importing anything from
+the SymPy namespace.}:
 

### 2016-08-17 rebuttal.tex
Address question 44 from the reviewer 3
Remove reviewer names for now
We do not know if they will opt for public review.
Clear some trailing whitespace
Fix some grammar
Fix some grammar/formatting

@@
 and SymPy have exp callable, so it depends which symbols are
-imported in your global namespace: exp function from the math
-library or exp class from the SymPy.  In later case a special
+imported in your global namespace: the \texttt{exp} function from the \texttt{math}
+module or \texttt{exp} class from SymPy.  In the later case a special
 symbolic constant will be produced.
@@
 \begin{quote}
-Symbols or symbolic constants, like $e$ or $\pi$ are another examples of
+Symbols or symbolic constants, like $e$ or $\pi$, are other examples of
 leaf nodes.
@@
 
-\section{Reviewer 1 (James Davenport)}
+\section{Reviewer 1}
 \subsection{Basic reporting}
@@
 
-\section{Reviewer 2 (Michael Croucher)}
+\section{Reviewer 2}
 
@@
 
-\section{Reviewer 3 (Vedran Sego)}
+\section{Reviewer 3}
 
@@
 \begin{solution}
-
+Footnote was added right before the final example from the Gruntz algorithm section.
 \end{solution}

### 2016-08-17 rebuttal.tex
Some more grammar fixes

@@
 imported in your global namespace: the \texttt{exp} function from the \texttt{math}
-module or \texttt{exp} class from SymPy.  In the later case a special
+module or the \texttt{exp} class from SymPy.  In the later case a special
 symbolic constant will be produced.

### 2016-08-17 rebuttal.tex
Some more for the editor comment 4

--- 2016-08-17 rebuttal.tex
@@
 \begin{solution}
-It's unclear what you are asking for.  Both Python standard library
-and SymPy have exp callable, so it depends which symbols are
-imported in your global namespace: the \texttt{exp} function from the \texttt{math}
-module or \texttt{exp} class from SymPy.  In the later case a special
-symbolic constant will be produced.
+  It's unclear what you are asking for. Both Python standard library and SymPy
+  have exp callable, so it depends which symbols are imported in your global
+  namespace: the \texttt{exp} function from the \texttt{math} module or the
+  \texttt{exp} class from SymPy. In the later case a special symbolic constant
+  will be produced.
 
@@
 \begin{solution}
-Yes, included.  Second example in that section address computation of
-eigenvalues.  Let us know if we should expand this.
+Yes, they are included.  The second example in that section addresses the computation of
+eigenvalues.  Singular values can be computed with
+\texttt{A.singular\_values()}. We have not included this example, as the
+singular value expressions for our example matrix \texttt{A} are quite large,
+but we have added ``singular values'' to the list of features in the second
+paragraph of the section.
 \end{solution}

### 2016-08-17 matrices.tex
Some more for the editor comment 4

@@
 decomposition, Gauss-Jordan elimination, Cholesky decomposition, Moore-Penrose
-pseudoinverse, and adjugate matrix.
+pseudoinverse, singular values, and adjugate matrix.
 

### 2016-08-17 numerics.tex
Fix some whitespace in some examples

@@
 \subsection{The mpmath library}
+\label{sec:mpmath}
 
@@
 \texttt{complex}. The floating-point precision is controlled by a global
-context:\footnote{All examples in this section use mpmath version 0.19.}
+context:
 
@@
 >>> mpmath.mp.dps = 15
->>> mpmath.meijerg([[],[0]],[[-0.5,-1,-1.5],[]],10000)
+>>> mpmath.meijerg([[],[0]], [[-0.5,-1,-1.5],[]], 10000)
 mpf('2.4392576907199564e-94')
@@
 \begin{verbatim}
->>> meijerg([[],[0]],[[-S(1)/2,-1,-S(3)/2],[]],10000).evalf()
+>>> meijerg([[],[0]], [[-S(1)/2,-1,-S(3)/2],[]], 10000).evalf()
 2.43925769071996e-94

### 2016-08-17 rebuttal.tex
Add some more for editor comment 6

@@
 The mpmath library is not a SymPy submodule, but a separate library,
-that should be clear for a reader from the text.  From that, we believe,
-it should obvious that in general syntax (i.e. names for special
-functions, order of their arguments, etc) will be different, just
-as for any other external library.
+which we have indicated in the text. Most function names in mpmath are the same as SymPy, but it is not true in every case. We have
+added an example to the text showing an instance where they diverge
+(summations).
 \end{solution}

### 2016-08-17 numerics.tex
Add some more for editor comment 6

@@
 
+Most mpmath and SymPy functions use the same naming scheme, although this is
+not true in every case. For example, the symbolic SymPy summation expression
+\texttt{Sum(f(x), (x, a, b))} representing $\sum_{x=a}^b f(x)$ is represented
+in mpmath as \texttt{nsum(f, (a, b))}, where \texttt{f} is a numeric Python
+function.
+
 The mpmath library supports

### 2016-08-17 rebuttal.tex
Replace "done" with a complete sentence
Grammar
Fix some LaTeX quotes

@@
 \begin{solution}
-Done.
+We have done this.
 \end{solution}
@@
 \begin{solution}
-Now we use verbatim environment here.
+We now use a verbatim environment here.
 \end{solution}
@@
 This part was removed and footnote moved to the mentioned
-text in the section 2.2, "The Core".
+text in the section 2.2, ``The Core''.
 \end{solution}

### 2016-08-17 rebuttal.tex
Change up the Basic, Expr, Function paragraph a bit

@@
 \begin{solution}
-This paragraph now looks as:
+This paragraph now reads:
 \begin{quote}
 The typical way to create a custom SymPy object is to subclass an existing
-SymPy class, usually  \texttt{Basic}, \texttt{Expr}, or
-\texttt{Function}. As it was stated before, almost all SymPy classes used for
-expression trees should be subclasses of the base class \texttt{Basic}.
-\texttt{Expr} is the subclass for mathematical expressions (including \texttt{Symbol}) that can be
-added and multiplied together. Instances of \texttt{Expr} typically represent
-complex numbers, but may also include other ``rings'' like matrix expressions.
-Not all SymPy classes are subclasses of \texttt{Expr}. For instance, logic expressions such
-as \verb|And(x, y)| are subclasses of \texttt{Basic} but not of \texttt{Expr}.
+SymPy class, usually \texttt{Basic}, \texttt{Expr}, or \texttt{Function}. As
+it was stated before, all SymPy classes used for expression trees should be
+subclasses of the base class \texttt{Basic}. \texttt{Expr} is the
+\texttt{Basic} subclass for mathematical that can be added and multiplied
+together. The most commonly seen classes in SymPy are subclasses of
+\texttt{Expr}, including \texttt{Add}, \texttt{Mul}, and \texttt{Symbol}.
+Instances of \texttt{Expr} typically represent complex numbers, but may also
+include other ``rings'', like matrix expressions. Not all SymPy classes are
+subclasses of \texttt{Expr}. For instance, logic expressions, such as
+\verb|And(x, y)|, are subclasses of \texttt{Basic} but not of \texttt{Expr}.
 \end{quote}

### 2016-08-17 architecture.tex
Replace "stored" with "used as keys" as suggested by reviewer 3
Address 2nd question of the editor (I hope)
Remove outdated comment from paper
Fix some grammar
Some more grammar fixes
Change up the Basic, Expr, Function paragraph a bit

--- 2016-08-16 architecture.tex
@@
 hashed, that is used to implement caching in SymPy.
-
-%% I volunteer to write this section. --Aaron
-%%
-%% Representing symbolic expressions using Python objects
 
@@
 >>> expr.args[0].args
+()
+\end{verbatim}
+Symbols or symbolic constants, like $e$ or $\pi$, are examples of
+leaf nodes.
+\begin{verbatim}
+>>> exp(1)
+E
+>>> exp(1).args
+()
+>>> x.args
 ()
@@
 The typical way to create a custom SymPy object is to subclass an existing
-SymPy class, usually  \texttt{Basic}, \texttt{Expr}, or
-\texttt{Function}. As it was stated before, almost all SymPy classes used for
-expression trees should be subclasses of the base class \texttt{Basic}.
-\texttt{Expr} is the subclass for mathematical expressions (including \texttt{Symbol}) that can be
-added and multiplied together. Instances of \texttt{Expr} typically represent
-complex numbers, but may also include other ``rings'' like matrix expressions.
-Not all SymPy classes are subclasses of \texttt{Expr}. For instance, logic expressions such
-as \verb|And(x, y)| are subclasses of \texttt{Basic} but not of \texttt{Expr}.
+SymPy class, usually \texttt{Basic}, \texttt{Expr}, or \texttt{Function}. As
+it was stated before, all SymPy classes used for expression trees should be
+subclasses of the base class \texttt{Basic}. \texttt{Expr} is the
+\texttt{Basic} subclass for mathematical that can be added and multiplied
+together. The most commonly seen classes in SymPy are subclasses of
+\texttt{Expr}, including \texttt{Add}, \texttt{Mul}, and \texttt{Symbol}.
+Instances of \texttt{Expr} typically represent complex numbers, but may also
+include other ``rings'', like matrix expressions. Not all SymPy classes are
+subclasses of \texttt{Expr}. For instance, logic expressions, such as
+\verb|And(x, y)|, are subclasses of \texttt{Basic} but not of \texttt{Expr}.
 

### 2016-08-17 rebuttal.tex
Reinstate "dictionary of keys" and clarify some things
See also this discussion from GitHub
https://github.com/sympy/sympy-paper/pull/144#discussion_r63944359

@@
 \begin{solution}
-This sentence was omitted, just as for \texttt{polys} module.
+``Dictionary of Keys'' (DOK) refers specifically to a sparse matrix
+representation where entries are stored as \texttt{(row, column)} pairs
+mapping to the elements. See for instance \texttt{scipy.sparse.dok\_sparse}. We
+have fixed the capitalization of ``Dictionary of Keys'' and added the
+abbreviation (DOK), and stated explicitly what is meant by this. We have also
+made a similar note about List of Lists (LIL) earlier in the section.
 \end{solution}

### 2016-08-17 matrices.tex
Reinstate "dictionary of keys" and clarify some things
See also this discussion from GitHub
https://github.com/sympy/sympy-paper/pull/144#discussion_r63944359

@@
 \begin{verbatim}
->>> A = Matrix(2, 2, [x, x + y, y, x])
+>>> A = Matrix([[x, x + y], [y, x]])
 >>> A
@@
 
-Internally these matrices store the elements as lists of lists, making
-it a dense representation.\footnote{Similar to the polynomials module, dense
-  here means that all entries are stored in memory, contrasted with a sparse
+Internally these matrices store the elements as Lists of Lists (LIL), meaning
+the matrix is stored as a list of lists of entries (effectively, the
+input format used to create the matrix \texttt{A} above), making it a
+dense representation.\footnote{Similar to the polynomials module, dense here
+  means that all entries are stored in memory, contrasted with a sparse
   representation where only nonzero entries are stored.} For storing sparse
-matrices, the \verb|SparseMatrix| class can be used.
+matrices, the \verb|SparseMatrix| class can be used. Sparse matrices store
+their elements in Dictionary of Keys (DOK) format, meaning entries are stored
+as \texttt{(row, column)} pairs mapping to the elements.
 

### 2016-08-17 architecture.tex
Move footnotes to the right of punctuation

@@
 In SymPy every symbolic expression is an instance of a Python \texttt{Basic}
-class\footnote{Some
+class,\footnote{Some
   internal classes, such as those used in the polynomial module, do not follow
-  this rule for efficiency reasons.}, a superclass of all SymPy types providing
+  this rule for efficiency reasons.} a superclass of all SymPy types providing
 common methods to all SymPy tree-elements, such as traversals.  The children of a node in the
@@
 \texttt{negative}, \texttt{real}, \texttt{nonpositive}, \texttt{integer},
-\texttt{prime} and \texttt{commutative}\footnote{SymPy assumes that two
+\texttt{prime} and \texttt{commutative}.\footnote{SymPy assumes that two
   expressions $A$ and $B$ commute with each other multiplicatively, that is,
@@
   algorithms in SymPy require special consideration to work correctly with
-  noncommutative products.}. Assumptions on any object can be checked with the
+  noncommutative products.} Assumptions on any object can be checked with the
 \verb|is_|\texttt{\textit{assumption}} attributes, like \verb|t.is_positive|.

### 2016-08-17 gruntz.tex
Move footnotes to the right of punctuation

--- 2016-06-11 gruntz.tex
@@
 limit. In the above case the limit is $2$, as can be
-verified by SymPy:
+verified by SymPy:\footnote{To see intermediate steps, discussed above, interested
+readers can switch on debugging output by setting
+the environment variable \texttt{SYMPY\_DEBUG=True}, before importing anything from
+the SymPy namespace.}
 

### 2016-08-17 introduction.tex
Move footnotes to the right of punctuation

@@
 From here on, all examples in this paper assume that this statement has been
-executed\footnote{Three greater-than signs denote standard
-Python prompt for interactive session.}:
+executed:\footnote{Three greater-than signs denote standard
+Python prompt for interactive session.}
 

### 2016-08-17 rebuttal.tex
Fix some wording

@@
 \begin{solution}
-First your example, unfortunately, doesn't work due to automatic
-simplification (see last input line --- \texttt{exp(100)} cancel).
-Second example is good, but our looks to be more simple.
+  Your first example, unfortunately, doesn't work due to automatic
+  simplification (see the last input line above:  \texttt{exp(100)} cancels
+  automatically, before any numeric evaluation happens). Your second example
+  doesn't have this issue, but our example is simpler.
 

### 2016-08-17 numerics.tex
Should be "solving ODEs"

@@
 and numerical computation of limits, derivatives, integrals, infinite
-series, and solving ODE. All features work in arbitrary precision
+series, and solving ODEs. All features work in arbitrary precision
 and use algorithms that allow computing hundreds of digits rapidly

### 2016-08-18 rebuttal.tex
Should be "solving ODEs"
Address question 22 of the reviewer 3

@@
 \begin{solution}
-
+This sentence was replaced with:
+\begin{quote}
+Two-dimensional (2D) textual representation of the expression can
+be printed with monospace fonts via \verb|pprint|.
+\end{quote}
+We use here same terminology as in the Mathematica
+online documentation, e.g. OutputForm function.
 \end{solution}
@@
 \begin{solution}
-This was replaced with ``solving ODE''.
+This was replaced with ``solving ODEs''.
 \end{solution}

### 2016-08-18 printers.tex
Address question 22 of the reviewer 3

SymPy has a rich collection of expression printers.
By default, an interactive Python session will render the
\verb|str| form of an expression, which has been used in all the examples in
this paper so far. The \verb|str| form of an expression is valid Python and
roughly matches what a user would type to enter the expression.

\begin{verbatim}
>>> phi0 = Symbol('phi0')
>>> str(Integral(sqrt(phi0), phi0))
'Integral(sqrt(phi0), phi0)'
\end{verbatim}

Two-dimensional (2D) textual representation of the expression can
be printed with monospace fonts via \verb|pprint|.
Unicode characters are used for rendering mathematical symbols such as integral signs,
square roots, and parentheses. Greek letters and subscripts in symbol names
that have Unicode code points associated
are also rendered automatically.
% TODO cite unicode?

\noindent
\includegraphics[width=1\textwidth]{pprint.pdf}
Alternately, the \verb|use_unicode=False| flag can be set, which causes the
expression to be printed using only ASCII characters.
% TODO cite ASCII

\begin{verbatim}
>>> pprint(Integral(sqrt(phi0 + 1), phi0), use_unicode=False)
  /
 |
 |   __________
 | \/ phi0 + 1  d(phi0)
 |
/
\end{verbatim}

The function \verb|latex| returns a \LaTeX{} representation of an expression.
% TODO cite latex

\begin{verbatim}
>>> print(latex(Integral(sqrt(phi0 + 1), phi0)))
\int \sqrt{\phi_{0} + 1}\, d\phi_{0}
\end{verbatim}

Users are encouraged to run the \verb|init_printing| function at the beginning
of interactive sessions, which automatically enables the best pretty printing
supported by their environment. In the Jupyter Notebook or Qt
Console~\cite{perez2007ipython}, the \LaTeX{} printer is used to render
expressions using MathJax or \LaTeX{}, if it is installed on the system. The
2D text representation is used otherwise.

Other printers such as MathML are also available. SymPy uses an extensible
printer subsystem for customizing any given
printer, and allows custom objects to define their printing behavior for any
printer. The code generation functionality of SymPy
relies on this subsystem to convert expressions into code in various target
programming languages.

### 2016-08-18 rebuttal.tex
Address question 21 of the reviewer 3

@@
 \begin{solution}
-
+This footnote was added:
+\begin{quote}
+In SymPy, \texttt{str(expr) == repr(expr)}.  In other words,
+string representation of an expression designed to be both compact,
+human-readable and a valid Python code, that could be used to
+recreate one.  As it was noted before, use \texttt{srepr} function to
+see the exact form of an expression.
+\end{quote}
 \end{solution}

### 2016-08-18 printers.tex
Address question 21 of the reviewer 3

@@
 \verb|str| form of an expression, which has been used in all the examples in
-this paper so far. The \verb|str| form of an expression is valid Python and
+this paper so far. The \verb|str| form of an expression is valid
+Python\footnote{In SymPy, \texttt{str(expr) == repr(expr)}.  In other words,
+string representation of an expression designed to be both compact,
+human-readable and a valid Python code, that could be used to
+recreate one.  As it was noted before, use \texttt{srepr} function to
+see the exact form of an expression.} and
 roughly matches what a user would type to enter the expression.

### 2016-08-19 rebuttal.tex
Address question 35 of the reviewer 3

--- 2016-08-17 rebuttal.tex
@@
 \begin{questions}
-\question Nice paper - pity it seems totally out of scope. Why not J. Symbolic
+\question Nice paper --- pity it seems totally out of scope. Why not J. Symbolic
 Computation or some such?
@@
 \begin{solution}
-Citation was moved and footnote for Python interactive prompt was
-added where it was used first time.
-We believe that the paper assumes minimal familiarity with Python and
-that should be covered either with citation [25] or with the Python tutorial.
+  The citation was moved and a footnote for the Python interactive prompt was
+  added where it was used first time. We have also added a footnote indicating
+  a moderate familiarity with Python is assumed.
 \end{solution}
@@
 \begin{solution}
-
+We have updated the name and citation, and changed ``pure mathematics'' to
+``pure and applied mathematics''.
 \end{solution}
@@
 \begin{solution}
-
+This footnote was added:
+\begin{quote}
+  Many Python libraries distinguish the \texttt{str} form of an object, which
+  is meant to be human-readable, and the \texttt{repr} form, which is mean to
+  be valid Python that recreates the object. In SymPy, \texttt{str(expr) ==
+    repr(expr)}. In other words, the string representation of an expression is
+  designed to be compact, human-readable, and valid Python code that could be
+  used to recreate the expression. As it was noted in section~\ref{sec:core},
+  the \texttt{srepr} function prints the exact, verbose form of an expression.
+\end{quote}
 \end{solution}
@@
 \begin{solution}
-
+This sentence was replaced with:
+\begin{quote}
+A two-dimensional (2D) textual representation of the expression can
+be printed with monospace fonts via \verb|pprint|.
+\end{quote}
+We use here the same terminology as in the Mathematica
+online documentation, e.g., the
+\href{https://reference.wolfram.com/language/ref/OutputForm.html}{\texttt{OutputForm}}
+function:
+\begin{quote}
+\texttt{OutputForm[expr]}: prints as a two-dimensional representation of expr using only keyboard characters.
+\end{quote}
 \end{solution}
@@
 \begin{solution}
-
+  % TODO: Should we rename it to "Physics submodules"?
+(see also the discussion at~\ref{rev1point37}).
 \end{solution}
@@
 \begin{solution}
-Word ``package'' replaced with ``submodule''.
+The word ``package'' replaced with ``submodule'' (see also point~\ref{rev1point37}).
 \end{solution}
@@
 \begin{solution}
-
+This was reworded to:
+\begin{quote}
+to solve problems in quantum mechanics
+\end{quote}
 \end{solution}
@@
 \begin{solution}
-This sentence was removed.
+This paper focuses on the architecture of SymPy, and we consider this to be a
+core feature of SymPy's architecture, hence, it's inclusion in the conclusion.
+We have also discussed the different points in this sentence in-depth in the architecture section.
 \end{solution}
 
-\question All ``submodules'' should be replaced by ``modules'' (examples: lines 657, 662).
+\question All ``submodules'' should be replaced by ``modules'' (examples:
+lines 657, 662).
+\label{rev1point37}
 \begin{solution}
-This was done.
+We are unclear how the reviewer is making distinctions between ``modules'',
+``submodules'', and ``packages''. Our use coincides with the definitions used
+in the official Python documentation (see
+\url{https://docs.python.org/3/tutorial/modules.html#packages} and
+\url{https://docs.python.org/3/glossary.html#term-package}). The distinction
+between (sub)module and (sub)package in terms of Python libraries seems very
+technical. To clarify things, we have modified the manuscript to use
+``package'' only when referring to a library external to SymPy (such as
+mpmath and Xy-pic).
+
+To avoid confusion, and to be consistent with the definitions used by the
+official Python documentation, we have modified the manuscript to use
+``submodule'' to refer to any module that is a dotted submodule of
+\texttt{sympy}, such as \texttt{sympy.physics.quantum}. This is to emphasize
+that these modules all live within the \texttt{sympy} namespace, as ``module''
+can suggest a separate library.
 \end{solution}
 
-\question The sentences in lines 657--661 should swap places, because ``areas of mathematics'' are discrete mathematics, concrete mathematics, etc., while simplifying expressions, performing common calculus operations, pretty print- ing expressions, etc.\ belongs to common operations (``other areas'' is also fine, albeit slightly wrong).
+\question The sentences in lines 657--661 should swap places, because ``areas of mathematics'' are discrete mathematics, concrete mathematics, etc., while simplifying expressions, performing common calculus operations, pretty printing expressions, etc.\ belongs to common operations (``other areas'' is also fine, albeit slightly wrong).
 \begin{solution}
+The ``areas of mathematics'' was intended to encompass all mathematical
+operations. We have rewritten the first sentence as
+\begin{quote}
+SymPy supports a wide array of mathematical facilities.
+\end{quote}
+We have also changed ``other included areas'' to ``other supported facilities'' in the third sentence.
 
+The order of the subsequent sentences was chosen to match the order they were
+presented in the paper, so we have left them as-is.
 \end{solution}

### 2016-08-19 domain_specific.tex
Address question 35 of the reviewer 3

@@
-SymPy includes several packages that allow users to solve domain specific
-problems. For example, a comprehensive physics package is included that is
+SymPy includes several submodules that allow users to solve domain specific
+problems. For example, a comprehensive physics submodule is included that is
 useful for solving problems in mechanics, optics, and quantum
@@
 mechanics. This is in turn separated into two distinct components:
-vector algebra symbolics and mechanics.
+vector algebra and mechanics.
 
@@
 
-The \verb|sympy.physics.mechanics| package utilizes the \texttt{sympy.\allowbreak{}physics.\allowbreak{}vector} package
+The \verb|sympy.physics.mechanics| submodule utilizes the \texttt{sympy.\allowbreak{}physics.\allowbreak{}vector} submodule
 to populate time-aware particle and rigid-body objects to fully describe the
@@
 of motion. These equations of motion abide by Newton's laws of motion and can
-handle arbitrary kinematic constraints or complex loads. The package
+handle arbitrary kinematic constraints or complex loads. The submodule
 offers two automated methods for formulating the equations of motion based on
@@
 
-The \verb|sympy.physics.quantum| package has extensive capabilities for
-performing symbolic quantum mechanics, using Python objects to represent the
+The \verb|sympy.physics.quantum| submodule has extensive capabilities to
+solve problems in quantum mechanics, using Python objects to represent the
 different mathematical objects relevant in quantum theory~\cite{Sakurai2010}:

### 2016-08-19 rebuttal.tex
More consistent use of "module", "submodule", and "package"

--- 2016-08-18 rebuttal.tex
@@
 \begin{solution}
-This sentence was replaced with:
-\begin{quote}
-Two-dimensional (2D) textual representation of the expression can
-be printed with monospace fonts via \verb|pprint|.
-\end{quote}
-We use here same terminology as in the Mathematica
-online documentation, e.g. OutputForm function.
+
 \end{solution}
@@
 \begin{solution}
-
+We have left this as is (see the discussion at~\ref{rev1point37}).
 \end{solution}
@@
 
-\question All ``submodules'' should be replaced by ``modules'' (examples: lines 657, 662).
+\question All ``submodules'' should be replaced by ``modules'' (examples:
+lines 657, 662).
+\label{rev1point37}
 \begin{solution}
-This was done.
+We are unclear how the reviewer is making distinctions between ``modules'',
+``submodules'', and ``packages''. Our use coincides with the definitions used
+in the official Python documentation (see
+\href{https://docs.python.org/3/tutorial/modules.html#packages} and
+\href{https://docs.python.org/3/glossary.html#term-package}). The distinction
+between (sub)module and (sub)package in terms of Python libraries seems very
+technical. To clarify things, we have modified the manuscript to use
+``package'' only when referring to a library external to SymPy (such as
+mpmath and Xy-pic).
+
+To avoid confusion, and to be consistent with the definitions used by the
+official Python documentation, we have modified the manuscript to use
+``submodule'' to refer to any module that is a dotted submodule of
+\texttt{sympy}, such as \texttt{sympy.physics.quantum}. This is to emphasize
+that these modules all live within the \texttt{sympy} namespace, as ``module''
+can suggest a separate library.
 \end{solution}

### 2016-08-19 architecture.tex
More consistent use of "module", "submodule", and "package"

@@
 class,\footnote{Some
-  internal classes, such as those used in the polynomial module, do not follow
+  internal classes, such as those used in the polynomial submodule, do not follow
   this rule for efficiency reasons.} a superclass of all SymPy types providing
@@
 returns a string representation of an expression as valid Python code\footnote{
-The \texttt{dotprint} function from the \texttt{sympy.printing.dot} package
+The \texttt{dotprint} function from the \texttt{sympy.printing.dot} submodule
 prints output to dot format, which can be rendered with Graphviz to

### 2016-08-19 categories.tex
More consistent use of "module", "submodule", and "package"

SymPy includes a submodule for dealing with categories---abstract mathematical
objects representing classes of structures as classes of objects (points) and
morphisms (arrows) between the objects. It was
designed with the following two goals in mind:

\begin{enumerate}
\item automatic typesetting of diagrams given by a collection of
  objects and of morphisms between them, and
\item specification and semi-automatic derivation of properties
  using commutative diagrams.
\end{enumerate}

As of version 1.0, SymPy only implements the first goal, while a partially
working draft of implementation of the second goal is available at
\url{https://github.com/scolobb/sympy/tree/ct4-commutativity}.

In order to achieve the two goals, the submodule \texttt{sympy.categories} defines
several classes representing some of the essential concepts: objects, morphisms,
categories, and diagrams.  In category theory, the inner structure of objects is
often discarded in the favor of studying the properties of morphisms, so the
class \texttt{Object} is essentially a synonym of the class \texttt{Symbol}.
There are several morphism classes which do not have a particular internal
structure either, though an exception is \texttt{CompositeMorphism}, which
essentially stores a list of morphisms.

The class \texttt{Diagram} captures the properties of morphisms. This class
stores a family of morphisms, the corresponding source and target objects,
and, possibly, some properties of the morphisms. Generally, no restrictions
are imposed on what the properties may be---for example, one might use strings
of the form ``forall'', ``exists'', ``unique'', etc. Furthermore, the
morphisms of a diagram are grouped into \textit{premises} and
\textit{conclusions} in order to be able to represent logical implications of
the form ``for a collection of morphisms $P$ with properties $p:P\to \Omega$
(the premises), there exists a collection of morphisms $C$ with properties
$c:C\to \Omega$ (the conclusions)'', where $\Omega$ is the universal
collection of properties. Finally, the class \texttt{Category} includes a
collection of diagrams which are deemed commutative and which therefore define
the properties of this category.

Automatic typesetting of diagrams takes a \texttt{Diagram} and produces \LaTeX{}
code using the \texttt{Xy-pic} package.  Typesetting is done in two stages:
layout and generation of \texttt{Xy-pic} code.  The layout stage is taken care
of by the class \texttt{DiagramGrid}, which takes a \texttt{Diagram} and lays out
the objects in a grid, trying to reduce the average length of the arrows in the
final picture.  By default, \texttt{DiagramGrid} uses a series of triangle-based
heuristics to produce a rectangular grid.  A linear layout can also be imposed.
Furthermore, groups of objects can be given; in this case, the groups will be
treated as atomic cells, and the member objects will be typeset independently of
the other objects.

The second phase of diagram typesetting consists in actually drawing the picture
and is carried out by the class \texttt{XypicDiagramDrawer}.  An example of a
diagram automatically typeset by \texttt{DiagramgGrid} and
\texttt{XypicDiagramDrawer} in given in Figure~\ref{fig:cat:loops}.
\begin{figure}[h]
  \centerline{
    \xymatrix{
      A \ar[r]_{f} \ar@/^3mm/[rr]^{h_{2}} \ar@(u,l)[]^{l_{A}} \ar@/^3mm/@(l,d)[]^{n_{A}} & B \ar[d]^{g} & D \ar[l]^{k} \ar@/_7mm/[ll]_{h} \ar@/_11mm/[ll]_{h_{1}} \ar@(r,u)[]^{l_{D}} \ar@/^3mm/@(d,r)[]^{n_{D}} \\
      & C \ar@(l,d)[]^{l_{C}} \ar@/^3mm/@(d,r)[]^{n_{C}} &
    }
  }
  \caption{An automatically typeset commutative diagram}\label{fig:cat:loops}
\end{figure}

As far as the second main goal of \texttt{sympy.categories} is concerned, the principal idea
consists in automatically deciding whether a diagram is commutative or not,
given a collection of ``axioms'': diagrams known to be commutative. The
implementation is based on graph embeddings (injective maps): whenever an
embedding of a commutative diagram into a given diagram is found, one
concludes that the subdiagram is commutative. Deciding commutativity of the
whole diagram is therefore based (theoretically) on finding a ``cover'' of the
target diagram by embeddings of the axioms. The na\"{\i}ve implementation
proved to be prohibitively slow; a better optimized version is therefore in
order, as well as application of heuristics.

### 2016-08-19 conclusion_and_future_work.tex
Removed sentence, see question 36
More consistent use of "module", "submodule", and "package"

@@
 
-SymPy expressions are immutable trees of Python objects. SymPy uses Python both
+SymPy uses Python both
 as the internal language and the user language. This permits users to access to
@@
 
-SymPy has modules for many areas of mathematics. This includes functions for
+SymPy has submodules for many areas of mathematics. This includes functions for
 simplifying expressions, performing common calculus operations, pretty printing
@@
 generation, tensors, Lie algebras, cryptography, and special functions.
-Additionally, SymPy contains modules targeting certain specific domains,
+Additionally, SymPy contains submodules targeting certain specific domains,
 such as classical mechanics and quantum mechanics.  This breadth of domains has
@@
 alternative core for SymPy), improving the assumptions system, and improving the
-solvers module.
+solvers submodule.
 

### 2016-08-19 diophantine.tex
More consistent use of "module", "submodule", and "package"

@@
 Currently, the following five types of Diophantine equations can be solved using
-SymPy's Diophantine module ($a_1, \ldots, a_{n+1}$, $a$, $b$, $c$, $d$, $e$,
+SymPy's Diophantine submodule ($a_1, \ldots, a_{n+1}$, $a$, $b$, $c$, $d$, $e$,
 $f$, and $k$ are explicitly given rational constants):
@@
 The following examples illustrate some of the
-basic functionalities of the Diophantine module.
+basic functionalities of the Diophantine submodule.
 

### 2016-08-19 domain_specific.tex
More consistent use of "module", "submodule", and "package"

@@
 
-The \verb|sympy.physics.quantum| submodule has extensive capabilities to
-solve problems in quantum mechanics, using Python objects to represent the
+The \verb|sympy.physics.quantum| submodule has extensive capabilities for
+performing symbolic quantum mechanics, using Python objects to represent the
 different mathematical objects relevant in quantum theory~\cite{Sakurai2010}:

### 2016-08-19 logic.tex
More consistent use of "module", "submodule", and "package"

SymPy supports construction and manipulation of boolean expressions
through the \texttt{sympy.logic} submodule. SymPy symbols can be used as
propositional variables and subsequently be replaced with \texttt{True}
or \texttt{False} values. Many functions for manipulating boolean
expressions have been implemented in the \texttt{sympy.logic} submodule.

\subsection{Constructing boolean expressions}

A boolean variable can be declared as a SymPy \verb|Symbol|. Python operators
\texttt{\&}, \texttt{\textbar{}} and \texttt{\textasciitilde{}} are overridden
when using SymPy objects to use the SymPy functionality for logical
\texttt{And}, \texttt{Or}, and \texttt{Not}. Other logic functions are also
integrated into SymPy, including \texttt{Xor} and \texttt{Implies}, which are
constructed with \texttt{\^{}} and \texttt{\textgreater{}\textgreater{}},
respectively. Expressions can therefore be constructed either by using
the shortcut operator notation or by directly creating the relevant
objects: \verb|And()|,
\verb|Or()|, \verb|Not()|, \verb|Xor()|, \verb|Implies()|, \verb|Nand()|,
\verb|Nor()|, etc.:

\begin{verbatim}
>>> e = (x & y) | z
>>> e.subs({x: True, y: True, z: False})
True
\end{verbatim}

\subsection{CNF and DNF}

Any boolean expression can be converted to conjunctive normal form, disjunctive
normal form, or negation normal form. The API also exposes methods to check if
a boolean expression is in any of the aforementioned forms.

\begin{verbatim}
>>> from sympy.logic.boolalg import is_dnf, is_cnf
>>> to_cnf((x & y) | z)
And(Or(x, z), Or(y, z))
>>> to_dnf(x & (y | z))
Or(And(x, y), And(x, z))
>>> is_cnf((x | y) & z)
True
>>> is_dnf((x & y) | z)
True
\end{verbatim}

\subsection{Simplification and Equivalence}

The \texttt{sympy.logic} submodule supports simplification of given boolean expression by making
deductions from the expression. Equivalence of two logical expressions can also
be checked. In the case of equivalence, the function \texttt{bool\_map}
can be used to show which variables of the first expression correspond
to which variables of the second one.

\begin{verbatim}
>>> a, b, c = symbols('a b c')
>>> e = a & (~a | ~b) & (a | c)
>>> simplify(e)
And(Not(b), a)
>>> e1 = a & (b | c)
>>> e2 = (x & y) | (x & z)
>>> bool_map(e1, e2)
(And(Or(b, c), a), {a: x, b: y, c: z})
\end{verbatim}

\subsection{SAT solving}

The submodule also supports satisfiability (SAT) checking of a given boolean
expression. If an expression is satisfiable, it is possible to return
a variable assignment which satisfies it. The API also supports
listing all possible assignments.
The SAT solver has a clause learning DPLL algorithm implemented with a watch
literal scheme and VSIDS heuristic~\cite{moskewicz2008method}.

\begin{verbatim}
>>> satisfiable(a & (~a | b) & (~b | c) & ~c)
False
>>> satisfiable(a & (~a | b) & (~b | c) & c)
{a: True, b: True, c: True}
\end{verbatim}

### 2016-08-19 solvers.tex
More consistent use of "module", "submodule", and "package"

SymPy has equation solvers that can handle ordinary differential
equations, recurrence relationships, Diophantine equations, and algebraic
equations. There is also
rudimentary support for simple partial differential equations.

There are two functions for solving algebraic equations in SymPy:
\texttt{solve} and \texttt{solveset}.
\texttt{solveset} has several design changes with respect to the older
\texttt{solve} function. This distinction is present in order to resolve the
usability issues with the
previous \texttt{solve} function API while maintaining backward compatibility
with earlier versions of SymPy.
\texttt{solveset} only requires essential input information from the user.
The function signatures of \texttt{solve} and \texttt{solveset} are
\begin{verbatim}
solve(f, *symbols, **flags)
solveset(f, symbol, domain=S.Complexes)
\end{verbatim}
The \texttt{domain} parameter is typically either \texttt{S.Complexes} (the
default) or \texttt{S.Reals}; the latter causes \texttt{solveset} to only return real solutions.

An important difference between the two functions is that the output API of \texttt{solve}
varies with input (sometimes returning a Python list and sometimes a Python dictionary) whereas
\texttt{solveset} always returns
a SymPy set object.

Both functions implicitly assume that expressions are equal to 0. For
instance, \texttt{solveset(x - 1, x)} solves $x - 1 = 0$ for $x$.

\texttt{solveset} is under active development as a planned replacement for
\texttt{solve}. There are certain features which are implemented in
\texttt{solve} that are not yet implemented in \texttt{solveset}, including
multivariate systems, and some transcendental equations.

More examples of \texttt{solveset} and \texttt{solve} can be found in the
supplement.

### 2016-08-19 stats.tex
More consistent use of "module", "submodule", and "package"

%% Stats
The \verb|sympy.stats| submodule provides random variable types and methods for
computing of statistical properties of expressions involving random
variables, which can be either continuous or discrete, the latter ones being
further divided into finite and infinite. The variables are associated
with probability densities on corresponding domains and internally defined
in terms of probability spaces.
Apart from the possibility of defining the random variables from user supplied
density distribution, SymPy provides definitions of most common
distributions, including \texttt{Uniform}, \texttt{Poisson}, \texttt{Normal},
\texttt{Binomial}, \texttt{Bernoulli}, and many others.

Properties of random expressions can be calculated using, e.g.,
\texttt{expectation} (abbreviated \texttt{E}) and \texttt{variance} to
calculate expectation and variance. Internally, these functions generate
integrals and summations, which are automatically evaluated. The evaluation
can be suppressed using \texttt{evaluate=False} keyword argument.

Conditions on random variables can be defined with inequalities, equalities,
and logical operators and their overall probabilities are obtained using
\texttt{P}. The features can be illustrated on a model of two dice throws:
\begin{verbatim}
>>> from sympy.stats import Die, P, E
>>> X, Y = Die("X"), Die("Y")
>>> P(Eq(X, 6) & Eq(Y, 6))
1/36
>>> P(X>Y)
5/12
\end{verbatim}
The conditions can also be supplied as a second parameter to \texttt{E},
\texttt{P}, and other methods to calculate the property given the condition:
\begin{verbatim}
>>> E(X, X+Y<5)
5/3
\end{verbatim}

Using the facilities of the \texttt{sympy.stats} submodule, one can, for
example, calculate
%Using the above-mentioned features, it is easy to calculate, for example
the well known properties of maxwellian velocity distribution
\begin{verbatim}
>>> from sympy.stats import Maxwell, density
>>> kT, m, x = symbols("kT m x", positive=True)
>>> v = Maxwell("v", sqrt(kT/m))
>>> E(v)                                # mean velocity
2*sqrt(2)*sqrt(kT)/(sqrt(pi)*sqrt(m))
>>> E(v, evaluate=False)                # unevaluated mean velocity
Integral(sqrt(2)*m**(3/2)*v**3*exp(-m*v**2/(2*kT))/(sqrt(pi)*kT**(3/2)),
(v, 0, oo))
>>> E(m*v**2/2)                         # mean energy
3*kT/2
>>> solve(density(v)(x).diff(x), x)[0]  # most probable velocity
sqrt(2)*sqrt(kT)/sqrt(m)
\end{verbatim}

More information on the \texttt{sympy.stats} submodule can be found in
\cite{StatsMRocklin}.

### 2016-08-19 matrices.tex
Missed some files for the submodules fix

@@
 input format used to create the matrix \texttt{A} above), making it a
-dense representation.\footnote{Similar to the polynomials module, dense here
+dense representation.\footnote{Similar to the polynomials submodule, dense here
   means that all entries are stored in memory, contrasted with a sparse
@@
 
-When symbolic matrices are combined with the assumptions module for logical
+When symbolic matrices are combined with the assumptions submodule for logical
 inference, they provide powerful reasoning over invertibility,

### 2016-08-19 rebuttal.tex
Fix response for reviewer 3 point 30
Use \url instead of \href

@@
 \begin{solution}
-We have left this as is (see the discussion at~\ref{rev1point37}).
+  % TODO: Should we rename it to "Physics submodules"?
+(see also the discussion at~\ref{rev1point37}).
 \end{solution}
@@
 in the official Python documentation (see
-\href{https://docs.python.org/3/tutorial/modules.html#packages} and
-\href{https://docs.python.org/3/glossary.html#term-package}). The distinction
+\url{https://docs.python.org/3/tutorial/modules.html#packages} and
+\url{https://docs.python.org/3/glossary.html#term-package}). The distinction
 between (sub)module and (sub)package in terms of Python libraries seems very

### 2016-08-19 rebuttal.tex
Fix em-dash
Fix hyphenation

@@
 
-\question The sentences in lines 657--661 should swap places, because ``areas of mathematics'' are discrete mathematics, concrete mathematics, etc., while simplifying expressions, performing common calculus operations, pretty print- ing expressions, etc.\ belongs to common operations (``other areas'' is also fine, albeit slightly wrong).
+\question The sentences in lines 657--661 should swap places, because ``areas of mathematics'' are discrete mathematics, concrete mathematics, etc., while simplifying expressions, performing common calculus operations, pretty printing expressions, etc.\ belongs to common operations (``other areas'' is also fine, albeit slightly wrong).
 \begin{solution}

### 2016-08-19 rebuttal.tex
Add a link to the Mathematica docs for OutputForm
Quote the OutputForm docs

@@
 \end{quote}
-We use here same terminology as in the Mathematica
-online documentation, e.g. OutputForm function.
+We use here the same terminology as in the Mathematica
+online documentation, e.g., the
+\href{https://reference.wolfram.com/language/ref/OutputForm.html}{\texttt{OutputForm}}
+function:
+\begin{quote}
+\texttt{OutputForm[expr]}: prints as a two-dimensional representation of expr using only keyboard characters.
+\end{quote}
 \end{solution}

### 2016-08-19 rebuttal.tex
Add an article
Clarify and cleanup str/repr footnote. Add a \ref to the srepr section.

@@
 \begin{quote}
-In SymPy, \texttt{str(expr) == repr(expr)}.  In other words,
-string representation of an expression designed to be both compact,
-human-readable and a valid Python code, that could be used to
-recreate one.  As it was noted before, use \texttt{srepr} function to
-see the exact form of an expression.
+  Many Python libraries distinguish the \texttt{str} form of an object, which
+  is meant to be human-readable, and the \texttt{repr} form, which is mean to
+  be valid Python that recreates the object. In SymPy, \texttt{str(expr) ==
+    repr(expr)}. In other words, the string representation of an expression is
+  designed to be compact, human-readable, and valid Python code that could be
+  used to recreate the expression. As it was noted in section~\ref{sec:core},
+  the \texttt{srepr} function prints the exact, verbose form of an expression.
 \end{quote}
@@
 \begin{quote}
-Two-dimensional (2D) textual representation of the expression can
+A two-dimensional (2D) textual representation of the expression can
 be printed with monospace fonts via \verb|pprint|.

### 2016-08-19 printers.tex
Add an article
Clarify and cleanup str/repr footnote. Add a \ref to the srepr section.

--- 2016-08-18 printers.tex
@@
-SymPy has a rich collection of expression printers.
-By default, an interactive Python session will render the
-\verb|str| form of an expression, which has been used in all the examples in
-this paper so far. The \verb|str| form of an expression is valid Python and
-roughly matches what a user would type to enter the expression.
+SymPy has a rich collection of expression printers. By default, an interactive
+Python session will render the \verb|str| form of an expression, which has
+been used in all the examples in this paper so far. The \verb|str| form of an
+expression is valid Python and roughly matches what a user would type to enter
+the expression.\footnote{Many Python libraries distinguish the \texttt{str}
+  form of an object, which is meant to be human-readable, and the
+  \texttt{repr} form, which is mean to be valid Python that recreates the
+  object. In SymPy, \texttt{str(expr) == repr(expr)}. In other words, the
+  string representation of an expression is designed to be compact,
+  human-readable, and valid Python code that could be used to recreate the
+  expression. As it was noted in section~\ref{sec:core}, the \texttt{srepr}
+  function prints the exact, verbose form of an expression.}
 
@@
 
-Two-dimensional (2D) textual representation of the expression can
+A two-dimensional (2D) textual representation of the expression can
 be printed with monospace fonts via \verb|pprint|.

### 2016-08-19 introduction.tex
The mpmath version could affect everything, not just the numerics section
Sage -> SageMath (reviewer 3, point 2)

@@
 and tools to support research within a variety of domains, such as
-Sage~\cite{SAGE} (pure mathematics),
+SageMath~\cite{sagemath} (pure and applied mathematics),
 yt~\cite{2011ApJS..192....9T} (astronomy and astrophysics),
@@
 in section~\ref{sec:conclusion}.  All examples in this paper use SymPy
-version 1.0, examples in section~\ref{sec:mpmath} use mpmath
-version 0.19.
+version 1.0 and mpmath version 0.19.
 

### 2016-08-19 introduction.tex
Clarify reviewer 3 point 1 a bit more

@@
 Python is a dynamically typed programming language that has a focus on
-ease of use and readability. Due in part to this focus, it has become a popular
+ease of use and readability.\footnote{This paper assumes a moderate
+  familiarity with the Python programming language.} Due in part to this focus, it has become a popular
 language for scientific
@@
 From here on, all examples in this paper assume that this statement has been
-executed:\footnote{Three greater-than signs denote standard
-Python prompt for interactive session.}
+executed:\footnote{The three greater-than signs denote the user input for the
+  Python interactive session, with the result, if there is one, shown on the
+  next line.}
 

### 2016-08-19 rebuttal.tex
Revert "Removed sentence, see question 36"
This reverts commit 214a695f591f0da43888338bd01ab63af14ecfe3.
Indicate why the sentence is in the conclusion
Get rid of "areas of mathematics" in conclusion (reviewer 3, point 38)

@@
 \begin{solution}
-This sentence was removed.
+This paper focuses on the architecture of SymPy, and we consider this to be a
+core feature of SymPy's architecture, hence, it's inclusion in the conclusion.
+We have also discussed the different points in this sentence in-depth in the architecture section.
 \end{solution}
@@
 \begin{solution}
+The ``areas of mathematics'' was intended to encompass all mathematical
+operations. We have rewritten the first sentence as
+\begin{quote}
+SymPy supports a wide array of mathematical facilities.
+\end{quote}
+We also changed ``other included areas'' to ``other supported facilities'' in the third sentence.
 
+The order of the subsequence sentences was chosen to match the order they were
+presented in the paper, so we have left them as-is.
 \end{solution}

### 2016-08-19 conclusion_and_future_work.tex
Revert "Removed sentence, see question 36"
This reverts commit 214a695f591f0da43888338bd01ab63af14ecfe3.
Get rid of "areas of mathematics" in conclusion (reviewer 3, point 38)

@@
 
-SymPy uses Python both
+SymPy expressions are immutable trees of Python objects. SymPy uses Python both
 as the internal language and the user language. This permits users to access to
@@
 
-SymPy has submodules for many areas of mathematics. This includes functions for
+SymPy supports a wide array of mathematical facilities. This includes functions for
 simplifying expressions, performing common calculus operations, pretty printing
-expressions, solving equations, and representing symbolic matrices. Other included
-areas
-are discrete math, concrete math, plotting, geometry, statistics,
+expressions, solving equations, and representing symbolic matrices. Other supported
+facilities
+include discrete math, concrete math, plotting, geometry, statistics,
 polynomials, sets, series, vectors, combinatorics, group theory, code

### 2016-08-19 rebuttal.tex
Typo
Grammar

@@
 \end{quote}
-We also changed ``other included areas'' to ``other supported facilities'' in the third sentence.
+We have also changed ``other included areas'' to ``other supported facilities'' in the third sentence.
 
-The order of the subsequence sentences was chosen to match the order they were
+The order of the subsequent sentences was chosen to match the order they were
 presented in the paper, so we have left them as-is.

### 2016-08-20 rebuttal.tex
Add mpfr reference, this address editor comment 4
FYI
https://groups.google.com/d/msg/sympy/f4t2ORLEKMM/Yu1Y1VcN5ewJ

@@
 \begin{solution}
-  % TODO
+Major advantage is that mpmath is a pure Python library.  It's slow c.f.
+MPFR, but not too much to evaluate any reasonably complicated single
+expression at interactive speed, which covers most use cases in SymPy.
+
+We mention this at the beginning of the section.
 \end{solution}

### 2016-08-20 numerics.tex
More consistent use of "module", "submodule", and "package"
Add mpfr reference, this address editor comment 4
FYI
https://groups.google.com/d/msg/sympy/f4t2ORLEKMM/Yu1Y1VcN5ewJ

@@
 supplied by the mpmath library~\cite{mpmath}. Originally, it was developed as a SymPy
-module but has subsequently been moved to a standalone pure-Python package.
+submodule but has subsequently been moved to a standalone pure-Python package.
 The basic datatypes in mpmath are \texttt{mpf} and \texttt{mpc}, which
@@
 
-Like SymPy, mpmath is a pure Python library.
-Internally, mpmath represents a floating-point number
-${(-1)}^s x \cdot 2^y$ by a tuple $(s, x, y, b)$ where
+Like SymPy, mpmath is a pure Python library.  This is it's major advantage
+over the GNU MPFR~\cite{Fousse:2007:MMB:1236463.1236468},
+which is much more feature-rich and fast.  Internally, mpmath represents
+a floating-point number ${(-1)}^s x \cdot 2^y$ by a tuple $(s, x, y, b)$ where
 $x$ and $y$ are arbitrary-size Python integers

### 2016-08-20 features.tex
Fix grammar error
Address question 19 of the reviewer 3
Table 1 extended to include submodule names.

@@
 
-\begin{longtable}[htbc]{p{0.20\linewidth}p{0.73\linewidth}}
+\begin{longtable}[htbc]{p{0.3\linewidth}p{0.7\linewidth}}
 \caption{SymPy Features and Descriptions\label{features-table}}\\
 \toprule
-\textbf{Feature} & \textbf{Description} \\
+\textbf{Feature(submodules)} & \textbf{Description} \\
 \midrule
-Calculus & Algorithms for computing derivatives, integrals, and limits.\\
+Calculus (core, series, integrals) & Algorithms for computing derivatives, integrals, and limits.\\
 
-Category Theory & Representation of objects, morphisms, and diagrams. Tools
+Category Theory (categories) & Representation of objects, morphisms, and diagrams. Tools
 for drawing diagrams with Xy-pic.\\
 
-Code Generation & Generation of compilable and executable code in a
+Code Generation (printing, codegen) & Generation of compilable and executable code in a
 variety of different programming languages from expressions directly. Target
@@
 
-Combinatorics \& Group Theory & Permutations, combinations,
+Combinatorics \& Group Theory (combinatorics) & Permutations, combinations,
 partitions, subsets, various permutation groups (such as polyhedral, Rubik,
@@
 
-Concrete Math & Summation, products, tools for determining whether summation
+Concrete Math (concrete) & Summation, products, tools for determining whether summation
 and product expressions are convergent, absolutely convergent, hypergeometric,
@@
 
-Cryptography & Block and stream ciphers, including shift, Affine,
+Cryptography (crypto) & Block and stream ciphers, including shift, Affine,
 substitution, Vigen\`{e}re's, Hill's, bifid, RSA, Kid RSA,
@@
 
-Differential Geometry & Representations of manifolds, metrics, tensor
+Differential Geometry (diffgeom) & Representations of manifolds, metrics, tensor
 products, and coordinate systems in Riemannian and pseudo-Riemannian
@@
 
-Geometry & Representations of 2D geometrical entities, such as lines and
+Geometry (geometry) & Representations of 2D geometrical entities, such as lines and
 circles. Enables queries on these entities, such as asking the area of an
@@
 
-Lie Algebras & Representations of Lie algebras and root systems.\\
+Lie Algebras (liealgebras) & Representations of Lie algebras and root systems.\\
 
-Logic & Boolean expressions, equivalence testing, satisfiability, and normal
+Logic (logic) & Boolean expressions, equivalence testing, satisfiability, and normal
 forms.\\
 
-Matrices & Tools for creating matrices of symbols and expressions.
+Matrices (matrices) & Tools for creating matrices of symbols and expressions.
 Both sparse and dense representations, as well as symbolic linear
@@
 
-Matrix Expressions & Matrices with symbolic dimensions (unspecified entries).
+Matrix Expressions (matrices.expressions) & Matrices with symbolic dimensions (unspecified entries).
 Block matrices.\\
 
-Number Theory & Prime number generation, primality testing, integer
+Number Theory (ntheory) & Prime number generation, primality testing, integer
 factorization, continued fractions, Egyptian fractions, modular arithmetic,
@@
 
-Plotting & Hooks for visualizing expressions via matplotlib~\cite{Hunter:2007}
+Plotting (plotting) & Hooks for visualizing expressions via matplotlib~\cite{Hunter:2007}
 or as text drawings when lacking a graphical back-end. 2D function plotting,
@@
 
-Polynomials & Polynomial algebras over various coefficient domains.
+Polynomials (polys) & Polynomial algebras over various coefficient domains.
 Functionality ranges from simple operations (e.g., polynomial division) to
@@
 
-Printing & Functions for printing SymPy expressions in the terminal with ASCII
+Printing (printing) & Functions for printing SymPy expressions in the terminal with ASCII
 or Unicode characters and converting SymPy expressions to \LaTeX{} and
@@
 
-Quantum Mechanics & Quantum states, bra--ket notation, operators, basis sets,
+Quantum Mechanics (physics.quantum) & Quantum states, bra--ket notation, operators, basis sets,
 representations, tensor products, inner products, outer products, commutators,
@@
 
-Series & Series expansion, sequences, and limits of sequences.
+Series (series) & Series expansion, sequences, and limits of sequences.
 This includes Taylor, Laurent, and Puiseux series as well as special series, such
@@
 
-Sets & Representations of empty, finite, and infinite sets (including
-special sets such as for all natural, integer, and complex numbers). Operations
+Sets (sets) & Representations of empty, finite, and infinite sets (including
+special sets such as the natural, integer, and complex numbers). Operations
 on sets such as union, intersection, Cartesian product, and building sets from
@@
 
-Simplification & Functions for manipulating and simplifying expressions.
+Simplification (simplify) & Functions for manipulating and simplifying expressions.
 Includes algorithms for simplifying hypergeometric functions, trigonometric
@@
 
-Solvers & Functions for symbolically solving equations, systems
+Solvers (solvers) & Functions for symbolically solving equations, systems
 of equations, both linear and non-linear, inequalities, ordinary differential
@@
 
-Special Functions & Implementations of a number of well known special functions,
+Special Functions (functions) & Implementations of a number of well known special functions,
 including Dirac delta, Gamma, Beta, Gauss error functions, Fresnel integrals,
@@
 
-Statistics & Support for a random variable type as well as the ability to
+Statistics (stats) & Support for a random variable type as well as the ability to
 declare this variable from prebuilt distribution functions such as
@@
 
-Tensors & Symbolic manipulation of indexed objects.\\
+Tensors (tensor) & Symbolic manipulation of indexed objects.\\
 
-Vectors & Basic operations on vectors and differential calculus with respect
+Vectors (vector) & Basic operations on vectors and differential calculus with respect
 to 3D Cartesian coordinate systems.\\

### 2016-08-22 rebuttal.tex
Address question 19 of the reviewer 3
Table 1 extended to include submodule names.
Cleanup MPFR bit. Note that it's LGPL.

--- 2016-08-19 rebuttal.tex
@@
 \begin{solution}
-  % TODO
+
+  The major advantage is that mpmath is a pure Python library. It's slower
+  than MPFR, but not too much to evaluate any reasonably complicated single
+  expression at interactive speed, which covers most use cases in SymPy.
+
+We mention this at the beginning of the section. Another advantage is that
+mpmath is BSD licensed, like SymPy, which is important for the scientific
+Python ecosystem (MPFR is licensed under the LGPL).
+
+It is also worth noting that none of the referee materials we received mention MPFR\@.
 \end{solution}
@@
 \begin{solution}
+The IEEE defines the term feature in IEEE 829 as "A distinguishing
+characteristic of a software item (e.g., performance, portability,
+or functionality)."  We believe, that this definition fits our needs.
 
+Table 1 was extended to include submodule names.
 \end{solution}

### 2016-08-22 numerics.tex
Cleanup MPFR bit. Note that it's LGPL.

@@
 
-Like SymPy, mpmath is a pure Python library.  This is it's major advantage
-over the GNU MPFR~\cite{Fousse:2007:MMB:1236463.1236468},
-which is much more feature-rich and fast.  Internally, mpmath represents
+Like SymPy, mpmath is a pure Python library.  A design decision of SymPy is to
+keep it and its required dependencies pure Python. This is a primary advantage
+ of mpmath over other multiple precision libraries such as GNU MPFR~\cite{Fousse:2007:MMB:1236463.1236468},
+which is much more feature-rich and fast.  Like SymPy, mpmath is also BSD
+licensed (GNU MPFR is licensed under the GNU Lesser General Public License~\cite{rosen2005open}).
+
+Internally, mpmath represents
 a floating-point number ${(-1)}^s x \cdot 2^y$ by a tuple $(s, x, y, b)$ where

### 2016-08-22 rebuttal.tex
Cleanup in rebuttal

@@
 \begin{solution}
-The IEEE defines the term feature in IEEE 829 as "A distinguishing
+IEEE 829 defines the term feature as ``A distinguishing
 characteristic of a software item (e.g., performance, portability,
-or functionality)."  We believe, that this definition fits our needs.
+or functionality).''  We believe, that this definition fits our needs.
 

### 2016-08-22 rebuttal.tex
Rework the intro text

@@
 
-Please read below our answer about the
-questions made by you and reviewers. I hope that you agree
-with all our comments. \bigskip
+We thank you and the referees for your most helpful comments and suggestions.
+The manuscript is greatly improved as a result.
+
+Below we give your and reviewers' comments on white background and our
+replies on gray background. \bigskip
 

### 2016-08-22 conclusion_and_future_work.tex
Fix a typo

@@
 direction is \href{https://github.com/symengine/symengine}{SymEngine}, a
-C++ symbolic manipulation library that is planned to be usable as a
+C++ symbolic manipulation library that is planned to be usable as an
 alternative core for SymPy), improving the assumptions system, and improving the

### 2016-08-22 architecture.tex
Clarify and cleanup str/repr footnote. Add a \ref to the srepr section.
Add a dedicated paragraph about Speed
And move all SymEngine discussio in there. The conclusion was simplified.

@@
 \subsection{The Core}
-
+\label{sec:core}
 A computer algebra system stores mathematical expressions as data
@@
 above listing, such as evaluation at rational points and series expansion.
+
+\subsection{Speed}
+
+Due to being written in pure Python, SymPy's speed is generally slower compared
+with its commercial competitors. For many applications and uses of SymPy, that
+is not a problem, as SymPy is able to return the answer quickly enough, but for
+some applications that require handling of very long expressions and/or lots of
+small expressions, the speed becomes a problem.
+
+For this reason, a new library called SymEngine~\cite{SymEngine} was started.
+It is a pure C++ library with thin wrappers to other languages (Python, Ruby,
+Julia, ...) whose aim is to be the fastest manipulation library.  Preliminary
+benchmarks suggest that SymEngine is as fast or faster than the commercial or
+open source competitors.
+
+SymPy recently started to use SymEngine as an optional backend, initially in
+\texttt{sympy.mechanics} only. The plan is to allow more algorithms in SymPy to
+take advantage of the speed of SymEngine.

### 2016-08-22 conclusion_and_future_work.tex
Add a dedicated paragraph about Speed
And move all SymEngine discussio in there. The conclusion was simplified.

--- 2016-08-19 conclusion_and_future_work.tex
@@
 Some of the planned future work for SymPy includes work on improving code
-generation, improvements to the speed of SymPy (one area of work in this
-direction is \href{https://github.com/symengine/symengine}{SymEngine}, a
-C++ symbolic manipulation library that is planned to be usable as a
-alternative core for SymPy), improving the assumptions system, and improving the
-solvers submodule.
+generation, improvements to the speed of SymPy using SymEngine, improving the
+assumptions system, and improving the solvers submodule.
 

### 2016-08-22 authors/list_latex.py
Add institutions' details into footnotes

@@ -12,6 +12,12 @@ with open("../authors.tex", "w", encoding='utf-8') as f:
     for n, author in enumerate(author_list):
         f.write((u"\\author[%d]{%s}%%\n" % (n+1, author["name"])))
     for n, author in enumerate(author_list):
-        f.write(u"\\affil[%d]{%s, %s (\\email{%s}).}%%\n" \
-                % (n+1, author["institution"], author["institution_address_siam"],
-                    author["email"]))
+        if "institution_explanation" in author:
+            institution_explanation = u" %s" % author["institution_explanation"]
+        else:
+            institution_explanation = u""
+        f.write(u"\\affil[%d]{%s, %s (\\email{%s}).%s}%%\n" \
+                % (n+1, author["institution"],
+                    author["institution_address_siam"],
+                    author["email"],
+                    institution_explanation))

### 2016-08-22 acknowledgements.tex
Reorganize the acknowledgements.tex
We only thank the institutions, but do not explain what they are, per
reviewer's comment.

@@
 % People should add their funding agency acknowledgements here.
+All authors thank the Google Summer of Code for its financial support of
+students who contributed to SymPy.
+
+The author of this paper Ondřej Čertík thanks the Los Alamos National
+Laboratory for its financial support.
+
+The author of this paper Richard P. Muller thanks Sandia National Laboratory
+for its financial support.
+
 The author of this paper Francesco Bonazzi thanks the Deutsche

### 2016-08-22 rebuttal.tex
Update the rebuttal.tex
Revert "Cleanup acknowledgements section"
This reverts commit 5ef75ae24816e907f2fe47927b3d7684dbf75b6e.
Add \usepackage[T1]{fontenc} to rebuttal.tex
This makes the underscores in texttt render as actual underscores.
Explain what was done with the acknowledgements

@@
 \begin{solution}
-  % TODO
+A new paragraph titled "Speed" was added to the manuscript, which discusses the
+speed issue and lays out a path forward. The conclusion was simplified.
 \end{solution}
@@
 \begin{solution}
-These details were removed.
+The institutions's details were moved to the footnotes in the head of the
+paper, and the acknowledgements only thank the institutions.
 \end{solution}
@@
 \begin{solution}
-Now we use \texttt{symbols('a b c d')}, like in other places.
+
 \end{solution}

### 2016-08-22 comparison_with_mma.tex
CASs -> CASes
Reviewer 3, comment 3.

Wolfram Mathematica is a popular proprietary CAS that
features highly advanced algorithms, has a core written in
C++~\cite{Wolfram2003book}, and interprets its own programming language,
Wolfram Language.

% M-expressions

Analogous to Lisp S-expressions,
Mathematica uses its own style of M-expres\-sions,
which are arrays of either atoms or other M-expressions.
The first element of the expression identifies the type of the expression
and is indexed by zero, and the first argument is indexed starting with one.
In SymPy, expression arguments are stored in a Python tuple
(that is, an immutable array),
while the expression type is identified by the type of the object storing the
expression.

% Attributes

Mathematica can associate attributes to its atoms.
Attributes may define mathematical properties and behavior of the nodes
associated to the atom.
In SymPy, the usage of static class fields is roughly similar to Mathematica's
attributes, though other programming patterns may also be used to achieve an
equivalent behavior such as class inheritance.

% Expression mutability

Unlike SymPy, Mathematica's expressions are mutable:
one can change parts of the expression tree without the need of
creating a new object.
The mutability of Mathematica expressions allows for a lazy updating of any references
to a given data structure.

% * comparison with Mathematica: commutativity, associative expressions, one-identity. Advantage of SymPy: multiplicative commutativity defined on symbols.
% Products and commutativity

Products in Mathematica are determined by some built in node types, such as
\texttt{Times}, \texttt{Dot}, and others.  \texttt{Times} is a representation of
the \texttt{*} operator, and is always meant to represent a commutative product
operator.  The other notable product is \texttt{Dot}, which represents the
\texttt{.} operator.  This product represents matrix multiplication. It is not
commutative.  Unlike
Mathematica, SymPy determines commutativity with respect to multiplication from
the expression type of the factors.  Mathematica puts the \texttt{Orderless} attribute
on the expression type.

% Associative expressions.

Regarding associative expressions,
SymPy handles associativity of sums and products by automatically flattening them,
Mathematica specifies the \texttt{Flat} attribute on the expression type.

% One identity

% Pattern matching

Mathematica relies heavily on pattern matching---even the so-called equivalent of function declaration is in reality
the definition of a pattern generating an expression tree transformation
on input expressions.
%
Mathematica's pattern matching is sensitive to
associative, commutative, and
one-identity properties of its expression tree
nodes.
%
SymPy has various ways to perform pattern matching.
All of them play a lesser role in the CAS than in Mathematica
and are basically available as a tool to rewrite expressions.
The differential equation solver in SymPy somewhat relies on pattern matching to
identify differential equation types, but it is envisaged to replace
that strategy with analysis of Lie symmetries in the future.
Mathematica's real advantage is the ability to add (at runtime) new overloading
to the expression builder or specific subnodes.
Consider for example:
\begin{verbatim}
In[1]:= Unprotect[Plus]
Out[1]= {Plus}

In[2]:= Sin[x_]^2 + Cos[y_]^2 := 1

In[3]:= x + Sin[t]^2 + y + Cos[t]^2
Out[3]= 1 + x + y
\end{verbatim}
This expression in Mathematica defines a substitution rule that overloads
the functionality of the \texttt{Plus} node (the node for additions in Mathematica).
A symbol with a trailing underscore is treated as a wildcard.
Although one may wish to keep this identity unevaluated, this example clearly
illustrates the potential to define
one's own immediate transformation rules.
In SymPy, the operations constructing the addition node in the expression tree
are Python class constructors
and cannot be modified at runtime.\footnote{Nonetheless, Python supports monkey
patching but it is a discouraged programming pattern.}
The way SymPy deals with extending the missing runtime overloadability functionality
is by subclassing the node types: subclasses may redefine the class constructor
to yield the proper extended functionality.

%% TODO list:
% * comparison with Mathematica: MatrixExp, product not always commutative, type inheritance (polymorphism) and advantage in unifying the product symbol * for symbols and matrices, pattern matching vs. single dispatch.

% Type inheritance and polymorphism

Unlike SymPy, Mathematica does not support type inheritance or poly\-morph\-ism~\cite{Fateman1992}.
% cite examples of class inheritance in SymPy:
%
SymPy relies heavily on class inheritance, but for the most part,
class inheritance is used to make sure that SymPy objects inherit the proper
methods and implement the basic hashing system.
%There are also cases where inheritance is used to extend the mathematical meaning of an expression.

% Matrices

While Mathematica interprets nested lists as matrices whenever
the sublists have the same length, matrices in SymPy are a type in their own
right, allowing ordinary operators and functions (like multiplication and
exponentiation) to be used as they traditionally are in mathematics.
\begin{verbatim}
>>> exp(Matrix([[1, 1],[0, 2]])) * Matrix([a, b])
Matrix([
[E*a + b*(-E + exp(2))],
[             b*exp(2)]])
\end{verbatim}

Using the standard multiplication in Mathematica performs an element-wise
product and
calling the exponential function \texttt{Exp} on a matrix
returns an element-wise exponentiation of its elements.

% * comparison with Mathematica: avoid misspelling variables through forced declaration (check that you can't do it in Mathematica).
% * evaluate=False vs HoldForm

Unevaluated expressions in Mathematica can be achieved in various ways,
most commonly with the \texttt{HoldForm} or \texttt{Hold} nodes,
that block the evaluation of subnodes by the parser.
Such a node cannot be expressed in Python because of greedy evaluation.
Whenever needed in SymPy, it is necessary to add the parameter \texttt{evaluate=False}
to all subnodes.

% * comparison with Mathematica: == is structural (not mathematical) equality

In Mathematica, the operator \texttt{==} returns a boolean whenever it is able
to immediately evaluate the truth of the equality, otherwise it returns an
\texttt{Equal} expression.  In SymPy, \texttt{==} means structural equality and
is always guaranteed to return a boolean expression.  To express a mathematical equality in
SymPy it is necessary to explicitly construct an instance of the \texttt{Equality}
class.

% * comparison with Mathematica: polynomial module.
% * comparison with Mathematica: space is product, ** vs ^

SymPy, in accordance with Python (and unlike the usual programming convention),
uses \texttt{**} to express the power operator, while Mathematica uses the more
common \verb|^|.

% * comparison with Mathematica: ( ) is Sequence, functions are generally uppercase.
% * comparsion with Mathematica: table of comparison?
% * comparison with Mathematica: Wolfram language has loads of operator overloading, functional paradigm.

SymPy's use of floating-point numbers is similar to that of most
other CASes, including Maple and Maxima.
By contrast, Mathematica uses a form
of significance arithmetic~\cite{Sofroniou2005precise} for approximate numbers.
This offers further protection against numerical errors,
although it comes with its own set of problems
(for a critique of significance arithmetic, see Fateman~\cite{Fateman1992}).
Internally, SymPy's \texttt{evalf} method works similarly to Mathematica's
significance arithmetic, but the semantics are isolated from the rest of the system.

### 2016-08-23 numerics.tex
Use Title Case consistently in section titles
Consistently hyphenate floating-point

@@
-Floating point numbers in SymPy are implemented by the \texttt{Float} class,
+Floating-point numbers in SymPy are implemented by the \texttt{Float} class,
 which represents an arbitrary-precision binary floating-point number by
@@
 
-\subsection{The mpmath library}
+\subsection{The mpmath Library}
 \label{sec:mpmath}

### 2016-08-23 numerics.tex
Add a small intro paragraph to the numerics section (review 3, point 24)

@@
+While SymPy primarily focuses on symbolics, it is impossible to have a
+complete symbolic system without the ability to numerically evaluate
+expressions. Many operations directly use numerical evaluation, such as
+plotting a function, or solving an equation numerically. Beyond this, certain
+purely symbolic operations require numerical evaluation to effectively
+compute. For instance, determining the truth value of $e + 1 > \pi$ is most
+conveniently done by numerically evaluating both sides of the inequality and
+checking which is larger.
+
+\subsection{Floating-Point Numbers}
 Floating-point numbers in SymPy are implemented by the \texttt{Float} class,

### 2016-08-23 domain_specific.tex
Remove unnecessary hyphenation

--- 2016-08-19 domain_specific.tex
@@
 The \verb|sympy.physics.vector| submodule provides reference frame-, time-, and
-space-aware vector and dyadic objects that allow for three-di\-men\-sional
+space-aware vector and dyadic objects that allow for three-dimensional
 operations such as addition, subtraction, scalar multiplication, inner and

### 2016-08-23 rebuttal.tex
Put the removed text back in
This was accidentally removed in
b0d4c326ab43ae1fcada6724eafa7a3bd0cde92c
Add a small intro paragraph to the numerics section (review 3, point 24)
Address reviewer 3 point 33

@@
 \usepackage{textcomp}
+\usepackage[T1]{fontenc}
 
@@
 \begin{solution}
-
+We have added an introduction and a subsection ``Floating-Point Numbers''
+(which, for mathematical clarity, we consider to be distinct from ``real
+numbers'').
 \end{solution}
@@
 \begin{solution}
-
+We have changed ``both of these objects can be written\ldots'' to ``the vector
+and dyadic objects can both be written\ldots''.
 \end{solution}

### 2016-08-23 domain_specific.tex
Address reviewer 3 point 33

@@
 
-The \verb|sympy.physics.vector| submodule provides reference frame-, time-, and
-space-aware vector and dyadic objects that allow for three-dimensional
+The \verb|sympy.physics.vector| submodule provides reference frame-, time-,
+and space-aware vector and dyadic objects that allow for three-dimensional
 operations such as addition, subtraction, scalar multiplication, inner and
-outer products, and cross products. Both of these objects can be written in
-very compact notation that make it easy to express the vectors and dyadics in
-terms of multiple reference frames with arbitrarily defined relative
-orientations. The vectors are used to specify the positions, velocities, and
-accelerations of points; orientations, angular velocities, and angular
-accelerations of reference frames; and forces and torques. The dyadics are
-essentially reference frame-aware $3 \times 3$ tensors~\cite{tai1997generalized}.
-The vector and dyadic
-objects can be used for any one-, two-, or three-dimensional vector algebra, and
-they provide a strong framework for building physics and engineering tools.
+outer products, and cross products. The vector and dyadic objects can both be
+written in very compact notation that make it easy to express the vectors and
+dyadics in terms of multiple reference frames with arbitrarily defined
+relative orientations. The vectors are used to specify the positions,
+velocities, and accelerations of points; orientations, angular velocities, and
+angular accelerations of reference frames; and forces and torques. The dyadics
+are essentially reference frame-aware $3 \times 3$
+tensors~\cite{tai1997generalized}. The vector and dyadic objects can be used
+for any one-, two-, or three-dimensional vector algebra, and they provide a
+strong framework for building physics and engineering tools.
 

### 2016-08-24 rebuttal.tex
Remove carriage return from rebuttal.tex
Apparently this messes up GitHub's diff highlighting.
You can directly import _assume_defined if you use a "from" import

@@
 \begin{solution}
-There is no simple way to access this in code (the simplest way is
-\texttt{sys.\allowbreak{}modules[\allowbreak{}\textquotesingle{}sympy.core.assumptions\textquotesingle{}].\allowbreak{}\_assume\_defined}, which works
-around some technical issues with Python's import system, and uses a
-private method). The recommended list is in the SymPy documentation, at
+  To access this in code, the only way is
+\begin{verbatim}
+from sympy.core.assumptions import _assume_defined
+\end{verbatim}
+  but as this is a private variable (it begins with an underscore), it is not
+  considered part of the public API, so we do not wish to recommend it in the
+  paper. The recommended list is in the SymPy documentation, at
   \url{http://docs.sympy.org/latest/modules/core.html#module-sympy.core.assumptions}.
@@
 Kluyver, T., Ragan-Kelley, B., P{\'e}rez, F.,
-Granger, B., Bussonnier, M., Frederic, J., Kelley,
-K., Hamrick, J., Grout,
+Granger, B., Bussonnier, M., Frederic, J., Kelley, K., Hamrick, J., Grout,
 J., Corlay, S., et al. (2016). Jupyter notebooks---a publishing format for

### 2016-08-24 features.tex
Shorten features table, so it doesn't extend into the margin
Space before parenthesis
Explain how the submodules in the table work
Format submodules in the table in typewriter text

--- 2016-08-16 features.tex
@@
 
-Although SymPy's extensive feature set cannot be covered in-depth in this paper,
-calculus and other bedrock areas are discussed in their own subsections.
-Additionally,
-Table~\ref{features-table} gives a compact listing of all major capabilities
-present in the SymPy codebase. This grants a sampling from the breadth of
-topics and application domains that SymPy services. Unless stated otherwise,
-all features noted in Table~\ref{features-table} are symbolic in nature.
-Numeric features are discussed in Section~\ref{sec:numerics}.
+Although SymPy's extensive feature set cannot be covered in-depth in this
+paper, calculus and other bedrock areas are discussed in their own
+subsections. Additionally, Table~\ref{features-table} gives a compact listing
+of all major capabilities present in the SymPy codebase. This grants a
+sampling from the breadth of topics and application domains that SymPy
+services. Unless stated otherwise, all features noted in
+Table~\ref{features-table} are symbolic in nature. The submodules in
+parentheses are the \texttt{sympy} submodules where the features are
+implemented (like \texttt{sympy.core}). Numeric features are discussed in
+Section~\ref{sec:numerics}.
 
-\begin{longtable}[htbc]{p{0.20\linewidth}p{0.73\linewidth}}
+\begin{longtable}[htbc]{p{0.30\linewidth}p{0.63\linewidth}}
 \caption{SymPy Features and Descriptions\label{features-table}}\\
 \toprule
-\textbf{Feature} & \textbf{Description} \\
+\textbf{Feature (submodules)} & \textbf{Description} \\
 \midrule
-Calculus & Algorithms for computing derivatives, integrals, and limits.\\
+Calculus (\texttt{core}, \texttt{series}, \texttt{integrals}) & Algorithms for computing derivatives, integrals, and limits.\\
 
-Category Theory & Representation of objects, morphisms, and diagrams. Tools
+Category Theory (\texttt{categories}) & Representation of objects, morphisms, and diagrams. Tools
 for drawing diagrams with Xy-pic.\\
 
-Code Generation & Generation of compilable and executable code in a
+Code Generation (\texttt{printing}, \texttt{codegen}) & Generation of compilable and executable code in a
 variety of different programming languages from expressions directly. Target
@@
 
-Combinatorics \& Group Theory & Permutations, combinations,
+Combinatorics \& Group Theory (\texttt{combinatorics}) & Permutations, combinations,
 partitions, subsets, various permutation groups (such as polyhedral, Rubik,
@@
 
-Concrete Math & Summation, products, tools for determining whether summation
+Concrete Math (\texttt{concrete}) & Summation, products, tools for determining whether summation
 and product expressions are convergent, absolutely convergent, hypergeometric,
@@
 
-Cryptography & Block and stream ciphers, including shift, Affine,
+Cryptography (\texttt{crypto}) & Block and stream ciphers, including shift, Affine,
 substitution, Vigen\`{e}re's, Hill's, bifid, RSA, Kid RSA,
@@
 
-Differential Geometry & Representations of manifolds, metrics, tensor
+Differential Geometry (\texttt{diffgeom}) & Representations of manifolds, metrics, tensor
 products, and coordinate systems in Riemannian and pseudo-Riemannian
@@
 
-Geometry & Representations of 2D geometrical entities, such as lines and
+Geometry (\texttt{geometry}) & Representations of 2D geometrical entities, such as lines and
 circles. Enables queries on these entities, such as asking the area of an
@@
 
-Lie Algebras & Representations of Lie algebras and root systems.\\
+Lie Algebras (\texttt{liealgebras}) & Representations of Lie algebras and root systems.\\
 
-Logic & Boolean expressions, equivalence testing, satisfiability, and normal
+Logic (\texttt{logic}) & Boolean expressions, equivalence testing, satisfiability, and normal
 forms.\\
 
-Matrices & Tools for creating matrices of symbols and expressions.
+Matrices (\texttt{matrices}) & Tools for creating matrices of symbols and expressions.
 Both sparse and dense representations, as well as symbolic linear
@@
 
-Matrix Expressions & Matrices with symbolic dimensions (unspecified entries).
+Matrix Expressions (\texttt{matrices.expressions}) & Matrices with symbolic dimensions (unspecified entries).
 Block matrices.\\
 
-Number Theory & Prime number generation, primality testing, integer
+Number Theory (\texttt{ntheory}) & Prime number generation, primality testing, integer
 factorization, continued fractions, Egyptian fractions, modular arithmetic,
@@
 
-Plotting & Hooks for visualizing expressions via matplotlib~\cite{Hunter:2007}
+Plotting (\texttt{plotting}) & Hooks for visualizing expressions via matplotlib~\cite{Hunter:2007}
 or as text drawings when lacking a graphical back-end. 2D function plotting,
@@
 
-Polynomials & Polynomial algebras over various coefficient domains.
+Polynomials (\texttt{polys}) & Polynomial algebras over various coefficient domains.
 Functionality ranges from simple operations (e.g., polynomial division) to
@@
 
-Printing & Functions for printing SymPy expressions in the terminal with ASCII
+Printing (\texttt{printing}) & Functions for printing SymPy expressions in the terminal with ASCII
 or Unicode characters and converting SymPy expressions to \LaTeX{} and
@@
 
-Quantum Mechanics & Quantum states, bra--ket notation, operators, basis sets,
+Quantum Mechanics (\texttt{physics.quantum}) & Quantum states, bra--ket notation, operators, basis sets,
 representations, tensor products, inner products, outer products, commutators,
@@
 
-Series & Series expansion, sequences, and limits of sequences.
+Series (\texttt{series}) & Series expansion, sequences, and limits of sequences.
 This includes Taylor, Laurent, and Puiseux series as well as special series, such
@@
 
-Sets & Representations of empty, finite, and infinite sets (including
-special sets such as for all natural, integer, and complex numbers). Operations
+Sets (\texttt{sets}) & Representations of empty, finite, and infinite sets (including
+special sets such as the natural, integer, and complex numbers). Operations
 on sets such as union, intersection, Cartesian product, and building sets from
@@
 
-Simplification & Functions for manipulating and simplifying expressions.
+Simplification (\texttt{simplify}) & Functions for manipulating and simplifying expressions.
 Includes algorithms for simplifying hypergeometric functions, trigonometric
@@
 
-Solvers & Functions for symbolically solving equations, systems
+Solvers (\texttt{solvers}) & Functions for symbolically solving equations, systems
 of equations, both linear and non-linear, inequalities, ordinary differential
@@
 
-Special Functions & Implementations of a number of well known special functions,
+Special Functions (\texttt{functions}) & Implementations of a number of well known special functions,
 including Dirac delta, Gamma, Beta, Gauss error functions, Fresnel integrals,
@@
 
-Statistics & Support for a random variable type as well as the ability to
+Statistics (\texttt{stats}) & Support for a random variable type as well as the ability to
 declare this variable from prebuilt distribution functions such as
@@
 
-Tensors & Symbolic manipulation of indexed objects.\\
+Tensors (\texttt{tensor}) & Symbolic manipulation of indexed objects.\\
 
-Vectors & Basic operations on vectors and differential calculus with respect
+Vectors (\texttt{vector}) & Basic operations on vectors and differential calculus with respect
 to 3D Cartesian coordinate systems.\\

### 2016-08-24 features.tex
Try putting "sympy." on all the submodules in the table
I'm not sure if I like this. Feel free to revert this commit.

--- 2016-08-16 features.tex
@@
 
-Although SymPy's extensive feature set cannot be covered in-depth in this paper,
-calculus and other bedrock areas are discussed in their own subsections.
-Additionally,
-Table~\ref{features-table} gives a compact listing of all major capabilities
-present in the SymPy codebase. This grants a sampling from the breadth of
-topics and application domains that SymPy services. Unless stated otherwise,
-all features noted in Table~\ref{features-table} are symbolic in nature.
-Numeric features are discussed in Section~\ref{sec:numerics}.
+Although SymPy's extensive feature set cannot be covered in-depth in this
+paper, calculus and other bedrock areas are discussed in their own
+subsections. Additionally, Table~\ref{features-table} gives a compact listing
+of all major capabilities present in the SymPy codebase. This grants a
+sampling from the breadth of topics and application domains that SymPy
+services. Unless stated otherwise, all features noted in
+Table~\ref{features-table} are symbolic in nature. Numeric features are
+discussed in Section~\ref{sec:numerics}.
 
-\begin{longtable}[htbc]{p{0.20\linewidth}p{0.73\linewidth}}
+\begin{longtable}[htbc]{p{0.30\linewidth}p{0.63\linewidth}}
 \caption{SymPy Features and Descriptions\label{features-table}}\\
 \toprule
-\textbf{Feature} & \textbf{Description} \\
+\textbf{Feature (submodules)} & \textbf{Description} \\
 \midrule
-Calculus & Algorithms for computing derivatives, integrals, and limits.\\
+Calculus (\texttt{sympy.core}, \texttt{sympy.series}, \texttt{sympy.integrals}) & Algorithms for computing derivatives, integrals, and limits.\\
 
-Category Theory & Representation of objects, morphisms, and diagrams. Tools
+Category Theory (\texttt{sympy.categories}) & Representation of objects, morphisms, and diagrams. Tools
 for drawing diagrams with Xy-pic.\\
 
-Code Generation & Generation of compilable and executable code in a
+Code Generation (\texttt{sympy.printing}, \texttt{sympy.codegen}) & Generation of compilable and executable code in a
 variety of different programming languages from expressions directly. Target
@@
 
-Combinatorics \& Group Theory & Permutations, combinations,
+Combinatorics \& Group Theory (\texttt{sympy.combinatorics}) & Permutations, combinations,
 partitions, subsets, various permutation groups (such as polyhedral, Rubik,
@@
 
-Concrete Math & Summation, products, tools for determining whether summation
+Concrete Math (\texttt{sympy.concrete}) & Summation, products, tools for determining whether summation
 and product expressions are convergent, absolutely convergent, hypergeometric,
@@
 
-Cryptography & Block and stream ciphers, including shift, Affine,
+Cryptography (\texttt{sympy.crypto}) & Block and stream ciphers, including shift, Affine,
 substitution, Vigen\`{e}re's, Hill's, bifid, RSA, Kid RSA,
@@
 
-Differential Geometry & Representations of manifolds, metrics, tensor
+Differential Geometry (\texttt{sympy.diffgeom}) & Representations of manifolds, metrics, tensor
 products, and coordinate systems in Riemannian and pseudo-Riemannian
@@
 
-Geometry & Representations of 2D geometrical entities, such as lines and
+Geometry (\texttt{sympy.geometry}) & Representations of 2D geometrical entities, such as lines and
 circles. Enables queries on these entities, such as asking the area of an
@@
 
-Lie Algebras & Representations of Lie algebras and root systems.\\
+Lie Algebras (\texttt{sympy.liealgebras}) & Representations of Lie algebras and root systems.\\
 
-Logic & Boolean expressions, equivalence testing, satisfiability, and normal
+Logic (\texttt{sympy.logic}) & Boolean expressions, equivalence testing, satisfiability, and normal
 forms.\\
 
-Matrices & Tools for creating matrices of symbols and expressions.
+Matrices (\texttt{sympy.matrices}) & Tools for creating matrices of symbols and expressions.
 Both sparse and dense representations, as well as symbolic linear
@@
 
-Matrix Expressions & Matrices with symbolic dimensions (unspecified entries).
+Matrix Expressions (\texttt{sympy.matrices.expressions}) & Matrices with symbolic dimensions (unspecified entries).
 Block matrices.\\
 
-Number Theory & Prime number generation, primality testing, integer
+Number Theory (\texttt{sympy.ntheory}) & Prime number generation, primality testing, integer
 factorization, continued fractions, Egyptian fractions, modular arithmetic,
@@
 
-Plotting & Hooks for visualizing expressions via matplotlib~\cite{Hunter:2007}
+Plotting (\texttt{sympy.plotting}) & Hooks for visualizing expressions via matplotlib~\cite{Hunter:2007}
 or as text drawings when lacking a graphical back-end. 2D function plotting,
@@
 
-Polynomials & Polynomial algebras over various coefficient domains.
+Polynomials (\texttt{sympy.polys}) & Polynomial algebras over various coefficient domains.
 Functionality ranges from simple operations (e.g., polynomial division) to
@@
 
-Printing & Functions for printing SymPy expressions in the terminal with ASCII
+Printing (\texttt{sympy.printing}) & Functions for printing SymPy expressions in the terminal with ASCII
 or Unicode characters and converting SymPy expressions to \LaTeX{} and
@@
 
-Quantum Mechanics & Quantum states, bra--ket notation, operators, basis sets,
+Quantum Mechanics (\texttt{sympy.physics.quantum}) & Quantum states, bra--ket notation, operators, basis sets,
 representations, tensor products, inner products, outer products, commutators,
@@
 
-Series & Series expansion, sequences, and limits of sequences.
+Series (\texttt{sympy.series}) & Series expansion, sequences, and limits of sequences.
 This includes Taylor, Laurent, and Puiseux series as well as special series, such
@@
 
-Sets & Representations of empty, finite, and infinite sets (including
-special sets such as for all natural, integer, and complex numbers). Operations
+Sets (\texttt{sympy.sets}) & Representations of empty, finite, and infinite sets (including
+special sets such as the natural, integer, and complex numbers). Operations
 on sets such as union, intersection, Cartesian product, and building sets from
@@
 
-Simplification & Functions for manipulating and simplifying expressions.
+Simplification (\texttt{sympy.simplify}) & Functions for manipulating and simplifying expressions.
 Includes algorithms for simplifying hypergeometric functions, trigonometric
@@
 
-Solvers & Functions for symbolically solving equations, systems
+Solvers (\texttt{sympy.solvers}) & Functions for symbolically solving equations, systems
 of equations, both linear and non-linear, inequalities, ordinary differential
@@
 
-Special Functions & Implementations of a number of well known special functions,
+Special Functions (\texttt{sympy.functions}) & Implementations of a number of well known special functions,
 including Dirac delta, Gamma, Beta, Gauss error functions, Fresnel integrals,
@@
 
-Statistics & Support for a random variable type as well as the ability to
+Statistics (\texttt{sympy.stats}) & Support for a random variable type as well as the ability to
 declare this variable from prebuilt distribution functions such as
@@
 
-Tensors & Symbolic manipulation of indexed objects.\\
+Tensors (\texttt{sympy.tensor}) & Symbolic manipulation of indexed objects.\\
 
-Vectors & Basic operations on vectors and differential calculus with respect
+Vectors (\texttt{sympy.vector}) & Basic operations on vectors and differential calculus with respect
 to 3D Cartesian coordinate systems.\\

### 2016-08-24 rebuttal.tex
Fix some grammar

@@
 \begin{solution}
-IEEE 829 defines the term feature as ``A distinguishing
+IEEE 829 defines the term feature as ``a distinguishing
 characteristic of a software item (e.g., performance, portability,

### 2016-08-26 rebuttal.tex
Use correct quotes
Better response for SymPy Gamma figure

@@
 \begin{solution}
-Now we use \texttt{symbols('a b c d')}, like in other places.
+Now we use \texttt{symbols(\textquotesingle{}a b c d\textquotesingle{})}, like in other places.
 \end{solution}
@@
 \begin{solution}
-We believe this is not a good idea.  SymPy Gamma currently
-don't export output as Jupyter notebook or other external
-format, but HTML.
+We believe this is not a good idea.  This figure is meant to be an archival
+capture, not software subject to being re-rendered. SymPy Gamma currently only
+renders as HTML\@. SymPy Gamma doesn't export output as \LaTeX{} or other external
+formats.
 
-URL to the project site was added.
+A URL to the project site was added.
 \end{solution}

### 2016-08-26 architecture.tex
Make it clear that SymPy 1.0 doesn't use SymEngine
sympy.mechanics -> sympy.physics.mechanics
Fix some minor LaTeX issues

@@
 class method \texttt{eval}, which returns a canonical form of the function
-application (usually an instance of some other class, i.e. a \texttt{Number})
+application (usually an instance of some other class, i.e., a \texttt{Number})
 or \texttt{None}, if for given arguments that function should not be
@@
 It is a pure C++ library with thin wrappers to other languages (Python, Ruby,
-Julia, ...) whose aim is to be the fastest manipulation library.  Preliminary
+Julia, \ldots) whose aim is to be the fastest manipulation library.  Preliminary
 benchmarks suggest that SymEngine is as fast or faster than the commercial or
@@
 
-SymPy recently started to use SymEngine as an optional backend, initially in
-\texttt{sympy.mechanics} only. The plan is to allow more algorithms in SymPy to
-take advantage of the speed of SymEngine.
+The development branch of SymPy recently started to use SymEngine as an
+optional backend, initially in \texttt{sympy.physics.mechanics} only. The plan is
+to allow more algorithms in SymPy to take advantage of the speed of SymEngine.

### 2016-08-26 rebuttal.tex
Fix some chktex issues

@@
 \begin{solution}
-A new paragraph titled "Speed" was added to the manuscript, which discusses the
+A new paragraph titled ``Speed'' was added to the manuscript, which discusses the
 speed issue and lays out a path forward. The conclusion was simplified.
@@
 \begin{questions}
-\question Generally good. A part that confused me is the assertion (footnote 3) that "If
+\question Generally good. A part that confused me is the assertion (footnote 3) that ``If
 A and B are Symbols created with commutative=False then SymPy will keep A·B
-and B·A distinct." Does that mean that BOTH of them must be created this way,
+and B·A distinct.'' Does that mean that BOTH of them must be created this way,
 and that A and x (if x is created normally) will commute? Is there any way to
@@
 \begin{questions}
-\question The referee is asked to comment whether the research is withinthe scope of the
-journal, defined as "PeerJ is an Open Access, peer-reviewed, scholarly
+\question The referee is asked to comment whether the research is within the scope of the
+journal, defined as ``PeerJ is an Open Access, peer-reviewed, scholarly
 journal. It considers articles in the Biological Sciences, Medical Sciences,
@@
 articles in those areas have clear applicability to the core areas of
-Biological, Medical or Health sciences)."
+Biological, Medical or Health sciences).''
 
@@
 
+\begin{quote}
 The measure parameter of the simplify function lets specify the Python function used to determine how
+\end{quote}
 
-should this be..
-
+should this be
+\begin{quote}
 The measure parameter of the simplify function lets the user specify the Python function used to determine how
-
+\end{quote}
 \begin{solution}

### 2016-08-26 acknowledgements.tex
Fix a typo
Fix some chktex issues

@@
 
-The author of this paper Richard P. Muller thanks Sandia National Laboratory
-for its financial support.
+The author of this paper Richard P. Muller thanks Sandia National Laboratories
+for their financial support.
 
@@
 Forschungsgemeinschaft (DFG) for its financial support via the International
-Research Training Group 1524 "Self-Assembled Soft Matter Nano-Structures at
-Interfaces."
+Research Training Group 1524 ``Self-Assembled Soft Matter Nano-Structures at
+Interfaces.''

### 2016-08-30 rebuttal.tex
Use CAS's (reviewer 3 comment 3)

@@
 \begin{solution}
-  % TODO: I recall we looked this up and the spelling consensus agreed with
-  % the way we have it. We should double check (perhaps we can email the
-  % editor to ask if the journal has a convention).
+  Different style guides appear to disagree on the proper way to pluralize
+  non-plural acronyms that end in ``S''. We have been informed by the PeerJ
+  publishing staff that either form (``CASs'' or ``CAS's'') is acceptable, so
+  long as we are consistent. We have chosen ``CAS's'' as suggested, and have
+  changed all instances in the manuscript.
 \end{solution}

### 2016-09-11 rebuttal.tex
Fix reviewer 3 comment 7
reviewer 1 point 2 (Carette's citation replaced)

--- 2016-08-26 rebuttal.tex
@@
 \begin{solution}
-  % TODO. I agree that [8] is a bad reference, because we are not actually
-  % using their definition. The aim was just to have a reference for
-  % "simplification is not well-defined", but that paper aims to define it.
-  % What is a better reference here? Perhaps Moses [19] from the Carette
-  % paper? Perhaps we can remove the reference (the referenced statement is
-  % "it must be emphasized that simplification is not a rigorously defined
-  % mathematical operation").
+We agree that this is a bad reference, because we are not actually
+using their definition.  Yes, that article rigorously define
+``simplification'', but it's just one approach.  This citation
+was replaced with a better reference, that illustrate our point.
 \end{solution}
@@
 \begin{solution}
-  % TODO: We had decided not to use first person. We should perhaps email the editor
-  % to ask about this.
+We had decided not to use first person.  We prefer to be
+consistent here, unless this does violate some journal policy,
+which is not the case (see, for example,
+\url{https://peerj.com/articles/cs-80/}).
 \end{solution}

### 2016-09-11 rebuttal.tex
Renamed "Domain Specific" section per reviewer 3 point 30
Address point 39 of the reviewer 3

@@
 \begin{solution}
-  % TODO: Should we rename it to "Physics submodules"?
-(see also the discussion at~\ref{rev1point37}).
+This section was renamed to "Physics submodule".  (See also the
+discussion at~\ref{rev1point37} about module vs package).
 \end{solution}
@@
 \begin{solution}
+Physics is a very wide subject, so it would be too rough description to mention
+it in the way you suggest, as a single domain.  Also, we list here only two of mentioned above
+submodules in the \texttt{sympy.physics}, so \emph{there are} others.
 
+We believe the current wording does make sense.
 \end{solution}

### 2016-09-11 supplement.tex
Use Title Case consistently in section titles
Add an introduction for the supplement; address point 43 of the reviewer 3

@@
 
+Section \ref{suppsec:Gruntz} discusses the Gruntz algorithm, used to
+calculate limits in the SymPy.  Sections \ref{suppsec:Series}--\ref{suppsec:Ten}
+discuss in depth some selected submodules.  Section \ref{suppsec:numsimpl}
+discusses numerical simplification.  Section \ref{suppsec:examples} provides
+additional examples for topics, discussed in the main paper.  In sections
+\ref{sympy-gamma} and \ref{sympy-live} introduced the SymPy Gamma
+and the SymPy Live projects.  Finally, section \ref{comp-mma} do
+brief comparison of the SymPy with the Wolfram Mathematica and the
+section \ref{other-proj} lists some projects that depend on SymPy.
+
 \section{Limits: The Gruntz Algorithm}
-
+\label{suppsec:Gruntz}
 \input{gruntz.tex}
@@
 \section{Series}
-
+\label{suppsec:Series}
 \input{series.tex}
@@
 \section{Logic}
-
+\label{suppsec:Logic}
 \input{logic}
@@
 \section{Diophantine Equations}
-
+\label{suppsec:Dioph}
 \input{diophantine}
@@
 \section{Sets}
+\label{suppsec:Sets}
 \input{sets}
 
+\section{Statistics}
+\label{suppsec:Stats}
+\input{stats}
+
 \section{Category Theory}
+\label{suppsec:Cat}
 \input{categories}
+
+\section{Tensors}
+\label{suppsec:Ten}
+\input{tensors}
+
+\section{Numerical Simplification}
+\label{suppsec:numsimpl}
+\input{nsimplify}
+
+\section{Examples}
+\label{suppsec:examples}
+\input{examples}
 
@@
 \section{Comparison with Mathematica}
-
+\label{comp-mma}
 \input{comparison_with_mma}
 
-\section{Other Projects that use SymPy}
-
+\section{Other Projects that Use SymPy}
+\label{other-proj}
 \input{other_projects_that_use_sympy}
-
-\section{Tensors}
-
-\input{tensors}
-
-\section{Numerical simplification}
-
-\input{nsimplify}
-
-\section{Examples}
-
-\input{examples}
-
-\section{References}
 

### 2016-09-11 other_projects_that_use_sympy.tex
Clarify "usage" of Sympy: that use -> that depend on
Ondrej (and perhaps Aaron) won't mention the Diofant project
among SymPy's users, so we must clarify that we mean by "using".

There are several projects that depend on SymPy as a library for implementing
a part of their functionality. Some of them are listed below:

\begin{itemize}
\item
  \href{http://cadabra.science/index.html}{\textbf{Cadabra}}: Cadabra is
  a CAS designed specifically for the
  resolution of problems encountered in field theory.
\item
  \href{http://octave.sourceforge.net/symbolic/}{\textbf{Octave Symbolic}}:
  The Octave-Forge Symbolic package adds symbolic calculation features
  to GNU Octave. These include common CAS tools such
  as algebraic operations, calculus, equation solving, Fourier and
  Laplace transforms, variable precision arithmetic, and other features.
\item
  \href{https://github.com/jverzani/SymPy.jl}{\textbf{SymPy.jl}}:
  Provides a Julia interface to SymPy using PyCall.
\item
  \href{https://mathics.github.io/}{\textbf{Mathics}}: Mathics is a
  free, general-purpose online CAS featuring Mathematica compatible
  syntax and functions. It is backed by highly extensible Python code,
  relying on SymPy for most mathematical tasks.
\item
  \href{http://mathpix.com/}{\textbf{Mathpix}}: An iOS App, that detects handwritten math as input, and uses
  SymPy Gamma to evaluate the math input and generate the relevant
  steps to solve the problem.
\item
  \href{http://openrave.org/docs/0.8.2/openravepy/ikfast/}{\textbf{IKFast}}:
  IKFast is a robot kinematics compiler provided by
  \href{http://openrave.org/}{OpenRAVE}. It analytically solves robot inverse
  kinematics equations and generates optimized C++ files. It uses SymPy for
  its internal symbolic mathematics.
\item
  \href{http://www.sagemath.org/}{\textbf{Sage}}: A CAS, visioned to be
  a viable free open source alternative to Magma, Maple, Mathematica and
  MATLAB\@. Sage includes many open source mathematical libraries, including
  SymPy.
\item
  \href{https://cloud.sagemath.com}{\textbf{SageMathCloud}}:
  SageMathCloud is a web-based cloud computing and course management
  platform for computational mathematics.
\item
  \href{http://www.pydy.org/}{\textbf{PyDy}}: Multibody Dynamics with
  Python.
\item
  \href{https://github.com/brombo/galgebra}{\textbf{galgebra}}:
  Geometric algebra (previously \texttt{sympy.galgebra}).
\item
  \href{http://yt-project.org/}{\textbf{yt}}: Python package for
  analyzing and visualizing volumetric data (\texttt{yt.units} uses SymPy).
\item
  \href{http://sfepy.org/}{\textbf{SfePy}}: Simple finite elements in
  Python, see section~\ref{sfepy}.
\item
  \href{http://quameon.sourceforge.net/}{\textbf{Quameon}}: Quantum
  Monte Carlo in Python.
\item
  \href{http://lcapy.elec.canterbury.ac.nz/}{\textbf{Lcapy}}:
  Experimental Python package for teaching linear circuit analysis.
\item
  \href{http://digitalcommons.calpoly.edu/cgi/viewcontent.cgi?article=1072\&context=physsp/}{\textbf{Quantum
  Programming in Python}}: Quantum 1D Simple Harmonic Oscillator and
  Quantum Mapping Gate.
\item
  \href{http://mech.fsv.cvut.cz/~stransky/software/latexexpr/doc/}{\textbf{LaTeX
  Expression project}}: Easy \LaTeX{} typesetting of algebraic expressions
  in symbolic form with automatic substitution and result computation.
\item
  \href{https://www.researchgate.net/publication/260585491_Symbolic_Statistics_with_SymPy/}{\textbf{Symbolic
  statistical modeling}}: Adding statistical operations to complex
  physical models.
\end{itemize}

\subsection{SfePy}\label{sfepy}

\href{http://sfepy.org/}{\textbf{SfePy}} (Simple finite elements in Python),
cf.~\cite{cimrman2014sfepy}, is a Python package for solving partial
differential equations (PDEs) in 1D, 2D and 3D by the finite element (FE)
method~\cite{Zienkiewicz2013FEM}. SymPy is used within this package mostly for
code generation and testing, namely:

\begin{itemize}
\item generation of the hierarchical FE basis module, involving generation and
  symbolic differentiation of 1D Legendre and Lobatto polynomials, constructing
  the FE basis polynomials~\cite{Solin2003hpfem} and generating the C code;
\item generation of symbolic conversion formulas for various groups of elastic
  constants~\cite{Fung1993contmech}:  provide any two of the Young's modulus,
  Poisson's ratio, bulk modulus, Lam\'{e}'s first parameter, shear modulus
  (Lam\'{e}'s second parameter) or longitudinal wave modulus and get the other
  ones;
\item simple physical unit conversions, generation of consistent unit sets;
\item testing FE solutions using method of manufactured (analytical)
  solutions:  the differential operator of a PDE is symbolically applied and a symbolic
  right-hand side is created, evaluated in quadrature points, and subsequently
  used to obtain a numerical solution that is then compared to the analytical
  one;
\item testing accuracy of 1D, 2D and 3D numerical quadrature formulas
  (cf.~\cite{Abramowitz1964handbook}) by generating polynomials of suitable
  orders, integrating them, and comparing the results with those obtained by
  the numerical quadrature.
\end{itemize}

### 2016-09-11 rebuttal.tex
Add an introduction for the supplement; address point 43 of the reviewer 3
Correct typo in reviewer 2 example
Add major points from reviewer 3

@@
 from sympy import *
-x,y,z=symbols(′xyz′)
+x,y,z=symbols("x,y,z")
 test=array(123)
@@
 \section{Reviewer 3}
+
+\begin{questions}
+\question I suggest a major rewrite of the paper, to improve the structure and group differently approached subjects. It would probably be best to:
+\begin{enumerate}
+\item Create a new section (following the introduction of SymPy) on projects that use SymPy, and put in it the materials currently available in supplement's sections 8, 9, 11.
+The comparison with Mathematica (supplement's section 10) should be moved either to the introduction as a section, or right after the introduction as its own section, but it should not be in the middle of the description of SymPy-powered projects.
+\item  This should be followed by the list of SymPy packages and modules (currently section 3) and descriptions of selected modules (currently sections 3.1, 3.3, 3.5, supplement's sections 5, 7, and 9).
+\item Now, basic usage can be given as its own section (currently done in sections 2.1 and 2.3), followed by introductions on usage of various modules (currently in sections 3.2, 3.4, 3.6, 4, 5, and supplement's sections 2, 3, 4, 6).
+\item In-depth architecture (most of the so far unmentioned sections) can be given either as its own section, or made into a supplement of its own (as it is naturally far more technical and less interesting to general audience).
+\item The current conclusion works fine as the finishing section of the paper.
+\end{enumerate}
+\begin{solution}
+\end{solution}
+\end{questions}
 
@@
 \begin{solution}
-
+We add an introduction.
 \end{solution}

### 2016-09-12 other_projects_that_use_sympy.tex
Drop special SfePy section

@@
 \item
-  \href{http://sfepy.org/}{\textbf{SfePy}}: Simple finite elements in
-  Python, see section~\ref{sfepy}.
+  \href{http://sfepy.org/}{\textbf{SfePy}} (Simple finite elements in Python),
+  cf.~\cite{cimrman2014sfepy}, is a Python package for solving partial
+  differential equations (PDEs) in 1D, 2D and 3D by the finite element (FE)
+  method~\cite{Zienkiewicz2013FEM}. SymPy is used within this package mostly for
+  code generation and testing.
 \item
@@
 \end{itemize}
-
-\subsection{SfePy}\label{sfepy}
-
-\href{http://sfepy.org/}{\textbf{SfePy}} (Simple finite elements in Python),
-cf.~\cite{cimrman2014sfepy}, is a Python package for solving partial
-differential equations (PDEs) in 1D, 2D and 3D by the finite element (FE)
-method~\cite{Zienkiewicz2013FEM}. SymPy is used within this package mostly for
-code generation and testing, namely:
-
-\begin{itemize}
-\item generation of the hierarchical FE basis module, involving generation and
-  symbolic differentiation of 1D Legendre and Lobatto polynomials, constructing
-  the FE basis polynomials~\cite{Solin2003hpfem} and generating the C code;
-\item generation of symbolic conversion formulas for various groups of elastic
-  constants~\cite{Fung1993contmech}:  provide any two of the Young's modulus,
-  Poisson's ratio, bulk modulus, Lam\'{e}'s first parameter, shear modulus
-  (Lam\'{e}'s second parameter) or longitudinal wave modulus and get the other
-  ones;
-\item simple physical unit conversions, generation of consistent unit sets;
-\item testing FE solutions using method of manufactured (analytical)
-  solutions:  the differential operator of a PDE is symbolically applied and a symbolic
-  right-hand side is created, evaluated in quadrature points, and subsequently
-  used to obtain a numerical solution that is then compared to the analytical
-  one;
-\item testing accuracy of 1D, 2D and 3D numerical quadrature formulas
-  (cf.~\cite{Abramowitz1964handbook}) by generating polynomials of suitable
-  orders, integrating them, and comparing the results with those obtained by
-  the numerical quadrature.
-\end{itemize}

### 2016-09-12 architecture.tex
Move subsection "Basic usage" out of "Architecture"

@@
 user-facing code as appropriate.
-
-\subsection{Basic Usage}
-
-% symbols, various ways to declare them
-
-Symbolic variables, called symbols, must be defined and assigned to
-Python variables before they can be used. This is typically done through the
-\texttt{symbols} function, which may create multiple symbols in a single
-function call. For instance,
-\begin{verbatim}
->>> x, y, z = symbols('x y z')
-\end{verbatim}
-creates three symbols representing variables named $x$, $y$, and $z$. In this
-particular instance, these symbols are all assigned to Python variables of the
-same name. However, the user is free to assign them to different
-Python variables, while representing the same symbol, such as
-\texttt{a, b, c = symbols(\textquotesingle{}x y z\textquotesingle{})}.
-In order to minimize potential confusion, though, all examples in this paper will
-assume that
-the symbols \verb|x|, \verb|y|, and \verb|z| have been assigned to Python variables
-identical to their symbolic names.
-
-Expressions are created from symbols using Python's mathematical syntax.  For
-instance, the following Python code creates the expression $(x^2 - 2x + 3)/y$.
-Note that the expression remains unevaluated: it is represented symbolically.
-
-\begin{verbatim}
->>> (x**2 - 2*x + 3)/y
-(x**2 - 2*x + 3)/y
-\end{verbatim}
-
-Importantly, SymPy expressions are immutable. This simplifies the design of
-SymPy by allowing expression interning. It also enables expressions to be
-hashed, that is used to implement caching in SymPy.
 

### 2016-09-12 features.tex
Left-justify the first column of the features table
See http://tex.stackexchange.com/a/52247/2111.
Move subsection "Basic usage" out of "Architecture"

@@
 
-\begin{longtable}[htbc]{p{0.30\linewidth}p{0.63\linewidth}}
+\begin{longtable}[htbc]{>{\raggedright}p{0.30\linewidth}p{0.63\linewidth}}
 \caption{SymPy Features and Descriptions\label{features-table}}\\
@@
 
+\subsection{Basic Usage}
+
+% symbols, various ways to declare them
+
+Symbolic variables, called symbols, must be defined and assigned to
+Python variables before they can be used. This is typically done through the
+\texttt{symbols} function, which may create multiple symbols in a single
+function call. For instance,
+\begin{verbatim}
+>>> x, y, z = symbols('x y z')
+\end{verbatim}
+creates three symbols representing variables named $x$, $y$, and $z$. In this
+particular instance, these symbols are all assigned to Python variables of the
+same name. However, the user is free to assign them to different
+Python variables, while representing the same symbol, such as
+\texttt{a, b, c = symbols(\textquotesingle{}x y z\textquotesingle{})}.
+In order to minimize potential confusion, though, all examples in this paper will
+assume that
+the symbols \verb|x|, \verb|y|, and \verb|z| have been assigned to Python variables
+identical to their symbolic names.
+
+Expressions are created from symbols using Python's mathematical syntax.  For
+instance, the following Python code creates the expression $(x^2 - 2x + 3)/y$.
+Note that the expression remains unevaluated: it is represented symbolically.
+
+\begin{verbatim}
+>>> (x**2 - 2*x + 3)/y
+(x**2 - 2*x + 3)/y
+\end{verbatim}
+
+Importantly, SymPy expressions are immutable. This simplifies the design of
+SymPy by allowing expression interning. It also enables expressions to be
+hashed, that is used to implement caching in SymPy.
+
 \subsection{Simplification}

### 2016-09-12 architecture.tex
Move assumption stuff from architecture to features.

@@
 operators and numeric literals.
-
-%%
-%% Assumptions
-\subsection{Assumptions}
-\label{sec:assumptions}
-
-SymPy performs logical inference through its assumptions system. The
-assumptions system allows users to specify that symbols have certain common
-mathematical properties, such as being positive, imaginary, or integral. SymPy
-is careful to never perform simplifications on an expression unless the
-assumptions allow them. For instance, the identity $\sqrt{t^2} = t$ holds if
-$t$ is nonnegative ($t\ge 0$). However, for general complex $t$, no
-such identity holds.
-
-By default, SymPy performs all calculations assuming that symbols are
-complex valued. This assumption makes it easier to treat mathematical problems
-in full generality.
-\begin{verbatim}
->>> t = Symbol('t')
->>> sqrt(t**2)
-sqrt(t**2)
-\end{verbatim}
-
-By assuming the most general case, that \texttt{t} is complex by default, SymPy
-avoids performing mathematically invalid operations. However, in many cases
-users will wish to simplify expressions containing terms like $\sqrt{t^2}$.
-
-Assumptions are set on \texttt{Symbol} objects when they are created. For
-instance \texttt{Symbol(\textquotesingle{}t\textquotesingle{}, positive=True)} will create a symbol named
-\texttt{t} that is assumed to be positive.
-\begin{verbatim}
->>> t = Symbol('t', positive=True)
->>> sqrt(t**2)
-t
-\end{verbatim}
-
-Some of the common assumptions that SymPy allows are \texttt{positive},
-\texttt{negative}, \texttt{real}, \texttt{nonpositive}, \texttt{integer},
-\texttt{prime} and \texttt{commutative}.\footnote{SymPy assumes that two
-  expressions $A$ and $B$ commute with each other multiplicatively, that is,
-  $A\cdot B=B\cdot A$, unless they both have \texttt{commutative=False}. Many
-  algorithms in SymPy require special consideration to work correctly with
-  noncommutative products.} Assumptions on any object can be checked with the
-\verb|is_|\texttt{\textit{assumption}} attributes, like \verb|t.is_positive|.
-
-Assumptions are only needed to restrict a domain so that certain
-simplifications can be performed. They are not required to make the domain match
-the input of a function. For instance, one can create the object
-$\sum_{n=0}^m f(n)$ as \verb|Sum(f(n), (n, 0, m))| without setting
-\texttt{integer=True} when creating the Symbol object \texttt{n}.
-
-The assumptions system additionally has deductive capabilities. The
-assumptions use a three-valued logic using the Python built in objects
-\texttt{True}, \texttt{False}, and \texttt{None}. Note that \texttt{False}
-is returned if the SymPy object doesn't or can't have the assumption.
-For example, both \texttt{I.is\_real} and \texttt{I.is\_prime} return
-\texttt{False} for the imaginary unit \texttt{I}.
-
-\texttt{None} represents the
-``unknown'' case. This could mean that given assumptions do not unambiguously
-specify the truth of an attribute. For instance,
-\texttt{Symbol(\textquotesingle{}x\textquotesingle{}, real=True).is\_positive}
-will give \texttt{None} because a real symbol might be positive or negative.
-The \texttt{None} could also mean that not enough is known or implemented to
-compute the given fact. For instance, \texttt{(pi + E).is\_irrational} gives
-\texttt{None}, because determining whether $\pi + e$ is rational or irrational
-is an open problem in mathematics~\cite{lang1966introduction}.
-
-Basic implications between the facts are used to deduce assumptions. For
-instance, the assumptions system knows that being an integer implies being
-rational.
-\begin{verbatim}
->>> i = Symbol('i', integer=True)
->>> i.is_rational
-True
-\end{verbatim}
-Furthermore, expressions compute the assumptions on themselves
-based on the assumptions of their arguments. For instance, if \texttt{x} and
-\texttt{y} are both created with \texttt{positive=True}, then
-\texttt{(x + y).is\_positive} will be \texttt{True} whereas
-\texttt{(x - y).is\_positive} will be \texttt{None}.
 

### 2016-09-12 features.tex
Move assumption stuff from architecture to features.

@@
 
+%% Assumptions
+\subsection{Assumptions}
+\label{sec:assumptions}
+
+SymPy performs logical inference through its assumptions system. The
+assumptions system allows users to specify that symbols have certain common
+mathematical properties, such as being positive, imaginary, or integral. SymPy
+is careful to never perform simplifications on an expression unless the
+assumptions allow them. For instance, the identity $\sqrt{t^2} = t$ holds if
+$t$ is nonnegative ($t\ge 0$). However, for general complex $t$, no
+such identity holds.
+
+By default, SymPy performs all calculations assuming that symbols are
+complex valued. This assumption makes it easier to treat mathematical problems
+in full generality.
+\begin{verbatim}
+>>> t = Symbol('t')
+>>> sqrt(t**2)
+sqrt(t**2)
+\end{verbatim}
+
+By assuming the most general case, that \texttt{t} is complex by default, SymPy
+avoids performing mathematically invalid operations. However, in many cases
+users will wish to simplify expressions containing terms like $\sqrt{t^2}$.
+
+Assumptions are set on \texttt{Symbol} objects when they are created. For
+instance \texttt{Symbol(\textquotesingle{}t\textquotesingle{}, positive=True)} will create a symbol named
+\texttt{t} that is assumed to be positive.
+\begin{verbatim}
+>>> t = Symbol('t', positive=True)
+>>> sqrt(t**2)
+t
+\end{verbatim}
+
+Some of the common assumptions that SymPy allows are \texttt{positive},
+\texttt{negative}, \texttt{real}, \texttt{nonpositive}, \texttt{integer},
+\texttt{prime} and \texttt{commutative}.\footnote{SymPy assumes that two
+  expressions $A$ and $B$ commute with each other multiplicatively, that is,
+  $A\cdot B=B\cdot A$, unless they both have \texttt{commutative=False}. Many
+  algorithms in SymPy require special consideration to work correctly with
+  noncommutative products.} Assumptions on any object can be checked with the
+\verb|is_|\texttt{\textit{assumption}} attributes, like \verb|t.is_positive|.
+
+Assumptions are only needed to restrict a domain so that certain
+simplifications can be performed. They are not required to make the domain match
+the input of a function. For instance, one can create the object
+$\sum_{n=0}^m f(n)$ as \verb|Sum(f(n), (n, 0, m))| without setting
+\texttt{integer=True} when creating the Symbol object \texttt{n}.
+
+The assumptions system additionally has deductive capabilities. The
+assumptions use a three-valued logic using the Python built in objects
+\texttt{True}, \texttt{False}, and \texttt{None}. Note that \texttt{False}
+is returned if the SymPy object doesn't or can't have the assumption.
+For example, both \texttt{I.is\_real} and \texttt{I.is\_prime} return
+\texttt{False} for the imaginary unit \texttt{I}.
+
+\texttt{None} represents the
+``unknown'' case. This could mean that given assumptions do not unambiguously
+specify the truth of an attribute. For instance,
+\texttt{Symbol(\textquotesingle{}x\textquotesingle{}, real=True).is\_positive}
+will give \texttt{None} because a real symbol might be positive or negative.
+The \texttt{None} could also mean that not enough is known or implemented to
+compute the given fact. For instance, \texttt{(pi + E).is\_irrational} gives
+\texttt{None}, because determining whether $\pi + e$ is rational or irrational
+is an open problem in mathematics~\cite{lang1966introduction}.
+
+Basic implications between the facts are used to deduce assumptions. For
+instance, the assumptions system knows that being an integer implies being
+rational.
+\begin{verbatim}
+>>> i = Symbol('i', integer=True)
+>>> i.is_rational
+True
+\end{verbatim}
+Furthermore, expressions compute the assumptions on themselves
+based on the assumptions of their arguments. For instance, if \texttt{x} and
+\texttt{y} are both created with \texttt{positive=True}, then
+\texttt{(x + y).is\_positive} will be \texttt{True} whereas
+\texttt{(x - y).is\_positive} will be \texttt{None}.
+
 \subsection{Simplification}

### 2016-09-12 rebuttal.tex
Add some thoughts about major rewrite

@@
 \begin{solution}
+  FIXME.
+  \begin{enumerate}
+    \item I believe we shouldn't move ``Other Projects'' section to the
+      main paper content.  It either should stay in the supplement or we
+      should do a moderate cleanup for this section.  E.g. list only few
+      projects here and put other in the supplement.  Or/And show here
+      only project names.
+
+    \item Logical place for above section - at the end of the
+      paper, right before the "conclusion" section.
+
+    \item We should either drop the Mathematica section or discuss
+      other CAS too, which make this stuff suitable only as part
+      of the suppliment.  I prefer the first variant.
+  \end{enumerate}
 \end{solution}

### 2016-09-22 rebuttal.tex
Grammar

@@
 We agree that this is a bad reference, because we are not actually
-using their definition.  Yes, that article rigorously define
+using their definition.  Yes, that article rigorously defines
 ``simplification'', but it's just one approach.  This citation
-was replaced with a better reference, that illustrate our point.
+was replaced with a better reference [28], that illustrates our point.
 \end{solution}

### 2016-09-22 rebuttal.tex
Grammar

@@
 \begin{solution}
-Physics is a very wide subject, so it would be too rough description to mention
-it in the way you suggest, as a single domain.  Also, we list here only two of mentioned above
-submodules in the \texttt{sympy.physics}, so \emph{there are} others.
+Physics is a very wide subject, so it would be too rough a description to mention
+it in the way you suggest, as a single domain.  Also, we list here only two of
+the above mentioned
+submodules in \texttt{sympy.physics}, but there are others.
 

### 2016-09-22 supplement.tex
Clarify "usage" of Sympy: that use -> that depend on
Ondrej (and perhaps Aaron) won't mention the Diofant project
among SymPy's users, so we must clarify that we mean by "using".
Add an intro sentence to the supplement, and re-arrange the paragraph
Grammar

@@
 
+The supplementary material take a deeper look at certain topics in SymPy for
+which there was not enough room to discuss in the paper.
+Section \ref{suppsec:Gruntz} discusses the Gruntz algorithm, used to
+calculate limits in the SymPy.  Sections \ref{suppsec:Series}--\ref{suppsec:Ten}
+discuss in depth some selected submodules.  Section \ref{suppsec:numsimpl}
+discusses numerical simplification.  Section \ref{suppsec:examples} provides
+additional examples for topics, discussed in the main paper.  In sections
+\ref{sympy-gamma} and \ref{sympy-live} the SymPy Gamma
+and the SymPy Live projects are introduced.  Finally, section \ref{comp-mma} do
+brief comparison of the SymPy with the Wolfram Mathematica and the
+section \ref{other-proj} lists some projects that depend on SymPy.
+
 As in the paper, all examples in the supplement assume that the following
@@
 \end{verbatim}
-
-Section \ref{suppsec:Gruntz} discusses the Gruntz algorithm, used to
-calculate limits in the SymPy.  Sections \ref{suppsec:Series}--\ref{suppsec:Ten}
-discuss in depth some selected submodules.  Section \ref{suppsec:numsimpl}
-discusses numerical simplification.  Section \ref{suppsec:examples} provides
-additional examples for topics, discussed in the main paper.  In sections
-\ref{sympy-gamma} and \ref{sympy-live} introduced the SymPy Gamma
-and the SymPy Live projects.  Finally, section \ref{comp-mma} do
-brief comparison of the SymPy with the Wolfram Mathematica and the
-section \ref{other-proj} lists some projects that depend on SymPy.
 
@@
 
-\section{Other Projects that Use SymPy}
+\section{Other Projects that depend on SymPy}
 \label{other-proj}

### 2016-09-22 supplement.tex
Grammar

@@
 \ref{sympy-gamma} and \ref{sympy-live} the SymPy Gamma
-and the SymPy Live projects are introduced.  Finally, section \ref{comp-mma} do
-brief comparison of the SymPy with the Wolfram Mathematica and the
+and the SymPy Live projects are introduced.  Finally, section \ref{comp-mma}
+has a
+brief comparison of SymPy with Wolfram Mathematica, and
 section \ref{other-proj} lists some projects that depend on SymPy.

### 2016-09-22 features.tex
Move basic usage to the top of the section, and add an intro paragraph
I feel this is more motivating, to jump in with some usage, before the
detailed listing of features.

@@
 % Features to discuss in-depth:
+This section gives a basic introduction of SymPy, and lists its features.
+A few features---assumptions, simplification, calculus, polynomials, printers,
+solvers, and matrices---are core components of SymPy and are discussed in
+depth. Many other features are discussed in depth in the supplementary
+material.
+
+\subsection{Basic Usage}
+
+% symbols, various ways to declare them
+
+Symbolic variables, called symbols, must be defined and assigned to
+Python variables before they can be used. This is typically done through the
+\texttt{symbols} function, which may create multiple symbols in a single
+function call. For instance,
+\begin{verbatim}
+>>> x, y, z = symbols('x y z')
+\end{verbatim}
+creates three symbols representing variables named $x$, $y$, and $z$. In this
+particular instance, these symbols are all assigned to Python variables of the
+same name. However, the user is free to assign them to different
+Python variables, while representing the same symbol, such as
+\texttt{a, b, c = symbols(\textquotesingle{}x y z\textquotesingle{})}.
+In order to minimize potential confusion, though, all examples in this paper will
+assume that
+the symbols \verb|x|, \verb|y|, and \verb|z| have been assigned to Python variables
+identical to their symbolic names.
+
+Expressions are created from symbols using Python's mathematical syntax.  For
+instance, the following Python code creates the expression $(x^2 - 2x + 3)/y$.
+Note that the expression remains unevaluated: it is represented symbolically.
+
+\begin{verbatim}
+>>> (x**2 - 2*x + 3)/y
+(x**2 - 2*x + 3)/y
+\end{verbatim}
+
+Importantly, SymPy expressions are immutable. This simplifies the design of
+SymPy by allowing expression interning. It also enables expressions to be
+hashed, that is used to implement caching in SymPy.
+
+\subsection{List of Features}
 
@@
 
-\subsection{Basic Usage}
-
-% symbols, various ways to declare them
-
-Symbolic variables, called symbols, must be defined and assigned to
-Python variables before they can be used. This is typically done through the
-\texttt{symbols} function, which may create multiple symbols in a single
-function call. For instance,
-\begin{verbatim}
->>> x, y, z = symbols('x y z')
-\end{verbatim}
-creates three symbols representing variables named $x$, $y$, and $z$. In this
-particular instance, these symbols are all assigned to Python variables of the
-same name. However, the user is free to assign them to different
-Python variables, while representing the same symbol, such as
-\texttt{a, b, c = symbols(\textquotesingle{}x y z\textquotesingle{})}.
-In order to minimize potential confusion, though, all examples in this paper will
-assume that
-the symbols \verb|x|, \verb|y|, and \verb|z| have been assigned to Python variables
-identical to their symbolic names.
-
-Expressions are created from symbols using Python's mathematical syntax.  For
-instance, the following Python code creates the expression $(x^2 - 2x + 3)/y$.
-Note that the expression remains unevaluated: it is represented symbolically.
-
-\begin{verbatim}
->>> (x**2 - 2*x + 3)/y
-(x**2 - 2*x + 3)/y
-\end{verbatim}
-
-Importantly, SymPy expressions are immutable. This simplifies the design of
-SymPy by allowing expression interning. It also enables expressions to be
-hashed, that is used to implement caching in SymPy.
-
 %% Assumptions

### 2016-09-22 rebuttal.tex
Grammar
Remove spaces around em dashes

@@
 \question References reasonable, though [8], an excellent reference, makes almost
-precisely the opposite point to that for which it is cited --- ``simplification
+precisely the opposite point to that for which it is cited---``simplification
 is not well-defined''. I think the authors are trying to follow [8]'s
@@
 \begin{solution}
-We add an introduction.
+We have added an introduction.
 \end{solution}

### 2016-09-22 rebuttal.tex
Start addressing reviewer 3 rewrite comments

@@
 \begin{solution}
-  FIXME.
+  We have made the following changes to the paper
+
   \begin{enumerate}
-    \item I believe we shouldn't move ``Other Projects'' section to the
-      main paper content.  It either should stay in the supplement or we
-      should do a moderate cleanup for this section.  E.g. list only few
-      projects here and put other in the supplement.  Or/And show here
-      only project names.
+  \item We have renamed the ``Features'' section to ``Overview of
+    Capabilities''.
+    \item We have renamed the ``Domain Specific Submodules'' section to
+      ``Physics Submodule'' (see point~\ref{rev3point30}).
+  \item We have moved the ``Architecture'' section to the end, after the
+    ``Phusics Submodule'' section.
+  \item We have moved the ``Basic Usage'' subsection to the beginning of the
+    ``Overview of Capabilities'' section, and added an intro paragraph to the
+    ``Overview of Capabilities'' section.
+  \item We have moved the ``Assumptions'' subsection to the ``Overview of
+    Capabilities'' subsection, after the features table.
+  \end{enumerate}
 
-    \item Logical place for above section - at the end of the
-      paper, right before the "conclusion" section.
-
-    \item We should either drop the Mathematica section or discuss
-      other CAS too, which make this stuff suitable only as part
-      of the suppliment.  I prefer the first variant.
-  \end{enumerate}
+We disagree with the reviewer on a few points
 \end{solution}
@@
 \begin{solution}
+\label{rev3point30}
 This section was renamed to "Physics submodule".  (See also the

### 2016-09-22 rebuttal.tex
Move the "Comparison with Mathematica" section to the end of the supplement
Some more rebuttal on the rewrite

@@
 \begin{solution}
+\label{editorpoint1}
 This sentence was added to the introduction:
@@
     Capabilities'' subsection, after the features table.
+  \item The ``Comparison with Mathematica'' section has been moved to the end of
+    the supplement.
   \end{enumerate}
 
-We disagree with the reviewer on a few points
+We disagree with the reviewer on a few points. We do not consider SymPy Live
+or SymPy Gamma to be key components of SymPy. The paper goal of the paper is
+to primarily be about the SymPy library. While SymPy Live and SymPy Gamma are
+maintained by the SymPy community, we consider them to be separate projects. A
+mention of SymPy Live has been added to the introduction (see editor
+point~\ref{editorpoint1}), as it may assist an interested reader in trying the
+examples from the paper.
+
+The ``Other Projects that
+Depend on SymPy'' section (which has been renamed from ``Other Projects that
+Use SymPy'') is {\color{red}TODO: Add reasons here, or else move this to the
+paper as the reviewer requests.}
+
+The ``Comparison with Mathematica'' section has been placed in the supplement
+because we do not feel that a treatment in the main manuscript would be fair
+without an equal comparison to other principle computer algebra systems, such
+as Maple, SageMath, and Maxima. Unfortunately, our authorship is inexpert in
+these systems, so a complete and fair comparison is impossible.
+
 \end{solution}

### 2016-09-22 solvers.tex
Add a note about sympy.sets and the supplement to the solvers section

@@
 \end{verbatim}
-The \texttt{domain} parameter is typically either \texttt{S.Complexes} (the
+The \texttt{domain} parameter can be any set from the \texttt{sympy.sets}
+module (see the supplementary material for details on \texttt{sympy.sets}),
+but is typically either \texttt{S.Complexes} (the
 default) or \texttt{S.Reals}; the latter causes \texttt{solveset} to only return real solutions.

### 2016-09-22 calculus.tex
Add a bit about series and the supplement to the calculus section

SymPy provides all the basic operations of calculus, such as calculating
limits, derivatives, integrals, or summations.

Limits are computed with the \verb|limit| function, using the Gruntz
algorithm~\cite{Gruntz1996limits} for computing symbolic limits and heuristics (a description of the Gruntz algorithm may be found in the supplement).
For example, the following computes
$\lim\limits_{x\to \infty} x\sin(\frac{1}{x})=1$. Note that SymPy denotes
$\infty$ as \verb|oo|.
\begin{verbatim}
>>> limit(x*sin(1/x), x, oo)
1
\end{verbatim}
As a more complex example, SymPy computes \[\lim\limits_{x\to 0}{\left(2 e^{\frac{1 - \cos{\left (x \right )}}{\sin{\left (x \right )}}} -
  1\right)}^{\frac{\sinh{\left (x \right )}}{\operatorname{atan}^{2}{\left (x
      \right )}}} = e.\]
\begin{verbatim}
>>> limit((2*E**((1-cos(x))/sin(x))-1)**(sinh(x)/atan(x)**2), x, 0)
E
\end{verbatim}

Derivatives are computed with the \verb|diff| function, which recursively uses
the various differentiation rules.
\begin{verbatim}
>>> diff(sin(x)*exp(x), x)
exp(x)*sin(x) + exp(x)*cos(x)
\end{verbatim}

Integrals are calculated with the \verb|integrate| function. SymPy
implements a combination of the Risch
algorithm~\cite{bronstein2005integration}, table lookups, a reimplementation
of Manuel Bronstein's ``Poor Man's Integrator''~\cite{Bronstein2005pmint}, and
an algorithm for computing integrals based on Meijer G-functions~\cite{Roach1996hyper,roach1997meijerg}. These allow
SymPy to compute a wide variety of indefinite and definite integrals. The
Meijer G-function algorithm and the Risch algorithm are respectively
demonstrated below by the computation of \[\int_{0}^{\infty} e^{-s t}\log{\left (t \right )}\, dt = - \frac{ \log{\left (s \right )} + \gamma}{s}\] and \[\int \frac{- 2 x^{2} \left(\log{\left (x \right )} + 1\right) e^{x^{2}} + {\left(e^{x^{2}} + 1\right)}^{2}}{x {\left(e^{x^{2}} + 1\right)}^{2} \left(\log{\left (x \right )} + 1\right)}\, dx = \log{\left (\log{\left (x \right )} + 1 \right )} + \frac{1}{e^{x^{2}} + 1}.\]
\begin{verbatim}
>>> s, t = symbols('s t', positive=True)
>>> integrate(exp(-s*t)*log(t), (t, 0, oo)).simplify()
-(log(s) + EulerGamma)/s
>>> integrate((-2*x**2*(log(x) + 1)*exp(x**2) +
... (exp(x**2) + 1)**2)/(x*(exp(x**2) + 1)**2*(log(x) + 1)), x)
log(log(x) + 1) + 1/(exp(x**2) + 1)
\end{verbatim}

Summations are computed with \verb|summation|  using a combination of Gosper's
algorithm~\cite{gosper1978decision}, an algorithm that uses Meijer
G-functions~\cite{Roach1996hyper,roach1997meijerg}, and heuristics. Products
are computed with \verb|product| function via a suite of heuristics.
% TODO: Are there other summation algorithms implemented?
% TODO: A good summation example or two
\begin{verbatim}
>>> i, n = symbols('i n')
>>> summation(2**i, (i, 0, n - 1))
2**n - 1
>>> summation(i*factorial(i), (i, 1, n))
n*factorial(n) + factorial(n) - 1
\end{verbatim}

Series expansions are computed with the \verb|series| function. This example computes the power series of $\sin(x)$ around $x=0$ up to $x^6$.
\begin{verbatim}
>>> series(sin(x), x, 0, 6)
x - x**3/6 + x**5/120 + O(x**6)
\end{verbatim}
The supplementary material discusses series expansions methods in more depth.

Integrals, derivatives, summations, products, and limits that cannot be
computed return unevaluated objects. These can also be created directly if the
user chooses.
\begin{verbatim}
>>> integrate(x**x, x)
Integral(x**x, x)
>>> Sum(2**i, (i, 0, n - 1))
Sum(2**i, (i, 0, n - 1))
\end{verbatim}

### 2016-09-22 solvers.tex
Add reference to Diophantine supplement

@@
 SymPy has equation solvers that can handle ordinary differential
-equations, recurrence relationships, Diophantine equations, and algebraic
+equations, recurrence relationships, Diophantine equations\footnote{See the
+  supplementary material for an in depth discussion on the Diophantine submodule.}, and algebraic
 equations. There is also

### 2016-09-22 rebuttal.tex
More rewrite rebuttal

@@
 these systems, so a complete and fair comparison is impossible.
+
+We are unclear why the reviewer has called out specific sections over others
+for his suggested reorganization. We deep all the sections in section 3 (now
+section 2) to be important core components of the library. Regarding the supplement
+sections (previously) 5, 7, and 9---``Sets'', ``Category Theory'', and ``SymPy
+Live'', respesctively---only set sets module is potentially applicable to the
+wider audience of the main paper. However, in SymPy 1.0, the sets module is
+still in a relatively infant stage. It is mentioned already in conjunction
+with the \texttt{solveset} function. We have added some additional text to the
+solvers subsection regarding \texttt{sympy.sets}. The category theory module
+in SymPy is very domain specific, and independent of other SymPy modules. We
+consider the module too niche to warrant expanding upon in the main
+manuscript. SymPy Live, as has been discussed above, is not a SymPy submodule.
+It is a separate project from the SymPy library. It would be inappropriate to
+list it alongside the features of SymPy.
+
+We also consider the supplement sections (previously) 2, 3, 4, and
+6---''Series'', ``Logic'', ``Diophantine Equations'', and ``Statistics'',
+respectively---to be in depth beyond the goals of the main manuscript. The
+``Series'' section discusses relatively advanced series expansion methods. We
+have added a paragraph detailing basic series expansion to the ``Calculus''
+subsection, with a reference to the more detailed section in the supplement.
+The ``Logic'' section is alluded to in the ``Assumptions'' subsection, but is
+not discussed in depth in the main manuscript. While the logic submodule is
+indeed important for certain parts of SymPy, particularly the assumptions, we
+do not consider it to be of interest to the general reader. Furthermore, on a
+technical note, the assumptions discussed in the manuscript are the so-called
+``old assumptions'', and do not use the logic module discussed in the
+supplement. The so-called ``new assumptions'', mentioned briefly in the future
+work section, does use it, but we have opted to not discuss this module in
+depth, as it is still in development and not fully usable.
+
+The ``Solvers'' subsection already mentions that SymPy can solve Diophantine
+equations. We consider this specific application to be niche enough to keep
+the discussion in the supplement. We have added a reference to the supplement
+to the ``Solvers'' subsection. The ``Statistics'' module is an independent
+submodule (no other SymPy submodules depend on it). We do not consider it to
+be a core module of SymPy.
 

### 2016-09-22 rebuttal.tex
Architecture rebuttal

@@
 
+We feel the architecture is one of the most important parts of the paper. The
+goal of the paper is to discuss the architecture core features of SymPy. We
+therefore strongly feel that this section belongs in the main manuscript, and
+not the supplementary material. We have followed the reviewer's advice and
+moved the architecture section to the end of the paper, to make the paper
+easier to follow.
 \end{solution}

### 2016-09-23 other_projects_that_use_sympy.tex
Remove some of the "other projects that use SymPy"

@@
   Experimental Python package for teaching linear circuit analysis.
-\item
-  \href{http://digitalcommons.calpoly.edu/cgi/viewcontent.cgi?article=1072\&context=physsp/}{\textbf{Quantum
-  Programming in Python}}: Quantum 1D Simple Harmonic Oscillator and
-  Quantum Mapping Gate.
-\item
-  \href{http://mech.fsv.cvut.cz/~stransky/software/latexexpr/doc/}{\textbf{LaTeX
-  Expression project}}: Easy \LaTeX{} typesetting of algebraic expressions
-  in symbolic form with automatic substitution and result computation.
-\item
-  \href{https://www.researchgate.net/publication/260585491_Symbolic_Statistics_with_SymPy/}{\textbf{Symbolic
-  statistical modeling}}: Adding statistical operations to complex
-  physical models.
 \end{itemize}

### 2016-09-23 rebuttal.tex
Skip failing rebuttal "doctest"
Spelling
Be more forceful

@@
 
+% This depends on matplotlib. We want to make sure the rest of the paper
+% doesn't depend on it, so skip.
+% no-doctest
 \begin{verbatim}
@@
   \item We have moved the ``Architecture'' section to the end, after the
-    ``Phusics Submodule'' section.
+    ``Physics Submodule'' section.
   \item We have moved the ``Basic Usage'' subsection to the beginning of the
@@
 or SymPy Gamma to be key components of SymPy. The paper goal of the paper is
-to primarily be about the SymPy library. While SymPy Live and SymPy Gamma are
-maintained by the SymPy community, we consider them to be separate projects. A
+to be about the SymPy library. While SymPy Live and SymPy Gamma are
+maintained by the SymPy community, they are separate projects. A
 mention of SymPy Live has been added to the introduction (see editor

### 2016-09-23 rebuttal.tex
Remove sentence

@@
 
-We are unclear why the reviewer has called out specific sections over others
-for his suggested reorganization. We deep all the sections in section 3 (now
-section 2) to be important core components of the library. Regarding the supplement
-sections (previously) 5, 7, and 9---``Sets'', ``Category Theory'', and ``SymPy
-Live'', respesctively---only set sets module is potentially applicable to the
-wider audience of the main paper. However, in SymPy 1.0, the sets module is
-still in a relatively infant stage. It is mentioned already in conjunction
-with the \texttt{solveset} function. We have added some additional text to the
-solvers subsection regarding \texttt{sympy.sets}. The category theory module
-in SymPy is very domain specific, and independent of other SymPy modules. We
-consider the module too niche to warrant expanding upon in the main
-manuscript. SymPy Live, as has been discussed above, is not a SymPy submodule.
-It is a separate project from the SymPy library. It would be inappropriate to
-list it alongside the features of SymPy.
+We deem all the sections in section 3 (now section 2) to be important core
+components of the library. Regarding the supplement sections (previously) 5,
+7, and 9---``Sets'', ``Category Theory'', and ``SymPy Live'',
+respectively---only set sets module is potentially applicable to the wider
+audience of the main paper. However, in SymPy 1.0, the sets module is still in
+a relatively infant stage. It is mentioned already in conjunction with the
+\texttt{solveset} function. We have added some additional text to the solvers
+subsection regarding \texttt{sympy.sets}. The category theory module in SymPy
+is very domain specific, and independent of other SymPy modules. We consider
+the module too niche to warrant expanding upon in the main manuscript. SymPy
+Live, as has been discussed above, is not a SymPy submodule. It is a separate
+project from the SymPy library. It would be inappropriate to list it alongside
+the features of SymPy.
 
@@
 \begin{solution}
-We are unclear how the reviewer is making distinctions between ``modules'',
-``submodules'', and ``packages''. Our use coincides with the definitions used
-in the official Python documentation (see
+  Our use of ``modules'', ``submodules'', and ``packages'' coincides with the
+  definitions used in the official Python documentation (see
   \url{https://docs.python.org/3/tutorial/modules.html#packages} and

### 2016-09-23 rebuttal.tex
Remove a sentence
Remove sentence

@@
 \begin{solution}
-  It's unclear what you are asking for. Both Python standard library and SymPy
-  have exp callable, so it depends which symbols are imported in your global
-  namespace: the \texttt{exp} function from the \texttt{math} module or the
-  \texttt{exp} class from SymPy. In the later case a special symbolic constant
-  will be produced.
+  Both Python standard library and SymPy have exp callable, so it depends
+  which symbols are imported in your global namespace: the \texttt{exp}
+  function from the \texttt{math} module or the \texttt{exp} class from SymPy.
+  In the later case a special symbolic constant will be produced.
 
@@
 examples from the paper.
+
+% TODO For projects that use SymPy
+% TODO: Make intro sentence longer.
+% TODO: Make it into a table
+% TODO: Project name, domain, description
 
@@
   \url{https://docs.python.org/3/tutorial/modules.html#packages} and
-  \url{https://docs.python.org/3/glossary.html#term-package}). The distinction
-  between (sub)module and (sub)package in terms of Python libraries seems very
-  technical. To clarify things, we have modified the manuscript to use
-  ``package'' only when referring to a library external to SymPy (such as
-  mpmath and Xy-pic).
+  \url{https://docs.python.org/3/glossary.html#term-package}). To clarify
+  things, we have modified the manuscript to use ``package'' only when
+  referring to a library external to SymPy (such as mpmath and Xy-pic).
 

### 2016-09-23 rebuttal.tex
Grammar

@@
 7, and 9---``Sets'', ``Category Theory'', and ``SymPy Live'',
-respectively---only set sets module is potentially applicable to the wider
+respectively---only the sets module is potentially applicable to the wider
 audience of the main paper. However, in SymPy 1.0, the sets module is still in

### 2016-09-23 rebuttal.tex
Split something into an itemize

@@
 7, and 9---``Sets'', ``Category Theory'', and ``SymPy Live'',
-respectively---only the sets module is potentially applicable to the wider
-audience of the main paper. However, in SymPy 1.0, the sets module is still in
+respectively:
+\begin{itemize}
+\item In SymPy 1.0, the sets module is still in
 a relatively infant stage. It is mentioned already in conjunction with the
 \texttt{solveset} function. We have added some additional text to the solvers
-subsection regarding \texttt{sympy.sets}. The category theory module in SymPy
+subsection regarding \texttt{sympy.sets}.
+
+\item The category theory module in SymPy
 is very domain specific, and independent of other SymPy modules. We consider
-the module too niche to warrant expanding upon in the main manuscript. SymPy
+the module too niche to warrant expanding upon in the main manuscript.
+
+\item SymPy
 Live, as has been discussed above, is not a SymPy submodule. It is a separate
@@
 the features of SymPy.
+\end{itemize}
 

### 2016-09-23 rebuttal.tex
Use the right quotes

@@
 We also consider the supplement sections (previously) 2, 3, 4, and
-6---''Series'', ``Logic'', ``Diophantine Equations'', and ``Statistics'',
+6---``Series'', ``Logic'', ``Diophantine Equations'', and ``Statistics'',
 respectively---to be in depth beyond the goals of the main manuscript. The

### 2016-09-23 rebuttal.tex
Itemize some text

@@
 6---``Series'', ``Logic'', ``Diophantine Equations'', and ``Statistics'',
-respectively---to be in depth beyond the goals of the main manuscript. The
-``Series'' section discusses relatively advanced series expansion methods. We
+respectively---to be in depth beyond the goals of the main manuscript.
+
+\begin{itemize}
+\item The ``Series'' section discusses relatively advanced series expansion methods. We
 have added a paragraph detailing basic series expansion to the ``Calculus''
 subsection, with a reference to the more detailed section in the supplement.
-The ``Logic'' section is alluded to in the ``Assumptions'' subsection, but is
+
+\item The ``Logic'' section is alluded to in the ``Assumptions'' subsection, but is
 not discussed in depth in the main manuscript. While the logic submodule is
@@
 
-The ``Solvers'' subsection already mentions that SymPy can solve Diophantine
+\item The ``Solvers'' subsection already mentions that SymPy can solve Diophantine
 equations. We consider this specific application to be niche enough to keep
 the discussion in the supplement. We have added a reference to the supplement
-to the ``Solvers'' subsection. The ``Statistics'' module is an independent
+to the ``Solvers'' subsection.
+
+\item The ``Statistics'' module is an independent
 submodule (no other SymPy submodules depend on it). We do not consider it to
 be a core module of SymPy.
+\end{itemize}
 

### 2016-09-23 rebuttal.tex
Spell out the submodule name

@@
 technical note, the assumptions discussed in the manuscript are the so-called
-``old assumptions'', and do not use the logic module discussed in the
+``old assumptions'', and do not use the \texttt{sympy.logic} submodule discussed in the
 supplement. The so-called ``new assumptions'', mentioned briefly in the future

### 2016-09-23 rebuttal.tex
Spell out section name

@@
 ``old assumptions'', and do not use the \texttt{sympy.logic} submodule discussed in the
-supplement. The so-called ``new assumptions'', mentioned briefly in the future
-work section, does use it, but we have opted to not discuss this module in
+supplement. The so-called ``new assumptions'', mentioned briefly in the
+``Conclusion and Future
+Work'' section, does use it, but we have opted to not discuss this module in
 depth, as it is still in development and not fully usable.

### 2016-09-23 rebuttal.tex
Todos
Make the items more parallel
Grammar
Reword

@@
 
+% TODO: Note that old assumptions do not use sympy.logic
+% TODO: Reference the RETE algorithm
 \item The ``Logic'' section is alluded to in the ``Assumptions'' subsection, but is
@@
 
-\item The ``Solvers'' subsection already mentions that SymPy can solve Diophantine
-equations. We consider this specific application to be niche enough to keep
-the discussion in the supplement. We have added a reference to the supplement
-to the ``Solvers'' subsection.
+\item Regarding the ``Diophantine Solvers'' section, the ``Solvers''
+  subsection already mentions that SymPy can solve Diophantine Equations. This
+  is outside of the scope of the paper, because it's too specific. We have
+  added a reference to the supplement to the ``Solvers'' subsection.
 

### 2016-09-23 rebuttal.tex
Reword

@@
 \item The category theory module in SymPy
-is very domain specific, and independent of other SymPy modules. We consider
-the module too niche to warrant expanding upon in the main manuscript.
+is very domain specific, and independent of other SymPy modules. It is out of
+scope for an architecture paper.
 
@@
   subsection already mentions that SymPy can solve Diophantine Equations. This
-  is outside of the scope of the paper, because it's too specific. We have
+  is outside of the scope of the paper, because it's specific, and mostly
+  independent from other parts of SymPy. We have
   added a reference to the supplement to the ``Solvers'' subsection.

### 2016-09-23 rebuttal.tex
Some text

@@
 \item Regarding the ``Diophantine Solvers'' section, the ``Solvers''
-  subsection already mentions that SymPy can solve Diophantine Equations. This
-  is outside of the scope of the paper, because it's specific, and mostly
-  independent from other parts of SymPy. We have
+  subsection already mentions that SymPy can solve Diophantine Equations. The
+  in depth discussion in the supplement
+  is outside of the scope of the main paper, because it's specific, and
+  independent from parts of SymPy (with the exception of computing integer set
+  intersections in the \texttt{sympy.sets} submodule, no other part of SymPy
+  from solvers uses the Diophantine solvers). We have
   added a reference to the supplement to the ``Solvers'' subsection.

### 2016-09-23 rebuttal.tex
Missing word

--- 2016-09-23 rebuttal.tex
@@
 \item Regarding the ``Diophantine Solvers'' section, the ``Solvers''
-  subsection already mentions that SymPy can solve Diophantine Equations. This
-  is outside of the scope of the paper, because it's specific, and mostly
-  independent from other parts of SymPy. We have
+  subsection already mentions that SymPy can solve Diophantine Equations. The
+  in depth discussion in the supplement
+  is outside of the scope of the main paper, because it's specific, and
+  independent from other parts of SymPy (with the exception of computing integer set
+  intersections in the \texttt{sympy.sets} submodule, no other part of SymPy
+  from solvers uses the Diophantine solvers). We have
   added a reference to the supplement to the ``Solvers'' subsection.

### 2016-09-23 rebuttal.tex
Remove a sentence
Some words
Remove some words

@@
 
-\item The ``Statistics'' module is an independent
-submodule (no other SymPy submodules depend on it). We do not consider it to
-be a core module of SymPy.
+\item The ``Statistics'' module is not a core module and is independent (no
+  other SymPy submodules depend on it).
 \end{itemize}
 
-We feel the architecture is one of the most important parts of the paper. The
-goal of the paper is to discuss the architecture core features of SymPy. We
-therefore strongly feel that this section belongs in the main manuscript, and
-not the supplementary material. We have followed the reviewer's advice and
-moved the architecture section to the end of the paper, to make the paper
-easier to follow.
+The architecture is one of the most important parts of the paper. The goal of
+the paper is to discuss the architecture core features of SymPy. We therefore
+strongly feel that this section belongs in the main manuscript, and not the
+supplementary material. We have followed the reviewer's advice and moved the
+architecture section to the end of the paper, to make the paper easier to
+follow.
 \end{solution}

### 2016-09-23 rebuttal.tex
Change some words
Some wording

@@
 
-The architecture is one of the most important parts of the paper. The goal of
-the paper is to discuss the architecture core features of SymPy. We therefore
-strongly feel that this section belongs in the main manuscript, and not the
-supplementary material. We have followed the reviewer's advice and moved the
-architecture section to the end of the paper, to make the paper easier to
-follow.
+The ``Architecture'' section is a key component of the paper. The goal of the
+paper is to discuss the architecture and core features of SymPy. It therefore
+would be inappropriate to move this section to the supplement. We have
+followed the reviewer's advice and moved the architecture section to the end
+of the paper, to make the paper easier to follow.
 \end{solution}

### 2016-09-23 rebuttal.tex
Add name of manuscript to letter

@@
 
-We thank you and the referees for your most helpful comments and suggestions.
+We thank you and the referees for your most helpful comments and suggestions
+on our manuscript, ``SymPy: Symbolic Computing in Python''.
 The manuscript is greatly improved as a result.
 
-Below we give your and reviewers' comments on white background and our
+Below we give your and the reviewers' comments on white background and our
 replies on gray background. \bigskip

### 2016-09-23 rebuttal.tex
Some more of the letter

@@
 
-We thank you and the referees for your most helpful comments and suggestions
-on our manuscript, ``SymPy: Symbolic Computing in Python''.
-The manuscript is greatly improved as a result.
+We thank you and the referees for your thoughtful comments and suggestions on
+our manuscript, ``SymPy: Symbolic Computing in Python''. The manuscript is
+greatly improved as a result. We have endeavored to address your comments and
+concerns point by point below. Below we give your and the reviewers' comments
+on white background and our replies on gray background. \bigskip
 
-Below we give your and the reviewers' comments on white background and our
-replies on gray background. \bigskip
+Reviewer 1 had concerns that the paper was out of scope, but the editor has
+confirmed that it is in scope. Many of the points that reviewer 3 raised were
+in regards to the relevance of various sections. The intended scope of this
+paper was to be an architecture paper for SymPy's 1.0 release. We have
+examined reviewer 3's points through this lens. In light of this, which
+contents were previously in the supplement vs.\ the main manuscript was
+confusing. To address this, we have made some restructurings of the
+manuscript, to make the paper more clear to a general audience.
 

### 2016-09-23 rebuttal.tex
Delete sentence
Add missing vertical space
Fix some phrasing
Phrasing

@@
 We thank you and the referees for your thoughtful comments and suggestions on
-our manuscript, ``SymPy: Symbolic Computing in Python''. The manuscript is
-greatly improved as a result. We have endeavored to address your comments and
-concerns point by point below. Below we give your and the reviewers' comments
-on white background and our replies on gray background. \bigskip
+our manuscript, ``SymPy: Symbolic Computing in Python''. We have endeavored to
+address your comments and concerns point by point below. Below we give your
+and the reviewers' comments on white background and our replies on gray
+background. \bigskip
 
@@
 examined reviewer 3's points through this lens. In light of this, which
-contents were previously in the supplement vs.\ the main manuscript was
-confusing. To address this, we have made some restructurings of the
-manuscript, to make the paper more clear to a general audience.
+contents were previously in the supplement vs.\ the main manuscript may have
+been confusing to a general audience. To address this, we have restructured
+the manuscript to make the paper more readable to those not already experts in
+SymPy or symbolic mathematics. \bigskip
 

### 2016-09-23 other_projects_that_use_sympy.tex
Add SymPy Live and SymPy Gamma to the projects that depend on SymPy section

@@
 \begin{itemize}
+\item SymPy Live: SymPy Live an online Python shell, which uses the Google
+App Engine to executes SymPy code. It is integrated in the SymPy
+documentation examples at \href{http://docs.sympy.org}{http://docs.sympy.org}.
+SymPy Live is maintained by the SymPy community.
+\item SymPy Gamma: SymPy Gamma is a web application that executes and displays
+  results for SymPy expressions, in a fashon similar to that of Wolfram|Alpha.
+  SymPy Gamma is maintained by the SymPy community.
+  See the supplementary material for more information about SymPy Gamma.
 \item

### 2016-09-23 test-paper.py
Rename other_projects_that_use_sympy.tex to projects_that_depend_on_sympy.tex

@@ -26,7 +26,7 @@ files = ["introduction.tex",
          "gamma.tex",
          "live.tex",
          "comparison_with_mma.tex",
-         "other_projects_that_use_sympy.tex",
+         "projects_that_depend_on_sympy.tex",
          "tensors.tex",
          "polys.tex"
         ]

### 2016-09-23 test-paper.py
Remove references to live.tex

@@ -24,7 +24,6 @@ files = ["introduction.tex",
          "sets.tex",
          "stats.tex",
          "gamma.tex",
-         "live.tex",
          "comparison_with_mma.tex",
          "projects_that_depend_on_sympy.tex",
          "tensors.tex",

### 2016-09-23 test-paper.py
Fix order of files in tests and add missing files

@@ -5,29 +5,33 @@ from sympy import pprint_use_unicode, init_printing
 import __future__
 import sys
 
-files = ["introduction.tex",
-         "architecture.tex",
-         "numerics.tex",
-         "features.tex",
-         "simplification.tex",
-         "calculus.tex",
-         "printers.tex",
-         "solvers.tex",
-         "matrices.tex",
-         "domain_specific.tex",
+files = [
+    "introduction.tex",
+    "features.tex",
+    "simplification.tex",
+    "calculus.tex",
+    "polys.tex",
+    "printers.tex",
+    "solvers.tex",
+    "matrices.tex",
+    "numerics.tex",
+    "domain_specific.tex",
+    "architecture.tex",
+    "projects_that_depend_on_sympy.tex",
 
-         "supplement.tex",
-         "gruntz.tex",
-         "series.tex",
-         "logic.tex",
-         "diophantine.tex",
-         "sets.tex",
-         "stats.tex",
-         "gamma.tex",
-         "comparison_with_mma.tex",
-         "projects_that_depend_on_sympy.tex",
-         "tensors.tex",
-         "polys.tex"
+    "supplement.tex",
+    "gruntz.tex",
+    "series.tex",
+    "logic.tex",
+    "diophantine.tex",
+    "sets.tex",
+    "stats.tex",
+    "categories.tex",
+    "tensors.tex",
+    "nsimplify.tex",
+    "examples.tex",
+    "gamma.tex",
+    "comparison_with_mma.tex",
         ]
 
 output_file = "test_full_paper.py"

### 2016-09-23 comparison_with_mma.tex
Use CAS's (reviewer 3 comment 3)
Fix failing Matrix doctest

@@
 \begin{verbatim}
->>> exp(Matrix([[1, 1],[0, 2]])) * Matrix([a, b])
+>>> a, b = symbols('a b')
+>>> exp(Matrix([[1, 1], [0, 2]])) * Matrix([a, b])
 Matrix([
@@
 SymPy's use of floating-point numbers is similar to that of most
-other CASes, including Maple and Maxima.
+other CAS's, including Maple and Maxima.
 By contrast, Mathematica uses a form

### 2016-09-23 assumptions.tex
Move assumptions and basic usage into their own files

--- 2016-09-12 features.tex
@@
-% Features to discuss in-depth:
-
-Although SymPy's extensive feature set cannot be covered in-depth in this
-paper, calculus and other bedrock areas are discussed in their own
-subsections. Additionally, Table~\ref{features-table} gives a compact listing
-of all major capabilities present in the SymPy codebase. This grants a
-sampling from the breadth of topics and application domains that SymPy
-services. Unless stated otherwise, all features noted in
-Table~\ref{features-table} are symbolic in nature. Numeric features are
-discussed in Section~\ref{sec:numerics}.
-
-\begin{longtable}[htbc]{>{\raggedright}p{0.30\linewidth}p{0.63\linewidth}}
-\caption{SymPy Features and Descriptions\label{features-table}}\\
-\toprule
-\textbf{Feature (submodules)} & \textbf{Description} \\
-\midrule
-Calculus (\texttt{sympy.core}, \texttt{sympy.series}, \texttt{sympy.integrals}) & Algorithms for computing derivatives, integrals, and limits.\\
-
-Category Theory (\texttt{sympy.categories}) & Representation of objects, morphisms, and diagrams. Tools
-for drawing diagrams with Xy-pic.\\
-
-Code Generation (\texttt{sympy.printing}, \texttt{sympy.codegen}) & Generation of compilable and executable code in a
-variety of different programming languages from expressions directly. Target
-languages include C, Fortran, Julia, JavaScript,
-Mathematica, MATLAB and Octave, Python, and Theano.\\
-
-Combinatorics \& Group Theory (\texttt{sympy.combinatorics}) & Permutations, combinations,
-partitions, subsets, various permutation groups (such as polyhedral, Rubik,
-symmetric, and others), Gray codes~\cite{Nijenhuis1978combinatorial},
-and Prufer sequences~\cite{biggs1976graph}.\\
-
-Concrete Math (\texttt{sympy.concrete}) & Summation, products, tools for determining whether summation
-and product expressions are convergent, absolutely convergent, hypergeometric,
-and for determining other properties; computation of Gosper's normal form~\cite{petkovvsek1996bak} for two univariate polynomials.\\
-
-Cryptography (\texttt{sympy.crypto}) & Block and stream ciphers, including shift, Affine,
-substitution, Vigen\`{e}re's, Hill's, bifid, RSA, Kid RSA,
-linear-feedback shift registers, and Elgamal encryption.\\
-
-Differential Geometry (\texttt{sympy.diffgeom}) & Representations of manifolds, metrics, tensor
-products, and coordinate systems in Riemannian and pseudo-Riemannian
-geometries~\cite{FunctionalDifferentialGeometry}.\\
-% TODO: Someone verify that this is a good summary of the diffgeom module
-
-Geometry (\texttt{sympy.geometry}) & Representations of 2D geometrical entities, such as lines and
-circles. Enables queries on these entities, such as asking the area of an
-ellipse, checking for collinearity of a set of
-points, or finding the intersection between objects.\\
-
-Lie Algebras (\texttt{sympy.liealgebras}) & Representations of Lie algebras and root systems.\\
-
-Logic (\texttt{sympy.logic}) & Boolean expressions, equivalence testing, satisfiability, and normal
-forms.\\
-
-Matrices (\texttt{sympy.matrices}) & Tools for creating matrices of symbols and expressions.
-Both sparse and dense representations, as well as symbolic linear
-algebraic operations (e.g., inversion and factorization), are
-supported.\\
-
-Matrix Expressions (\texttt{sympy.matrices.expressions}) & Matrices with symbolic dimensions (unspecified entries).
-Block matrices.\\
-
-Number Theory (\texttt{sympy.ntheory}) & Prime number generation, primality testing, integer
-factorization, continued fractions, Egyptian fractions, modular arithmetic,
-quadratic residues, partitions, binomial and multinomial coefficients,
-prime number tools, hexidecimal digits of $\pi$, and integer factorization. \\
-
-Plotting (\texttt{sympy.plotting}) & Hooks for visualizing expressions via matplotlib~\cite{Hunter:2007}
-or as text drawings when lacking a graphical back-end. 2D function plotting,
-3D function
-plotting, and 2D implicit function plotting are supported.\\
-
-Polynomials (\texttt{sympy.polys}) & Polynomial algebras over various coefficient domains.
-Functionality ranges from simple operations (e.g., polynomial division) to
-advanced computations (e.g., Gr\"obner bases~\cite{adams1994introduction} and multivariate
-factorization over algebraic number domains).\\
-
-Printing (\texttt{sympy.printing}) & Functions for printing SymPy expressions in the terminal with ASCII
-or Unicode characters and converting SymPy expressions to \LaTeX{} and
-MathML.\\
-
-Quantum Mechanics (\texttt{sympy.physics.quantum}) & Quantum states, bra--ket notation, operators, basis sets,
-representations, tensor products, inner products, outer products, commutators,
-anticommutators, and specific quantum system implementations.\\
-
-Series (\texttt{sympy.series}) & Series expansion, sequences, and limits of sequences.
-This includes Taylor, Laurent, and Puiseux series as well as special series, such
-as Fourier and formal power series.\\
-
-Sets (\texttt{sympy.sets}) & Representations of empty, finite, and infinite sets (including
-special sets such as the natural, integer, and complex numbers). Operations
-on sets such as union, intersection, Cartesian product, and building sets from
-other sets are supported.\\
-
-Simplification (\texttt{sympy.simplify}) & Functions for manipulating and simplifying expressions.
-Includes algorithms for simplifying hypergeometric functions, trigonometric
-expressions, rational functions, combinatorial functions, square root
-denesting, and common subexpression elimination.\\
-
-Solvers (\texttt{sympy.solvers}) & Functions for symbolically solving equations, systems
-of equations, both linear and non-linear, inequalities, ordinary differential
-equations, partial differential equations, Diophantine equations, and
-recurrence relations.\\
-
-Special Functions (\texttt{sympy.functions}) & Implementations of a number of well known special functions,
-including Dirac delta, Gamma, Beta, Gauss error functions, Fresnel integrals,
-Exponential integrals, Logarithmic integrals, Trigonometric integrals, Bessel,
-Hankel, Airy, B-spline, Riemann Zeta, Dirichlet eta, polylogarithm, Lerch
-transcendent, hypergeometric, elliptic integrals, Mathieu, Jacobi polynomials,
-Gegenbauer polynomial, Chebyshev polynomial, Legendre polynomial, Hermite
-polynomial, Laguerre polynomial, and
-spherical harmonic functions.\\
-
-Statistics (\texttt{sympy.stats}) & Support for a random variable type as well as the ability to
-declare this variable from prebuilt distribution functions such as
-Normal, Exponential, Coin, Die, and other custom distributions~\cite{StatsMRocklin}.\\
-
-Tensors (\texttt{sympy.tensor}) & Symbolic manipulation of indexed objects.\\
-
-Vectors (\texttt{sympy.vector}) & Basic operations on vectors and differential calculus with respect
-to 3D Cartesian coordinate systems.\\
-\bottomrule
-
-\end{longtable}
-
-\subsection{Basic Usage}
-
-% symbols, various ways to declare them
-
-Symbolic variables, called symbols, must be defined and assigned to
-Python variables before they can be used. This is typically done through the
-\texttt{symbols} function, which may create multiple symbols in a single
-function call. For instance,
-\begin{verbatim}
->>> x, y, z = symbols('x y z')
-\end{verbatim}
-creates three symbols representing variables named $x$, $y$, and $z$. In this
-particular instance, these symbols are all assigned to Python variables of the
-same name. However, the user is free to assign them to different
-Python variables, while representing the same symbol, such as
-\texttt{a, b, c = symbols(\textquotesingle{}x y z\textquotesingle{})}.
-In order to minimize potential confusion, though, all examples in this paper will
-assume that
-the symbols \verb|x|, \verb|y|, and \verb|z| have been assigned to Python variables
-identical to their symbolic names.
-
-Expressions are created from symbols using Python's mathematical syntax.  For
-instance, the following Python code creates the expression $(x^2 - 2x + 3)/y$.
-Note that the expression remains unevaluated: it is represented symbolically.
-
-\begin{verbatim}
->>> (x**2 - 2*x + 3)/y
-(x**2 - 2*x + 3)/y
-\end{verbatim}
-
-Importantly, SymPy expressions are immutable. This simplifies the design of
-SymPy by allowing expression interning. It also enables expressions to be
-hashed, that is used to implement caching in SymPy.
-
-%% Assumptions
-\subsection{Assumptions}
-\label{sec:assumptions}
-
 SymPy performs logical inference through its assumptions system. The
@@
 compute the given fact. For instance, \texttt{(pi + E).is\_irrational} gives
-\texttt{None}, because determining whether $\pi + e$ is rational or irrational
+\texttt{None}---indeed, determining whether $\pi + e$ is rational or irrational
 is an open problem in mathematics~\cite{lang1966introduction}.
@@
 \texttt{(x - y).is\_positive} will be \texttt{None}.
-
-\subsection{Simplification}
-
-\input{simplification}
-
-\subsection{Calculus}
-\label{sec:calculus}
-\input{calculus}
-
-\subsection{Polynomials}
-\input{polys}
-
-\subsection{Printers}
-
-\input{printers.tex}
-
-% Solvers (regular equations, maybe also mention other types of solvers like ODEs/recurrence/Diophantine)
-\subsection{Solvers}
-\input{solvers.tex}
-
-% Matrices (worth emphasizing that they are symbolic)
-\subsection{Matrices}
-
-\input{matrices}

### 2016-09-23 basic_usage.tex
Move assumptions and basic usage into their own files

--- 2016-09-12 features.tex
@@
-% Features to discuss in-depth:
-
-Although SymPy's extensive feature set cannot be covered in-depth in this
-paper, calculus and other bedrock areas are discussed in their own
-subsections. Additionally, Table~\ref{features-table} gives a compact listing
-of all major capabilities present in the SymPy codebase. This grants a
-sampling from the breadth of topics and application domains that SymPy
-services. Unless stated otherwise, all features noted in
-Table~\ref{features-table} are symbolic in nature. Numeric features are
-discussed in Section~\ref{sec:numerics}.
-
-\begin{longtable}[htbc]{>{\raggedright}p{0.30\linewidth}p{0.63\linewidth}}
-\caption{SymPy Features and Descriptions\label{features-table}}\\
-\toprule
-\textbf{Feature (submodules)} & \textbf{Description} \\
-\midrule
-Calculus (\texttt{sympy.core}, \texttt{sympy.series}, \texttt{sympy.integrals}) & Algorithms for computing derivatives, integrals, and limits.\\
-
-Category Theory (\texttt{sympy.categories}) & Representation of objects, morphisms, and diagrams. Tools
-for drawing diagrams with Xy-pic.\\
-
-Code Generation (\texttt{sympy.printing}, \texttt{sympy.codegen}) & Generation of compilable and executable code in a
-variety of different programming languages from expressions directly. Target
-languages include C, Fortran, Julia, JavaScript,
-Mathematica, MATLAB and Octave, Python, and Theano.\\
-
-Combinatorics \& Group Theory (\texttt{sympy.combinatorics}) & Permutations, combinations,
-partitions, subsets, various permutation groups (such as polyhedral, Rubik,
-symmetric, and others), Gray codes~\cite{Nijenhuis1978combinatorial},
-and Prufer sequences~\cite{biggs1976graph}.\\
-
-Concrete Math (\texttt{sympy.concrete}) & Summation, products, tools for determining whether summation
-and product expressions are convergent, absolutely convergent, hypergeometric,
-and for determining other properties; computation of Gosper's normal form~\cite{petkovvsek1996bak} for two univariate polynomials.\\
-
-Cryptography (\texttt{sympy.crypto}) & Block and stream ciphers, including shift, Affine,
-substitution, Vigen\`{e}re's, Hill's, bifid, RSA, Kid RSA,
-linear-feedback shift registers, and Elgamal encryption.\\
-
-Differential Geometry (\texttt{sympy.diffgeom}) & Representations of manifolds, metrics, tensor
-products, and coordinate systems in Riemannian and pseudo-Riemannian
-geometries~\cite{FunctionalDifferentialGeometry}.\\
-% TODO: Someone verify that this is a good summary of the diffgeom module
-
-Geometry (\texttt{sympy.geometry}) & Representations of 2D geometrical entities, such as lines and
-circles. Enables queries on these entities, such as asking the area of an
-ellipse, checking for collinearity of a set of
-points, or finding the intersection between objects.\\
-
-Lie Algebras (\texttt{sympy.liealgebras}) & Representations of Lie algebras and root systems.\\
-
-Logic (\texttt{sympy.logic}) & Boolean expressions, equivalence testing, satisfiability, and normal
-forms.\\
-
-Matrices (\texttt{sympy.matrices}) & Tools for creating matrices of symbols and expressions.
-Both sparse and dense representations, as well as symbolic linear
-algebraic operations (e.g., inversion and factorization), are
-supported.\\
-
-Matrix Expressions (\texttt{sympy.matrices.expressions}) & Matrices with symbolic dimensions (unspecified entries).
-Block matrices.\\
-
-Number Theory (\texttt{sympy.ntheory}) & Prime number generation, primality testing, integer
-factorization, continued fractions, Egyptian fractions, modular arithmetic,
-quadratic residues, partitions, binomial and multinomial coefficients,
-prime number tools, hexidecimal digits of $\pi$, and integer factorization. \\
-
-Plotting (\texttt{sympy.plotting}) & Hooks for visualizing expressions via matplotlib~\cite{Hunter:2007}
-or as text drawings when lacking a graphical back-end. 2D function plotting,
-3D function
-plotting, and 2D implicit function plotting are supported.\\
-
-Polynomials (\texttt{sympy.polys}) & Polynomial algebras over various coefficient domains.
-Functionality ranges from simple operations (e.g., polynomial division) to
-advanced computations (e.g., Gr\"obner bases~\cite{adams1994introduction} and multivariate
-factorization over algebraic number domains).\\
-
-Printing (\texttt{sympy.printing}) & Functions for printing SymPy expressions in the terminal with ASCII
-or Unicode characters and converting SymPy expressions to \LaTeX{} and
-MathML.\\
-
-Quantum Mechanics (\texttt{sympy.physics.quantum}) & Quantum states, bra--ket notation, operators, basis sets,
-representations, tensor products, inner products, outer products, commutators,
-anticommutators, and specific quantum system implementations.\\
-
-Series (\texttt{sympy.series}) & Series expansion, sequences, and limits of sequences.
-This includes Taylor, Laurent, and Puiseux series as well as special series, such
-as Fourier and formal power series.\\
-
-Sets (\texttt{sympy.sets}) & Representations of empty, finite, and infinite sets (including
-special sets such as the natural, integer, and complex numbers). Operations
-on sets such as union, intersection, Cartesian product, and building sets from
-other sets are supported.\\
-
-Simplification (\texttt{sympy.simplify}) & Functions for manipulating and simplifying expressions.
-Includes algorithms for simplifying hypergeometric functions, trigonometric
-expressions, rational functions, combinatorial functions, square root
-denesting, and common subexpression elimination.\\
-
-Solvers (\texttt{sympy.solvers}) & Functions for symbolically solving equations, systems
-of equations, both linear and non-linear, inequalities, ordinary differential
-equations, partial differential equations, Diophantine equations, and
-recurrence relations.\\
-
-Special Functions (\texttt{sympy.functions}) & Implementations of a number of well known special functions,
-including Dirac delta, Gamma, Beta, Gauss error functions, Fresnel integrals,
-Exponential integrals, Logarithmic integrals, Trigonometric integrals, Bessel,
-Hankel, Airy, B-spline, Riemann Zeta, Dirichlet eta, polylogarithm, Lerch
-transcendent, hypergeometric, elliptic integrals, Mathieu, Jacobi polynomials,
-Gegenbauer polynomial, Chebyshev polynomial, Legendre polynomial, Hermite
-polynomial, Laguerre polynomial, and
-spherical harmonic functions.\\
-
-Statistics (\texttt{sympy.stats}) & Support for a random variable type as well as the ability to
-declare this variable from prebuilt distribution functions such as
-Normal, Exponential, Coin, Die, and other custom distributions~\cite{StatsMRocklin}.\\
-
-Tensors (\texttt{sympy.tensor}) & Symbolic manipulation of indexed objects.\\
-
-Vectors (\texttt{sympy.vector}) & Basic operations on vectors and differential calculus with respect
-to 3D Cartesian coordinate systems.\\
-\bottomrule
-
-\end{longtable}
-
-\subsection{Basic Usage}
-
 % symbols, various ways to declare them
@@
 hashed, that is used to implement caching in SymPy.
-
-\subsection{Simplification}
-
-\input{simplification}
-
-\subsection{Calculus}
-\label{sec:calculus}
-\input{calculus}
-
-\subsection{Polynomials}
-\input{polys}
-
-\subsection{Printers}
-
-\input{printers.tex}
-
-% Solvers (regular equations, maybe also mention other types of solvers like ODEs/recurrence/Diophantine)
-\subsection{Solvers}
-\input{solvers.tex}
-
-% Matrices (worth emphasizing that they are symbolic)
-\subsection{Matrices}
-
-\input{matrices}

### 2016-09-23 features.tex
I don't like this "because"
SymPy doesn't have a table of "open problems" or anything like that. The fact
just isn't implemented (there many are other facts that are known in
mathematics but aren't implemented in SymPy).
Move assumptions and basic usage into their own files

--- 2016-09-12 features.tex
@@
 % Features to discuss in-depth:
+This section gives a basic introduction of SymPy, and lists its features.
+A few features---assumptions, simplification, calculus, polynomials, printers,
+solvers, and matrices---are core components of SymPy and are discussed in
+depth. Many other features are discussed in depth in the supplementary
+material.
+
+\subsection{Basic Usage}
+\input{basic_usage.tex}
+
+\subsection{List of Features}
 
@@
 
-\subsection{Basic Usage}
-
-% symbols, various ways to declare them
-
-Symbolic variables, called symbols, must be defined and assigned to
-Python variables before they can be used. This is typically done through the
-\texttt{symbols} function, which may create multiple symbols in a single
-function call. For instance,
-\begin{verbatim}
->>> x, y, z = symbols('x y z')
-\end{verbatim}
-creates three symbols representing variables named $x$, $y$, and $z$. In this
-particular instance, these symbols are all assigned to Python variables of the
-same name. However, the user is free to assign them to different
-Python variables, while representing the same symbol, such as
-\texttt{a, b, c = symbols(\textquotesingle{}x y z\textquotesingle{})}.
-In order to minimize potential confusion, though, all examples in this paper will
-assume that
-the symbols \verb|x|, \verb|y|, and \verb|z| have been assigned to Python variables
-identical to their symbolic names.
-
-Expressions are created from symbols using Python's mathematical syntax.  For
-instance, the following Python code creates the expression $(x^2 - 2x + 3)/y$.
-Note that the expression remains unevaluated: it is represented symbolically.
-
-\begin{verbatim}
->>> (x**2 - 2*x + 3)/y
-(x**2 - 2*x + 3)/y
-\end{verbatim}
-
-Importantly, SymPy expressions are immutable. This simplifies the design of
-SymPy by allowing expression interning. It also enables expressions to be
-hashed, that is used to implement caching in SymPy.
+%% Assumptions
+\subsection{Assumptions}
+\label{sec:assumptions}
+\input{assumptions.tex}
 

### 2016-09-23 test-paper.py

@@ -8,6 +8,8 @@ import sys
 files = [
     "introduction.tex",
     "features.tex",
+    "basic_usage.tex",
+    "assumptions.tex",
     "simplification.tex",
     "calculus.tex",
     "polys.tex",

### 2016-09-23 stats.tex
Fix tests
We need to make sure that x, y, and z are never redefined with assumptions.

@@
 >>> from sympy.stats import Maxwell, density
->>> kT, m, x = symbols("kT m x", positive=True)
+>>> kT, m, t = symbols("kT m t", positive=True)
 >>> v = Maxwell("v", sqrt(kT/m))
@@
 3*kT/2
->>> solve(density(v)(x).diff(x), x)[0]  # most probable velocity
+>>> solve(density(v)(t).diff(t), t)[0]  # most probable velocity
 sqrt(2)*sqrt(kT)/sqrt(m)

### 2016-09-23 test-paper.py
Code quality

@@ -34,7 +34,7 @@ files = [
     "examples.tex",
     "gamma.tex",
     "comparison_with_mma.tex",
-        ]
+    ]
 
 output_file = "test_full_paper.py"
 

### 2016-09-23 assumptions.tex
Simplify some wording

@@
 
-\texttt{None} represents the
-``unknown'' case. This could mean that given assumptions do not unambiguously
-specify the truth of an attribute. For instance,
-\texttt{Symbol(\textquotesingle{}x\textquotesingle{}, real=True).is\_positive}
-will give \texttt{None} because a real symbol might be positive or negative.
-The \texttt{None} could also mean that not enough is known or implemented to
-compute the given fact. For instance, \texttt{(pi + E).is\_irrational} gives
-\texttt{None}---indeed, determining whether $\pi + e$ is rational or irrational
-is an open problem in mathematics~\cite{lang1966introduction}.
+\texttt{None} represents the ``unknown'' case. This could mean that given
+assumptions do not unambiguously specify the truth of an attribute. For
+instance, \texttt{Symbol(\textquotesingle{}x\textquotesingle{},
+  real=True).is\_positive} will give \texttt{None} because a real symbol might
+be positive or negative. The \texttt{None} could also mean that not enough is
+known or implemented to compute the given fact. For instance, \texttt{(pi +
+  E).is\_irrational} gives \texttt{None}---indeed, the rationality of
+$\pi + e$ is an open problem in mathematics~\cite{lang1966introduction}.
 

### 2016-09-25 rebuttal.tex
Final paragraph for letter
bigskip
Fix broken references

@@
 
+We thank you again for your consideration. Given these edits, we believe that
+the manuscript is now suitable for publication in PeerJ Computer Science. If
+there are further questions or concerns do not hesitate to contact us. \bigskip
+
 Best regards,\bigskip
 
-The authors
+Aaron Meurer
 
@@
 
-% TODO: Note that old assumptions do not use sympy.logic
-% TODO: Reference the RETE algorithm
-\item The ``Logic'' section is alluded to in the ``Assumptions'' subsection, but is
-not discussed in depth in the main manuscript. While the logic submodule is
-indeed important for certain parts of SymPy, particularly the assumptions, we
-do not consider it to be of interest to the general reader. Furthermore, on a
-technical note, the assumptions discussed in the manuscript are the so-called
-``old assumptions'', and do not use the \texttt{sympy.logic} submodule discussed in the
-supplement. The so-called ``new assumptions'', mentioned briefly in the
-``Conclusion and Future
-Work'' section, does use it, but we have opted to not discuss this module in
-depth, as it is still in development and not fully usable.
+\item The ``Logic'' section is alluded to in the ``Assumptions'' subsection,
+  but is not discussed in depth in the main manuscript. While the logic
+  submodule is indeed important for certain parts of SymPy, we do not consider
+  it to be of interest to the general reader. Furthermore, on a technical
+  note, the assumptions discussed in the manuscript are the so-called ``old
+  assumptions'', which do not use the \texttt{sympy.logic} submodule discussed
+  in the supplement (they use an implementation of the Rete algorithm, which
+  is separate from \texttt{sympy.logic}). The so-called ``new assumptions'',
+  mentioned briefly in the ``Conclusion and Future Work'' section, does use
+  it, but we have opted to not discuss this module in depth, as it is still in
+  development and not fully usable. We have added some text to the assumptions
+  section discussing this, and referencing the logic subsection of the
+  supplement.
 
@@
   designed to be compact, human-readable, and valid Python code that could be
-  used to recreate the expression. As it was noted in section~\ref{sec:core},
+  used to recreate the expression. As it was noted in section~5.1,
   the \texttt{srepr} function prints the exact, verbose form of an expression.

### 2016-09-25 supplement.tex
Move the "Comparison with Mathematica" section to the end of the supplement
Move projects that depend on SymPy list to the main paper
Remove the SymPy Live section from the supplement
We are just going to include it in the projects that use SymPy section. There
is not enough content here to warrant a separate supplement section.
Fix broken references

@@
 discusses numerical simplification.  Section \ref{suppsec:examples} provides
-additional examples for topics, discussed in the main paper.  In sections
-\ref{sympy-gamma} and \ref{sympy-live} the SymPy Gamma
+additional examples for topics, discussed in the main paper.  In section
+\ref{sympy-gamma} the SymPy Gamma
 and the SymPy Live projects are introduced.  Finally, section \ref{comp-mma}
 has a
-brief comparison of SymPy with Wolfram Mathematica, and
-section \ref{other-proj} lists some projects that depend on SymPy.
+brief comparison of SymPy with Wolfram Mathematica.
 
@@
 
-\section{SymPy Live}\label{sympy-live}
-
-\input{live}
-
 \section{Comparison with Mathematica}
@@
 
-\section{Other Projects that depend on SymPy}
-\label{other-proj}
-\input{other_projects_that_use_sympy}
-
 \bibliography{paper}

### 2016-09-26 projects_that_depend_on_sympy.tex
Add links for SymPy Live and Gamma

--- 2016-09-23 other_projects_that_use_sympy.tex
@@
 \begin{itemize}
+\item \href{http://live.sympy.org/}{\textbf{SymPy Live}}: SymPy Live an online
+  Python shell, which uses the Google App Engine to executes SymPy code. It is
+  integrated in the SymPy documentation examples at
+  \href{http://docs.sympy.org}{http://docs.sympy.org}. SymPy Live is
+  maintained by the SymPy community.
+\item \href{http://sympygamma.com/}{\textbf{SymPy Gamma}}: SymPy Gamma is a
+  web application that executes and displays results for SymPy expressions, in
+  a fashon similar to that of Wolfram|Alpha. SymPy Gamma is maintained by the
+  SymPy community. See the supplementary material for more information about
+  SymPy Gamma.
 \item

### 2016-09-26 projects_that_depend_on_sympy.tex
Add citation for Cadabra

@@
 \item
-  \href{http://cadabra.science/index.html}{\textbf{Cadabra}}: Cadabra is
-  a CAS designed specifically for the
-  resolution of problems encountered in field theory.
+  \href{http://cadabra.science/index.html}{\textbf{Cadabra}}~\cite{Peeters2007cadabra}:
+  Cadabra is a CAS designed specifically for the resolution of problems
+  encountered in field theory.
 \item

### 2016-09-26 projects_that_depend_on_sympy.tex
Fix name and URL for OctSymPy
Add an citation for OctSymPy
@cbm755 is this the right way to cite OctSymPy?

@@
 \item
-  \href{http://octave.sourceforge.net/symbolic/}{\textbf{Octave Symbolic}}:
-  The Octave-Forge Symbolic package adds symbolic calculation features
+  \href{https://github.com/cbm755/octsympy}{\textbf{OctSymPy}}~\cite{OctSymPy}:
+  The OctSymPy package adds symbolic calculation features
   to GNU Octave. These include common CAS tools such

### 2016-09-26 projects_that_depend_on_sympy.tex
Remove SageMathCloud from list of projects, since it somewhat duplicates SageMath
Add several citations for the other projects

@@
 \item
-  \href{https://github.com/jverzani/SymPy.jl}{\textbf{SymPy.jl}}:
+  \href{https://github.com/jverzani/SymPy.jl}{\textbf{SymPy.jl}}~\cite{SymPy.jl}:
   Provides a Julia interface to SymPy using PyCall.
 \item
-  \href{https://mathics.github.io/}{\textbf{Mathics}}: Mathics is a
+  \href{https://mathics.github.io/}{\textbf{Mathics}}~\cite{Mathics}: Mathics is a
   free, general-purpose online CAS featuring Mathematica compatible
   syntax and functions. It is backed by highly extensible Python code,
-  relying on SymPy for most mathematical tasks.
+  relying on SymPy for most mathematical tasks.[<8;32;30m
 \item
-  \href{http://mathpix.com/}{\textbf{Mathpix}}: An iOS App, that detects handwritten math as input, and uses
+  \href{http://mathpix.com/}{\textbf{Mathpix}}~\cite{Mathpix}: An iOS App, that detects handwritten math as input, and uses
   SymPy Gamma to evaluate the math input and generate the relevant
@@
 \item
-  \href{http://openrave.org/docs/0.8.2/openravepy/ikfast/}{\textbf{IKFast}}:
+  \href{http://openrave.org/docs/latest_stable/openravepy/ikfast/}{\textbf{IKFast}}~\cite{diankov2010ikfast}:
   IKFast is a robot kinematics compiler provided by
@@
 \item
-  \href{http://www.sagemath.org/}{\textbf{Sage}}: A CAS, visioned to be
+  \href{http://www.sagemath.org/}{\textbf{Sage}}~\cite{sagemath}: A CAS, visioned to be
   a viable free open source alternative to Magma, Maple, Mathematica and
@@
 \item
-  \href{https://cloud.sagemath.com}{\textbf{SageMathCloud}}:
-  SageMathCloud is a web-based cloud computing and course management
-  platform for computational mathematics.
-\item
-  \href{http://www.pydy.org/}{\textbf{PyDy}}: Multibody Dynamics with
+  \href{http://www.pydy.org/}{\textbf{PyDy}}~\cite{gede2013constrained}: Multibody Dynamics with
   Python.
 \item
-  \href{https://github.com/brombo/galgebra}{\textbf{galgebra}}:
+  \href{https://github.com/brombo/galgebra}{\textbf{galgebra}}~\cite{galgebra}:
   Geometric algebra (previously \texttt{sympy.galgebra}).

### 2016-09-26 projects_that_depend_on_sympy.tex
galgebra description
Add some more projects that use SymPy citations

@@
   \href{https://github.com/brombo/galgebra}{\textbf{galgebra}}~\cite{galgebra}:
-  Geometric algebra (previously \texttt{sympy.galgebra}).
+  A module for geometric algebra (previously \texttt{sympy.galgebra}).
 \item
-  \href{http://yt-project.org/}{\textbf{yt}}: Python package for
+  \href{http://yt-project.org/}{\textbf{yt}}~\cite{2011ApJS..192....9T}: Python package for
   analyzing and visualizing volumetric data (\texttt{yt.units} uses SymPy).
 \item
-  \href{http://sfepy.org/}{\textbf{SfePy}} (Simple finite elements in Python),
+  \href{http://sfepy.org/}{\textbf{SfePy}}~\cite{cimrman2014sfepy}: (Simple finite elements in Python),
   cf.~\cite{cimrman2014sfepy}, is a Python package for solving partial
@@
 \item
-  \href{http://quameon.sourceforge.net/}{\textbf{Quameon}}: Quantum
+  \href{http://quameon.sourceforge.net/}{\textbf{Quameon}}~\cite{quameon}: Quantum
   Monte Carlo in Python.
 \item
-  \href{http://lcapy.elec.canterbury.ac.nz/}{\textbf{Lcapy}}:
+  \href{http://lcapy.elec.canterbury.ac.nz/}{\textbf{Lcapy}}~\cite{lcapy}:
   Experimental Python package for teaching linear circuit analysis.

### 2016-09-26 projects_that_depend_on_sympy.tex
Convert other projects into a table
Still need to fill in the domain column

--- 2016-09-26 projects_that_depend_on_sympy.tex
@@
 There are several projects that depend on SymPy as a library for implementing
-a part of their functionality. Some of them are listed below:
+a part of their functionality. A selection of these projects are listed below:
 
-\begin{itemize}
-\item \href{http://live.sympy.org/}{\textbf{SymPy Live}}: SymPy Live an online
+\begin{longtable}[htbc]{>{\raggedright}p{0.14\linewidth}p{0.14\linewidth}p{0.63\linewidth}}
+\caption{Selected projects that depend on SymPy.\label{projects-table}}\\
+\toprule
+\textbf{Project name} & \textbf{Domain} & \textbf{Description} \\
+\midrule
+\href{http://live.sympy.org/}{\textbf{SymPy Live}} &  & SymPy Live an online
   Python shell, which uses the Google App Engine to executes SymPy code. It is
@@
   \href{http://docs.sympy.org}{http://docs.sympy.org}. SymPy Live is
-  maintained by the SymPy community.
-\item \href{http://sympygamma.com/}{\textbf{SymPy Gamma}}: SymPy Gamma is a
+  maintained by the SymPy community. \\
+
+\href{http://sympygamma.com/}{\textbf{SymPy Gamma}} &  & SymPy Gamma is a
   web application that executes and displays results for SymPy expressions, in
@@
   SymPy community. See the supplementary material for more information about
-  SymPy Gamma.
-\item
-  \href{http://cadabra.science/index.html}{\textbf{Cadabra}}~\cite{Peeters2007cadabra}:
+  SymPy Gamma. \\
+
+\href{http://cadabra.science/index.html}{\textbf{Cadabra}}~\cite{Peeters2007cadabra} &  &
   Cadabra is a CAS designed specifically for the resolution of problems
-  encountered in field theory.
-\item
-  \href{https://github.com/cbm755/octsympy}{\textbf{OctSymPy}}~\cite{OctSymPy}:
+  encountered in field theory. \\
+
+\href{https://github.com/cbm755/octsympy}{\textbf{OctSymPy}}~\cite{OctSymPy} &  &
   The OctSymPy package adds symbolic calculation features
@@
   as algebraic operations, calculus, equation solving, Fourier and
-  Laplace transforms, variable precision arithmetic, and other features.
-\item
-  \href{https://github.com/jverzani/SymPy.jl}{\textbf{SymPy.jl}}~\cite{SymPy.jl}:
-  Provides a Julia interface to SymPy using PyCall.
-\item
-  \href{https://mathics.github.io/}{\textbf{Mathics}}~\cite{Mathics}: Mathics is a
+  Laplace transforms, variable precision arithmetic, and other features. \\
+
+\href{https://github.com/jverzani/SymPy.jl}{\textbf{SymPy.jl}}~\cite{SymPy.jl} &  &
+  Provides a Julia interface to SymPy using PyCall. \\
+
+\href{https://mathics.github.io/}{\textbf{Mathics}}~\cite{Mathics} &  & Mathics is a
   free, general-purpose online CAS featuring Mathematica compatible
   syntax and functions. It is backed by highly extensible Python code,
-  relying on SymPy for most mathematical tasks.[<8;32;30m
-\item
-  \href{http://mathpix.com/}{\textbf{Mathpix}}~\cite{Mathpix}: An iOS App, that detects handwritten math as input, and uses
+  relying on SymPy for most mathematical tasks.[<8;32;30m \\
+
+\href{http://mathpix.com/}{\textbf{Mathpix}}~\cite{Mathpix} &  & An iOS App, that detects handwritten math as input, and uses
   SymPy Gamma to evaluate the math input and generate the relevant
-  steps to solve the problem.
-\item
-  \href{http://openrave.org/docs/latest_stable/openravepy/ikfast/}{\textbf{IKFast}}~\cite{diankov2010ikfast}:
+  steps to solve the problem. \\
+
+\href{http://openrave.org/docs/latest_stable/openravepy/ikfast/}{\textbf{IKFast}}~\cite{diankov2010ikfast} &  &
   IKFast is a robot kinematics compiler provided by
@@
   kinematics equations and generates optimized C++ files. It uses SymPy for
-  its internal symbolic mathematics.
-\item
-  \href{http://www.sagemath.org/}{\textbf{Sage}}~\cite{sagemath}: A CAS, visioned to be
+  its internal symbolic mathematics. \\
+
+\href{http://www.sagemath.org/}{\textbf{Sage}}~\cite{sagemath} &  & A CAS, visioned to be
   a viable free open source alternative to Magma, Maple, Mathematica and
   MATLAB\@. Sage includes many open source mathematical libraries, including
-  SymPy.
-\item
-  \href{http://www.pydy.org/}{\textbf{PyDy}}~\cite{gede2013constrained}: Multibody Dynamics with
-  Python.
-\item
-  \href{https://github.com/brombo/galgebra}{\textbf{galgebra}}~\cite{galgebra}:
-  Geometric algebra (previously \texttt{sympy.galgebra}).
-\item
-  \href{http://yt-project.org/}{\textbf{yt}}: Python package for
-  analyzing and visualizing volumetric data (\texttt{yt.units} uses SymPy).
-\item
-  \href{http://sfepy.org/}{\textbf{SfePy}} (Simple finite elements in Python),
+  SymPy. \\
+
+\href{http://www.pydy.org/}{\textbf{PyDy}}~\cite{gede2013constrained} &  & Multibody Dynamics with
+  Python. \\
+
+\href{https://github.com/brombo/galgebra}{\textbf{galgebra}}~\cite{galgebra} &  &
+  A module for geometric algebra (previously \texttt{sympy.galgebra}). \\
+
+\href{http://yt-project.org/}{\textbf{yt}}~\cite{2011ApJS..192....9T} &  & Python package for
+  analyzing and visualizing volumetric data (\texttt{yt.units} uses SymPy). \\
+
+\href{http://sfepy.org/}{\textbf{SfePy}}~\cite{cimrman2014sfepy} &  & (Simple finite elements in Python),
   cf.~\cite{cimrman2014sfepy}, is a Python package for solving partial
@@
   method~\cite{Zienkiewicz2013FEM}. SymPy is used within this package mostly for
-  code generation and testing.
-\item
-  \href{http://quameon.sourceforge.net/}{\textbf{Quameon}}: Quantum
-  Monte Carlo in Python.
-\item
-  \href{http://lcapy.elec.canterbury.ac.nz/}{\textbf{Lcapy}}:
-  Experimental Python package for teaching linear circuit analysis.
-\end{itemize}
+  code generation and testing. \\
+
+\href{http://quameon.sourceforge.net/}{\textbf{Quameon}}~\cite{quameon} &  & Quantum
+  Monte Carlo in Python. \\
+
+\href{http://lcapy.elec.canterbury.ac.nz/}{\textbf{Lcapy}}~\cite{lcapy} &  &
+  Experimental Python package for teaching linear circuit analysis. \\
+\bottomrule
+\end{longtable}

### 2016-09-26 projects_that_depend_on_sympy.tex
\ref the table
Change OctSymPy back to Octave Symbolic
As per @cbm755.

@@
 There are several projects that depend on SymPy as a library for implementing
-a part of their functionality. A selection of these projects are listed below:
+a part of their functionality. A selection of these projects are listed in
+Table~\ref{projects-table}.
 
@@
 
-\href{https://github.com/cbm755/octsympy}{\textbf{OctSymPy}}~\cite{OctSymPy} &  &
+\href{http://octave.sourceforge.net/symbolic}{\textbf{Octave Symbolic}}~\cite{OctSymPy} &  &
   The OctSymPy package adds symbolic calculation features

### 2016-09-26 projects_that_depend_on_sympy.tex
Fix Octave stuff

@@
 
-\href{http://octave.sourceforge.net/symbolic}{\textbf{Octave Symbolic}}~\cite{OctSymPy} &  &
+\href{https://github.com/cbm755/octsympy}{\textbf{GNU Octave Symbolic Package}}~\cite{OctSymPy} &  &
   The OctSymPy package adds symbolic calculation features

### 2016-09-26 projects_that_depend_on_sympy.tex
Sage -> SageMath

@@
 
-\href{http://www.sagemath.org/}{\textbf{Sage}}~\cite{sagemath} &  & A CAS, visioned to be
+\href{http://www.sagemath.org/}{\textbf{SageMath}}~\cite{sagemath} &  & A CAS, visioned to be
   a viable free open source alternative to Magma, Maple, Mathematica and

### 2016-09-26 assumptions.tex
Discuss Rete algorithm and new assumptions in the logic section
This should help our rebuttal point for keeping the supplement logic
subsection in the supplement.

@@
-SymPy performs logical inference through its assumptions system. The
-assumptions system allows users to specify that symbols have certain common
-mathematical properties, such as being positive, imaginary, or integral. SymPy
-is careful to never perform simplifications on an expression unless the
-assumptions allow them. For instance, the identity $\sqrt{t^2} = t$ holds if
-$t$ is nonnegative ($t\ge 0$). However, for general complex $t$, no
+The assumptions system allows users to specify that symbols have certain
+common mathematical properties, such as being positive, imaginary, or
+integral. SymPy is careful to never perform simplifications on an expression
+unless the assumptions allow them. For instance, the identity $\sqrt{t^2} = t$
+holds if $t$ is nonnegative ($t\ge 0$). However, for general complex $t$, no
 such identity holds.
@@
 
-Basic implications between the facts are used to deduce assumptions. For
-instance, the assumptions system knows that being an integer implies being
-rational.
+Basic implications between the facts are used to deduce assumptions.
+Deductions are made using the Rete
+algorithm~\cite{doorenbos1995production}.\footnote{For historical reasons,
+  this algorithm is distinct from the \texttt{sympy.logic} submodule, which is
+  discussed in the supplementary material. SymPy has an experimental
+  assumptions system which stores facts separate from objects, and uses
+  \texttt{sympy.logic} and a SAT solver for deduction. We will not discuss
+  this system here.} For instance, the assumptions system knows that being an
+integer implies being rational.
 \begin{verbatim}

### 2016-09-27 rebuttal.tex
Discuss Rete algorithm and new assumptions in the logic section
This should help our rebuttal point for keeping the supplement logic
subsection in the supplement.
Proofreading
This fixes section numbers and references, that out of sync
after reordering of sections.

@@
 
-We have added a sentence to the section 2.2, that shows symbolic
+We have added a sentence to the section 5.1, that shows symbolic
 constants as an example of leaf nodes:
@@
 Commutators are implemented in the \texttt{sympy.physics.quantum} module and
-are discussed in section 5.2.
+are discussed in section 4.2.
 
@@
 ``simplification'', but it's just one approach.  This citation
-was replaced with a better reference [28], that illustrates our point.
+was replaced with a better reference [32], that illustrates our point.
 \end{solution}
@@
 \begin{solution}
-We have added a footnote about the \texttt{dotprint} function (footnote 3),
+We have added a footnote about the \texttt{dotprint} function (footnote 12),
 which outputs expressions in the dot format that can be rendered with Graphviz.
@@
 This part was removed and footnote moved to the mentioned
-text in the section 2.2, ``The Core''.
+text in the section 5.1, ``The Core''.
 \end{solution}

### 2016-09-27 projects_that_depend_on_sympy.tex
Drop SymPy Live from table of projects, move to intro
* now we have the documentaion reference early
* save space
* we avoid some duplication

@@
 \midrule
-\href{http://live.sympy.org/}{\textbf{SymPy Live}} &  & SymPy Live an online
-  Python shell, which uses the Google App Engine to executes SymPy code. It is
-  integrated in the SymPy documentation examples at
-  \href{http://docs.sympy.org}{http://docs.sympy.org}. SymPy Live is
-  maintained by the SymPy community. \\
 

### 2016-09-27 projects_that_depend_on_sympy.tex
Compactify projects table
Domain column was removed.  I believe it can't be very usefull.  E.g.
for most entries there should be something like "CAS".  Free text
in the Description should be enough.

@@
 
-\begin{longtable}[htbc]{>{\raggedright}p{0.14\linewidth}p{0.14\linewidth}p{0.63\linewidth}}
+\begin{longtable}[htbc]{>{\raggedright}p{0.2\linewidth}p{0.8\linewidth}}
 \caption{Selected projects that depend on SymPy.\label{projects-table}}\\
 \toprule
-\textbf{Project name} & \textbf{Domain} & \textbf{Description} \\
+\textbf{Project name} & \textbf{Description} \\
 \midrule
 
-\href{http://sympygamma.com/}{\textbf{SymPy Gamma}} &  & SymPy Gamma is a
-  web application that executes and displays results for SymPy expressions, in
-  a fashon similar to that of Wolfram|Alpha. SymPy Gamma is maintained by the
-  SymPy community. See the supplementary material for more information about
-  SymPy Gamma. \\
+\href{http://sympygamma.com/}{\textbf{SymPy Gamma}} & An open source
+  analog of the Wolfram|Alpha.  See more in the supplementary material. \\
 
-\href{http://cadabra.science/index.html}{\textbf{Cadabra}}~\cite{Peeters2007cadabra} &  &
-  Cadabra is a CAS designed specifically for the resolution of problems
+\href{http://cadabra.science/index.html}{\textbf{Cadabra}}~\cite{Peeters2007cadabra} &
+  CAS designed specifically for the resolution of problems
   encountered in field theory. \\
 
-\href{https://github.com/cbm755/octsympy}{\textbf{GNU Octave Symbolic Package}}~\cite{OctSymPy} &  &
-  The OctSymPy package adds symbolic calculation features
-  to GNU Octave. These include common CAS tools such
-  as algebraic operations, calculus, equation solving, Fourier and
-  Laplace transforms, variable precision arithmetic, and other features. \\
+\href{https://github.com/cbm755/octsympy}{\textbf{GNU Octave Symbolic Package}}~\cite{OctSymPy} &
+  An implementation of a symbolic toolbox for Octave using SymPy. \\
 
-\href{https://github.com/jverzani/SymPy.jl}{\textbf{SymPy.jl}}~\cite{SymPy.jl} &  &
+\href{https://github.com/jverzani/SymPy.jl}{\textbf{SymPy.jl}}~\cite{SymPy.jl} &
   Provides a Julia interface to SymPy using PyCall. \\
 
-\href{https://mathics.github.io/}{\textbf{Mathics}}~\cite{Mathics} &  & Mathics is a
-  free, general-purpose online CAS featuring Mathematica compatible
-  syntax and functions. It is backed by highly extensible Python code,
-  relying on SymPy for most mathematical tasks.[<8;32;30m \\
+\href{https://mathics.github.io/}{\textbf{Mathics}}~\cite{Mathics} &
+  A free, online CAS featuring Mathematica compatible
+  syntax and functions. \\
 
-\href{http://mathpix.com/}{\textbf{Mathpix}}~\cite{Mathpix} &  & An iOS App, that detects handwritten math as input, and uses
+\href{http://mathpix.com/}{\textbf{Mathpix}}~\cite{Mathpix} & An iOS App, that detects handwritten math as input, and uses
   SymPy Gamma to evaluate the math input and generate the relevant
@@
 
-\href{http://openrave.org/docs/latest_stable/openravepy/ikfast/}{\textbf{IKFast}}~\cite{diankov2010ikfast} &  &
+\href{http://openrave.org/docs/latest_stable/openravepy/ikfast/}{\textbf{IKFast}}~\cite{diankov2010ikfast} &
   IKFast is a robot kinematics compiler provided by
-  \href{http://openrave.org/}{OpenRAVE}. It analytically solves robot inverse
-  kinematics equations and generates optimized C++ files. It uses SymPy for
-  its internal symbolic mathematics. \\
+  \href{http://openrave.org/}{OpenRAVE}. \\
 
-\href{http://www.sagemath.org/}{\textbf{SageMath}}~\cite{sagemath} &  & A CAS, visioned to be
-  a viable free open source alternative to Magma, Maple, Mathematica and
-  MATLAB\@. Sage includes many open source mathematical libraries, including
-  SymPy. \\
+\href{http://www.sagemath.org/}{\textbf{SageMath}}~\cite{sagemath} &
+  A free open-source mathematics software system, builds on top of many
+  existing open-source packages, including SymPy. \\
 
-\href{http://www.pydy.org/}{\textbf{PyDy}}~\cite{gede2013constrained} &  & Multibody Dynamics with
+\href{http://www.pydy.org/}{\textbf{PyDy}}~\cite{gede2013constrained} & Multibody Dynamics with
   Python. \\
 
-\href{https://github.com/brombo/galgebra}{\textbf{galgebra}}~\cite{galgebra} &  &
+\href{https://github.com/brombo/galgebra}{\textbf{galgebra}}~\cite{galgebra} &
   A module for geometric algebra (previously \texttt{sympy.galgebra}). \\
 
-\href{http://yt-project.org/}{\textbf{yt}}~\cite{2011ApJS..192....9T} &  & Python package for
-  analyzing and visualizing volumetric data (\texttt{yt.units} uses SymPy). \\
+\href{http://yt-project.org/}{\textbf{yt}}~\cite{2011ApJS..192....9T} & Python package for
+  analyzing and visualizing volumetric data. \\
 
-\href{http://sfepy.org/}{\textbf{SfePy}}~\cite{cimrman2014sfepy} &  & (Simple finite elements in Python),
-  cf.~\cite{cimrman2014sfepy}, is a Python package for solving partial
+\href{http://sfepy.org/}{\textbf{SfePy}}~\cite{cimrman2014sfepy} &
+  A Python package for solving partial
   differential equations (PDEs) in 1D, 2D and 3D by the finite element (FE)
-  method~\cite{Zienkiewicz2013FEM}. SymPy is used within this package mostly for
-  code generation and testing. \\
+  method~\cite{Zienkiewicz2013FEM}. \\
 
-\href{http://quameon.sourceforge.net/}{\textbf{Quameon}}~\cite{quameon} &  & Quantum
+\href{http://quameon.sourceforge.net/}{\textbf{Quameon}}~\cite{quameon} & Quantum
   Monte Carlo in Python. \\
 
-\href{http://lcapy.elec.canterbury.ac.nz/}{\textbf{Lcapy}}~\cite{lcapy} &  &
+\href{http://lcapy.elec.canterbury.ac.nz/}{\textbf{Lcapy}}~\cite{lcapy} &
   Experimental Python package for teaching linear circuit analysis. \\

### 2016-09-27 domain_specific.tex
Skip circuit_plot doctest
Compactify some symbol defs in Physics section

@@
 >>> from sympy.physics.vector import ReferenceFrame
->>> A = ReferenceFrame('A')
->>> B = ReferenceFrame('B')
->>> C = ReferenceFrame('C')
+>>> A, B, C = symbols('A B C', cls=ReferenceFrame)
 >>> B.orient(A, 'body', (pi, pi/3, pi/4), 'zxz')
@@
 >>> from sympy.physics.quantum import Ket, qapply
->>> A = Operator('A')
->>> B = Operator('B')
->>> C = Operator('C')
->>> D = Operator('D')
+>>> A, B, C, D = symbols('A B C D', cls=Operator)
 >>> a = Ket('a')
@@
 circuit plot for the sequence of gates (see Figure~\ref{fig-circuitplot-qft}).
+% This depends on matplotlib. We want to make sure the rest of the paper
+% doesn't depend on it, so skip.
+% no-doctest
 \begin{verbatim}

### 2016-09-27 architecture.tex
Add missing symbols declaration, fix tests
Consistent non-usage of commas in symbols()
Drop another assumption handler for gamma
We have one example, why second?

@@
 \begin{verbatim}
+>>> x, y = symbols('x y')
 >>> expr = x*y + 2
@@
 function $\Gamma(x)$ from SymPy, which evaluates itself on positive integer
-arguments, has the positive and real assumptions defined, can be rewritten in
+arguments, has the real assumption defined, can be rewritten in
 terms of factorial with \verb|gamma(x).rewrite(factorial)|, and can be
@@
             return factorial(arg - 1)
-
-    def _eval_is_positive(self):
-        x = self.args[0]
-        if x.is_positive:
-            return True
-        elif x.is_noninteger:
-            return floor(x).is_even
 

### 2016-09-27 rebuttal.tex
Add automatic cross-references to the paper (in rebuttal)
See more at:
http://tex.stackexchange.com/questions/3060/referencing-labels-from-other-files

XXX: Section and footnotes numbers are correct, but my pdf viewer
(evince) doesn't open a new document if I click on them.  I hope,
this is a bug.

@@
 \usepackage{xcolor}
+
+\usepackage{xr-hyper}
+\usepackage{xr}
+\externaldocument[P-]{paper}
+
 \usepackage{hyperref}
@@
 
-We have added a sentence to the section 5.1, that shows symbolic
+We have added a sentence to the section~\ref{P-sec:core}, that shows symbolic
 constants as an example of leaf nodes:
@@
 Commutators are implemented in the \texttt{sympy.physics.quantum} module and
-are discussed in section 4.2.
+are discussed in section~\ref{P-sec:quantum}.
 
@@
 \begin{solution}
-We have added a footnote about the \texttt{dotprint} function (footnote 12),
+We have added a footnote about the \texttt{dotprint} function (footnote~\ref{P-note:dotprint}),
 which outputs expressions in the dot format that can be rendered with Graphviz.
@@
 
-We deem all the sections in section 3 (now section 2) to be important core
+We deem all the sections in section 3 (now section~\ref{P-sec:features}) to be important core
 components of the library. Regarding the supplement sections (previously) 5,
@@
 \begin{solution}
-  The citation was moved and a footnote for the Python interactive prompt was
+The citation was moved and a footnote~\ref{P-note:prompt} for the Python interactive prompt was
   added where it was used first time. We have also added a footnote indicating
@@
 This part was removed and footnote moved to the mentioned
-text in the section 5.1, ``The Core''.
+text in the section~\ref{P-sec:core}, ``The Core''.
 \end{solution}
@@
   designed to be compact, human-readable, and valid Python code that could be
-  used to recreate the expression. As it was noted in section~5.1,
+  used to recreate the expression. As it was noted in section~\ref{P-sec:core},
   the \texttt{srepr} function prints the exact, verbose form of an expression.

### 2016-09-27 architecture.tex
Add automatic cross-references to the paper (in rebuttal)
See more at:
http://tex.stackexchange.com/questions/3060/referencing-labels-from-other-files

XXX: Section and footnotes numbers are correct, but my pdf viewer
(evince) doesn't open a new document if I click on them.  I hope,
this is a bug.

@@
 returns a string representation of an expression as valid Python code\footnote{
+\label{note:dotprint}
 The \texttt{dotprint} function from the \texttt{sympy.printing.dot} submodule

### 2016-09-27 introduction.tex
CASs -> CASes
Reviewer 3, comment 3.
Use CAS's (reviewer 3 comment 3)
Reorder section descriptions in the intro
Drop SymPy Live from table of projects, move to intro
* now we have the documentaion reference early
* save space
* we avoid some duplication
Add automatic cross-references to the paper (in rebuttal)
See more at:
http://tex.stackexchange.com/questions/3060/referencing-labels-from-other-files

XXX: Section and footnotes numbers are correct, but my pdf viewer
(evince) doesn't open a new document if I click on them.  I hope,
this is a bug.

@@
 
-Unlike many CASs, SymPy does not invent its own programming language. Python
+Unlike many CAS's, SymPy does not invent its own programming language. Python
 itself is used both for the internal implementation and end user
@@
 The remainder of this paper discusses key components of the SymPy library.
-Section~\ref{sec:architecture} discusses the architecture of SymPy.
 Section~\ref{sec:features} enumerates the features of SymPy and takes a closer
-look at some of the important ones. The section~\ref{sec:numerics}
-looks at the numerical features of SymPy and its dependency library, mpmath.
+look at some of the important ones. The section~\ref{sec:numerics} looks at
+the numerical features of SymPy and its dependency library, mpmath.
 Section~\ref{sec:domain_specific} looks at the domain specific physics
-submodules for performing symbolic and numerical calculations in classical mechanics
-and quantum mechanics. Conclusions and future directions for SymPy are given
-in section~\ref{sec:conclusion}.  All examples in this paper use SymPy
-version 1.0 and mpmath version 0.19.
+submodules for performing symbolic and numerical calculations in classical
+mechanics and quantum mechanics. Section~\ref{sec:architecture} discusses the
+architecture of SymPy. Conclusions and future directions for SymPy are given
+in section~\ref{sec:conclusion}. All examples in this paper use SymPy version
+1.0 and mpmath version 0.19.
 
@@
 From here on, all examples in this paper assume that this statement has been
-executed:\footnote{The three greater-than signs denote the user input for the
+executed:\footnote{\label{note:prompt} The three greater-than signs denote the user input for the
   Python interactive session, with the result, if there is one, shown on the
@@
 
-All examples could be tested on the \href{http://live.sympy.org}{SymPy
-Live} instance, that is an online Python shell, which uses the Google
-App Engine to execute SymPy code.
+All examples could be tested on the \href{http://live.sympy.org}{SymPy Live}
+instance, that is an online Python shell, which uses the Google App Engine to
+execute SymPy code.  The SymPy Live is integrated in the SymPy documentation
+system at \href{http://docs.sympy.org}{http://docs.sympy.org}.

### 2016-09-27 rebuttal.tex
Document in rebuttal.tex "other projects" section

@@
     the supplement.
+  \item The ``Other Projects that
+      Depend on SymPy'' section (which has been renamed from ``Other Projects that
+      Use SymPy'') has been moved from the supplement to the main paper content.
   \end{enumerate}
@@
 examples from the paper.
-
-% TODO For projects that use SymPy
-% TODO: Make intro sentence longer.
-% TODO: Make it into a table
-% TODO: Project name, domain, description
-
-The ``Other Projects that
-Depend on SymPy'' section (which has been renamed from ``Other Projects that
-Use SymPy'') is {\color{red}TODO: Add reasons here, or else move this to the
-paper as the reviewer requests.}
 

### 2016-09-27 numerics.tex
small change about mpfr

@@
  of mpmath over other multiple precision libraries such as GNU MPFR~\cite{Fousse:2007:MMB:1236463.1236468},
-which is much more feature-rich and fast.  Like SymPy, mpmath is also BSD
+which is faster.  Like SymPy, mpmath is also BSD
 licensed (GNU MPFR is licensed under the GNU Lesser General Public License~\cite{rosen2005open}).

### 2016-09-27 rebuttal.tex
Remove acknowledgements section
This metadata should go in the funding statement on the PeerJ site, which will
then be part of the final document created by PeerJ. See
https://github.com/sympy/sympy-paper/issues/205 and
https://github.com/sympy/sympy-paper/pull/206#issuecomment-241577396.
Split the cover letter into a separate document so we can us the letter class

@@
 
-\usepackage{xr-hyper}
 \usepackage{xr}
@@
 
+\usepackage{pdfpages}
+
 \unframedsolutions
@@
 
-Dear Editors, \bigskip
-
-We thank you and the referees for your thoughtful comments and suggestions on
-our manuscript, ``SymPy: Symbolic Computing in Python''. We have endeavored to
-address your comments and concerns point by point below. Below we give your
-and the reviewers' comments on white background and our replies on gray
-background. \bigskip
-
-Reviewer 1 had concerns that the paper was out of scope, but the editor has
-confirmed that it is in scope. Many of the points that reviewer 3 raised were
-in regards to the relevance of various sections. The intended scope of this
-paper was to be an architecture paper for SymPy's 1.0 release. We have
-examined reviewer 3's points through this lens. In light of this, which
-contents were previously in the supplement vs.\ the main manuscript may have
-been confusing to a general audience. To address this, we have restructured
-the manuscript to make the paper more readable to those not already experts in
-SymPy or symbolic mathematics. \bigskip
-
-We thank you again for your consideration. Given these edits, we believe that
-the manuscript is now suitable for publication in PeerJ Computer Science. If
-there are further questions or concerns do not hesitate to contact us. \bigskip
-
-Best regards,\bigskip
-
-Aaron Meurer
-
-\pagebreak
+\includepdf{coverletter.pdf}
 

### 2016-09-27 coverletter.tex
Split the cover letter into a separate document so we can us the letter class

\documentclass{letter}
\usepackage{hyperref}
\signature{Aaron Meurer}
\address{address}
\begin{document}

\begin{letter}{address}
\opening{Dear Editors,}

Dear Editors, \bigskip

We thank you and the referees for your thoughtful comments and suggestions on
our manuscript, ``SymPy: Symbolic Computing in Python''. We have endeavored to
address your comments and concerns point by point below. Below we give your
and the reviewers' comments on white background and our replies on gray
background. \bigskip

Reviewer 1 had concerns that the paper was out of scope, but the editor has
confirmed that it is in scope. Many of the points that reviewer 3 raised were
in regards to the relevance of various sections. The intended scope of this
paper was to be an architecture paper for SymPy's 1.0 release. We have
examined reviewer 3's points through this lens. In light of this, which
contents were previously in the supplement vs.\ the main manuscript may have
been confusing to a general audience. To address this, we have restructured
the manuscript to make the paper more readable to those not already experts in
SymPy or symbolic mathematics. \bigskip

We thank you again for your consideration. Given these edits, we believe that
the manuscript is now suitable for publication in PeerJ Computer Science. If
there are further questions or concerns do not hesitate to contact us. \bigskip

Best regards,\bigskip

\end{letter}
\end{document}

### 2016-09-28 coverletter.tex
Clean up the letter formatting a bit
Add my signature
Uses method from http://texblog.org/2010/05/04/latex-letter-signature-as-image/
Make my signature smaller
Add the ERGS image to the cover letter
See https://tex.stackexchange.com/questions/42492/put-an-image-header-in-documentclassletter
Add `please`

@@
 \documentclass{letter}
+
+\usepackage{graphicx}
+
 \usepackage{hyperref}
-\signature{Aaron Meurer}
-\address{address}
+% \signature{Aaron Meurer}
+\address{Aaron Meurer \\
+ERGS: \url{http://www.ergs.sc.edu/} \\
+Nuclear Eng., Mechanical Eng. Dept. \\
+University of South Carolina \\
+541 Main Street, St. 011 \\
+Columbia, SC 29201 \\
+\url{asmeurer@gmail.com}}
+
+\date{September 31, 2016} % stand-in
 \begin{document}
 
-\begin{letter}{address}
+\begin{letter}{}
 \opening{Dear Editors,}
-
-Dear Editors, \bigskip
 
@@
 and the reviewers' comments on white background and our replies on gray
-background. \bigskip
+background.
 
@@
 the manuscript to make the paper more readable to those not already experts in
-SymPy or symbolic mathematics. \bigskip
+SymPy or symbolic mathematics.
 
@@
 the manuscript is now suitable for publication in PeerJ Computer Science. If
-there are further questions or concerns do not hesitate to contact us. \bigskip
+there are further questions or concerns please do not hesitate to contact us.
 
-Best regards,\bigskip
+% Main text
+\closing{Best regards,\\
+\fromsig{\includegraphics[scale=0.07]{aaron-signature.png}} \\
+\fromname{Aaron Meurer}
+}
 

### 2016-09-28 rebuttal.tex
Add the full text from reviewer 3

@@
 \begin{questions}
-\question I suggest a major rewrite of the paper, to improve the structure and group differently approached subjects. It would probably be best to:
+\question The paper introduces SymPy, a well-known pure Python library for
+symbolic computation. It covers many aspects of the library: architecture,
+basic usage, overview of modules, a more detailed look into some of them, and
+physics application.
+
+The writing style is mostly clear and easy to understand. Some example code is
+provided to further explain the use of various module, classes, and functions.
+Unfortunately, the paper is very unconnected and disorganized, with some
+unnecessary repetition and some things left undefined. Further, parts of the
+supplement should be moved to the main paper and vice versa. All of this makes
+the paper look more like a collection of unconnected or, at best, very loosely
+connected parts, instead of a meaningful whole.
+
+Another big problem of the paper is the lack of aim. Some parts of SymPy are
+covered at informative level (short descriptions of the elements related to
+some subject), some at the beginners level (basic usage examples), while some
+go deep in internal SymPy implementation of certain features. This structure
+leaves the impression that different, yet mutually intermingled sections aim
+at different audiences.
+
+Moreover, these vastly differently approached elements come in no specific
+order, and with no obvious reason why each of them is picked to be covered at
+all and, specifically, at the chosen level of complexity in the approach.
+
+I suggest a major rewrite of the paper, to improve the structure and group differently approached subjects. It would probably be best to:
 \begin{enumerate}

### 2016-09-28 coverletter.tex
Add the sentence back in
It was removed in 59dcf547dfd813fe91a377c6ddcfda6462f66dbf.
Fix my office number
Remove confusing sentence from letter

@@
 University of South Carolina \\
-541 Main Street, St. 011 \\
+541 Main Street, St. 009 \\
 Columbia, SC 29201 \\
@@
 paper was to be an architecture paper for SymPy's 1.0 release. We have
-examined reviewer 3's points through this lens. In light of this, which
-contents were previously in the supplement vs.\ the main manuscript may have
-been confusing to a general audience. To address this, we have restructured
-the manuscript to make the paper more readable to those not already experts in
-SymPy or symbolic mathematics.
+examined reviewer 3's points through this lens. To address this, we have
+restructured the manuscript to make the paper more readable to those not
+already experts in SymPy or symbolic mathematics.
 
-We thank you again for your consideration. Given these edits, we believe that
-the manuscript is now suitable for publication in PeerJ Computer Science. If
-there are further questions or concerns please do not hesitate to contact us.
+We thank you again for your consideration. Given these edits, the manuscript
+has been greatly improved and we believe that it is now suitable for
+publication in PeerJ Computer Science. If there are further questions or
+concerns please do not hesitate to contact us.
 

### 2016-09-28 rebuttal.tex
Note that it was moved from the supplement
Note the section number
Update mpfr rebuttal

@@
 \label{editorpoint1}
-This sentence was added to the introduction:
+Section (previously) 9 from the supplement was edited and moved to the introduction:
 \begin{quote}
@@
   than MPFR, but not too much to evaluate any reasonably complicated single
-  expression at interactive speed, which covers most use cases in SymPy.
+  expression at interactive speed, which covers most use cases in SymPy. We
+  have added a sentence about this to the numerics section.
 
@@
 Python ecosystem (MPFR is licensed under the LGPL).
-
-It is also worth noting that none of the referee materials we received mention MPFR\@.
 \end{solution}

### 2016-09-28 projects_that_depend_on_sympy.tex
Fix the width of the projects that depend on SymPy table
Add citation for SymPy Gamma
Move citations to the text in the projects table

@@
 
-\begin{longtable}[htbc]{>{\raggedright}p{0.2\linewidth}p{0.8\linewidth}}
+\begin{longtable}[htbc]{>{\raggedright}p{0.2\linewidth}p{0.74\linewidth}}
 \caption{Selected projects that depend on SymPy.\label{projects-table}}\\
@@
 \href{http://sympygamma.com/}{\textbf{SymPy Gamma}} & An open source
-  analog of the Wolfram|Alpha.  See more in the supplementary material. \\
+  analog of Wolfram|Alpha that uses SymPy~\cite{SymPyGamma}.  See more in the supplementary material. \\
 
-\href{http://cadabra.science/index.html}{\textbf{Cadabra}}~\cite{Peeters2007cadabra} &
+\href{http://cadabra.science/index.html}{\textbf{Cadabra}} &
   CAS designed specifically for the resolution of problems
-  encountered in field theory. \\
+  encountered in field theory~\cite{Peeters2007cadabra}. \\
 
-\href{https://github.com/cbm755/octsympy}{\textbf{GNU Octave Symbolic Package}}~\cite{OctSymPy} &
-  An implementation of a symbolic toolbox for Octave using SymPy. \\
+\href{https://github.com/cbm755/octsympy}{\textbf{GNU Octave Symbolic Package}} &
+  An implementation of a symbolic toolbox for Octave using SymPy~\cite{OctSymPy}. \\
 
-\href{https://github.com/jverzani/SymPy.jl}{\textbf{SymPy.jl}}~\cite{SymPy.jl} &
-  Provides a Julia interface to SymPy using PyCall. \\
+\href{https://github.com/jverzani/SymPy.jl}{\textbf{SymPy.jl}} &
+  Provides a Julia interface to SymPy using PyCall~\cite{SymPy.jl}. \\
 
-\href{https://mathics.github.io/}{\textbf{Mathics}}~\cite{Mathics} &
+\href{https://mathics.github.io/}{\textbf{Mathics}} &
   A free, online CAS featuring Mathematica compatible
-  syntax and functions. \\
+  syntax and functions~\cite{Mathics}. \\
 
-\href{http://mathpix.com/}{\textbf{Mathpix}}~\cite{Mathpix} & An iOS App, that detects handwritten math as input, and uses
+\href{http://mathpix.com/}{\textbf{Mathpix}} & An iOS App, that detects handwritten math as input, and uses
   SymPy Gamma to evaluate the math input and generate the relevant
-  steps to solve the problem. \\
+  steps to solve the problem~\cite{Mathpix}. \\
 
-\href{http://openrave.org/docs/latest_stable/openravepy/ikfast/}{\textbf{IKFast}}~\cite{diankov2010ikfast} &
+\href{http://openrave.org/docs/latest_stable/openravepy/ikfast/}{\textbf{IKFast}} &
   IKFast is a robot kinematics compiler provided by
-  \href{http://openrave.org/}{OpenRAVE}. \\
+  \href{http://openrave.org/}{OpenRAVE}~\cite{diankov2010ikfast}. \\
 
-\href{http://www.sagemath.org/}{\textbf{SageMath}}~\cite{sagemath} &
+\href{http://www.sagemath.org/}{\textbf{SageMath}} &
   A free open-source mathematics software system, builds on top of many
-  existing open-source packages, including SymPy. \\
+  existing open-source packages, including SymPy~\cite{sagemath}. \\
 
-\href{http://www.pydy.org/}{\textbf{PyDy}}~\cite{gede2013constrained} & Multibody Dynamics with
-  Python. \\
+\href{http://www.pydy.org/}{\textbf{PyDy}} & Multibody Dynamics with
+  Python~\cite{gede2013constrained}. \\
 
-\href{https://github.com/brombo/galgebra}{\textbf{galgebra}}~\cite{galgebra} &
-  A module for geometric algebra (previously \texttt{sympy.galgebra}). \\
+\href{https://github.com/brombo/galgebra}{\textbf{galgebra}} &
+  A module for geometric algebra (previously \texttt{sympy.galgebra})~\cite{galgebra}. \\
 
-\href{http://yt-project.org/}{\textbf{yt}}~\cite{2011ApJS..192....9T} & Python package for
-  analyzing and visualizing volumetric data. \\
+\href{http://yt-project.org/}{\textbf{yt}} & Python package for
+  analyzing and visualizing volumetric data~\cite{2011ApJS..192....9T}. \\
 
-\href{http://sfepy.org/}{\textbf{SfePy}}~\cite{cimrman2014sfepy} &
+\href{http://sfepy.org/}{\textbf{SfePy}} &
   A Python package for solving partial
   differential equations (PDEs) in 1D, 2D and 3D by the finite element (FE)
-  method~\cite{Zienkiewicz2013FEM}. \\
+  method~\cite{Zienkiewicz2013FEM,cimrman2014sfepy}. \\
 
-\href{http://quameon.sourceforge.net/}{\textbf{Quameon}}~\cite{quameon} & Quantum
-  Monte Carlo in Python. \\
+\href{http://quameon.sourceforge.net/}{\textbf{Quameon}} & Quantum
+  Monte Carlo in Python~\cite{quameon}. \\
 
-\href{http://lcapy.elec.canterbury.ac.nz/}{\textbf{Lcapy}}~\cite{lcapy} &
-  Experimental Python package for teaching linear circuit analysis. \\
+\href{http://lcapy.elec.canterbury.ac.nz/}{\textbf{Lcapy}} &
+  Experimental Python package for teaching linear circuit analysis~\cite{lcapy}. \\
 \bottomrule

### 2016-09-28 rebuttal.tex
Fix some formatting

@@
 \question Generally good. A part that confused me is the assertion (footnote 3) that ``If
-A and B are Symbols created with commutative=False then SymPy will keep A·B
-and B·A distinct.'' Does that mean that BOTH of them must be created this way,
-and that A and x (if x is created normally) will commute? Is there any way to
+$A$ and $B$ are Symbols created with \texttt{commutative=False} then SymPy will keep
+$A\cdot B$
+and $B\cdot A$ distinct.'' Does that mean that BOTH of them must be created this way,
+and that $A$ and $x$ (if $x$ is created normally) will commute? Is there any way to
 declare commutators? How does one guarantee that other pieces of code, e.g.

### 2016-09-28 architecture.tex
new performance

@@
 
-\subsection{Speed}
+\subsection{Performance}
 
-Due to being written in pure Python, SymPy's speed is generally slower compared
-with its commercial competitors. For many applications and uses of SymPy, that
-is not a problem, as SymPy is able to return the answer quickly enough, but for
-some applications that require handling of very long expressions and/or lots of
-small expressions, the speed becomes a problem.
+Due to being written in pure Python without the use of extension modules,
+SymPy's performance characteristics are generally poorer than
+its commercial competitors. For many applications,
+the performance of SymPy, as measured by clock cycles, memory usage, and memory
+layout, is sufficient.
+However, the boundaries for when SymPy's pure Python strategy becomes
+insufficient are when the user requires handling of very long expressions or many
+small expressions. Where this boundray lies depends on the system at hand, but tends
+to be within the range of $O(10^4) - O(10^6)$ symbols for modern computers.
 
-For this reason, a new library called SymEngine~\cite{SymEngine} was started.
-It is a pure C++ library with thin wrappers to other languages (Python, Ruby,
-Julia, \ldots) whose aim is to be the fastest manipulation library.  Preliminary
-benchmarks suggest that SymEngine is as fast or faster than the commercial or
+For this reason, a new project called SymEngine~\cite{SymEngine} has been started.
+The aim of this poject is to develop a library with better performance
+characteristics for symbolic manipulation. SymEngine is a pure C++ library,
+which allows it fine grain control over the memory layout of expressions.
+SymEngine has thin wrappers to other languages (Python, Ruby,
+Julia, etc.) whose aim is to be the fastest manipulation library. Preliminary
+benchmarks suggest that SymEngine is as performant as its commercial and
 open source competitors.
 
-The development branch of SymPy recently started to use SymEngine as an
-optional backend, initially in \texttt{sympy.physics.mechanics} only. The plan is
-to allow more algorithms in SymPy to take advantage of the speed of SymEngine.
+SymPy has recently started to use SymEngine as an
+optional backend, initially in \texttt{sympy.physics.mechanics} only.
+Future work witll involve
+allowing more algorithms in SymPy to use SymEngine as a backend.

### 2016-09-28 projects_that_depend_on_sympy.tex

@@
 \href{http://sympygamma.com/}{\textbf{SymPy Gamma}} & An open source
-  analog of Wolfram|Alpha that uses SymPy~\cite{SymPyGamma}.  See more in the supplementary material. \\
+  analog of Wolfram|Alpha that uses SymPy~\cite{SymPyGamma}.  There is more
+  information about SymPy Gamma supplementary material. \\
 
 \href{http://cadabra.science/index.html}{\textbf{Cadabra}} &
-  CAS designed specifically for the resolution of problems
+  A CAS designed specifically for the resolution of problems
   encountered in field theory~\cite{Peeters2007cadabra}. \\
@@
 \href{https://github.com/jverzani/SymPy.jl}{\textbf{SymPy.jl}} &
-  Provides a Julia interface to SymPy using PyCall~\cite{SymPy.jl}. \\
+  A Julia interface to SymPy, provided using PyCall~\cite{SymPy.jl}. \\
 
@@
 
-\href{http://mathpix.com/}{\textbf{Mathpix}} & An iOS App, that detects handwritten math as input, and uses
+\href{http://mathpix.com/}{\textbf{Mathpix}} & An iOS App that detects handwritten math as input and uses
   SymPy Gamma to evaluate the math input and generate the relevant
@@
 \href{http://openrave.org/docs/latest_stable/openravepy/ikfast/}{\textbf{IKFast}} &
-  IKFast is a robot kinematics compiler provided by
+  A robot kinematics compiler provided by
   \href{http://openrave.org/}{OpenRAVE}~\cite{diankov2010ikfast}. \\
@@
 \href{http://www.sagemath.org/}{\textbf{SageMath}} &
-  A free open-source mathematics software system, builds on top of many
+  A free open-source mathematics software system, which builds on top of many
   existing open-source packages, including SymPy~\cite{sagemath}. \\
@@
 
-\href{http://yt-project.org/}{\textbf{yt}} & Python package for
+\href{http://yt-project.org/}{\textbf{yt}} & A Python package for
   analyzing and visualizing volumetric data~\cite{2011ApJS..192....9T}. \\
@@
   A Python package for solving partial
-  differential equations (PDEs) in 1D, 2D and 3D by the finite element (FE)
+  differential equations (PDEs) in 1D, 2D, and 3D by the finite element (FE)
   method~\cite{Zienkiewicz2013FEM,cimrman2014sfepy}. \\
@@
 \href{http://lcapy.elec.canterbury.ac.nz/}{\textbf{Lcapy}} &
-  Experimental Python package for teaching linear circuit analysis~\cite{lcapy}. \\
+  An experimental Python package for teaching linear circuit analysis~\cite{lcapy}. \\
 \bottomrule

### 2016-09-28 architecture.tex
Fix line wrapping

@@
 small expressions. Where this boundray lies depends on the system at hand, but tends
-to be within the range of $O(10^4) - O(10^6)$ symbols for modern computers.
+to be within the range of $10^4$--$10^6$ symbols for modern computers.
 
@@
 SymPy has recently started to use SymEngine as an
-optional backend, initially in \texttt{sympy.physics.mechanics} only.
+optional backend, initially in \texttt{sympy.\allowbreak{}physics.\allowbreak{}mechanics} only.
 Future work witll involve

### 2016-09-28 rebuttal.tex
speed -> perf
Cross-reference the section

@@
 \begin{solution}
-A new paragraph titled ``Speed'' was added to the manuscript, which discusses the
+Section~\ref{P-sec:performance}, titled ``Performance'' was added to the manuscript, which discusses the
 speed issue and lays out a path forward. The conclusion was simplified.

### 2016-09-28 rebuttal.tex
Fix broken external links.
And order does matter.

@@
 
-\usepackage{xr}
+\usepackage{xr-hyper}
+\usepackage{hyperref}
 \externaldocument[P-]{paper}
 
-\usepackage{hyperref}
 \usepackage{upquote}
@@
 \usepackage{pdfpages}
+
+\hypersetup{
+  colorlinks=true,
+  linkcolor=black, % color of internal links. Black because the cross-document
+                   % links don't actually work.
+  urlcolor=blue % color of external links
+}
 
@@
   \url{http://docs.sympy.org/latest/modules/core.html#module-sympy.core.assumptions}.
+  We have opened an \href{https://github.com/sympy/sympy/issues/11539}{issue}
+  about this in our issue tracker.
 \end{solution}
@@
 We disagree with the reviewer on a few points. We do not consider SymPy Live
-or SymPy Gamma to be key components of SymPy. The paper goal of the paper is
+or SymPy Gamma to be key components of SymPy. The goal of the paper is
 to be about the SymPy library. While SymPy Live and SymPy Gamma are

### 2016-09-28 rebuttal.tex
Add a link to the issue
Colorize the links in the rebuttal
minor paper

--- 2016-09-28 rebuttal.tex
@@
 \begin{solution}
-A new paragraph titled ``Speed'' was added to the manuscript, which discusses the
+A section titled ``Performance'' was added to the manuscript, which discusses the
 speed issue and lays out a path forward. The conclusion was simplified.

### 2016-09-28 architecture.tex
Cross-reference the section

@@
 \subsection{Performance}
+\label{sec:performance}
 
@@
 SymEngine has thin wrappers to other languages (Python, Ruby,
-Julia, etc.) whose aim is to be the fastest manipulation library. Preliminary
+Julia, etc.) whose aim is to be the fastest symbolic manipulation library. Preliminary
 benchmarks suggest that SymEngine is as performant as its commercial and
@@
 optional backend, initially in \texttt{sympy.\allowbreak{}physics.\allowbreak{}mechanics} only.
-Future work witll involve
+Future work will involve
 allowing more algorithms in SymPy to use SymEngine as a backend.

### 2016-09-28 matrices.tex
Add SciPy reference for LIL

@@
 
-Internally these matrices store the elements as Lists of Lists (LIL), meaning
+Internally these matrices store the elements as Lists of Lists (LIL)~\cite{scipy}, meaning
 the matrix is stored as a list of lists of entries (effectively, the

### 2016-09-28 rebuttal.tex
Set cross-document link color to black
Fix minus sign

--- 2016-09-28 rebuttal.tex
@@
   colorlinks=true,
-  linkcolor=black, % color of internal links. Black because the cross-document
-                   % links don't actually work.
+  filecolor=black, % color of cross-file links. Black because they only work
+                   % if the paper sits alongside the rebuttal as "paper.pdf".
   urlcolor=blue % color of external links
@@
 
-\question I suggest a better example for lines 459--460: a computation of $(e^{100} +1)− e^{100}$:
+\question I suggest a better example for lines 459--460: a computation of
+$(e^{100} +1) - e^{100}$:
 

### 2016-09-28 coverletter.tex

@@
 \usepackage{graphicx}
+\usepackage{fancyhdr}
+\usepackage{geometry}
 
@@
 \date{September 31, 2016} % stand-in
+
+\geometry{headheight = 0.6in}
+\fancypagestyle{empty}{\fancyhf{}\fancyhead[L]{\includegraphics[height=1in, keepaspectratio=true]{logo-curves.png}}}
+
 \begin{document}
@@
 our manuscript, ``SymPy: Symbolic Computing in Python''. We have endeavored to
-address your comments and concerns point by point below. Below we give your
+address your comments and concerns point by point below. We give your
 and the reviewers' comments on white background and our replies on gray

### 2016-09-29 rebuttal.tex
Clear an unnecessary sentence
Modify SymPy Live paragraph

@@
 \begin{quote}
-All examples could be tested on the \href{http://live.sympy.org}{SymPy
-Live} instance, that is an online Python shell, which uses the Google
-App Engine to execute SymPy code.
+  All the examples in this paper can be tested on
+  \href{http://live.sympy.org}{SymPy Live}, an online Python shell that uses
+  the Google App Engine to execute SymPy code. SymPy Live is also integrated
+  in the SymPy documentation at
+  \href{http://docs.sympy.org}{http://docs.sympy.org}.
 \end{quote}
@@
 
-Applying \texttt{evalf()} has nothing related to \emph{symbolic} computation,
-that method is for numerical evaluation, that should be clear from
-the sentence below ``The evalf method converts a constant symbolic expression
-to a Float\dots''.  Numerical evaluation of the whole expression should
+Numerical evaluation of the whole expression should
 always solve cancellation problem, as it was stated in the sentence

### 2016-09-29 introduction.tex
Modify SymPy Live paragraph

@@
 
-All examples could be tested on the \href{http://live.sympy.org}{SymPy Live}
-instance, that is an online Python shell, which uses the Google App Engine to
-execute SymPy code.  The SymPy Live is integrated in the SymPy documentation
-system at \href{http://docs.sympy.org}{http://docs.sympy.org}.
+All the examples in this paper can be tested on
+\href{http://live.sympy.org}{SymPy Live}, an online Python shell that uses the
+Google App Engine~\cite{ciurana2009google} to execute SymPy code. SymPy Live
+is also integrated in the SymPy documentation at
+\href{http://docs.sympy.org}{http://docs.sympy.org}.

### 2016-09-29 rebuttal.tex
Grammar

@@
 \begin{solution}
-  Both Python standard library and SymPy have exp callable, so it depends
+  Both the Python standard library and SymPy have an exp callable, so it depends
   which symbols are imported in your global namespace: the \texttt{exp}
-  function from the \texttt{math} module or the \texttt{exp} class from SymPy.
+  function from the \texttt{math} module or the \texttt{exp} class from \texttt{sympy}.
   In the later case a special symbolic constant will be produced.

### 2016-09-29 rebuttal.tex
Wording

@@
 
-We mention this at the beginning of the section. Another advantage is that
-mpmath is BSD licensed, like SymPy, which is important for the scientific
-Python ecosystem (MPFR is licensed under the LGPL).
+  We've also mentioned that another advantage is that mpmath is BSD licensed,
+  like SymPy, which is important for the scientific Python ecosystem (MPFR is
+  licensed under the LGPL).
 \end{solution}

### 2016-09-29 rebuttal.tex
Fix reference number
List the issue number

@@
 ``simplification'', but it's just one approach.  This citation
-was replaced with a better reference [32], that illustrates our point.
+was replaced with a better reference [34], that illustrates our point.
 \end{solution}
@@
   \url{http://docs.sympy.org/latest/modules/core.html#module-sympy.core.assumptions}.
-  We have opened an \href{https://github.com/sympy/sympy/issues/11539}{issue}
-  about this in our issue tracker.
+  We have opened \href{https://github.com/sympy/sympy/issues/11539}{issue 11539}
+  to track this in our issue tracker.
 \end{solution}

### 2016-09-29 rebuttal.tex
More about SymPy Gamma in the rebuttal
Fix some formatting

@@
 point~\ref{editorpoint1}), as it may assist an interested reader in trying the
-examples from the paper.
+examples from the paper. A mention of SymPy Gamma has been added to the ``Projects that
+Depends on SymPy'' table (Table~\ref{P-projects-table}) (see below). We have
+left the more in-depth discussion of SymPy Gamma in the supplement.
 
@@
 Further, the paper assumes a moderate familiarity with Python (for example, Python's console, OOP, and exceptions), and this should be specified.
-There should be a short note on the used Python console (\texttt{>>>} is the prompt, with the results of computation following immediately in the lines after it).
+There should be a short note on the used Python console (\verb|>>>| is the prompt, with the results of computation following immediately in the lines after it).
 The citation [25] from line 65 should be moved next to ``Python'' in line 70.

### 2016-09-29 supplement.tex
Fix supplement ~ in front of \ref

@@
 which there was not enough room to discuss in the paper.
-Section \ref{suppsec:Gruntz} discusses the Gruntz algorithm, used to
-calculate limits in the SymPy.  Sections \ref{suppsec:Series}--\ref{suppsec:Ten}
-discuss in depth some selected submodules.  Section \ref{suppsec:numsimpl}
-discusses numerical simplification.  Section \ref{suppsec:examples} provides
+Section~\ref{suppsec:Gruntz} discusses the Gruntz algorithm, used to
+calculate limits in the SymPy.  Sections~\ref{suppsec:Series}--\ref{suppsec:Ten}
+discuss in depth some selected submodules.  Section~\ref{suppsec:numsimpl}
+discusses numerical simplification.  Section~\ref{suppsec:examples} provides
 additional examples for topics, discussed in the main paper.  In section

### 2016-09-29 rebuttal.tex
Add the supplement as an external document
Name the section numbers in the reviewer 3 rewrite rebuttal

@@
 \externaldocument[P-]{paper}
+\externaldocument[S-]{supplement}
 
@@
   \item We have renamed the ``Features'' section to ``Overview of
-    Capabilities''.
+    Capabilities'', now section~\ref{P-sec:features}.
   \item We have renamed the ``Domain Specific Submodules'' section to
-      ``Physics Submodule'' (see point~\ref{rev3point30}).
+      ``Physics Submodule'', now section~\ref{P-sec:domain_specific} (see point~\ref{rev3point30}).
   \item We have moved the ``Architecture'' section to the end, after the
-    ``Physics Submodule'' section.
+    ``Physics Submodule'' section. It is now section~\ref{P-sec:architecture}.
   \item We have moved the ``Basic Usage'' subsection to the beginning of the
-    ``Overview of Capabilities'' section, and added an intro paragraph to the
+    ``Overview of Capabilities'' section (it is now section~\ref{P-sec:basic-usage}), and added an intro paragraph to the
     ``Overview of Capabilities'' section.
   \item We have moved the ``Assumptions'' subsection to the ``Overview of
-    Capabilities'' subsection, after the features table.
+    Capabilities'' subsection, after the features table. It is now section~\ref{P-sec:assumptions}.
   \item The ``Comparison with Mathematica'' section has been moved to the end of
-    the supplement.
+    the supplement, now section~\ref{S-suppsec:comp-mma}.
   \item The ``Other Projects that
       Depend on SymPy'' section (which has been renamed from ``Other Projects that
-      Use SymPy'') has been moved from the supplement to the main paper content.
+      Use SymPy'') has been moved from the supplement to the main paper
+      content, now section~\ref{sec:other-proj}.
   \end{enumerate}

### 2016-09-29 rebuttal.tex
Fix a \ref

@@
       Use SymPy'') has been moved from the supplement to the main paper
-      content, now section~\ref{sec:other-proj}.
+      content, now section~\ref{P-sec:other-proj}.
   \end{enumerate}

### 2016-09-29 supplement.tex
Rename \label{comp-mma} to \label{suppsec:comp-mma}
Fix Mathematica section \ref
Rename SymPy Gamma label

@@
 additional examples for topics, discussed in the main paper.  In section
-\ref{sympy-gamma} the SymPy Gamma
-and the SymPy Live projects are introduced.  Finally, section \ref{comp-mma}
+\ref{suppsec:sympy-gamma} the SymPy Gamma
+and the SymPy Live projects are introduced.  Finally, section~\ref{suppsec:comp-mma}
 has a
@@
 
-\section{SymPy Gamma}\label{sympy-gamma}
-
+\section{SymPy Gamma}
+\label{suppsec:sympy-gamma}
 \input{gamma}
@@
 \section{Comparison with Mathematica}
-\label{comp-mma}
+\label{suppsec:comp-mma}
 \input{comparison_with_mma}

### 2016-09-29 rebuttal.tex
Add a x-ref to the Gamma supplement section to the rebuttal

@@
 examples from the paper. A mention of SymPy Gamma has been added to the ``Projects that
-Depends on SymPy'' table (Table~\ref{P-projects-table}) (see below). We have
-left the more in-depth discussion of SymPy Gamma in the supplement.
+Depends on SymPy'' table (Table~\ref{P-projects-table}), now part of the main
+paper (see above). We have
+left the more in-depth discussion of SymPy Gamma in the supplement (section~\ref{S-suppsec:sympy-gamma}).
 

### 2016-09-29 rebuttal.tex
Fix SymPy Live rebuttal
We removed the section entirely.

@@
 
-\item SymPy
-Live, as has been discussed above, is not a SymPy submodule. It is a separate
-project from the SymPy library. It would be inappropriate to list it alongside
-the features of SymPy.
+\item The SymPy Live section was short and would be mostly duplicated by the
+  mention that has been added to the introduction in the paper. We have thus
+  removed it.
 \end{itemize}

### 2016-09-29 rebuttal.tex
Better word
Change "in-depth" to "in depth" everywhere
Except for one place where it was in the reviewer's text.

@@
 paper (see above). We have
-left the more in-depth discussion of SymPy Gamma in the supplement (section~\ref{S-suppsec:sympy-gamma}).
+left the more in depth discussion of SymPy Gamma in the supplement (section~\ref{S-suppsec:sympy-gamma}).
 
@@
 \item The SymPy Live section was short and would be mostly duplicated by the
-  mention that has been added to the introduction in the paper. We have thus
+  paragraph that has been added to the introduction in the paper. We have thus
   removed it.
@@
 core feature of SymPy's architecture, hence, it's inclusion in the conclusion.
-We have also discussed the different points in this sentence in-depth in the architecture section.
+We have also discussed the different points in this sentence in depth in the architecture section.
 \end{solution}

### 2016-09-29 features.tex
Add a section label for the basic usage section
Change "in-depth" to "in depth" everywhere
Except for one place where it was in the reviewer's text.

@@
-% Features to discuss in-depth:
+% Features to discuss in depth:
 This section gives a basic introduction of SymPy, and lists its features.
@@
 \subsection{Basic Usage}
+\label{sec:basic-usage}
 \input{basic_usage.tex}
@@
 
-Although SymPy's extensive feature set cannot be covered in-depth in this
+Although SymPy's extensive feature set cannot be covered in depth in this
 paper, calculus and other bedrock areas are discussed in their own

### 2016-09-29 rebuttal.tex
Punctuation

@@
 6---``Series'', ``Logic'', ``Diophantine Equations'', and ``Statistics'',
-respectively---to be in depth beyond the goals of the main manuscript.
+respectively---to be in depth beyond the goals of the main manuscript:
 
@@
   development and not fully usable. We have added some text to the assumptions
-  section discussing this, and referencing the logic subsection of the
+  section discussing this and referencing the logic subsection of the
   supplement.

### 2016-09-29 rebuttal.tex
Xref a footnote in the rebuttal

@@
 \begin{solution}
-The citation was moved and a footnote~\ref{P-note:prompt} for the Python interactive prompt was
-  added where it was used first time. We have also added a footnote indicating
+The citation was moved and footnote~\ref{P-note:prompt} for the Python interactive prompt was
+  added where it was used first time. We have also added footnote~\ref{P-note:python} indicating
   a moderate familiarity with Python is assumed.

### 2016-09-29 introduction.tex
Xref a footnote in the rebuttal

@@
 Python is a dynamically typed programming language that has a focus on
-ease of use and readability.\footnote{This paper assumes a moderate
+ease of use and readability.\footnote{\label{note:python}This paper assumes a moderate
   familiarity with the Python programming language.} Due in part to this focus, it has become a popular

### 2016-09-29 rebuttal.tex
Fix a typo
Fix some reviewer numbering

@@
 It might, however, encourage bad practice. It may also lead to a poor user experience for newbies.
-\label{rev1point1}
+\label{rev2point1}
 
@@
 We have added a footnote discussing the concerns with \texttt{import *} (see
-reviewer 3, point~\ref{rev1point1}).
+reviewer 2, point~\ref{rev2point1}).
 \end{solution}
@@
 This section was renamed to "Physics submodule".  (See also the
-discussion at~\ref{rev1point37} about module vs package).
+discussion at~\ref{rev3point37} about module vs package).
 \end{solution}
@@
 \begin{solution}
-The word ``package'' replaced with ``submodule'' (see also point~\ref{rev1point37}).
+The word ``package'' replaced with ``submodule'' (see also point~\ref{rev3point37}).
 \end{solution}
@@
 lines 657, 662).
-\label{rev1point37}
+\label{rev3point37}
 \begin{solution}

### 2016-09-29 rebuttal.tex
Grammar
XRef the assumptions footnote to the rebuttal

@@
   section discussing this and referencing the logic subsection of the
-  supplement.
+  supplement (see footnote~\ref{P-note:assumptions}).
 
@@
 \begin{solution}
-See next answer.
+See the next answer.
 \end{solution}
@@
 SymPy by allowing expression interning. It also enables expressions to be
-hashed, that is used to implement caching in SymPy.
+hashed, which is used to implement caching in SymPy.
 \end{quote}

### 2016-09-29 assumptions.tex
Better wording
XRef the assumptions footnote to the rebuttal

@@
 Deductions are made using the Rete
-algorithm~\cite{doorenbos1995production}.\footnote{For historical reasons,
+algorithm~\cite{doorenbos1995production}.\footnote{\label{note:assumptions}For historical reasons,
   this algorithm is distinct from the \texttt{sympy.logic} submodule, which is
-  discussed in the supplementary material. SymPy has an experimental
+  discussed in the supplementary material. SymPy also has an experimental
   assumptions system which stores facts separate from objects, and uses

### 2016-09-29 rebuttal.tex
Missing word

@@
   subclasses of the base class \texttt{Basic}. \texttt{Expr} is the
-\texttt{Basic} subclass for mathematical that can be added and multiplied
-together. The most commonly seen classes in SymPy are subclasses of
-\texttt{Expr}, including \texttt{Add}, \texttt{Mul}, and \texttt{Symbol}.
+  \texttt{Basic} subclass for mathematical objects that can be added and
+  multiplied together. The most commonly seen classes in SymPy are subclasses
+  of \texttt{Expr}, including \texttt{Add}, \texttt{Mul}, and \texttt{Symbol}.
   Instances of \texttt{Expr} typically represent complex numbers, but may also

### 2016-09-29 rebuttal.tex
Logic is also mentioned in the extensibility section

@@
 \item The ``Logic'' section is alluded to in the ``Assumptions'' subsection,
+  and the module is briefly mentioned in the ``Extensibility'' section,
   but is not discussed in depth in the main manuscript. While the logic

### 2016-09-29 rebuttal.tex
Note that we renamed the features section

@@
 characteristic of a software item (e.g., performance, portability,
-or functionality).''  We believe, that this definition fits our needs.
+or functionality).''  We believe that this definition fits our needs. Note
+that we have renamed section (now)~\ref{P-sec:features} to ``Overview of Capabilities''.
 
-Table 1 was extended to include submodule names.
+Table~\ref{P-features-table} was extended to include submodule names.
 \end{solution}

### 2016-09-29 rebuttal.tex
Replace repr quote in rebuttal with cross reference

@@
 \begin{solution}
-This footnote was added:
-\begin{quote}
-  Many Python libraries distinguish the \texttt{str} form of an object, which
-  is meant to be human-readable, and the \texttt{repr} form, which is mean to
-  be valid Python that recreates the object. In SymPy, \texttt{str(expr) ==
-    repr(expr)}. In other words, the string representation of an expression is
-  designed to be compact, human-readable, and valid Python code that could be
-  used to recreate the expression. As it was noted in section~\ref{P-sec:core},
-  the \texttt{srepr} function prints the exact, verbose form of an expression.
-\end{quote}
+See the added footnote~\ref{P-note:repr}.
 \end{solution}

### 2016-09-29 printers.tex
Replace repr quote in rebuttal with cross reference

@@
 expression is valid Python and roughly matches what a user would type to enter
-the expression.\footnote{Many Python libraries distinguish the \texttt{str}
+the expression.\footnote{\label{note:repr}Many Python libraries distinguish the \texttt{str}
   form of an object, which is meant to be human-readable, and the

### 2016-09-29 rebuttal.tex
Cross reference the floating point section in the rebuttal

@@
 \begin{solution}
-We have added an introduction and a subsection ``Floating-Point Numbers''
+We have added an introduction and a subsection~\ref{P-sec:floating-point}, ``Floating-Point Numbers''
 (which, for mathematical clarity, we consider to be distinct from ``real

### 2016-09-29 numerics.tex
Cross reference the floating point section in the rebuttal

@@
 \subsection{Floating-Point Numbers}
+\label{sec:floating-point}
 Floating-point numbers in SymPy are implemented by the \texttt{Float} class,

### 2016-09-29 rebuttal.tex
Complete sentence
Fix quotes and crossref module number

@@
 \begin{solution}
-Mpmath version now mentioned in the introduction.
+The mpmath version is now mentioned in the introduction.
 \end{solution}
@@
 \label{rev3point30}
-This section was renamed to "Physics submodule".  (See also the
+This section (section~\ref{P-sec:domain_specific}) was renamed to ``Physics submodule''.  (See also the
 discussion at~\ref{rev3point37} about module vs package).

### 2016-09-29 rebuttal.tex
Be consistent about "submodule" in the rebuttal

@@
 
-Commutators are implemented in the \texttt{sympy.physics.quantum} module and
+Commutators are implemented in the \texttt{sympy.physics.quantum} submodule and
 are discussed in section~\ref{P-sec:quantum}.
@@
 \begin{itemize}
-\item In SymPy 1.0, the sets module is still in
+\item In SymPy 1.0, the sets submodule is still in
 a relatively infant stage. It is mentioned already in conjunction with the
@@
 
-\item The category theory module in SymPy
-is very domain specific, and independent of other SymPy modules. It is out of
+\item The category theory submodule in SymPy
+is very domain specific, and independent of other SymPy submodules. It is out of
 scope for an architecture paper.
@@
 \item The ``Logic'' section is alluded to in the ``Assumptions'' subsection,
-  and the module is briefly mentioned in the ``Extensibility'' section,
+  and the submodule is briefly mentioned in the ``Extensibility'' section,
   but is not discussed in depth in the main manuscript. While the logic
@@
   mentioned briefly in the ``Conclusion and Future Work'' section, does use
-  it, but we have opted to not discuss this module in depth, as it is still in
+  it, but we have opted to not discuss this submodule in depth, as it is still in
   development and not fully usable. We have added some text to the assumptions
@@
 \texttt{sympy}, such as \texttt{sympy.physics.quantum}. This is to emphasize
-that these modules all live within the \texttt{sympy} namespace, as ``module''
+that these submodules all live within the \texttt{sympy} namespace, as ``module''
 can suggest a separate library.

### 2016-09-29 conclusion_and_future_work.tex
CASs -> CASes
Reviewer 3, comment 3.
Match the conclusion to the current paper sections

@@
 disciplines. This allows SymPy to be used in a first-class way with other
-Python projects, including the scientific Python stack. Unlike many other CASs, SymPy
+Python projects, including the scientific Python stack. Unlike many other CAS's, SymPy
 is designed to be used in an extensible way: both as an end-user
@@
 
-SymPy supports a wide array of mathematical facilities. This includes functions for
-simplifying expressions, performing common calculus operations, pretty printing
+SymPy supports a wide array of mathematical facilities. This includes
+functions for assuming and deducing common mathematical facts,
+simplifying expressions, performing common calculus operations, polynomial
+manipulations, pretty printing
 expressions, solving equations, and representing symbolic matrices. Other supported
@@
 generation, tensors, Lie algebras, cryptography, and special functions.
-Additionally, SymPy contains submodules targeting certain specific domains,
+Additionally, SymPy contains submodules targeting certain specific physics domains,
 such as classical mechanics and quantum mechanics.  This breadth of domains has

### 2016-09-29 conclusion_and_future_work.tex
Rearrange conclusion to match the current order of the paper

@@
 disciplines. This allows SymPy to be used in a first-class way with other
-Python projects, including the scientific Python stack. Unlike many other CAS's, SymPy
-is designed to be used in an extensible way: both as an end-user
-application and as a library.
-
-SymPy expressions are immutable trees of Python objects. SymPy uses Python both
-as the internal language and the user language. This permits users to access to
-the same
-methods that the library implements in order to extend it for their needs.
-Additionally, SymPy has a powerful assumptions
-system for declaring and deducing mathematical properties of expressions.
+Python projects, including the scientific Python stack.
 
@@
 functions for assuming and deducing common mathematical facts,
-simplifying expressions, performing common calculus operations, polynomial
-manipulations, pretty printing
+simplifying expressions, performing common calculus operations, manipulating polynomials, pretty printing
 expressions, solving equations, and representing symbolic matrices. Other supported
@@
 include discrete math, concrete math, plotting, geometry, statistics,
-polynomials, sets, series, vectors, combinatorics, group theory, code
+sets, series, vectors, combinatorics, group theory, code
 generation, tensors, Lie algebras, cryptography, and special functions.
-Additionally, SymPy contains submodules targeting certain specific physics domains,
+SymPy has strong support for arbitrary precision numerics, backed by the
+mpmath package. Additionally, SymPy contains submodules targeting certain specific physics domains,
 such as classical mechanics and quantum mechanics.  This breadth of domains has
 been engendered by a strong and vibrant user community.
-Anecdotally, these users likely chose SymPy because of its ease of access.
+Anecdotally, many of these users chose SymPy because of its ease of access.
+SymPy a dependency of many external projects across a wide
+spectrum of domains.
+
+SymPy expressions are immutable trees of Python objects. Unlike many other
+CAS's, SymPy is designed to be used in an extensible way: both as an end-user
+application and as a library. SymPy uses Python both as the internal language
+and the user language. This permits users to access to the same methods used
+by the library itself in order to extend it for their needs.
 

### 2016-09-29 rebuttal.tex
Match the conclusion to the current paper sections
Update point 39 of the rebuttal

@@
 The order of the subsequent sentences was chosen to match the order they were
-presented in the paper, so we have left them as-is.
+presented in the paper. We have modified them to match the current ordering in
+the paper.
 \end{solution}
@@
 \begin{solution}
-Physics is a very wide subject, so it would be too rough a description to mention
-it in the way you suggest, as a single domain.  Also, we list here only two of
-the above mentioned
-submodules in \texttt{sympy.physics}, but there are others.
-
-We believe the current wording does make sense.
+  We have changed the text to ``certain specific physics domains''. Also, we
+  list here only two of the above mentioned submodules in
+  \texttt{sympy.physics}, but there are others.
 \end{solution}

### 2016-09-29 rebuttal.tex
Clarify that the acknowledgments section was removed

@@
 \begin{solution}
-The institutions's details were moved to the footnotes in the head of the
-paper, and the acknowledgements only thank the institutions.
+  The institution details were moved to the footnotes in the head of the
+  paper. The acknowledgments section was removed, as per the PeerJ staff
+  requested technical changes (this information is to be listed separately as
+  a funding statement deceleration as part of the submission process).
 \end{solution}

### 2016-09-29 gruntz.tex
Remove spaces around em dashes
Cross reference gruntz footnote in rebuttal

--- 2016-08-17 gruntz.tex
@@
 \end{equation}
-First, the set of most rapidly varying subexpressions is determined ---
-the so-called \textit{mrv set}. For~\eqref{gruntz_example_fn}, the mrv set
+First, the set of most rapidly varying subexpressions is determined---the
+so-called \textit{mrv set}. For~\eqref{gruntz_example_fn}, the mrv set
 $\{e^x, e^{-x}, e^{x+2e^{-x}}\}$ is obtained. These are all subexpressions of%
@@
 limit. In the above case the limit is $2$, as can be
-verified by SymPy\footnote{To see intermediate steps, discussed above, interested
-reader can switch on debugging output by setting
-\texttt{SYMPY\_DEBUG} environment variable, before importing anything from
-the SymPy namespace.}:
+verified by SymPy:\footnote{\label{suppnote:gruntz}To see the intermediate steps, discussed above, interested
+readers can switch on debugging output by setting
+the environment variable \texttt{SYMPY\_DEBUG=True}, before importing anything from
+the SymPy namespace.}
 

### 2016-09-29 gamma.tex
Add a citation for Google App Engine
Use sympygamma.com instead of gamma.sympy.org
Shouldn't use colon

@@
-\href{http://gamma.sympy.org}{SymPy Gamma} is a simple web application that runs on Google App Engine.
-It executes and displays the results of SymPy expressions as well as
-additional related computations, in a fashion similar to that of
-Wolfram\textbar{}Alpha. For instance, entering an integer will display
-its prime factors, digits in the base-10 expansion, and a factorization
-diagram. Entering a function will display its docstring; in general,
-entering an arbitrary expression will display its derivative, integral,
-series expansion, plot, and roots.
+\href{http://sympygamma.com}{SymPy Gamma} is a simple web application that
+runs on the Google App Engine~\cite{ciurana2009google}. It executes and
+displays the results of SymPy expressions as well as additional related
+computations, in a fashion similar to that of Wolfram\textbar{}Alpha. For
+instance, entering an integer will display its prime factors, digits in the
+base-10 expansion, and a factorization diagram. Entering a function will
+display its docstring; in general, entering an arbitrary expression will
+display its derivative, integral, series expansion, plot, and roots.
 
@@
   SymPy Gamma displays integration and differentiation steps in detail, which
-  can be viewed in Figure~\ref{fig:integralsteps}:\par
+  can be viewed in Figure~\ref{fig:integralsteps}.\par
   {

### 2016-09-29 rebuttal.tex
Cross reference gruntz footnote in rebuttal
Add link to gamma page for tan(x)

@@
 \begin{solution}
-Footnote was added right before the final example from the Gruntz algorithm section.
+Footnote~\ref{S-suppnote:gruntz} was added right before the final example from the Gruntz algorithm section.
 \end{solution}
@@
 
-A URL to the project site was added.
+A link to the project site was added to the text ``SymPy Gamma''. The text for
+the figure caption now links to the page that generated it.
 \end{solution}

### 2016-09-30 introduction.tex
Better word
Add sentence for projects that depend on SymPy to intro
Wording

@@
 
-As a library, SymPy does not have a built-in graphical user interface (GUI).
+Being a library, SymPy does not have a built-in graphical user interface (GUI).
 However, SymPy exposes a rich interactive display system, and supports
@@
 mechanics and quantum mechanics. Section~\ref{sec:architecture} discusses the
-architecture of SymPy. Conclusions and future directions for SymPy are given
+architecture of SymPy. Section~\ref{sec:other-proj} looks at a selection of
+packages that depend on SymPy. Conclusions and future directions for SymPy are given
 in section~\ref{sec:conclusion}. All examples in this paper use SymPy version
@@
 Google App Engine~\cite{ciurana2009google} to execute SymPy code. SymPy Live
-is also integrated in the SymPy documentation at
+is also integrated into the SymPy documentation at
 \href{http://docs.sympy.org}{http://docs.sympy.org}.

### 2016-09-30 architecture.tex
Missing word
Move immutability paragraph to the architecture section

@@
 
+Importantly, SymPy expressions are immutable. This simplifies the design of
+SymPy by allowing expression interning. It also enables expressions to be
+hashed, which is used to implement caching in SymPy.
+
 Python allows classes to override mathematical operators. The Python
@@
 subclasses of the base class \texttt{Basic}. \texttt{Expr} is the
-\texttt{Basic} subclass for mathematical that can be added and multiplied
-together. The most commonly seen classes in SymPy are subclasses of
+\texttt{Basic} subclass for mathematical objects that can be added and
+multiplied together. The most commonly seen classes in SymPy are subclasses of
 \texttt{Expr}, including \texttt{Add}, \texttt{Mul}, and \texttt{Symbol}.

### 2016-09-30 features.tex
Define "bedrock areas"

@@
 Although SymPy's extensive feature set cannot be covered in depth in this
-paper, calculus and other bedrock areas are discussed in their own
-subsections. Additionally, Table~\ref{features-table} gives a compact listing
+paper, bedrock areas, that is, those areas that are used throughout the
+library, are discussed in their own
+subsections below. Additionally, Table~\ref{features-table} gives a compact listing
 of all major capabilities present in the SymPy codebase. This grants a

### 2016-09-30 features.tex
Add sympy.calculus to the list of calculus submodules

@@
 \midrule
-Calculus (\texttt{sympy.core}, \texttt{sympy.series}, \texttt{sympy.integrals}) & Algorithms for computing derivatives, integrals, and limits.\\
+Calculus (\texttt{sympy.core}, \texttt{sympy.calculus}, \texttt{sympy.integrals}, \texttt{sympy.series}) & Algorithms for computing derivatives, integrals, and limits.\\
 

### 2016-09-30 assumptions.tex
Use "integer" instead of "integral" (same word the assumptions system uses)

@@
 common mathematical properties, such as being positive, imaginary, or
-integral. SymPy is careful to never perform simplifications on an expression
+integer. SymPy is careful to never perform simplifications on an expression
 unless the assumptions allow them. For instance, the identity $\sqrt{t^2} = t$

### 2016-09-30 assumptions.tex
Can we fix this overflowing text?
Maybe we should just leave it and let the journal figure it out.
Grammar
Parenthetical

@@
 
+% TODO: What can we do about the overflowing text here?
 Some of the common assumptions that SymPy allows are \texttt{positive},
@@
   real=True).is\_positive} will give \texttt{None} because a real symbol might
-be positive or negative. The \texttt{None} could also mean that not enough is
+be positive or negative. \texttt{None} could also mean that not enough is
 known or implemented to compute the given fact. For instance, \texttt{(pi +
@@
 \texttt{y} are both created with \texttt{positive=True}, then
-\texttt{(x + y).is\_positive} will be \texttt{True} whereas
-\texttt{(x - y).is\_positive} will be \texttt{None}.
+\texttt{(x + y).is\_positive} will be \texttt{True} (whereas
+\texttt{(x - y).is\_positive} will be \texttt{None}).

### 2016-09-30 simplification.tex
reviewer 1 point 2 (Carette's citation replaced)
Don't use longtable for the simplification table
It should only be used for tables that can't fit on one page.

@@
 It must be emphasized that simplification is not a rigorously defined
-mathematical operation~\cite{Carette2004understanding}.
+mathematical operation~\cite{Moses1971simplification}.
 The \texttt{simplify} function applies several simplification routines along
@@
 
-\begin{longtable}[htbc]{lp{0.83\linewidth}}
-\caption{Some SymPy Simplification Functions\label{simplify-table}}\\
+\begin{center}
+\begin{table}[htbc]
+\caption{Some SymPy Simplification Functions}\label{simplify-table}
+\begin{tabular}{lp{0.83\linewidth}}
 \toprule
@@
 \bottomrule
-\end{longtable}
+\end{tabular}
+\end{table}
+\end{center}

### 2016-09-30 calculus.tex
Use exp instead of E** in the limit example
Wording

@@
 \begin{verbatim}
->>> limit((2*E**((1-cos(x))/sin(x))-1)**(sinh(x)/atan(x)**2), x, 0)
+>>> limit((2*exp((1-cos(x))/sin(x))-1)**(sinh(x)/atan(x)**2), x, 0)
 E
@@
 
-Summations are computed with \verb|summation|  using a combination of Gosper's
-algorithm~\cite{gosper1978decision}, an algorithm that uses Meijer
-G-functions~\cite{Roach1996hyper,roach1997meijerg}, and heuristics. Products
-are computed with \verb|product| function via a suite of heuristics.
+Summations are computed with the \verb|summation| function, which uses a
+combination of Gosper's algorithm~\cite{gosper1978decision}, an algorithm that
+uses Meijer G-functions~\cite{Roach1996hyper,roach1997meijerg}, and
+heuristics. Products are computed with \verb|product| function via a suite of
+heuristics.
 % TODO: Are there other summation algorithms implemented?

### 2016-09-30 matrices.tex
Remove incorrect "and"

@@
 matrix addition, multiplication, exponentiation, computing determinants,
-solving linear systems, and computing inverses using LU decomposition, LDL
+solving linear systems, computing inverses using LU decomposition, LDL
 decomposition, Gauss-Jordan elimination, Cholesky decomposition, Moore-Penrose

### 2016-09-30 matrices.tex
Clarify matrices features list
I think the last items are all supposed to be related to computing inverses
(except for singular values, which was added later).

@@
 matrix addition, multiplication, exponentiation, computing determinants,
-solving linear systems, computing inverses using LU decomposition, LDL
-decomposition, Gauss-Jordan elimination, Cholesky decomposition, Moore-Penrose
-pseudoinverse, singular values, and adjugate matrix.
+solving linear systems, singular values, and computing inverses using LU
+decomposition, LDL decomposition, Gauss-Jordan elimination, Cholesky
+decomposition, Moore-Penrose pseudoinverse, or adjugate matrices.
 

### 2016-09-30 matrices.tex
Clarify that Dictionary of Keys uses a Python dictionary

@@
 their elements in Dictionary of Keys (DOK) format, meaning entries are stored
-as \texttt{(row, column)} pairs mapping to the elements.
+as a \texttt{dict} of \texttt{(row, column)} pairs mapping to the elements.
 

### 2016-09-30 domain_specific.tex
Add automatic cross-references to the paper (in rebuttal)
See more at:
http://tex.stackexchange.com/questions/3060/referencing-labels-from-other-files

XXX: Section and footnotes numbers are correct, but my pdf viewer
(evince) doesn't open a new document if I click on them.  I hope,
this is a bug.
"domain specific" to "domain specific physics"
Clearer wording

@@
 SymPy includes several submodules that allow users to solve domain specific
-problems. For example, a comprehensive physics submodule is included that is
-useful for solving problems in mechanics, optics, and quantum
+physics problems. For example, a comprehensive physics submodule is included
+that is useful for solving problems in mechanics, optics, and quantum
 mechanics along with support for manipulating physical quantities with units.
@@
 operations such as addition, subtraction, scalar multiplication, inner and
-outer products, and cross products. The vector and dyadic objects can both be
+outer products, and cross products. The vector and dyadic objects both can be
 written in very compact notation that make it easy to express the vectors and
@@
 \subsection{Quantum Mechanics}
+\label{sec:quantum}
 

### 2016-09-30 domain_specific.tex
It should be Wigner D-matrix

@@
 eigenstates can be represented in any basis and for any quantum numbers.
-A rotation operator representing the Wigner-D matrix, which may be defined
+A rotation operator representing the Wigner D-matrix, which may be defined
 symbolically or numerically, is also implemented to rotate spin eigenstates.

### 2016-09-30 architecture.tex
Reword
Clearer wording

@@
 \label{sec:core}
-A computer algebra system stores mathematical expressions as data
-structures.  For example, the mathematical expression $x + y$ is represented as
-a tree with three nodes, $+$, $x$, and $y$, where $x$ and $y$ are ordered
-children of $+$.  As users manipulate
-mathematical expressions with traditional mathematical syntax, the CAS
-manipulates the underlying data structures.  Automated optimizations and
+A computer algebra system stores mathematical expressions as data structures.
+For example, the mathematical expression $x + y$ is represented as a tree with
+three nodes, $+$, $x$, and $y$, where $x$ and $y$ are ordered children of $+$.
+As users manipulate mathematical expressions with traditional mathematical
+syntax, the CAS manipulates the underlying data structures. Symbolic
 computations such as integration, simplification, etc.\ are all functions that
@@
 
-Importantly, SymPy expressions are immutable. This simplifies the design of
-SymPy by allowing expression interning. It also enables expressions to be
-hashed, which is used to implement caching in SymPy.
+Another important property of SymPy expressions is that they are immutable.
+This simplifies the design of SymPy by allowing expression interning. It also
+enables expressions to be hashed, which is used to implement caching in SymPy.
 

### 2016-09-30 architecture.tex
Remove ambiguity that could make it sound like Basic is part of Python

@@
 
-In SymPy every symbolic expression is an instance of a Python \texttt{Basic}
+In SymPy every symbolic expression is an instance of the class \texttt{Basic}
 class,\footnote{Some

### 2016-09-30 architecture.tex
Remove redundant word

--- 2016-09-30 architecture.tex
@@
 
-In SymPy every symbolic expression is an instance of a Python \texttt{Basic}
-class,\footnote{Some
-  internal classes, such as those used in the polynomial submodule, do not follow
-  this rule for efficiency reasons.} a superclass of all SymPy types providing
-common methods to all SymPy tree-elements, such as traversals.  The children of a node in the
-tree are held in the \texttt{args} attribute.  A terminal or leaf node in the
-expression tree has empty \texttt{args}.
+In SymPy every symbolic expression is an instance of the class
+\texttt{Basic},\footnote{Some internal classes, such as those used in the
+  polynomial submodule, do not follow this rule for efficiency reasons.} a
+superclass of all SymPy types providing common methods to all SymPy
+tree-elements, such as traversals. The children of a node in the tree are held
+in the \texttt{args} attribute. A terminal or leaf node in the expression tree
+has empty \texttt{args}.
 

### 2016-09-30 architecture.tex
Definite article

@@
 \texttt{Basic},\footnote{Some internal classes, such as those used in the
-  polynomial submodule, do not follow this rule for efficiency reasons.} a
+  polynomial submodule, do not follow this rule for efficiency reasons.} the
 superclass of all SymPy types providing common methods to all SymPy

### 2016-09-30 architecture.tex
SymPy never uses the phrase "terminal node"
Remove "so"

@@
 tree-elements, such as traversals. The children of a node in the tree are held
-in the \texttt{args} attribute. A terminal or leaf node in the expression tree
+in the \texttt{args} attribute. A leaf node in the expression tree
 has empty \texttt{args}.
@@
 By order of operations, the parent of the expression tree for \texttt{expr} is
-an addition, so it is of type \texttt{Add}. The child nodes of \texttt{expr} are
+an addition. It is of type \texttt{Add}. The child nodes of \texttt{expr} are
 \texttt{2} and \texttt{x*y}.

### 2016-09-30 architecture.tex
"other examples" (one was just shown)
Hashing is useful for any dict, not just caching

@@
 \end{verbatim}
-Symbols or symbolic constants, like $e$ or $\pi$, are examples of
+Symbols or symbolic constants, like $e$ or $\pi$, are other examples of
 leaf nodes.
@@
 This simplifies the design of SymPy by allowing expression interning. It also
-enables expressions to be hashed, which is used to implement caching in SymPy.
+enables expressions to be hashed, which allows expressions to be used as keys
+in Python dictionaries, and is used to implement caching in SymPy.
 

### 2016-09-30 architecture.tex
The design isn't just simplified by interning
Add another reference to the supplement

@@
 Another important property of SymPy expressions is that they are immutable.
-This simplifies the design of SymPy by allowing expression interning. It also
+This simplifies the design of SymPy, and enables expression interning. It also
 enables expressions to be hashed, which allows expressions to be used as keys
@@
 subclasses of \texttt{Expr}. For instance, logic expressions, such as
-\verb|And(x, y)|, are subclasses of \texttt{Basic} but not of \texttt{Expr}.
+\verb|And(x, y)|, are subclasses of \texttt{Basic} but not of
+\texttt{Expr}.\footnote{See the supplement for more information on the
+  \texttt{sympy.logic} submodule.}
 

### 2016-09-30 architecture.tex
Better wording

@@
 like $f(x)$. Subclasses of \texttt{Function} should define a
-class method \texttt{eval}, which returns a canonical form of the function
-application (usually an instance of some other class, i.e., a \texttt{Number})
-or \texttt{None}, if for given arguments that function should not be
+class method \texttt{eval}, which returns an evaluated value for the function
+application (usually an instance of some other class, i.e., a \texttt{Number}),
+or \texttt{None} if for given arguments that function should not be
 automatically evaluated.

### 2016-09-30 architecture.tex
e.g., not i.e.

--- 2016-09-30 architecture.tex
@@
 like $f(x)$. Subclasses of \texttt{Function} should define a
-class method \texttt{eval}, which returns a canonical form of the function
-application (usually an instance of some other class, i.e., a \texttt{Number})
-or \texttt{None}, if for given arguments that function should not be
+class method \texttt{eval}, which returns an evaluated value for the function
+application (usually an instance of some other class, e.g., a \texttt{Number}),
+or \texttt{None} if for given arguments that function should not be
 automatically evaluated.

### 2016-09-30 architecture.tex
Better wording

--- 2016-09-30 architecture.tex
@@
 like $f(x)$. Subclasses of \texttt{Function} should define a
-class method \texttt{eval}, which returns a canonical form of the function
-application (usually an instance of some other class, i.e., a \texttt{Number})
-or \texttt{None}, if for given arguments that function should not be
+class method \texttt{eval}, which returns an evaluated value for the function
+application (usually an instance of some other class, e.g., a \texttt{Number}),
+or \texttt{None} if for the given arguments it should not be
 automatically evaluated.

### 2016-09-30 architecture.tex
Grammar

@@
 \texttt{diff} on its \texttt{args}. (Subclasses of \texttt{Function} should
-implement \texttt{fdiff} method instead, it returns the derivative of the function
+implement the \texttt{fdiff} method instead; it returns the derivative of the function
 without considering the chain rule.) The most common

### 2016-09-30 architecture.tex
Better wording

@@
 
-As an example of the notions presented in this section,
-Listing~\ref{fig:gamma-example} presents a minimal version of the gamma
-function $\Gamma(x)$ from SymPy, which evaluates itself on positive integer
-arguments, has the real assumption defined, can be rewritten in
-terms of factorial with \verb|gamma(x).rewrite(factorial)|, and can be
-differentiated. \texttt{self.func} is used throughout
-instead of referencing \texttt{gamma} explicitly so that potential subclasses
-of \texttt{gamma} can reuse the methods.
+Listing~\ref{fig:gamma-example} presents an example of extensibility in SymPy.
+It gives a minimal version of the gamma function $\Gamma(x)$ from SymPy, which
+evaluates itself on positive integer arguments, has the real assumption
+defined, can be rewritten in terms of factorial with
+\verb|gamma(x).rewrite(factorial)|, and can be differentiated.
+\texttt{self.func} is used throughout instead of referencing \texttt{gamma}
+explicitly so that potential subclasses of \texttt{gamma} can reuse the
+methods.
 

### 2016-09-30 architecture.tex
Better wording

@@
 
-Listing~\ref{fig:gamma-example} presents an example of extensibility in SymPy.
-It gives a minimal version of the gamma function $\Gamma(x)$ from SymPy, which
-evaluates itself on positive integer arguments, has the real assumption
-defined, can be rewritten in terms of factorial with
-\verb|gamma(x).rewrite(factorial)|, and can be differentiated.
-\texttt{self.func} is used throughout instead of referencing \texttt{gamma}
-explicitly so that potential subclasses of \texttt{gamma} can reuse the
-methods.
+Listing~\ref{fig:gamma-example} presents an example of this extensibility. It
+gives a stripped down version of the \texttt{gamma} function $\Gamma(x)$ from
+SymPy. The methods defined allow it to evaluate itself on positive integer
+arguments, define the real assumption, allow it to be rewritten in terms of
+factorial (with \verb|gamma(x).rewrite(factorial)|), and allow it to be
+differentiated. \texttt{self.func} is used throughout instead of referencing
+\texttt{gamma} explicitly so that potential subclasses of \texttt{gamma} can
+reuse the methods.
 

### 2016-09-30 architecture.tex
Sort the imports

@@
 \begin{lstlisting}[caption={A minimal implementation of \texttt{sympy.gamma}.},label=fig:gamma-example]
-from sympy import Integer, Function, floor, factorial, polygamma
+from sympy import floor, Function, Integer, factorial, polygamma
 

### 2016-09-30 architecture.tex
Fix syntax and unnecessary import in gamma example

@@
 \begin{lstlisting}[caption={A minimal implementation of \texttt{sympy.gamma}.},label=fig:gamma-example]
-from sympy import floor, Function, Integer, factorial, polygamma
+from sympy import Function, Integer, factorial, polygamma
 
-class gamma(Function)
+class gamma(Function):
     @classmethod

### 2016-09-30 architecture.tex
"fine-grained"
I think these should be two sentences (previous wording did not make sense)

@@
 characteristics for symbolic manipulation. SymEngine is a pure C++ library,
-which allows it fine grain control over the memory layout of expressions.
+which allows it fine-grained control over the memory layout of expressions.
 SymEngine has thin wrappers to other languages (Python, Ruby,
-Julia, etc.) whose aim is to be the fastest symbolic manipulation library. Preliminary
+Julia, etc.). Its aim is to be the fastest symbolic manipulation library. Preliminary
 benchmarks suggest that SymEngine is as performant as its commercial and

### 2016-09-30 architecture.tex
Make it clear that SymEngine is not used in SymPy 1.0

@@
 
-SymPy has recently started to use SymEngine as an
+The development branch SymPy has recently started to use SymEngine as an
 optional backend, initially in \texttt{sympy.\allowbreak{}physics.\allowbreak{}mechanics} only.

### 2016-09-30 architecture.tex
Fix wording

--- 2016-09-30 architecture.tex
@@
 
-SymPy has recently started to use SymEngine as an
+The development version of SymPy has recently started to use SymEngine as an
 optional backend, initially in \texttt{sympy.\allowbreak{}physics.\allowbreak{}mechanics} only.

### 2016-09-30 projects_that_depend_on_sympy.tex
"module" -> "package" (based on what we said in the rebuttal)
The other ones say "Python package" so this one should too

@@
 \href{https://github.com/brombo/galgebra}{\textbf{galgebra}} &
-  A module for geometric algebra (previously \texttt{sympy.galgebra})~\cite{galgebra}. \\
+  A Python package for geometric algebra (previously \texttt{sympy.galgebra})~\cite{galgebra}. \\
 

### 2016-09-30 assumptions.tex
Fix overflow (assumptions)

@@
 
-% TODO: What can we do about the overflowing text here?
-Some of the common assumptions that SymPy allows are \texttt{positive},
+Some of the common assumptions are
 \texttt{negative}, \texttt{real}, \texttt{nonpositive}, \texttt{integer},
@@
   algorithms in SymPy require special consideration to work correctly with
-  noncommutative products.} Assumptions on any object can be checked with the
+  noncommutative products.} Assumptions on any SymPy object can be checked with the
 \verb|is_|\texttt{\textit{assumption}} attributes, like \verb|t.is_positive|.

### 2016-09-30 supplement.tex
Fix supplement intro for SymPy Gamma
Add color to the links in the paper and supplement
What do people think? We can remove this if people don't like it. I wanted to
add it because some things that were links were subtle, and probably wouldn't
be noticed by most readers.
Unnecessary comma

@@
 \usepackage{url}
-\usepackage{hyperref}
 
-% this is required for all the \url{} commands in the bib file
-%\usepackage{hyperref}
+\hypersetup{
+  colorlinks=true,
+  urlcolor=blue,
+  citecolor=black
+}
 
@@
 discusses numerical simplification.  Section~\ref{suppsec:examples} provides
-additional examples for topics, discussed in the main paper.  In section
+additional examples for topics discussed in the main paper.  In section
 \ref{suppsec:sympy-gamma} the SymPy Gamma
-and the SymPy Live projects are introduced.  Finally, section~\ref{suppsec:comp-mma}
+project is discussed.  Finally, section~\ref{suppsec:comp-mma}
 has a

### 2016-09-30 gruntz.tex
Grammar, and use \eqref

@@
 the relation $f(x)\sim {1\over f(x)}$ can be used.  In the
-considered case, only item $\omega=e^{-x}$ can be accepted.
+considered case, only the item $\omega=e^{-x}$ can be accepted.
 
@@
 Then the rewritten subexpressions are
-substituted back into $f(x)$ in~(\ref{gruntz_example_fn})
+substituted back into $f(x)$ in~\eqref{gruntz_example_fn}
 and the result is expanded with respect to~$\omega$:
@@
 limit. In the above case the limit is $2$, as can be
-verified by SymPy:\footnote{\label{suppnote:gruntz}To see the intermediate steps, discussed above, interested
+verified by SymPy:\footnote{\label{suppnote:gruntz}To see the intermediate steps discussed above, interested
 readers can switch on debugging output by setting
@@
 equal to $\lim\limits_{x\to\infty} C_0(x)$. The expression $C_0(x)$ is always
-simpler than original $f(x)$, same is true for limits, arising in the
-rewrite stage~(\ref{gruntz_rewrite}), so the algorithm converges. A proof of this and further
+simpler than original $f(x)$, and the same is true for limits arising in the
+rewrite stage~\eqref{gruntz_rewrite}, so the algorithm converges. A proof of this and further
 details on the algorithm are given in Gruntz's PhD thesis~\cite{Gruntz1996limits}.

### 2016-09-30 diophantine.tex
Consistent non-usage of commas in symbols()
List the unknown variables for completeness

@@
 SymPy's Diophantine submodule ($a_1, \ldots, a_{n+1}$, $a$, $b$, $c$, $d$, $e$,
-$f$, and $k$ are explicitly given rational constants):
+$f$, and $k$ are explicitly given rational constants, $x_1, \ldots, x_{n+1}$,
+$x$, $y$, and $z$ are unknown variables):
 
@@
 
->>> x1, x2, x3, x4, x5, x6 = symbols('x1, x2, x3, x4, x5, x6')
+>>> x1, x2, x3, x4, x5, x6 = symbols('x1 x2 x3 x4 x5 x6')
 >>> diophantine(9*x1**2 + 16*x2**2 + x3**2 + 49*x4**2 + 4*x5**2 - 25*x6**2)

### 2016-09-30 sets.tex
Wording

%% Sets

SymPy supports representation of a wide variety of mathematical sets. This is
achieved by first defining abstract representations of atomic set classes and
then combining and transforming them using various set operations.

Each of the set classes inherits from the base class \texttt{Set} and defines
methods to check membership and calculate unions, intersections, and set
differences. When these methods are not able to evaluate to atomic set
classes, they are represented as abstract unevaluated objects.

SymPy has the following atomic set classes:

\begin{itemize}

    \item \verb|EmptySet| represents the empty set $\emptyset$.

    \item \verb|UniversalSet| is an abstract ``universal set'' of which
      everything is a member. The union of the universal set with any set
      gives the universal set and the intersection gives the other set itself.

    \item \verb|FiniteSet| is functionally equivalent to Python's built
      in \texttt{set} object. Its members can be any SymPy object including
      other sets.

    \item \verb|Integers| represents the set of integers $\mathbb{Z}$.

    \item \verb|Naturals| represents the set of natural numbers $\mathbb{N}$,
      i.e., the set of positive integers.

    \item \verb|Naturals0| represents the set of whole numbers $\mathbb{N}_0$,
      which are all the non-negative integers.

    \item \verb|Range| represents a range of integers. A range is defined by
      specifying a start value, an end value, and a step size. The enumeration
      of a \texttt{Range} object is functionally equivalent to Python's
      \texttt{range} except it supports infinite endpoints, allowing the
      representation of infinite ranges.

    \item \verb|Interval| represents an interval of real numbers. It is
      defined by giving the start and the end points and by specifying if the interval is open
      or closed on the respective ends.

\end{itemize}

%% Operations

In addition to unevaluated classes for the basic \texttt{Union},
\texttt{Intersection}, and \texttt{Complement} set operations, SymPy has the
following set classes.

\begin{itemize}

    \item \verb|ProductSet| defines the Cartesian product of two
        or more sets. The product set is useful when representing higher
        dimensional spaces. For example, to represent a three-dimensional space,
        SymPy uses the Cartesian product of three real sets.

      \item \verb|ImageSet| represents the image of a function when applied to
        a particular set. The image set of a function $F$ with respect to a set
        $S$ is $\{ F(x) \mid x \in S \}$. SymPy uses image sets to represent sets
        of infinite solutions of equations such as $\sin(x)=0$.

      \item \verb|ConditionSet| represents a subset of a set whose members
        satisfy a particular condition. The subset of set $S$
        given by the condition $H$ is $\{x \mid H(x), x \in S \}$. SymPy uses
        condition sets to represent the set of solutions of equations and
        inequalities, where the equation or the inequality is the condition and
        the set is the domain over which it is being solved.

\end{itemize}

A few other classes are implemented as special cases of the classes described
above. The set of real numbers, \verb|Reals|, is implemented as a special case
of \verb|Interval|. \verb|ComplexRegion|
is implemented as a special case of \verb|ImageSet|. \verb|ComplexRegion|
supports both polar and rectangular representation of regions on the complex
plane.

### 2016-09-30 categories.tex
Add a citation for Xy-pic
Fix typo

@@
 Automatic typesetting of diagrams takes a \texttt{Diagram} and produces \LaTeX{}
-code using the \texttt{Xy-pic} package.  Typesetting is done in two stages:
+code using the \texttt{Xy-pic} package~\cite{rose1999xy}.  Typesetting is done in two stages:
 layout and generation of \texttt{Xy-pic} code.  The layout stage is taken care
@@
 diagram automatically typeset by \texttt{DiagramgGrid} and
-\texttt{XypicDiagramDrawer} in given in Figure~\ref{fig:cat:loops}.
+\texttt{XypicDiagramDrawer} is given in Figure~\ref{fig:cat:loops}.
 \begin{figure}[h]

### 2016-09-30 gamma.tex
Add link to gamma page for tan(x)
Wording

@@
 \item
-  SymPy Gamma displays integration and differentiation steps in detail, which
-  can be viewed in Figure~\ref{fig:integralsteps}.\par
+  SymPy Gamma displays integration and differentiation steps in detail, as
+  demonstrated in Figure~\ref{fig:integralsteps}.\par
   {
@@
     \includegraphics[width=0.7\textwidth]{integral_steps.png}
-    \captionof{figure}{Integral steps of $\tan (x)$}\label{fig:integralsteps}
+    \captionof{figure}{\href{http://www.sympygamma.com/input/?i=tan(x)}{Integral steps of $\tan (x)$}}\label{fig:integralsteps}
     \par

### 2016-09-30 gamma.tex
Don't mention the Wolfram|Alpha paid features
Use MathJax instead of LaTeX for SymPy Gamma

@@
 \item
-  SymPy Gamma saves user search queries, and offers many such similar features
-  for free, which Wolfram\textbar{}Alpha only offers to its paid users.
+  SymPy Gamma saves user search queries.
 \end{itemize}
@@
 This parser converts the input query to the equivalent SymPy readable code,
-which is then processed by SymPy, and the result is finally printed
-with the built-in \LaTeX{} output and rendered by the SymPy Gamma web application.
+which is then processed by SymPy, and the result is finally printed with the
+built-in MathJax~\cite{cervone2012mathjax} output and rendered by the SymPy
+Gamma web application.

### 2016-09-30 comparison_with_mma.tex
I disagree that ** for exponentiation is "unlike the usual programming convention"
Several languages use this, not just Python. It's also quite common even for
languages that don't have an exponentiation operator for ^ to mean xor, not
exponentiation.

@@
 
-SymPy, in accordance with Python (and unlike the usual programming convention),
+SymPy, in accordance with Python,
 uses \texttt{**} to express the power operator, while Mathematica uses the more

### 2016-09-30 comparison_with_mma.tex
Grammar and wording

@@
 attributes, though other programming patterns may also be used to achieve an
-equivalent behavior such as class inheritance.
+equivalent behavior such, as class inheritance.
 
@@
 Unlike SymPy, Mathematica's expressions are mutable:
-one can change parts of the expression tree without the need of
-creating a new object.
+one can change parts of the expression tree without the needing to
+create a new object.
 The mutability of Mathematica expressions allows for a lazy updating of any references
@@
 Regarding associative expressions,
-SymPy handles associativity of sums and products by automatically flattening them,
+SymPy handles associativity of sums and products by automatically flattening them.
 Mathematica specifies the \texttt{Flat} attribute on the expression type.
@@
 are Python class constructors
-and cannot be modified at runtime.\footnote{Nonetheless, Python supports monkey
-patching but it is a discouraged programming pattern.}
+and cannot be modified at runtime.\footnote{Python does support monkey
+patching, but it is a discouraged programming pattern.}
 The way SymPy deals with extending the missing runtime overloadability functionality

### 2016-10-02 introduction.tex
I'm sorry, but you can't use verb use here.
Ondrej told me that "use" - means "depend on".  Please be consistent
and don't break your own rules.

@@
 computing and data science, with a broad ecosystem of
-libraries~\cite{oliphant2007python}. SymPy is itself used by many libraries
+libraries~\cite{oliphant2007python}. SymPy is itself used as a dependency
+by many libraries
 and tools to support research within a variety of domains, such as

### 2016-10-02 comparison_with_mma.tex
or -> "overloadability of the functionality" ?

@@
 patching, but it is a discouraged programming pattern.}
-The way SymPy deals with extending the missing runtime overloadability functionality
+The way SymPy deals with extending the missing runtime overloadability
 is by subclassing the node types: subclasses may redefine the class constructor

### 2016-10-05 matrices.tex

@@
 semi-definiteness, orthogonality, etc., which are valuable in the construction
-of numerical linear algebra systems.
+of numerical linear algebra systems~\cite{rocklin2013mathematically}.
 

### 2016-10-05 numerics.tex

@@
 Symbolic integration and summation often produce hypergeometric
-and Meijer G-function closed forms (see Subsection~\ref{sec:calculus});
+and Meijer G-function closed forms (see section~\ref{sec:calculus});
 numerical evaluation of such special functions is a useful complement

### 2016-10-05 polys.tex

% Polynomials

SymPy implements a suite of algorithms for polynomial manipulation,
which ranges from relatively simple algorithms for doing arithmetic of
polynomials, to advanced methods for factoring multivariate polynomials
into irreducibles, symbolically determining real and complex root isolation
intervals, or computing Gr\"{o}bner bases.

Polynomial manipulation is useful in its own right. Within SymPy, though, it is mostly used
indirectly as a tool in other areas of the library. In fact, many mathematical
problems in symbolic computing are first expressed using entities from the
symbolic core, preprocessed, and then transformed into a problem in the
polynomial algebra, where generic and efficient algorithms are used to solve
the problem.  The solutions to the original problem are subsequently recovered from the results.
This is a common scheme in symbolic integration or summation algorithms.

SymPy implements dense and sparse polynomial representations.\footnote{In a
dense representation, the coefficients for all terms up to the degree of each
variable are stored in memory. In a sparse representation, only the nonzero
coefficients are stored.} Both are used in the univariate and multivariate
cases. The dense representation is the default for univariate polynomials. For
multivariate polynomials, the choice of representation is based on the
application. The most common case for the sparse representation is algorithms
for computing Gr\"{o}bner bases (Buchberger, F4, and
F5)~\cite{Buchberger1965thesis,Faugere1999f4,Faugere2002f5}. This is because
different monomial orderings can be expressed easily in this representation.
However, algorithms for computing multivariate GCDs or factorizations, at
least those currently implemented in SymPy~\cite{paprocki2010thesis},
are better expressed when the representation is dense. The dense multivariate
representation is specifically a recursively-dense representation, where
polynomials in $K[x_0, x_1, \dotsc, x_n]$ are viewed as a polynomials in
$K[x_0][x_1]\dotso[x_n]$. Note that despite this, the coefficient domain $K$,
can be a multivariate polynomial domain as well. The dense recursive
representation in Python gets inefficient as the number of variables increases.

Some examples for the \texttt{sympy.polys} submodule can be found in the
supplement.

### 2016-10-05 stats.tex
Capitalization

@@
 The \verb|sympy.stats| submodule provides random variable types and methods for
-computing of statistical properties of expressions involving random
+computing statistical properties of expressions involving random
 variables, which can be either continuous or discrete, the latter ones being
@@
 in terms of probability spaces.
-Apart from the possibility of defining the random variables from user supplied
+Apart from the possibility of defining the random variables from a user supplied
 density distribution, SymPy provides definitions of most common
@@
 %Using the above-mentioned features, it is easy to calculate, for example
-the well known properties of maxwellian velocity distribution
+the well known properties of the Maxwellian velocity distribution.
 \begin{verbatim}

### 2016-10-05 authors/prettify.py
Allow authors/prettify.py to be run from any directory

@@ -2,10 +2,12 @@
 
 from __future__ import print_function
 
+from os.path import join, dirname
 from json import load, dumps
 from io import open
 
-author_list = load(open("authors.json"))
+authors_json = join(dirname(__file__), "authors.json")
+author_list = load(open(authors_json))
 for author in author_list:
     if "author_order" not in author:
         author["author_order"] = author["sympy_commits"]
@@ -14,7 +16,7 @@ for author in author_list:
     if author["author_order"] == author["sympy_commits"]:
         del author["author_order"]
 
-with open("authors.json", "w", encoding='utf-8') as f:
+with open(authors_json, "w", encoding='utf-8') as f:
     data = dumps(author_list, ensure_ascii=False, sort_keys=True, indent=4,
             separators=(',', ': '))
     f.write(data + "\n")

### 2016-10-05 authors/list_latex.py
Allow authors/list_latex.py to be run from any directory

@@ -2,13 +2,17 @@
 
 from __future__ import print_function
 
+
+from os.path import join, dirname, pardir
 from json import load
 from io import open
 
-author_list = load(open("authors.json"))
+authors_json = join(dirname(__file__), "authors.json")
+authors_tex = join(dirname(__file__), pardir, "authors.tex")
+author_list = load(open(authors_json))
 author_list = list(filter(lambda x: x["sympy_commits"] > 0, author_list))
 
-with open("../authors.tex", "w", encoding='utf-8') as f:
+with open(authors_tex, "w", encoding='utf-8') as f:
     for n, author in enumerate(author_list):
         f.write((u"\\author[%d]{%s}%%\n" % (n+1, author["name"])))
     for n, author in enumerate(author_list):

### 2016-10-05 authors/list_latex.py
Update authors/list_latex.py to use the peerj address

@@ -22,6 +22,6 @@ with open(authors_tex, "w", encoding='utf-8') as f:
             institution_explanation = u""
         f.write(u"\\affil[%d]{%s, %s (\\email{%s}).%s}%%\n" \
                 % (n+1, author["institution"],
-                    author["institution_address_siam"],
+                    author["institution_address_peerj"],
                     author["email"],
                     institution_explanation))

### 2016-10-05 authors/list_latex.py
Don't duplicate the institution

@@ -20,8 +20,8 @@ with open(authors_tex, "w", encoding='utf-8') as f:
             institution_explanation = u" %s" % author["institution_explanation"]
         else:
             institution_explanation = u""
-        f.write(u"\\affil[%d]{%s, %s (\\email{%s}).%s}%%\n" \
-                % (n+1, author["institution"],
-                    author["institution_address_peerj"],
-                    author["email"],
-                    institution_explanation))
+        f.write(u"\\affil[%d]{%s (\\email{%s}).%s}%%\n" \
+                % (n+1,
+                   author["institution_address_peerj"],
+                   author["email"],
+                   institution_explanation))

### 2016-10-05 authors/list_latex.py
Remove the emails and funding statements from the authors footnotes
Emails do not appear to be included in other PeerJ articles. The funding
statements will be provided to PeerJ via a special funding statement section
when submitting.

@@ -20,8 +20,7 @@ with open(authors_tex, "w", encoding='utf-8') as f:
             institution_explanation = u" %s" % author["institution_explanation"]
         else:
             institution_explanation = u""
-        f.write(u"\\affil[%d]{%s (\\email{%s}).%s}%%\n" \
+        f.write(u"\\affil[%d]{%s.}%%\n" \
                 % (n+1,
                    author["institution_address_peerj"],
-                   author["email"],
-                   institution_explanation))
+                   ))

### 2016-10-05 authors/list_latex.py
Remove unused variable

@@ -16,10 +16,6 @@ with open(authors_tex, "w", encoding='utf-8') as f:
     for n, author in enumerate(author_list):
         f.write((u"\\author[%d]{%s}%%\n" % (n+1, author["name"])))
     for n, author in enumerate(author_list):
-        if "institution_explanation" in author:
-            institution_explanation = u" %s" % author["institution_explanation"]
-        else:
-            institution_explanation = u""
         f.write(u"\\affil[%d]{%s.}%%\n" \
                 % (n+1,
                    author["institution_address_peerj"],

### 2016-10-05 authors/list_latex.py
Make institution_address_peerj a list
This way, multiple institutions are separate footnotes. Also fix list_latex.py
to output the correct structure for this (with deduplicated footnotes).

@@ -13,10 +13,17 @@ author_list = load(open(authors_json))
 author_list = list(filter(lambda x: x["sympy_commits"] > 0, author_list))
 
 with open(authors_tex, "w", encoding='utf-8') as f:
-    for n, author in enumerate(author_list):
-        f.write((u"\\author[%d]{%s}%%\n" % (n+1, author["name"])))
-    for n, author in enumerate(author_list):
-        f.write(u"\\affil[%d]{%s.}%%\n" \
-                % (n+1,
-                   author["institution_address_peerj"],
-                   ))
+    all_addresses = []
+    for author in author_list:
+        addresses = author["institution_address_peerj"]
+        for address in addresses:
+            if address not in all_addresses:
+                all_addresses.append(address)
+
+    for author in author_list:
+        addresses = author["institution_address_peerj"]
+        address_n = ["%d" % (all_addresses.index(address) + 1) for address in addresses]
+        f.write((u"\\author[%s]{%s}%%\n" % (', '.join(address_n), author["name"])))
+
+    for n, address in enumerate(all_addresses):
+        f.write(u"\\affil[%d]{%s.}%%\n" % (n+1, address))

### 2016-10-05 authors/list_latex.py
Remove period from institution addresses
This is how PeerJ formats it.

@@ -26,4 +26,4 @@ with open(authors_tex, "w", encoding='utf-8') as f:
         f.write((u"\\author[%s]{%s}%%\n" % (', '.join(address_n), author["name"])))
 
     for n, address in enumerate(all_addresses):
-        f.write(u"\\affil[%d]{%s.}%%\n" % (n+1, address))
+        f.write(u"\\affil[%d]{%s}%%\n" % (n+1, address))

### 2016-10-06 upload_travis.sh
Upload the diff files in upload_travis.sh

@@ -21,17 +21,25 @@ cd upload
 if [ "${TRAVIS_PULL_REQUEST}" = "false" ]; then
     export name=paper-${TRAVIS_BRANCH}.pdf
     export namesupplement=paper-${TRAVIS_BRANCH}-supplement.pdf
+    export namediff=paper-diff-${TRAVIS_BRANCH}.pdf
+    export namediffsupplement=paper-diff-${TRAVIS_BRANCH}-supplement.pdf
     export msg="Upload for branch ${TRAVIS_BRANCH} - ${commit_id}"
 else
     export name=paper-${TRAVIS_PULL_REQUEST}.pdf
     export namesupplement=paper-${TRAVIS_PULL_REQUEST}-supplement.pdf
+    export namediff=paper-diff-${TRAVIS_PULL_REQUEST}.pdf
+    export namediffsupplement=paper-diff-${TRAVIS_REQUEST}-supplement.pdf
     export msg="Upload for sympy/sympy-paper#${TRAVIS_PULL_REQUEST} - ${commit_id}"
 fi
 
 mv ../paper.pdf ${name}
 mv ../supplement.pdf ${namesupplement}
+mv ../paper-diff.pdf ${namediff}
+mv ../supplement-diff.pdf ${namediffsupplement}
 git add ${name}
 git add ${namesupplement}
+git add ${namediff}
+git add ${namesupplementdiff}
 git commit -m "${msg} - [skip ci]."
 
 PUSH_COUNTER=0

### 2016-10-06 upload_travis.sh
Fix uploading of the supplement diff

@@ -39,7 +39,7 @@ mv ../supplement-diff.pdf ${namediffsupplement}
 git add ${name}
 git add ${namesupplement}
 git add ${namediff}
-git add ${namesupplementdiff}
+git add ${namediffsupplement}
 git commit -m "${msg} - [skip ci]."
 
 PUSH_COUNTER=0

### 2016-10-06 supplement.tex
This fixes for me broken refs in pdf index
(to the "references" sections)
Apply grammar fixes suggested by @certik at #208

@@
 \vskip-36pt%
-{\raggedright\sffamily\bfseries\fontsize{20}{25}\selectfont SymPy: Symbolic Computing in Python\par}%
+{\raggedright\sffamily\bfseries\fontsize{20}{25}\selectfont SymPy: symbolic computing in Python\par}%
 \vskip10pt
@@
 
-The supplementary material take a deeper look at certain topics in SymPy for
+The supplementary material takes a deeper look at certain topics in SymPy for
 which there was not enough room to discuss in the paper.
@@
 
+\phantomsection
 \bibliography{paper}
